### 2017-09-13 simulation_studies/fdr_simulation_studies.Rmd

---
title: "Benchmarking FDR Methods"
author: "Rafalab Journal Club Members"
date: "`r Sys.Date()`"
output: 
    html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
        code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

The goal of this working group is to benchmark recent covariate-adjusted FDR methods. For 
complete details, check out our [benchmark-fdr GitHub repo](https://github.com/pkimes/benchmark-fdr).

In this R Markdown, we use a Monte Carlo simulation study  to evaluate different simulation scenarios:

1. The `All nulls` simulation: All tests are simulated from the null distribution (no difference). 
2. add here... 
3. add here...
4. etc

In all cases, we define

* $m$ = number of tests
* $M$ = Monte Carlo replications
*

```{r}
m = 20000 # number of tests
M = 1000 # number of MC reps
```

**Things to do**: add a summary of all the simulation studies we plan to do.

## Methods to evaluate

- Bonferroni 
- BH
- IHW
- ASH
- BL
- Scott
- locfdr
- mixfdr

**Things to do**: add details, links, etc of methods here

## Metrics to evaluate

- Number of discoveries
- Control FDR at $\alpha$
- Power
- Relationship between $pi_0$ and covariate
- Assumptions: unimodal, etc

**Things to do**: Better describe metrics to evaluate our simulation studies

# The `All nulls` simulation

Simulate two standard normals with mean 0, sd = 1. Calculate t-statistic. 
Independent covariate: Uniform between 0, 1. Apply different methods.

## Control for FDR and FWER at $\alpha$

In this section, we will plot the number of discoveries 
as a function of $\alpha$ (should be 0). We will also evaluate
methods control for false discovery rate (FDR) and family-wise 
error rate (FWER) at nominal $\alpha$ level ranging from [0.01, 0.1].

```{r, load-libraries, message=FALSE}
# library(genefilter)
library(dplyr)
library(ggplot2)
library(doParallel)

# library(IHW)
# library(ashr)
# library(qvalue)
# library(swfdr)
  ## to use IHWpaper::scott_fdrreg wrapper, we need to install FDRreg v2.0 (only available github):   
  ## devtools::install_github(repo= 'jgscott/FDRreg', subdir='R_pkg/') 
  ## however, I couldn't get this to install. So I did the v1.0 via install.packages("FDRreg") on CRAN
# library(FDRreg)

# notes: convert this into R code
# - Scott: t-test statistic, independent covariate
#   - assumption of test statistic follow a standard normal. this is good here. 
# - locfdr: zz = t-test statistic, nulltype (try both). 
# - mixfdr: ? do we want to include? 
```

Additional packages that you should install are: 
```{r, eval=FALSE}
library(IHW)
library(ashr)
library(qvalue)
library(swfdr)
  ## to use IHWpaper::scott_fdrreg wrapper, we need to install FDRreg v2.0 (only available github):   
  ## devtools::install_github(repo= 'jgscott/FDRreg', subdir='R_pkg/') 
  ## however, I couldn't get this to install. So I did the v1.0 via install.packages("FDRreg") on CRAN
library(FDRreg)
```

Load some helper functions: 
```{r simulate-null-data-helper-functions}
du_ttest_sim <- function(m, pi0, effect_size, n_samples=20, n_groups = 2, seed=NULL){
  if (!is.null(seed)) set.seed(seed)
    m0 <- ceiling(m*pi0)
    false_nulls <- sample(1:m, m-m0)
    z_table <- matrix(rnorm(n_samples*m), ncol=n_samples)
    z_table[false_nulls,(n_samples/2+1):n_samples] <- matrix(rnorm(n_samples/2*(m-m0), effect_size), ncol=n_samples/2)
    H <- rep(0,m)
    H[false_nulls] <- 1
    gF <- factor(rep(1:n_groups, each=n_samples/n_groups))
    t_test <- genefilter::rowttests(z_table, gF)
    sds <- genefilter::rowSds(z_table) # pooled var reduces to unpooled var b/c same sample sizes across groups 
    SE <- sds / sqrt(n_samples)
    simDf <- data.frame(H = H, t_stat = t_test$statistic, t_dm = t_test$dm, 
                      t_pval = t_test$p.value, sd = sds, SE = SE)
    return(simDf)
}

calculate_test_stats <- function(sim, adj_p, alpha){
  rejected <- adj_p <= alpha
  rjs <- sum(rejected) # number of rejected nulls
  false_rjs <- sum(sim$H == 0 & rejected) # number of false rejected nulls
  rj_ratio <- rjs/nrow(sim)
  FDP <- ifelse(rjs == 0, 0, false_rjs / rjs) # False Discovery Proportion
  power <- ifelse(sum(sim$H) == 0, NA, 
                  sum(sim$H == 1 & rejected) / sum(sim$H == 1)) # power
  FPR <- sum(sim$H == 0 & rejected) / sum(sim$H == 0)
 	FWER <- as.numeric(false_rjs > 0)
	df <- data.frame(n_rejects = rjs, FWER=FWER, FPR=FPR, 
	                 FDP=FDP, alpha = alpha) #  power=power, 
	return(df)
 }

# adapted from IHWpaper::scott_fdrreg()
scott_fdrreg_hickswrapper <- function(unadj_p, filterstat, df=3)
  {
	   # no automated way to choose function space over which we optimize
	   # so we just use bs(df=3) as done in their analysis
     b <- splines::bs(filterstat, df=df)

     Xs <- model.matrix( ~  b - 1)
     fdrreg_res <- FDRreg::FDRreg(z=qnorm(unadj_p), covars=Xs) # assumption of test statistic follow a standard normal
     adj_p <- fdrreg_res$FDR
     return(adj_p)
  }

```

Simulate an example "null" dataset. `H` is an indicator representing the true null (0) or alternative status (1). `t_stat` is the t-test statistic, `t_dm` is the difference in means, 
`t_pval` is the $p$-value, `sd` is the pooled standard deviation ($\sigma$), which reduces to the unpooled sd in this case, 
and `SE` is the standard error ($\frac{\sigma}{\sqrt{n}}$)
```{r example-simulate-null-data}
n_samples = 20
n_groups = 2

sim_df <- du_ttest_sim(m=m, pi0=1, effect_size=0, 
                       n_samples=n_samples, n_groups = 2)
head(sim_df)
```

Run following on odyssey and save output as RDS: 
```{r evaluate-methods-null-data, eval=FALSE}
alphas = seq(0.01, 0.1, length.out = 10)

nCores <- 4
registerDoParallel(cores = nCores)
workers <- getDoParWorkers()
backend <- getDoParName()
version <- getDoParVersion()

df.out <- foreach(i = 1:M) %dopar% 
    {        
        sim_df <- du_ttest_sim(m=m, pi0=1, effect_size=0, n_samples=n_samples, n_groups=2)

        # Bonferroni 
        adj_p <- p.adjust(sim_df$t_pval, method="bonferroni")
        df.bonf <- plyr::ldply(alphas, function(a){
                      calculate_test_stats(sim=sim_df, adj_p = adj_p, alpha = a) })
        # BH
        adj_p <- p.adjust(sim_df$t_pval, method="BH")
        df.bh <- plyr::ldply(alphas, function(a){
                      calculate_test_stats(sim=sim_df, adj_p = adj_p, alpha = a) })

        # qvalue (Storey)
        adj_p <- qvalue::qvalue(p=sim_df$t_pval)$qvalues
        df.qvalue <- plyr::ldply(alphas, function(a){
                      calculate_test_stats(sim=sim_df, adj_p = adj_p, alpha = a) })

        # IHW (Wolfgang)
        ihw_fdr <- IHW::ihw(pvalues = sim_df$t_pval, covariates = sim_df$sd, alpha =  0.1)
        adj_p <- IHW::adj_pvalues(ihw_fdr)
        df.ihw <- plyr::ldply(alphas, function(a){
                      calculate_test_stats(sim=sim_df, adj_p = adj_p, alpha = a) })

        # ash (Stephens)
        beta.ash = ashr::ash(betahat = sim_df$t_dm, sebetahat = sim_df$SE)
        adj_p <- ashr::get_svalue(beta.ash)
        df.ash <- plyr::ldply(alphas, function(a){
                      calculate_test_stats(sim=sim_df, adj_p = adj_p, alpha = a) })

        # swfdr::lm_pi0 (Boca-Leek) **think about multiple covariates** or **correlation structure**
        pi0x <- swfdr::lm_pi0(pValues = sim_df$t_pval, X = sim_df$sd, smooth.df =3)
        adj_p <- p.adjust(sim_df$t_pval, method="BH")*pi0x$pi0
        df.BL <- plyr::ldply(alphas, function(a){
                      calculate_test_stats(sim=sim_df, adj_p = adj_p, alpha = a) })

        # Scott et al. (2015) (FDRreg)
        adj_p <- scott_fdrreg_hickswrapper(unadj_p = sim_df$t_pval, filterstat = sim_df$sd, df=3)
        df.scott <- plyr::ldply(alphas, function(a){
                      calculate_test_stats(sim=sim_df, adj_p = adj_p, alpha = a) })

        # locfdr (Cai)

        # Summary table of FDP for each method at each alpha
        rbind(data.frame(df.bonf, "method" = "bonferroni"), 
              data.frame(df.bh, "method" = "bh"), 
              data.frame(df.qvalue, "method" = "qvalue"), 
              data.frame(df.ihw, "method" = "ihw"), 
              data.frame(df.ash, "method" = "ash"), 
              data.frame(df.BL, "method" = "boca-leek"), 
              data.frame(df.scott, "method" = "scott"))
  }

df.out <- dplyr::bind_rows(df.out)
saveRDS(df.out, file=file.path(workingPath, "nullSims-Results.RDS"))
```

Load and plot results from RDS file: 
```{r, eval=FALSE}
workingPath <- "/net/irizarryfs01/srv/export/irizarryfs01_backed_up/share_root/shicks/projects/benchmark-fdr"
readRDS(file=file.path(workingPath, "nullSims-Results.RDS"))

# Number of discoveries as a function of $\alpha$
df.out %>% dplyr::group_by(method, alpha) %>% 
      dplyr::summarize(n_rejects = mean(n_rejects), FDR = mean(FDP), 
                       FWER = mean(FWER), FPR = mean(FPR)) %>% 
    ggplot(aes(x = alpha, y = n_rejects, color = method)) + 
         geom_line() + geom_abline(linetype="dashed") + 
         xlab(expression(bold(paste("Nominal ",alpha)))) + ylab("Number of Discoveries") +
         theme(axis.title = element_text(face="bold") )

# False discovery rate as a function of $\alpha$
df.out %>% dplyr::group_by(method, alpha) %>% 
      dplyr::summarize(n_rejects = mean(n_rejects), FDR = mean(FDP), 
                       FWER = mean(FWER), FPR = mean(FPR)) %>% 
      ggplot(aes(x = alpha, y = FDR, color = method)) + 
         geom_line() + geom_abline(linetype="dashed") + 
         xlab(expression(bold(paste("Nominal ",alpha)))) + ylab("FDR") +
         theme(axis.title = element_text(face="bold") )

# Family wise error rate as a function of $\alpha$
df.out %>% dplyr::group_by(method, alpha) %>% 
      dplyr::summarize(n_rejects = mean(n_rejects), FDR = mean(FDP), 
                       FWER = mean(FWER), FPR = mean(FPR)) %>% 
      ggplot(aes(x = alpha, y = FWER, color = method)) + 
         geom_line() + geom_abline(linetype="dashed") + 
         xlab(expression(bold(paste("Nominal ",alpha)))) + ylab("FWER") +
         theme(axis.title = element_text(face="bold") )
```

## add here...