### 2017-09-15 simulation_studies/fdr_sims_alt.Rmd
Draft setup for analysis of single sim

---
title: "Benchmarking FDR Methods - Alternative Settings"
author: "Rafalab Journal Club Members"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    highlight: tango
    number_sections: true
    code_folding: hide
---

<!-- ######################################################################## -->
<!-- preliminaries                                                            --> 
<!-- ######################################################################## -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, install-benchmarked-methods, eval=FALSE}
## from CRAN
install.packages("ashr")

## from Bioconductor
BiocInstaller::biocLite("IHW")
BiocInstaller::biocLite("qvalue")
BiocInstaller::biocLite("swfdr")

## from GitHub
devtools::install_github('jgscott/FDRreg', subdir="R_pkg/")
devtools::install_github("nignatiadis/IHWpaper")
```

<!-- ######################################################################## -->
# Workspace Setup
<!-- ######################################################################## -->

```{r, load-libraries, message=FALSE}
library("genefilter")
library("dplyr")
library("ggplot2")
library("doParallel")

library("IHW")
library("ashr")
library("qvalue")
library("swfdr")
library("FDRreg")
library("IHWpaper")
```

```{r, load-helper-functions}
source("R/simulation-helpers.R")
```

```{r, set-location}
projdir <- "/n/irizarryfs01_backed_up/pkimes/project/project-02-benchmark-fdr"
```

<!-- ######################################################################## -->
# Overview
<!-- ######################################################################## -->

- Alternative Setttings
- Single Monte Carlo Simulations

In the alternative simulations, we consider:

- different magnitudes of effect sizes / difference in means
- different shapes of effect sizes / difference in means 
  - independent tests: (Boca-Leek): beta, normal, t, chi2; (ASH): spiky, near-normal, flat-top, skew, bimodal
  - correlated tests: (Boca-Leek): mutivar-normal, multivar-t (blocked covariance design)
- different sample sizes (e.g. impact on SE estimates)
  - same sample size for all tests
  - different sample sizes for some tests (i.e. for tests w/ different precision levels)
- different in relationship between pi_0 and independent covariate
- different correlation structure of multiple covariates
- different global $\pi_0$ values (between 0, 1) - setting 1 in Boca-Leek

```{r, define-base}
## setting 0: (base) null simulation setting
setting_base <- list(m = 2000, pi0 = 1, effect_size = 0, n_samples = 20, n_groups = 2)
```

<!-- ######################################################################## -->
# Alternative Setting 1: Varying Null Proportion ($\pi_0$)
<!-- ######################################################################## -->

First we define the set of simulation settings to be tested.

```{r, define-setting-1}
## create settings w/ varying pi0 (null proportions)
setting_alts <- lapply(seq(0, 1, by=.1),
                       function(p) { replace(setting_base,
                                             c("pi0", "effect_size"),
                                             c(p, 1.0))
                       })
names(setting_alts) <- paste0("altp0_", 0:10)

## verify settings are as expected
do.call(rbind, setting_alts)
```

```{r, run-setting-1}
## simulate single datasets for each setting
sim_dfl <- lapply(setting_alts, do.call, what=du_ttest_sim)

## calculate adjusted p-values and return pvals along w/ mets
adj_dfl <- lapply(sim_dfl, sim_runner, alphas=seq(0.01, 0.10, by=.01), pvals=TRUE)

## create 
sim_res <- cbind(sim_dfl$alteff_20, bb$alteff_20)
```

<!-- ------------------------------------------------------------------------ -->
## Visual Inspection of Single Monte Carlo Simulation
<!-- ------------------------------------------------------------------------ -->

```{r, ASH-svals-vs-qvals}
## plot ASH s-values against ASH q-values
gp <- ggplot(sim_res) +
    geom_point(aes(ashs, ashq), alpha=1/4) +
    expand_limits(x=c(0, 1), y=c(0, 1)) +
    theme_bw()
ggsave("plots/sims-01-null-ash_s_vs_q.png", gp, width=5, height=5)
```

```{r, effect-vs-adjp}
## base of plot for: 'effect size' vs. 'adjusted pval (method X)'
gp_base <- ggplot(sim_res, aes(x=effect_size, 
                               color=cut_width(SE, width=.05, boundary=0))) +
    expand_limits(y=c(0, 1)) + 
    scale_color_brewer("Binned StdErr", palette="Set1") + 
    theme_bw()

## 'effect size' vs. t-test p-values
gp0 <- gp_base + 
    geom_point(aes(y=pval), alpha=1/2) +
    ggtitle("dist. of unadjusted p-values (20k null tests)")

## 'effect size' vs. ASH adjusted p-values
gp1 <- gp_base + 
    geom_point(aes(y=ashs), alpha=1/2) +
    ggtitle("dist. of ASH s-values (20k null tests)")

## 'effect size' vs. IHW adjusted p-values
gp3 <- gp_base +
    geom_point(aes(y=ihw), alpha=1/2) + 
    ggtitle("dist. of IHW adjusted p-values (20k null tests)")

## 'effect size' vs. Boca-Leek adjusted p-values
gp4 <- gp_base +
    geom_point(aes(y=bl), alpha=1/2) +
    ggtitle("dist. of Boca-Leek adjusted p-values (20% alt tests; e.size=2)")

## plot together
png("plots/sims-01-null-ash.png", units="in", width=6, height=7, res=200)
pkimes::ggmulti(gp1, gp2, gp3, gp4, cols=2)
dev.off()
```

```{r}
gp <- ggplot(sim_res) +
    geom_point(aes(x=pval, y=ashs), alpha=1/10) +
    theme_bw() +
    xlab("unadjusted p-value") +
    ylab("ASH s-value")
ggsave("plots/sims-01-null-pval_vs_ash.png", gp, width=5, height=5)
```

```{r}
## distribution of rejected/non-rejected; sort reverse 
ggplot(sim_res) +
    geom_histogram(aes(x=qnorm(pval)*sign(test_statistic),
                       fill=cut_width(ashs, width=.05, boundary=0) %>%
                           factor(., levels=rev(levels(.)))),
                   color='black', binwidth=.1, boundary=0,
             position="stack") +
    scale_fill_brewer("ASH (s-values)", palette="Blues", direction=-1) +
    theme_bw()
```

<!-- ------------------------------------------------------------------------ -->
## Numerically Evaluate of Single Monte Carlo Simulation
<!-- ------------------------------------------------------------------------ -->

Analysis of metrics calculated for various FDR cutoffs.

<!-- ######################################################################## -->
# Alternative Setting 2: Varying Effect Size
<!-- ######################################################################## -->

First we define the set of simulation settings to be tested.

```{r}
## create settings w/ varying effect_size (mean difference)
setting_alts <- lapply(seq(0, 2, by=.2),
                       function(x) { replace(setting_base,
                                             c("pi0", "effect_size"),
                                             c(0.8, x))
                       })
names(setting_alts) <- paste0("alteff_", seq(0, 20, by=2))

## verify settings are as expected
do.call(rbind, setting_alts)
```

<!-- ------------------------------------------------------------------------ -->
## Visual Inspection of Single Monte Carlo Simulation
<!-- ------------------------------------------------------------------------ -->

<!-- ------------------------------------------------------------------------ -->
## Numerically Evaluate of Single Monte Carlo Simulation
<!-- ------------------------------------------------------------------------ -->

<!-- ######################################################################## -->
# Alternative Setting 3:
<!-- ######################################################################## -->