### 2017-09-15 datasets/ChIP-seq/ChIP-seq.Rmd
ChIP-seq

---
title: "ChIP-seq datasets for FDR benchmarking"
author: "Mingxiang Teng"
date: "9/14/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Summary

For FDR benchmark using ChIP-seq datasets, we may focus on peak calling and
differential peaks. Peak calling is the most common way that people analyze 
ChIP-seq since this technique was invented, though some groups started to care
about differential peaks in some complex experiment designs. For peak calling,
peak callers always report effect size (they call "enrichment signal") 
and p-value, but standard error. I think it is too complicate if we dig into
their code for generating standard error. For differential peaks, classical
pipeline would call peaks first, then recruit existing algorithms (*e.g.* DESeq2)
to test the covariant of interest, in a way that all effect size, p-value and 
standard error can be generated. However, this pipeline probably gives 
very similar patterns (distributions) of effect size / p-value to those
based on differential gene expression analysis. Thus, differential peak 
analysis might not be useful in our FDR benchmark.

### Oestrogen receptor binding case

This dataset have plenty of replicates so that we can focus on both peak
calling and differential peaks. And it is the first dataset people used to 
illustrate differential peak detection, where the tools DiffBind was 
introduced.

We can download the processed data from GEO.

```{bash, chipseq-download, eval=FALSE}
curl -O "https://www.ncbi.nlm.nih.gov/geo/download/?acc=GSE32222&format=file"
tar -xvf "?acc=GSE32222&format=file"
rm "?acc=GSE32222&format=file"
```

For one sample, extract the peak calling results.

```{bash, chipseq-unzip1, eval=FALSE}
gunzip "GSM798383_SLX-1201.250.s_4_SLX-1202.250.s_1_peaks.txt.gz"
rm *peaks.txt.gz
```

And load in to R.
```{r, chipseq-read1}
peaks <- read.table("~/GSM798383_SLX-1201.250.s_4_SLX-1202.250.s_1_peaks.txt", header=TRUE, sep='\t')
dim(peaks)
head(peaks)
```

Here, "X.10.log10.pvalue." is the -log10P, and fold_enrichment is the effect size.