### 2017-09-15 simulation_studies/fdr_sims_alt.Rmd
Add better plots

---
title: "Benchmarking FDR Methods - Alternative Settings"
author: "Rafalab Journal Club Members"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    highlight: tango
    number_sections: true
    code_folding: hide
---

<!-- ######################################################################## -->
<!-- preliminaries                                                            --> 
<!-- ######################################################################## -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, install-benchmarked-methods, eval=FALSE}
## from CRAN
install.packages("ashr")

## from Bioconductor
BiocInstaller::biocLite("IHW")
BiocInstaller::biocLite("qvalue")
BiocInstaller::biocLite("swfdr")

## from GitHub
devtools::install_github('jgscott/FDRreg', subdir="R_pkg/")
devtools::install_github("nignatiadis/IHWpaper")
```

<!-- ######################################################################## -->
## Workspace Setup
<!-- ######################################################################## -->

```{r, load-libraries, message=FALSE}
library("genefilter")
library("dplyr")
library("data.table")
library("ggplot2")

library("IHW")
library("ashr")
library("qvalue")
library("swfdr")
library("FDRreg")
library("IHWpaper")
```

```{r, load-helper-functions}
source("R/simulation-helpers.R")
```

```{r, set-location}
projdir <- "/n/irizarryfs01_backed_up/pkimes/project/project-02-benchmark-fdr"
```

<!-- ######################################################################## -->
## Overview
<!-- ######################################################################## -->

We first investigate a broad range of settings with only a **single Monte Carlo replication** under each condition. In each simulation, a large set of univariate two-sample $t$-tests are carried out based on raw data sampled from two mean-shifted Gaussian distributions. Each method for controlling FDR is applied to the results of the $t$-tests, and performance is compared using metrics include: FPR and power.

A range of simulations are considered by varying the following baseline parameters.

```{r, define-base}
## setting 0: (base) null simulation setting
setting_base <- list(m = 20000,         # number of hypothesis tests
                     pi0 = 1,           # proportion of null hypotheses
                     effect_size = 0,   # expected mean diff of non-null tests
                     n_samples = 20,    # total number of samples
                     n_groups = 2)      # number of groups in contrast
```

At a high level, we consider the following set of simulation settings:

- **no informative covariate**
  - *varying null proportion*
  - *varying sample size*
  - *varying effect size* (constant 'effect size')
  - *varying effect size* (random 'effect size')
    - independent tests: Beta, Normal, t, chi2, spiky, near-normal, flat-top, bimodal, skewed 
    - dependent tests: multivariate-Normal, multivariate-t
- **informative covariate**
  - *varying null proportion*
  - *varying sample size*
  - *varying effect size* (constant 'effect size')
  - *varying effect size* (random 'effect size')
    - independent tests: Beta, Normal, t, chi2, spiky, near-normal, flat-top, bimodal, skewed 
    - dependent tests: multivariate-Normal, multivariate-t
  - *varying covariate, pi0 relation*
    - single informative covariate
    - multiple informative covariates

In this analysis, we compare the following methods. For methods which accept user-specified covariates, under the "no informative covariate" settings, the pooled standard deviation was used specified to the methods as a dummy covariate.

```{r, define-plot-variables}
## string abb for each method (should match output of `sim_runner`)
methods_str <- c("pval", "bh", "qvalue", "ihw",
                 "ashs", "bl", "lfdr", "scott")

## description of each method
methods_desc <- list(pval = "unadjusted p-values",
                     bh = "BH adjusted p-values",
                     qvalue = "Storey's q-value",
                     ihw = "IHW adjusted p-values",
                     ashs = "ASH s-values",
                     bl = "Boca-Leek adjusted p-values",
                     lfdr = "local FDR",
                     scott = "FDRreg adjusted p-values")
```

<!-- ######################################################################## -->
## Setting 1.1: No Informative Covariate, Varying Null Proportion ($\pi_0$)
<!-- ######################################################################## -->

First we define the set of simulation settings to be tested. A range of $\pi_0$ values are considered between 0 and 1. For each setting, a fixed effect size (group mean differences) of `1.5` was used in the simulation of all non-null (alternative) tests.

```{r, define-setting-1}
## create settings w/ varying pi0 (null proportions)
setting_alts <- lapply(seq(0, 1, by=.1),
                       function(p) { replace(setting_base,
                                             c("pi0", "effect_size"),
                                             c(p, 1.5))
                       })
names(setting_alts) <- paste0("altp0_", 0:10)

## verify settings are as expected
ptab <- do.call(rbind.data.frame, setting_alts)
ptab$setting <- rownames(ptab)
setnames(ptab, "effect_size", "alt_effsize") 
ptab[] <- lapply(ptab, as.character)
ptab
```

Next, we perform one iteration of simulation for each setting and keep both the adjusted p-values (`pvals`) and the evaluation metrics (`stats`) computed for each method.

```{r, run-setting-1}
## simulate single datasets, calculate adj p-values
sim_dfl <- lapply(setting_alts, do.call, what=du_ttest_sim)
adj_dfl <- lapply(sim_dfl, sim_runner, alphas=seq(0.01, 0.10, by=.01), pvals=TRUE)

## combine results, settings for analysis
sim_pv <- mapply(function(x, y) { cbind(x, y$pvals) }, sim_dfl, adj_dfl, SIMPLIFY=FALSE)
sim_pv <- data.table::rbindlist(sim_pv, idcol="setting")
sim_pv <- merge(sim_pv, ptab, by="setting", all.x=TRUE)

sim_st <- lapply(adj_dfl, `[[`, "stats")
sim_st <- data.table::rbindlist(sim_st, fill=TRUE, idcol="setting")
sim_st <- merge(sim_st, ptab, by="setting", all.x=TRUE)
```

<!-- ------------------------------------------------------------------------ -->
### Visual Inspection of Single Monte Carlo Simulation
<!-- ------------------------------------------------------------------------ -->

All figures generated in this section will be saved with the following prefix.

```{r, define-plot-prefix-1}
plt_pfix <- "plots/sims-01_01-"
```

First, we take a look at the distribution of adjusted p-values as a function of the observed effect sizes over the `r setting_base$m` hypothesis tests. Each test is also colored by the corresponding standard error across the `r setting_base$n_samples` samples in the simulation. We plot this distribution across the values of pi0.

```{r, plot-effect-vs-adjp, fig.width=14, fig.height=10}
## set ID columns to keep when melting table for plot 
plotvars <- c("effect_size", "SE", "pi0")
pi0set <- c("0", "0.1", "0.3", "0.7", "0.9", "1")

## generate plot for subset of pi0 values
gp_all <- ggplot(sim_pv %>%
                 melt(id.vars=plotvars, measure.vars=methods_str,
                      variable.name='method', value.name='adjp') %>%
                 filter(pi0 %in% pi0set) %>%
                 mutate(method = plyr::revalue(method, unlist(methods_desc)),
                        pi0 = paste("pi0 =", pi0))) + 
    geom_point(aes(x=effect_size, y=adjp,
                   color=cut_width(SE, width=.05, boundary=0)),
               alpha=1/2) +
    scale_color_brewer("Binned StdErr", palette="Set1") + 
    expand_limits(y=c(0, 1)) +
    ylab("adjusted p-value") + 
    facet_grid(pi0 ~ method) + 
    theme_bw()
gp_all

## save to file
ggsave(paste0(plt_pfix, "effectsize_vs_adjustedp.png"), gp_all, width=14, height=10)
```

Next, we generate a similar plot, but look at the distribution of significant p-values as a function of the z-score, as was done in the original ASH paper. This was originally motivated by an interest in illustrating the effect of the unimodal assumption proposed with the ASH method.

```{r, plot-zscale-vs-adjp, fig.width=14, fig.height=10}
## set ID columns to keep when melting table for plot 
plotvars <- c("effect_size", "SE", "pi0", "pval", "test_statistic")
pi0set <- c("0", "0.1", "0.3", "0.7", "0.9", "1")

## helper; bin adjusted p-vals, flip order of bins
.p2disc <- function(x) {
    cut(x, breaks=c(0, .01, .05, .10, .20, 1), include.lowest=TRUE) %>%
        factor(., levels=rev(levels(.)))
}

## similar plot to Figure 1 of ASH paper
gp_all <- ggplot(sim_pv %>%
                 melt(id.vars=plotvars, measure.vars=methods_str,
                      variable.name='method', value.name='adjp') %>%
                 filter(pi0 %in% pi0set) %>%
                 mutate(method = plyr::revalue(method, unlist(methods_desc)),
                        pi0 = paste("pi0 =", pi0))) + 
    geom_histogram(aes(fill=.p2disc(adjp),
                       x=qnorm(pval)*sign(test_statistic)),
                   color='black', size=1/4, binwidth=.2, boundary=0,
                   position="stack") +
    xlab("z-score") + 
    viridis::scale_fill_viridis("Binned\nadj p-val", discrete=TRUE,
                                option="plasma", direction=-1,
                                begin=.1, end=.9, drop=FALSE) +
    facet_grid(pi0 ~ method) + 
    theme_bw()
gp_all

## save to file
ggsave(paste0(plt_pfix, "zscore_vs_adjustedp.png"), gp_all, width=14, height=10)
```

A similar plot, is simply to look at the pairwise plots of original p-values against the adjusted p-values generated by each method

```{r, plot-p-vs-adjp, fig.width=14, fig.height=10}
## set ID columns to keep when melting table for plot 
plotvars <- c("SE", "H", "pi0", "pval")
pi0set <- c("0", "0.1", "0.3", "0.7", "0.9", "1")

## color by SE
gp_all <- ggplot(bb <- sim_pv %>%
                 melt(id.vars=plotvars, measure.vars=methods_str,
                      variable.name='method', value.name='adjp') %>%
                 filter(pi0 %in% pi0set) %>%
                 mutate(method = plyr::revalue(method, unlist(methods_desc)),
                        pi0 = paste("pi0 =", pi0))) + 
    geom_point(aes(x=pval, y=adjp,
                   color=cut_width(SE, width=.05, boundary=0)),
               alpha=1/10) +
    scale_color_brewer("Binned StdErr", palette="Set1") + 
    expand_limits(x=c(0, 1), y=c(0, 1)) +
    xlab("unadjusted p-value") + 
    ylab("adjusted p-value") + 
    facet_grid(pi0 ~ method) + 
    theme_bw()
gp_all

## save to file
ggsave(paste0(plt_pfix, "pvalue_vs_adjustedp_bySE.png"), gp_all, width=14, height=10)

## color by H: 0/1
gp_all <- ggplot(bb <- sim_pv %>%
                 melt(id.vars=plotvars, measure.vars=methods_str,
                      variable.name='method', value.name='adjp') %>%
                 filter(pi0 %in% pi0set) %>%
                 mutate(method = plyr::revalue(method, unlist(methods_desc)),
                        pi0 = paste("pi0 =", pi0))) + 
    geom_point(aes(x=pval, y=adjp, color=factor(H)), alpha=1/10) +
    scale_color_brewer("null|alt", palette="Set1") + 
    expand_limits(x=c(0, 1), y=c(0, 1)) +
    xlab("unadjusted p-value") + 
    ylab("adjusted p-value") + 
    facet_grid(pi0 ~ method) + 
    theme_bw()
gp_all

## save to file
ggsave(paste0(plt_pfix, "pvalue_vs_adjustedp_byH.png"), gp_all, width=14, height=10)
ggsave(paste0(plt_pfix, "pvalue_vs_adjustedp_byH_zoom.png"),
       gp_all + coord_cartesian(x=c(0, .1), y=c(0, .1)),
       width=14, height=10)
```

<!-- ------------------------------------------------------------------------ -->
### Numeric Evaluate of Single Monte Carlo Simulation
<!-- ------------------------------------------------------------------------ -->

To be completed.

<!-- ######################################################################## -->
## Setting 1.2: No Informative Covariate, Varying Effect Size (constant)
<!-- ######################################################################## -->

First we define the set of simulation settings to be tested. A range of (constant) effect sizes are considered between 0 and 2. For each setting, a fixed null proportion of `pi0 = 0.2` was used. More complex settings with non-null effect sizes sampled from various distirbutions are considered in a later section.

```{r, define-setting-2}
## create settings w/ varying effect_size (mean difference)
setting_alts <- lapply(seq(0, 2, by=.2),
                       function(x) { replace(setting_base,
                                             c("pi0", "effect_size"),
                                             c(0.8, x))
                       })
names(setting_alts) <- paste0("alteff_", seq(0, 20, by=2))

## verify settings are as expected
ptab <- do.call(rbind.data.frame, setting_alts)
ptab$setting <- rownames(ptab)
setnames(ptab, "effect_size", "alt_effsize") 
ptab[] <- lapply(ptab, as.character)
ptab
```

Next, we perform one iteration of simulation for each setting and keep both the adjusted p-values (`pvals`) and the evaluation metrics (`stats`) computed for each method.

```{r, run-setting-2}
## simulate single datasets, calculate adj p-values
sim_dfl <- lapply(setting_alts, do.call, what=du_ttest_sim)
adj_dfl <- lapply(sim_dfl, sim_runner, alphas=seq(0.01, 0.10, by=.01), pvals=TRUE)

## combine results, settings for analysis
sim_pv <- mapply(function(x, y) { cbind(x, y$pvals) }, sim_dfl, adj_dfl, SIMPLIFY=FALSE)
sim_pv <- data.table::rbindlist(sim_pv, idcol="setting")
sim_pv <- merge(sim_pv, ptab, by="setting", all.x=TRUE)

sim_st <- lapply(adj_dfl, `[[`, "stats")
sim_st <- data.table::rbindlist(sim_st, fill=TRUE, idcol="setting")
sim_st <- merge(sim_st, ptab, by="setting", all.x=TRUE)
```

<!-- ------------------------------------------------------------------------ -->
### Visual Inspection of Single Monte Carlo Simulation
<!-- ------------------------------------------------------------------------ -->

All figures generated in this section will be saved with the following prefix.

```{r, define-plot-prefix-2}
plt_pfix <- "plots/sims-01_02-"
```

First, we take a look at the distribution of adjusted p-values as a function of the observed effect sizes over the `r setting_base$m` hypothesis tests. Each test is also colored by the corresponding standard error across the `r setting_base$n_samples` samples in the simulation. We plot this distribution across the values of `effect_size`.

```{r, plot-effect-vs-adjp-2, fig.width=14, fig.height=10}
## set ID columns to keep when melting table for plot 
plotvars <- c("effect_size", "H", "alt_effsize")
es0set <- c("0.2", "0.4", "0.8", "1.2", "1.6", "2")

## generate plot for subset of pi0 values
gp_all <- ggplot(sim_pv %>%
                 melt(id.vars=plotvars, measure.vars=methods_str,
                      variable.name='method', value.name='adjp') %>%
                 filter(alt_effsize %in% es0set) %>%
                 mutate(method = plyr::revalue(method, unlist(methods_desc)),
                        alt_effsize = paste("eff size =", alt_effsize))) + 
    geom_point(aes(x=effect_size, y=adjp, color=factor(H)), alpha=1/10) +
    scale_color_brewer("null|alt", palette="Set1") + 
    expand_limits(y=c(0, 1)) +
    ylab("p/q/s-value") + 
    facet_grid(alt_effsize ~ method) + 
    theme_bw()
gp_all

## save to file
ggsave(paste0(plt_pfix, "effectsize_vs_adjustedp.png"), gp_all, width=14, height=10)
```

Next, we generate a similar plot, but look at the distribution of significant p-values as a function of the z-score, as was done in the original ASH paper. This was originally motivated by an interest in illustrating the effect of the unimodal assumption proposed with the ASH method.

```{r, plot-zscale-vs-adjp-2, fig.width=14, fig.height=10}
## set ID columns to keep when melting table for plot 
plotvars <- c("H", "alt_effsize", "pval", "test_statistic")
es0set <- c("0.2", "0.4", "0.8", "1.2", "1.6", "2")

## helper; bin adjusted p-vals, flip order of bins
.p2disc <- function(x) {
    cut(x, breaks=c(0, .01, .05, .10, .20, 1), include.lowest=TRUE) %>%
        factor(., levels=rev(levels(.)))
}

## similar plot to Figure 1 of ASH paper
gp_all <- ggplot(sim_pv %>%
                 melt(id.vars=plotvars, measure.vars=methods_str,
                      variable.name='method', value.name='adjp') %>%
                 filter(alt_effsize %in% es0set) %>%
                 mutate(method = plyr::revalue(method, unlist(methods_desc)),
                        alt_effsize = paste("eff size =", alt_effsize))) + 
    geom_histogram(aes(fill=.p2disc(adjp),
                       x=qnorm(pval)*sign(test_statistic)),
                   color='black', size=1/4, binwidth=.2, boundary=0,
                   position="stack") +
    xlab("z-score") + 
    viridis::scale_fill_viridis("Binned\nadj p-val", discrete=TRUE,
                                option="plasma", direction=-1,
                                begin=.1, end=.9, drop=FALSE) +
    facet_grid(alt_effsize ~ method) + 
    theme_bw()
gp_all

## save to file
ggsave(paste0(plt_pfix, "zscore_vs_adjustedp.png"), gp_all, width=14, height=10)
```

A similar plot, is simply to look at the pairwise plots of original p-values against the adjusted p-values generated by each method

```{r, plot-p-vs-adjp-1, fig.width=14, fig.height=10}
## set ID columns to keep when melting table for plot 
plotvars <- c("SE", "H", "alt_effsize", "pval")
es0set <- c("0.2", "0.4", "0.8", "1.2", "1.6", "2")

## color by SE
gp_all <- ggplot(bb <- sim_pv %>%
                 melt(id.vars=plotvars, measure.vars=methods_str,
                      variable.name='method', value.name='adjp') %>%
                 filter(alt_effsize %in% es0set) %>%
                 mutate(method = plyr::revalue(method, unlist(methods_desc)),
                        alt_effsize = paste("eff size =", alt_effsize))) + 
    geom_point(aes(x=pval, y=adjp,
                   color=cut_width(SE, width=.05, boundary=0)),
               alpha=1/10) +
    scale_color_brewer("Binned StdErr", palette="Set1") + 
    expand_limits(x=c(0, 1), y=c(0, 1)) +
    xlab("unadjusted p-value") + 
    ylab("adjusted p-value") + 
    facet_grid(alt_effsize ~ method) + 
    theme_bw()
gp_all

## save to file
ggsave(paste0(plt_pfix, "pvalue_vs_adjustedp_bySE.png"), gp_all, width=14, height=10)
ggsave(paste0(plt_pfix, "pvalue_vs_adjustedp_bySE_zoom.png"),
       gp_all + coord_cartesian(x=c(0, .1), y=c(0, .1)),
       width=14, height=10)

## color by H: 0/1
gp_all <- ggplot(bb <- sim_pv %>%
                 melt(id.vars=plotvars, measure.vars=methods_str,
                      variable.name='method', value.name='adjp') %>%
                 filter(alt_effsize %in% es0set) %>%
                 mutate(method = plyr::revalue(method, unlist(methods_desc)),
                        alt_effsize = paste("eff size =", alt_effsize))) + 
    geom_point(aes(x=pval, y=adjp, color=factor(H)), alpha=1/10) +
    scale_color_brewer("null|alt", palette="Set1") + 
    expand_limits(x=c(0, 1), y=c(0, 1)) +
    xlab("unadjusted p-value") + 
    ylab("adjusted p-value") + 
    facet_grid(alt_effsize ~ method) + 
    theme_bw()
gp_all

## save to file
ggsave(paste0(plt_pfix, "pvalue_vs_adjustedp_byH.png"), gp_all, width=14, height=10)
ggsave(paste0(plt_pfix, "pvalue_vs_adjustedp_byH_zoom.png"),
       gp_all + coord_cartesian(x=c(0, .1), y=c(0, .1)),
       width=14, height=10)
```

<!-- ------------------------------------------------------------------------ -->
### Numeric Evaluation of Single Monte Carlo Simulation
<!-- ------------------------------------------------------------------------ -->

To be completed.