### 2017-09-28 datasets/GWAS/GWAS.Rmd
Add UpSet plots for GWAS results (Issue #13)

---
title: "GWAS datasets for FDR benchmarking"
author: "Keegan Korthauer"
date: "9/28/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Summary

Here are the GWAS datasets we will use to benchmark the FDR methods. For each one
we provide the script to download the data and read into R. We'd like to verify 
that each dataset contains the desired inputs:

- p-value
- effect size
- standard error
- additional covariates (optional)

Note that while running all the bash code in this document will download the data, 
some of it doesn't seem to work properly with knitting (e.g. the ftp download 
in the second dataset), and I'm not committing 
the data itself to this repo. Run the bash code in the terminal separately to 
download the data yourself, and see the knitted `.html` document if you do not 
wish to download the datasets but still wish to see the output.

We'll also run the benchmark methods on each dataset and report the number of 
discoveries for each one. At first, we'll demo on just a __**small subset of 
SNPs**__, since the computation is time-intensive. After that is working 
smoothly, we'll scale it up to the full set of SNPs.

### Workspace setup

To run the benchmark methods on the GWAS datasets, we'll use some of the helper 
and wrapper functions from the simulated data analysis. We also create and source
an analgous version of the `sim_runner` function to apply to the real data.

```{r, wkspace-setup}
# Note that swfdr package depends on R version 3.4 or higher
setwd("/n/irizarryfs01_backed_up/kkorthauer/FDR")
source("benchmark-fdr/simulation_studies/R/simulation-helpers.R")
source("benchmark-fdr/datasets/R/evaluation-helpers.R")
datdir <- "/n/irizarryfs01_backed_up/kkorthauer/FDR/DATA/GWAS/"
resdir <- "/n/irizarryfs01_backed_up/kkorthauer/FDR/RESULTS/GWAS/"

# wrangling & plotting tools
library(data.table)
library(readxl)
library(dplyr)
library(ggplot2)
library(cowplot)
library(UpSetR)

# benchmark methods
library(IHW)
library(ashr)
library(qvalue)
library(swfdr)
library(fdrtool)

# constant parameters (for UpSet Plots)
alpha.cutoff <- 0.05
```

### GWAS meta-analysis for BMI (included in Boca-Leek)

Here we download and analyze the GWAS BMI dataset analyzed in the Boca-Leek paper.

#### Data download

```{bash, GWAS1-download, eval=FALSE}
curl -O "http://portals.broadinstitute.org/collaboration/giant/images/3/3a/BMI.SNPadjSMK.zip"
```

Next we unzip the file, and delete the files we won't use (they provide results
subsetted by ancestry and sex), keeping only the file with all ancestries and
both sexes `BMI.SNPadjSMK.CombinedSexes.AllAncestry.txt`. Note that it seems that
the Boca-Leek paper actually used the 'EuropeanOnly' subset, but it's not clear
why (to me), and there's no reason we need to do the same.

```{bash, GWAS1-unzip, eval=FALSE}
unzip BMI.SNPadjSMK.zip
rm BMI.SNPadjSMK.zip
rm BMI.SNPadjSMK.Women.EuropeanOnly.txt
rm BMI.SNPadjSMK.Women.AllAncestry.txt
rm BMI.SNPadjSMK.Men.EuropeanOnly.txt
rm BMI.SNPadjSMK.Men.AllAncestry.txt
rm BMI.SNPadjSMK.CombinedSexes.EuropeanOnly.txt
```

Next, we'll read in the unzipped `.txt` file into R and verify that it contains
the necessary inputs to run the FDR benchmark comparisons.

```{r, GWAS1-verify-contents}
bmi <- fread(paste0(datdir, "BMI.SNPadjSMK.CombinedSexes.AllAncestry.txt"),
             header=TRUE)
dim(bmi)
head(bmi)
```

It looks like we have:

- `p_value`:p-value
- `effect`:effect size
- `stderr`:standard error
- additional covariates: 
    - `N`: Number of samples with this SNP - BMI association measured
    - `Freq_Allele1_HapMapCEU`: Allele Frequency of Allele1, as measured in HapMapCEU

for 2,458,133 SNPs.

#### Covariate Diagnostics

Here we examine whether the available covariates appear to be suitable for the  
assumptions of the benchmark methods. The main assumption used by many methods 
is that the covariate is independent of the p-value conditional on the null
hypothesis. In other words, the covariate itself is not informative for whether
the test should be rejected or not. If this is satisfied, the covariate still
may or may not be useful. To assess whether the covariate is useful or not, we 
look to see if there is an association between covariate values and p-values. 
This is because under the alternative, we expect to see an enrichment of 
rejections for certain values of the covariate (if it is indeed informative).
In this section we examine two diagnostic plots for each of the covariates to 
see if these conditions hold in this dataset.

First, we will look at histograms of p-values stratified by covariate values. To
stratify by continuous covariate values, we'll use quantiles.

```{r, GWAS1-diag1, fig.width=9, fig.height=2.5}
# Demo on a small subset of SNPs first (remove this later)
bmi <- bmi[sample(1:nrow(bmi), 100000),]

# First for the sample size covariate (N)
strat_hist(bmi, pval="p_value", covariate="N", maxy=3)

# Next for the MAF covariate (Freq_Allele1_HapMapCEU)
strat_hist(bmi %>% filter(!is.na(Freq_Allele1_HapMapCEU)),
           pval="p_value", covariate="Freq_Allele1_HapMapCEU", maxy=3)

```

Next we'll look at a scatter plot of p-values by the ranks of covariate values.

```{r, GWAS1-diag2, fig.width=7, fig.height=4}
# First for the sample size covariate (N)
rank_scatter(bmi, pval="p_value", covariate="N")

# Next for the MAF covariate (Freq_Allele1_HapMapCEU)
rank_scatter(bmi, pval="p_value", covariate="Freq_Allele1_HapMapCEU")

# Take proximity to 0.5 as the actual covariate (according to previous plot)
bmi <- bmi %>%
        mutate(MAF.5=-abs(Freq_Allele1_HapMapCEU-0.5))
rank_scatter(bmi, pval="p_value", covariate="MAF.5") 
```

#### Running benchmark methods

First, we'll standardize the data frame to use the wrapper function. For now, we'll
assume that there is only a single 'independent covariate' for those methods that 
utilize this, and we'll use the sample size (number of subjects with that SNP
measured). **Still to-do: consider multiple covariates for methods that allow this.**

```{r, GWAS1-dat}
bmi <- bmi %>% 
        rename(pval = p_value,
               SE = stderr,
               effect_size = effect,
               ind_covariate = N) %>%
        mutate(zscore = effect_size/SE)
```

Now, we'll run the `run_benchmarks` helper function on the `bmi` data frame for
several alpha values and plot the results. This is fairly computationally intensive 
for some of the benchmark methods (since there are more than 2.5 million tests)
and best run on a machine with lots of memory.

```{r, GWAS1-run, results="hide"}
cutoffs <- c(0.01, 0.025, 0.05, 0.10, 0.20)
resfile <- paste0(resdir, "bmi_results_",
                  nrow(bmi), ".RData")
if (!file.exists(resfile)){
  t1 <- proc.time()
  bmi_results <- run_benchmarks(dat=bmi, alphas=cutoffs, pvals=TRUE)
  save(bmi_results, file=resfile)
  message(paste0("Took ", round((proc.time()-t1)[3],1),
                 " seconds for ", nrow(bmi), " SNPs."))
}else{
  load(resfile)
}

# plot N rejects vs FDR cutoff
p <- ggplot(bmi_results$stats, aes(alpha, n_rejects, color=method)) +
  geom_point() +
  geom_line() +
  xlab("FDR cutoff") + ylab("Number of rejections") +
  ggtitle("BMI GWAS Metanalysis with Sample Size Covariate") +
  theme_classic()
print(p)

# log-scale y axis
print(p + scale_y_continuous(trans="log2") + 
        ylab("Number of rejections (log2 scale)"))
```

Next, we look beyond the number of rejections for each method and examine
the overlap between them. We'll use the Venn diagram alternative provided in the
UpsetR package.

```{r, GWAS1-upset, fig.width=9, fig.height=9}
sig_df <- bmi_results$pvals %>%
            mutate_all(function(x) 1*(x < alpha.cutoff))
upset(sig_df, nsets = 9, nintersects = 30, mb.ratio = c(0.5, 0.5),
      order.by = c("freq"), decreasing = c(TRUE), 
      sets.bar.color = "#56B4E9")
```

And we'll clean up the workspace before moving on to the next dataset.

```{r, GWAS1-cleanup}
rm(bmi)
rm(bmi_results)
```

### GWAS for Asthma

[This study by Moffatt et al. (2010, NEJM)](http://www.nejm.org/doi/full/10.1056/NEJMoa0906312?query=TOC) performed a large
GWAS on  10,365 people with asthma and 16,110 controls 23 individual studies.

#### Data Download

Let's download the data from the Gabriel site. The second file contains metadata
for the columns in the main data file.

```{bash, GWAS2-download, eval=FALSE}
curl -O "https://beaune.cng.fr/gabriel/gabriel_results.zip"
curl -O "https://beaune.cng.fr/gabriel/gabriel_results_description.xls"
```

Next we unzip the file and remove the zipped copy.

```{bash, GWAS2-unzip, eval=FALSE}
unzip gabriel_results.zip
rm gabriel_results.zip
```

Next, we'll read in the unzipped `.txt` file and the results desription `.xls` 
file into R and verify that it contains
the necessary inputs to run the FDR benchmark comparisons.

```{r, GWAS2-verify-contents}
# read in meta-data
meta <- read_xls(paste0(datdir, "gabriel_results_description.xls"))
head(meta)
dim(meta)

# read in data
asthma <- fread(paste0(datdir, 
                "gabriel_asthma_meta-analysis_36studies_format_repository_NEJM.txt"),
                header=TRUE)
dim(asthma)

# verify meta-data rows = data cols
nrow(meta) == ncol(asthma)

# remove study-specific columns
empty <- which(!is.na(meta$description))
asthma <- asthma %>%
            select(empty)
meta <- meta %>% 
            slice(empty)

# remove SNPs with missing effect size/pvalues
asthma <- asthma %>% 
            filter(!is.na(P_fix) & !is.na(theta_fix))

dim(asthma)
head(asthma)
```

It looks like we have:

- `P_fix` and `P_ran`:p-values for fixed and random effect models
- `theta_fix` and `theta_ran`:effect size for fixed and random effect models
- `se_theta_fix` and `se_theta_ran`:standard error for fixed and random effect models
- additional covariates: 
    - `no_studies_meta`: Number of studies included in the meta-analysis with this SNP
    - `freq_all_1_max` and `freq_all_1_min`: Maximum and Minimum Allele Frequency of Allele 1 across samples

for 546,182 SNPs.

#### Covariate Diagnostics

First, we will look at histograms of p-values stratified by covariate values. To
stratify by continuous covariate values, we'll use quantiles.

```{r, GWAS2-diag1, fig.width=9, fig.height=2.5}
# Demo on a small subset of SNPs first (remove this later)
asthma <- asthma[sample(1:nrow(asthma), 100000),]

# First for the sample size covariate (no_studies_meta)
strat_hist(asthma, pval="P_fix", covariate="no_studies_meta", maxy=1.3)

# Next for the MAF covariate (freq_all_1_max)
strat_hist(asthma %>% filter(!is.na(freq_all_1_max)),
           pval="P_fix", covariate="freq_all_1_max", maxy=1.3)
```

Next we'll look at a scatter plot of p-values by the ranks of covariate values.

```{r, GWAS2-diag2, fig.width=7, fig.height=4}
# First for the sample size covariate (N)
rank_scatter(asthma, pval="P_fix", covariate="no_studies_meta")

# Next for the MAF covariate (freq_all_1_max)
rank_scatter(asthma, pval="P_fix", covariate="freq_all_1_max")

# Take distance to 0.5 as the actual covariate 
asthma <- asthma %>%
        mutate(MAF.5=abs(freq_all_1_max-0.5))
rank_scatter(asthma, pval="P_fix", covariate="MAF.5") 
```

#### Running benchmark methods

First, we'll standardize the data frame to use the wrapper function. For now, we'll
assume that there is only a single 'independent covariate' for those methods that 
utilize this, and we'll use the sample size (number of studies with that SNP
measured). We'll also focus on the fixed effects tests first.
**Still to-do: consider multiple covariates for methods that allow this.**

```{r, GWAS2-dat}
asthma <- asthma %>% 
        rename(pval = P_fix,
               SE = se_theta_fix,
               effect_size = theta_fix,
               ind_covariate = no_studies_meta) %>%
        mutate(zscore = effect_size/SE)
```

Now, we'll run the `run_benchmarks` helper function on the `asthma` data frame for
several alpha values and plot the results. This is fairly computationally intensive 
for some of the benchmark methods
(since there are more than 500 thousand tests)
and best run on a machine with lots of memory.

```{r, GWAS2-run, results="hide"}
# use same cutoffs as for the bmi study (defined above)
resfile <- paste0(resdir, "asthma_results_",
                  nrow(asthma), ".RData")
if (!file.exists(resfile)){
  t1 <- proc.time()
  asthma_results <- run_benchmarks(dat=asthma, alphas=cutoffs, pvals=TRUE)
  save(asthma_results, file=resfile)
  message(paste0("Took ", round((proc.time()-t1)[3],1),
                 " seconds for ", nrow(asthma), " SNPs."))
}else{
  load(resfile)
}

# plot N rejects vs FDR cutoff
p <- ggplot(asthma_results$stats, aes(alpha, n_rejects, color=method)) +
  geom_point() +
  geom_line() +
  xlab("FDR cutoff") + ylab("Number of rejections") +
  ggtitle("Asthma GWAS Metanalysis with Sample Size Covariate") +
  theme_classic()
print(p)

# log-scale y axis
print(p + scale_y_continuous(trans="log2") + 
        ylab("Number of rejections (log2 scale)"))
```

Next, we look beyond the number of rejections for each method and examine
the overlap between them. We'll use the Venn diagram alternative provided in the
UpsetR package.

```{r, GWAS2-upset, fig.width=9, fig.height=9}
sig_df <- asthma_results$pvals %>%
            mutate_all(function(x) 1*(x < alpha.cutoff))
upset(sig_df, nsets = 9, nintersects = 30, mb.ratio = c(0.5, 0.5),
      order.by = c("freq"), decreasing = c(TRUE), 
      sets.bar.color = "#56B4E9")
```

And we'll clean up the workspace before moving on to the next dataset.

```{r, GWAS2-cleanup}
rm(asthma)
rm(asthma_results)
```

### GWAS for Lipid Levels

[This study by Willer et al. (2013, Nature Genetics)](http://www.nature.com/ng/journal/v45/n11/full/ng.2797.html) performed a large
GWAS on 188,577 and measured blood lipid levels (as well as several cardiovascular traits).

#### Data Download

Let's download the data from the University of Michigan site. The download page 
has a handy [metadata table](http://csg.sph.umich.edu//abecasis/public/lipids2013/)
that describes the contents of the columns in the results files.

There are several types of results files to download (in terms of which phenotype to analyze):

- LDL Cholesterol
- HDL Cholesterol
- Triglycerides
- Total Cholesterol

As well as two types of analyses:

- Joint analyses of GWAS and Metabochip
- analysis of Metabochip

We would like to look at the joint analysis of GWAS + Metabochip since we'd like 
to look at all available SNPs (not just those that have previously shown a possible
association with CVD). We could choose any or all of the traits, but I'll go ahead
and proceed with Triglycerides since I'm curious about the role of genetic variation 
on Triglyceride levels.

```{bash, GWAS3-download, eval=FALSE}
curl -O "http://csg.sph.umich.edu/abecasis/public/lipids2013/jointGwasMc_TG.txt.gz"
```

Next we uncompress the file.

```{bash, GWAS3-unzip, eval=FALSE}
gunzip jointGwasMc_TG.txt.gz 
```

Next, we'll read in the uncompressed `.txt` file into R and verify that it contains
the necessary inputs to run the FDR benchmark comparisons.

```{r, GWAS3-verify-contents}
# read in data
trig <- fread(paste0(datdir, "jointGwasMc_TG.txt"), header=TRUE)
dim(trig)
head(trig)
```

It looks like we have:

- `P_value`:p-value
- `beta`:effect size
- `SE`:standard error
- additional covariates: 
    - `N`: The number of individuals analyzed for this marker.
    - `Freq.A1.1000G.EUR`: Frequency of allele A1 from 1000G EUR sample.

for 2,439,432 SNPs.

#### Covariate Diagnostics

First, we will look at histograms of p-values stratified by covariate values. To
stratify by continuous covariate values, we'll use quantiles.

```{r, GWAS3-diag1, fig.width=9, fig.height=2.5}
# Demo on a small subset of SNPs first (remove this later)
trig <- trig[sample(1:nrow(trig), 100000),]

# First for the sample size covariate (no_studies_meta)
strat_hist(trig, pval="P-value", covariate="N", maxy=1.3)

# Next for the MAF covariate (Freq.A1.1000G.EUR)
strat_hist(trig %>% filter(!is.na(Freq.A1.1000G.EUR)),
           pval="P-value", covariate="Freq.A1.1000G.EUR", maxy=1.3)
```

Next we'll look at a scatter plot of p-values by the ranks of covariate values.

```{r, GWAS3-diag2, fig.width=7, fig.height=4}
# First for the sample size covariate (N)
rank_scatter(trig, pval="P-value", covariate="N")

# Next for the MAF covariate (Freq.A1.1000G.EUR)
rank_scatter(trig, pval="P-value", covariate="Freq.A1.1000G.EUR") 
```

#### Running benchmark methods

First, we'll standardize the data frame to use the wrapper function. For now, we'll
assume that there is only a single 'independent covariate' for those methods that 
utilize this, and we'll use the sample size (number of individuals with that SNP
measured). 
**Still to-do: consider multiple covariates for methods that allow this.**

Note that for this dataset, all reported effect sizes are positive. This is because
the alleles (A1 and A2) have been ordered such that the direction of effect are 
consistent. Since Scott assumes that the Z-scores are standard normal, we can't 
apply this method to a set of all positive Z-scores. To recover the direction of 
effect, we use the `Freq.A1.1000G.EUR` (allele frequency of A1) variable,
and use it to determine whether
`A1` (allele 1) is the reference allele or not. If the allele frequence of A1 is
greater than 0.5, we consider it the reference allele. Then we'll use the 
comparison of reference to minor as the positive direction. That is,
if A2 was reference,
then effect direction is negative. Otherwise it is positive. Since some of the
SNPs are missing the reference allele information, we cannot determine their 
direction of effect and they will be removed.

```{r, GWAS3-dat}
trig <- trig %>% 
        rename(pval = "P-value",
               SE = se,
               effect_size = beta,
               ind_covariate = N)%>%
        mutate(zscore = effect_size/SE)

# add effect size
trig <- trig %>% 
       mutate(A2.reference = Freq.A1.1000G.EUR < 0.5) %>%
       mutate(zscore = zscore * (1-2*A2.reference))

# remove SNPs with missing values for zscore (couldn't determine direction of 
# effect)
trig <- trig %>% 
  filter(!is.na(zscore))
```

Now, we'll run the `run_benchmarks` helper function on the `trig` data frame for
several alpha values and plot the results. This is fairly computationally intensive 
for some of the benchmark methods
(since there are more than 500 thousand tests)
and best run on a machine with lots of memory.

```{r, GWAS3-run, results="hide"}
# use same cutoffs as for the bmi study (defined above)
resfile <- paste0(resdir, "trig_results_",
                  nrow(trig), ".RData")
if (!file.exists(resfile)){
  t1 <- proc.time()
  trig_results <- run_benchmarks(dat=trig, alphas=cutoffs, pvals=TRUE)
  save(trig_results, file=resfile)
  message(paste0("Took ", round((proc.time()-t1)[3],1),
                 " seconds for ", nrow(trig), " SNPs."))
}else{
  load(resfile)
}

# plot N rejects vs FDR cutoff
p <- ggplot(trig_results$stats, aes(alpha, n_rejects, color=method)) +
  geom_point() +
  geom_line() +
  xlab("FDR cutoff") + ylab("Number of rejections") +
  ggtitle("Triglyceride GWAS Metanalysis with Sample Size Covariate") +
  theme_classic()
print(p)

# log-scale y axis
print(p + scale_y_continuous(trans="log2") + 
        ylab("Number of rejections (log2 scale)"))

```

Next, we look beyond the number of rejections for each method and examine
the overlap between them. We'll use the Venn diagram alternative provided in the
UpsetR package.

```{r, GWAS3-upset, fig.width=9, fig.height=9}
sig_df <- trig_results$pvals %>%
            mutate_all(function(x) 1*(x < alpha.cutoff))
upset(sig_df, nsets = 9, nintersects = 30, mb.ratio = c(0.5, 0.5),
      order.by = c("freq"), decreasing = c(TRUE), 
      sets.bar.color = "#56B4E9")
```

And we'll clean up the workspace before moving on to the next dataset.

```{r, GWAS3-cleanup}
rm(trig)
rm(trig_results)
```