### 2017-10-06 simulation_studies/fdr_simulation_steph.Rmd
Add diagnostic and FDR plots for All Null sims

---
title: "Benchmarking FDR Methods - Null settings"
author: "Rafalab Journal Club Members"
date: "`r Sys.Date()`"
output: 
    html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
        code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load helper functions
RPROJ <- list(PROJHOME = normalizePath(getwd()))
source(file.path(RPROJ$PROJHOME, "R/simulation-helpers.R"))
source(file.path(RPROJ$PROJHOME, "../datasets/R/evaluation-helpers.R"))

m = 20000 # number of tests
```

# Introduction

The goal of this working group is to benchmark recent covariate-adjusted FDR methods. For 
complete details, check out our [benchmark-fdr GitHub repo](https://github.com/pkimes/benchmark-fdr).

In this R Markdown, we use a Monte Carlo simulation study to evaluate 
the `All nulls` simulation (all tests are simulated from the null distribution
i.e. no difference). We define

* $m$ = number of tests
* $M$ = Monte Carlo replications

Methods to evaluate:

- Bonferroni 
- BH
- IHW
- ASH
- BL
- Scott
- locfdr
- mixfdr

Metrics to evaluate:

- Number of discoveries
- Control FDR at $\alpha$
- Power
- Relationship between $\pi_0$ and covariate
- Assumptions: unimodal, etc

Load libraries: 
```{r, load-libraries, message=FALSE}
library(dplyr)
library(ggplot2)
library(cowplot)
library(doParallel)
```

Additional packages that should be installed are: 
```{r, eval=FALSE}
library(genefilter)

library(IHW)
library(ashr)
library(qvalue)
library(swfdr)
library(fdrtool)
  ## to use IHWpaper::scott_fdrreg wrapper, need to install FDRreg v2.0 (only available github):   
  ## devtools::install_github(repo= 'jgscott/FDRreg', subdir='R_pkg/') 
library(FDRreg)
```

# The "All Nulls" simulation scenerio

Descriptive summary of simulation study:

- Simulate a "null" dataset using `du_ttest_sim()`:
    * Two standard normals with mean = 0, sd = 1. 
    * $\pi_0$ = 1
- Calculate t-statistic. 
- Independent covariate (independent of the p-value under the null hypothesis): 
    * Uninformative covariate: uniform between 0, 1. 
    * Informative covariate: pooled standard deviation
- Apply different methods to control for false discoveries

Generate example data set with an **uninformative covariate** and 
**informative covariate**: 
```{r example-simulate-null-data}
n_samples = 20
n_groups = 2

# uninformative covariate = uniform(0,1)
sim_df_uninform <- du_ttest_sim(m=m, pi0=1, effect_size=0, 
                       n_samples=n_samples, n_groups = 2, 
                       informative_ind_covariate = FALSE)

# informative covariate = pooled variance
sim_df_inform <- du_ttest_sim(m=m, pi0=1, effect_size=0, 
                       n_samples=n_samples, n_groups = 2, 
                       informative_ind_covariate = TRUE)

head(sim_df_inform)
```

Inside the simulated "null" dataset object:

- `H` is an indicator representing the true null (0) or alternative status (1)
- `test_statistic` is the t-test statistic
- `effect_size` is the difference in means
- `pval` is the $p$-value from the t-test 
- `ind_covariate` the independent covariate (can be informative or uninformative). Here it's informative
- `SE` is the standard error ($\frac{\sigma}{\sqrt{n}}$)

## Diagnostic plots

Need to check two things:

### Conditional independence property

1. Conditional independence property. Need to check if the covariate is independent of the $p$-value under the null hypothesis. In other words, the covariate itself is not informative for whether the test should be rejected or not. Can be checked by splitting histogram of p-values into groups based on covariate. If independent, the right tails of the distribution will be the same. If not independent, the tails may no longer look uniform.

#### Using the uninformative covariate

```{r example-diagnostic-plot-null-indepedence-uninform}
colnames(sim_df_uninform)[which(colnames(sim_df_uninform) == "pval")] <- "pvals"
strat_hist(sim_df_uninform, pval="pvals", covariate = "ind_covariate", maxy=2)
```

#### Using the informative covariate

```{r example-diagnostic-plot-null-indepedence-inform}
colnames(sim_df_inform)[which(colnames(sim_df_inform) == "pval")] <- "pvals"
strat_hist(sim_df_inform, pval="pvals", covariate = "ind_covariate", maxy=2)
```

**This does not mean the covariate is informative**. We use the second diagnostic plot to check this.

### Informative property

2. Informative property. Need to check if the covariate is informative or not. In other words, the covariate should be associated with the p-values. This is because under the alternative hypothesis, we expect to see an enrichment of rejections for certain values of the covariate (if it is indeed informative). If all histograms look the same, the covariate is uninformative, and its use will not lead to an increase in power.

#### Using the uninformative covariate

```{r example-diagnostic-plot-null-uninform}
rank_scatter(sim_df_uninform, pval="pvals", covariate = "ind_covariate", bins=100)
```

#### Using the informative covariate

```{r example-diagnostic-plot-null-inform}
rank_scatter(sim_df_inform, pval="pvals", covariate = "ind_covariate", bins=100)
```

Run following on odyssey and save output as RDS: 
```{r evaluate-methods-null-data, eval=FALSE}
library(dplyr)
library(ggplot2)
library(doParallel)

workingPath <- "/net/irizarryfs01/srv/export/irizarryfs01_backed_up/share_root/shicks/projects/benchmark-fdr"
source(file.path(workingPath, "simulation-helpers.R"))
source(file.path(workingPath, "evaluation-helpers.R"))

# set these to a smaller number temporarily 
m = 20000 # number of tests
M = 500 # number of MC reps

n_samples = 20
n_groups = 2

alphas = seq(0.01, 0.1, length.out = 10)

nCores <- 10
registerDoParallel(cores = nCores)
workers <- getDoParWorkers()
backend <- getDoParName()
version <- getDoParVersion()

# uninformative covariate (uniform[0,1])
df.out.uninform <- foreach(i = 1:M, .verbose=T) %dopar% {
  sim_df_uninform <- du_ttest_sim(m=m, pi0=1, effect_size=0, n_samples=n_samples, 
                                  n_groups=n_groups, informative_ind_covariate = FALSE)
  sim_runner(sim = sim_df_uninform, alphas = alphas, pvals = FALSE)
}

df.out.uninform <- dplyr::bind_rows(df.out.uninform)
saveRDS(df.out.uninform, file=file.path(workingPath, "Results-nullSims-uninform.RDS"))

# informative covariate (pooled variance)
df.out.inform <- foreach(i = 1:M, .verbose=T) %dopar% {
  sim_df_inform <- du_ttest_sim(m=m, pi0=1, effect_size=0, n_samples=n_samples, 
                                  n_groups=n_groups, informative_ind_covariate = TRUE)
  sim_runner(sim = sim_df_inform, alphas = alphas, pvals = FALSE)
}

df.out.inform <- dplyr::bind_rows(df.out.inform)
saveRDS(df.out.inform, file=file.path(workingPath, "Results-nullSims-inform.RDS"))
```

## Control for FDR and FWER at $\alpha$

Here we will plot the number of discoveries as a function 
of $\alpha$ (should be 0 false discoveries). We will also evaluate
methods control for false discovery rate (FDR) and family-wise 
error rate (FWER) at nominal $\alpha$ level ranging from [0.01, 0.1].

### Using the uninformative covariate
Load and plot results from RDS file: 
```{r, eval=TRUE}
# workingPath <- "/net/irizarryfs01/srv/export/irizarryfs01_backed_up/share_root/shicks/projects/benchmark-fdr"
# df.out.uninform <- readRDS(file=file.path(workingPath, "Results-nullSims-uninform.RDS"))
df.out.uninform <- readRDS(file=file.path(RPROJ$PROJHOME, "R/Results-nullSims-uninform.RDS"))

# Number of discoveries as a function of $\alpha$
df.out.uninform %>% dplyr::group_by(method, alpha) %>% 
      dplyr::summarize(n_rejects = mean(n_rejects), FDR = mean(FDP), 
                       FWER = mean(FWER), FPR = mean(FPR)) %>% 
    ggplot(aes(x = alpha, y = n_rejects, color = method)) + 
         geom_line() + geom_point() + geom_abline(linetype="dashed") + 
         xlab(expression(bold(paste("Nominal ",alpha)))) + ylab("Number of Discoveries") +
         theme(axis.title = element_text(face="bold") )

# False discovery rate as a function of $\alpha$
df.out.uninform %>% dplyr::group_by(method, alpha) %>%
      dplyr::summarize(n_rejects = mean(n_rejects), FDR = mean(FDP), 
                       FWER = mean(FWER), FPR = mean(FPR)) %>% 
      ggplot(aes(x = alpha, y = FDR, color = method)) + 
         geom_line() + geom_point() + geom_abline(linetype="dashed") + 
         xlab(expression(bold(paste("Nominal ",alpha)))) + ylab("FDR") +
         theme(axis.title = element_text(face="bold") )

# Family wise error rate as a function of $\alpha$
df.out.uninform %>% dplyr::group_by(method, alpha) %>% 
      dplyr::summarize(n_rejects = mean(n_rejects), FDR = mean(FDP), 
                       FWER = mean(FWER), FPR = mean(FPR)) %>% 
      ggplot(aes(x = alpha, y = FWER, color = method)) + 
         geom_line() + geom_point() + geom_abline(linetype="dashed") + 
         xlab(expression(bold(paste("Nominal ",alpha)))) + ylab("FWER") +
         theme(axis.title = element_text(face="bold") )
```

### Using the informative covariate
Load and plot results from RDS file: 
```{r, eval=TRUE}
# workingPath <- "/net/irizarryfs01/srv/export/irizarryfs01_backed_up/share_root/shicks/projects/benchmark-fdr"
# df.out.inform <- readRDS(file=file.path(workingPath, "Results-nullSims-inform.RDS"))
df.out.inform <- readRDS(file=file.path(RPROJ$PROJHOME, "R/Results-nullSims-inform.RDS"))

# Number of discoveries as a function of $\alpha$
df.out.inform %>% dplyr::group_by(method, alpha) %>% 
      dplyr::summarize(n_rejects = mean(n_rejects), FDR = mean(FDP), 
                       FWER = mean(FWER), FPR = mean(FPR)) %>% 
    ggplot(aes(x = alpha, y = n_rejects, color = method)) + 
         geom_line() + geom_point() + geom_abline(linetype="dashed") + 
         xlab(expression(bold(paste("Nominal ",alpha)))) + ylab("Number of Discoveries") +
         theme(axis.title = element_text(face="bold") )

# False discovery rate as a function of $\alpha$
df.out.inform %>% dplyr::group_by(method, alpha) %>%
      dplyr::summarize(n_rejects = mean(n_rejects), FDR = mean(FDP), 
                       FWER = mean(FWER), FPR = mean(FPR)) %>% 
      ggplot(aes(x = alpha, y = FDR, color = method)) + 
         geom_line() + geom_point() + geom_abline(linetype="dashed") + 
         xlab(expression(bold(paste("Nominal ",alpha)))) + ylab("FDR") +
         theme(axis.title = element_text(face="bold") )

# Family wise error rate as a function of $\alpha$
df.out.inform %>% dplyr::group_by(method, alpha) %>% 
      dplyr::summarize(n_rejects = mean(n_rejects), FDR = mean(FDP), 
                       FWER = mean(FWER), FPR = mean(FPR)) %>% 
      ggplot(aes(x = alpha, y = FWER, color = method)) + 
         geom_line() + geom_point() + geom_abline(linetype="dashed") + 
         xlab(expression(bold(paste("Nominal ",alpha)))) + ylab("FWER") +
         theme(axis.title = element_text(face="bold") )
```

# The "Varying $\pi_0$" simulation scenerio

Descriptive summary of simulation study:

- Simulate a dataset using `du_ttest_sim()`:
    * Two standard normals with difference in means of 1.5, sd = 1. 
    * $\pi_0$ varying between 0 and 1 (e.g. $\pi_0$= 0.7)
- Calculate t-statistic. 
- Independent covariate (independent of the p-value under the null hypothesis): 
    * Uninformative covariate: uniform between 0, 1. 
    * Informative covariate: pooled standard deviation
- Apply different methods to control for false discoveries

Generate example data set with an **uninformative covariate** and 
**informative covariate**: 
```{r example-simulate-varyingpi0-data}
n_samples = 20
n_groups = 2

# uninformative covariate = uniform(0,1)
sim_df_uninform <- du_ttest_sim(m=m, pi0=0.7, effect_size=1.5, 
                       n_samples=n_samples, n_groups = 2, 
                       informative_ind_covariate = FALSE)

# informative covariate = pooled variance
sim_df_inform <- du_ttest_sim(m=m, pi0=0.7, effect_size=1.5, 
                       n_samples=n_samples, n_groups = 2, 
                       informative_ind_covariate = TRUE)

head(sim_df_inform)
```

## Diagnostic plots

Need to check two things:

### Conditional independence property

1. Conditional independence property. Need to check if the covariate is independent of the $p$-value under the null hypothesis. In other words, the covariate itself is not informative for whether the test should be rejected or not. Can be checked by splitting histogram of p-values into groups based on covariate. If independent, the right tails of the distribution will be the same. If not independent, the tails may no longer look uniform.

#### Using the uninformative covariate

```{r example-diagnostic-plot-varyingpi0-indepedence-uninform, warning=FALSE}
colnames(sim_df_uninform)[which(colnames(sim_df_uninform) == "pval")] <- "pvals"
strat_hist(sim_df_uninform, pval="pvals", covariate = "ind_covariate", maxy=2)
```

#### Using the informative covariate

```{r example-diagnostic-plot-varyingpi0-indepedence-inform, warning=FALSE}
colnames(sim_df_inform)[which(colnames(sim_df_inform) == "pval")] <- "pvals"
strat_hist(sim_df_inform, pval="pvals", covariate = "ind_covariate", maxy=2)
```

**This does not mean the covariate is informative**. We use the second diagnostic plot to check this.

### Informative property

2. Informative property. Need to check if the covariate is informative or not. In other words, the covariate should be associated with the p-values. This is because under the alternative hypothesis, we expect to see an enrichment of rejections for certain values of the covariate (if it is indeed informative). If all histograms look the same, the covariate is uninformative, and its use will not lead to an increase in power.

#### Using the uninformative covariate

```{r example-diagnostic-plot-varyingpi0-uninform}
rank_scatter(sim_df_uninform, pval="pvals", covariate = "ind_covariate", bins=100)
```

#### Using the informative covariate

```{r example-diagnostic-plot-varyingpi0-inform}
rank_scatter(sim_df_inform, pval="pvals", covariate = "ind_covariate", bins=100)
```

Run following on odyssey and save output as RDS: 
```{r evaluate-methods-varyingpi0-data, eval=FALSE}
library(dplyr)
library(ggplot2)
library(doParallel)

workingPath <- "/net/irizarryfs01/srv/export/irizarryfs01_backed_up/share_root/shicks/projects/benchmark-fdr"
source(file.path(workingPath, "simulation-helpers.R"))
source(file.path(workingPath, "evaluation-helpers.R"))

# set these to a smaller number temporarily 
m = 20000 # number of tests
M = 500 # number of MC reps

n_samples = 20
n_groups = 2

alphas = seq(0.01, 0.1, length.out = 10)
pi0set <- c(0, 0.1, 0.3, 0.7, 0.9, 1)

nCores <- 10
registerDoParallel(cores = nCores)
workers <- getDoParWorkers()
backend <- getDoParName()
version <- getDoParVersion()

# uninformative covariate (uniform[0,1])
df.out.uninform <- foreach(i = 1:M, .verbose=T) %dopar% {
  dat <- NULL
  for(i in 1:length(pi0set)){
    sim_df_uninform <- du_ttest_sim(m=m, pi0=pi0set[i], effect_size=1.5, n_samples=n_samples, 
                                    n_groups=n_groups, informative_ind_covariate = FALSE)
    dat <- rbind(dat, data.frame(sim_runner(sim = sim_df_uninform, alphas = alphas, pvals = FALSE), 
                                 "pi0" = pi0set[i]))
  }
  dat
}

df.out.uninform <- dplyr::bind_rows(df.out.uninform)
saveRDS(df.out.uninform, file=file.path(workingPath, "Results-varyingpi0Sims-uninform.RDS"))

# informative covariate (pooled variance)
df.out.inform <- foreach(i = 1:M, .verbose=T) %dopar% {
    dat <- NULL
  for(i in 1:length(pi0set)){
    sim_df_inform <- du_ttest_sim(m=m, pi0=pi0set[i], effect_size=1.5, n_samples=n_samples, 
                                    n_groups=n_groups, informative_ind_covariate = TRUE)
    dat <- rbind(dat, data.frame(sim_runner(sim = sim_df_inform, alphas = alphas, pvals = FALSE), 
                                 "pi0" = pi0set[i]))
  }
  dat
}

df.out.inform <- dplyr::bind_rows(df.out.inform)
saveRDS(df.out.inform, file=file.path(workingPath, "Results-varyingpi0Sims-inform.RDS"))
```

## Control for FDR, FWER, Power at $\alpha$

Here we will plot the number of discoveries as a function 
of $\alpha$. We will also evaluate
methods control for false discovery rate (FDR), family-wise 
error rate (FWER) and Power at nominal $\alpha$ level 
ranging from [0.01, 0.1].

### Using the uninformative covariate
Load and plot results from RDS file: 
```{r, eval=FALSE}
# workingPath <- "/net/irizarryfs01/srv/export/irizarryfs01_backed_up/share_root/shicks/projects/benchmark-fdr"
# df.out.uninform <- readRDS(file=file.path(workingPath, "Results-varyingpi0Sims-uninform"))
df.out.uninform <- readRDS(file=file.path(RPROJ$PROJHOME, "R/Results-varyingpi0Sims-uninform"))

# Number of discoveries as a function of $\alpha$
df.out.uninform %>% dplyr::group_by(method, alpha, pi0) %>% 
      dplyr::summarize(n_rejects = mean(n_rejects), FDR = mean(FDP), 
                       FWER = mean(FWER), FPR = mean(FPR)) %>% 
    ggplot(aes(x = alpha, y = n_rejects, color = method)) + 
         geom_line() + geom_abline(linetype="dashed") + 
         xlab(expression(bold(paste("Nominal ",alpha)))) + ylab("Number of Discoveries") +
         theme(axis.title = element_text(face="bold") ) + 
  facet_grid(~ pi0)

# False discovery rate as a function of $\alpha$
df.out.uninform %>% dplyr::group_by(method, alpha, pi0) %>%
      dplyr::summarize(n_rejects = mean(n_rejects), FDR = mean(FDP), 
                       FWER = mean(FWER), FPR = mean(FPR)) %>% 
      ggplot(aes(x = alpha, y = FDR, color = method)) + 
         geom_line() + geom_abline(linetype="dashed") + 
         xlab(expression(bold(paste("Nominal ",alpha)))) + ylab("FDR") +
         theme(axis.title = element_text(face="bold") ) + 
  facet_grid(~ pi0)

# Family wise error rate as a function of $\alpha$
df.out.uninform %>% dplyr::group_by(method, alpha, pi0) %>% 
      dplyr::summarize(n_rejects = mean(n_rejects), FDR = mean(FDP), 
                       FWER = mean(FWER), FPR = mean(FPR)) %>% 
      ggplot(aes(x = alpha, y = FWER, color = method)) + 
         geom_line() + geom_abline(linetype="dashed") + 
         xlab(expression(bold(paste("Nominal ",alpha)))) + ylab("FWER") +
         theme(axis.title = element_text(face="bold") ) + 
  facet_grid(~ pi0)

# Power as a function of $\alpha$
df.out.uninform %>% dplyr::group_by(method, alpha, pi0) %>% 
      dplyr::summarize(n_rejects = mean(n_rejects), Power = mean(power)) %>% 
      ggplot(aes(x = alpha, y = Power, color = method)) + 
         geom_line() + 
         xlab(expression(bold(paste("Nominal ",alpha)))) + ylab("Power") +
         theme(axis.title = element_text(face="bold") ) + 
  facet_grid(~ pi0)

```

### Using the informative covariate
Load and plot results from RDS file: 
```{r, eval=FALSE}
# workingPath <- "/net/irizarryfs01/srv/export/irizarryfs01_backed_up/share_root/shicks/projects/benchmark-fdr"
# df.out.inform <- readRDS(file=file.path(workingPath, "Results-varyingpi0Sims-inform"))
df.out.inform <- readRDS(file=file.path(RPROJ$PROJHOME, "R/Results-varyingpi0Sims-inform"))

# Number of discoveries as a function of $\alpha$
df.out.inform %>% dplyr::group_by(method, alpha, pi0) %>% 
      dplyr::summarize(n_rejects = mean(n_rejects), FDR = mean(FDP), 
                       FWER = mean(FWER), FPR = mean(FPR)) %>% 
    ggplot(aes(x = alpha, y = n_rejects, color = method)) + 
         geom_line() + geom_abline(linetype="dashed") + 
         xlab(expression(bold(paste("Nominal ",alpha)))) + ylab("Number of Discoveries") +
         theme(axis.title = element_text(face="bold") ) + 
  facet_grid(~ pi0)

# False discovery rate as a function of $\alpha$
df.out.inform %>% dplyr::group_by(method, alpha, pi0) %>%
      dplyr::summarize(n_rejects = mean(n_rejects), FDR = mean(FDP), 
                       FWER = mean(FWER), FPR = mean(FPR)) %>% 
      ggplot(aes(x = alpha, y = FDR, color = method)) + 
         geom_line() + geom_abline(linetype="dashed") + 
         xlab(expression(bold(paste("Nominal ",alpha)))) + ylab("FDR") +
         theme(axis.title = element_text(face="bold") ) + 
  facet_grid(~ pi0)

# Family wise error rate as a function of $\alpha$
df.out.inform %>% dplyr::group_by(method, alpha, pi0) %>% 
      dplyr::summarize(n_rejects = mean(n_rejects), FDR = mean(FDP), 
                       FWER = mean(FWER), FPR = mean(FPR)) %>% 
      ggplot(aes(x = alpha, y = FWER, color = method)) + 
         geom_line() + geom_abline(linetype="dashed") + 
         xlab(expression(bold(paste("Nominal ",alpha)))) + ylab("FWER") +
         theme(axis.title = element_text(face="bold") ) + 
  facet_grid(~ pi0)

# Power as a function of $\alpha$
df.out.inform %>% dplyr::group_by(method, alpha, pi0) %>% 
      dplyr::summarize(n_rejects = mean(n_rejects), Power = mean(power)) %>% 
      ggplot(aes(x = alpha, y = Power, color = method)) + 
         geom_line() + 
         xlab(expression(bold(paste("Nominal ",alpha)))) + ylab("Power") +
         theme(axis.title = element_text(face="bold") ) + 
  facet_grid(~ pi0)

```