### 2017-11-10 datasets/GWAS/GWAS-SummarizedBenchmark.Rmd

---
title: "Comparison of FDR methods on GWAS with SummarizedBenchmark"
author: "Keegan Korthauer"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/n/irizarryfs01_backed_up/kkorthauer/FDR")
```

### Summary

In the previous document `GWAS.Rmd`, we identified three GWAS datasets for FDR
benchmarking. Here we will make use of the new `SummarizedBenchmark` package to
perform the benchmarking in a standardized way.

### Workspace setup

```{r, wkspace-setup, results='hide', message=FALSE, warning=FALSE}
# project directory & data/results folders
#setwd("/n/irizarryfs01_backed_up/kkorthauer/FDR")
datdir <- "/n/irizarryfs01_backed_up/kkorthauer/FDR/DATA/GWAS/"
resdir <- "/n/irizarryfs01_backed_up/kkorthauer/FDR/RESULTS/GWAS/"

# wrangling & plotting tools
library(data.table)
library(readxl)
library(dplyr)
library(ggplot2)
library(magrittr)
library(R.utils)
library(cowplot)

# benchmark methods
library(IHW)
library(ashr)
library(qvalue)
library(swfdr)
library(fdrtool)
library(FDRreg)

# comparison tools/functions
library(SummarizedBenchmark)
sourceDirectory("benchmark-fdr/datasets/R/")

# set up parallel backend
library(BiocParallel)
cores <- as.numeric(Sys.getenv("SLURM_NTASKS"))
multicoreParam <- MulticoreParam(workers = cores)
```

### GWAS meta-analysis for BMI (included in Boca-Leek)

Here we download and analyze the GWAS BMI dataset analyzed in the Boca-Leek paper.

#### Data download

```{bash, GWAS1-download, eval=FALSE}
cd /n/irizarryfs01_backed_up/kkorthauer/FDR/DATA/GWAS/
curl -O "http://portals.broadinstitute.org/collaboration/giant/images/3/3a/BMI.SNPadjSMK.zip"
```

Next we unzip the file, and delete the files we won't use (they provide results
subsetted by ancestry and sex), keeping only the file with European ancestry and
both sexes `BMI.SNPadjSMK.CombinedSexes.EuropeanOnly.txt`. Note that
the Boca-Leek paper also used the 'EuropeanOnly' subset, likely because using 
all ancestries would require that we assess the impact of population stratification
and adjust for it if present. For simplicity, we follow suit and use the homogeneous 
subset to avoid the impact of population stratification on our results.

```{bash, GWAS1-unzip, eval=FALSE}
unzip BMI.SNPadjSMK.zip
rm BMI.SNPadjSMK.zip
rm BMI.SNPadjSMK.Women.EuropeanOnly.txt
rm BMI.SNPadjSMK.Women.AllAncestry.txt
rm BMI.SNPadjSMK.Men.EuropeanOnly.txt
rm BMI.SNPadjSMK.Men.AllAncestry.txt
rm BMI.SNPadjSMK.CombinedSexes.AllAncestry.txt
```

Next, we'll read in the unzipped `.txt` file into R and verify that it contains
the necessary inputs to run the FDR benchmark comparisons.

```{r, GWAS1-verify-contents}
bmi <- fread(paste0(datdir, "BMI.SNPadjSMK.CombinedSexes.EuropeanOnly.txt"),
             header=TRUE)
dim(bmi)
head(bmi)
```

It looks like we have:

- `p_value`:p-value
- `effect`:effect size
- `stderr`:standard error
- additional covariates: 
    - `N`: Number of samples with this SNP - BMI association measured
    - `Freq_Allele1_HapMapCEU`: Allele Frequency of Allele1, as measured in HapMapCEU

for `r nrow(bmi)` SNPs.

Before moving on, we'll rename the relevant columns so that they will use common terms
across the different datasets. We'll also add a 'test_statistic' column.

```{r, GWAS1-reformat}
bmi <- bmi %>% 
        dplyr::rename(pval = p_value,
               SE = stderr,
               effect_size = effect,
               ind_covar_N = N,
               ind_covar_AF = Freq_Allele1_HapMapCEU) %>%
        mutate(test_statistic = qnorm(exp(log(pval)-log(2)), lower.tail=FALSE) * sign(effect_size)) %>%
        select(pval, SE, effect_size, ind_covar_N, ind_covar_AF, test_statistic) 
```

#### Check Covariate Diagnostics

Here we look to see if the covariates do indeed look informative. First we look at 
the sample size covariate.

```{r, GWAS1-diag, fig.width=10, fig.height=3.5}
strat_hist(bmi, pvalue="pval", covariate="ind_covar_N", maxy=3)
rank_scatter(bmi, pvalue="pval", covariate="ind_covar_N")
```

Next we look at the allele frequency covariate.

```{r, GWAS1-diag2, fig.width=10, fig.height=3.5}
strat_hist(bmi, pvalue="pval", covariate="ind_covar_AF", maxy=3)
rank_scatter(bmi, pvalue="pval", covariate="ind_covar_AF")
```

#### Set up BenchDesign object

First, we'll create an object of `BenchDesign` class to hold the data and 
add the benchmark methods to the `BenchDesign` object.

```{r, GWAS1-benchdesign}
bd <- initializeBenchDesign()
```

#### Run buildBench for the sample size covariate and examine results

Now, we're ready to construct the `SummarizedBenchmark` object, which will run
the functions specified in each method (these are actually sourced in from the
helper scripts).

```{r, GWAS1-sb, results="hide", message=FALSE}
resN <- paste0(resdir, "bmi_N_summarizedBenchmark_",
                  nrow(bmi), ".RData")

duration <- NA
if (!file.exists(resN)){
  t1 <- proc.time()
  sb <- bd %>% buildBench(data=(bmi %>% mutate(ind_covariate=ind_covar_N)), 
                          ftCols = c("ind_covar_N"),
                          parallel=TRUE, BPPARAM=multicoreParam)
  metadata(sb)$data_download_link <- "http://portals.broadinstitute.org/collaboration/giant/images/3/3a/BMI.SNPadjSMK.zip"
  save(sb, file=resN)
  duration <- round((proc.time()-t1)[3]/60,1)
}else{
  load(resN)
}
```

```{r, echo=FALSE}
if(!is.na(duration)){
  message("This step took ", duration, " minutes for ", nrow(bmi), " SNPs using ",
        cores, " cores.")
}
```

Next, we'll add the default performance metric for q-value assays. First, we have
to rename the assay to 'qvalue'.

```{r, GWAS1-metrics}
# rename assay to qvalue
assayNames(sb) <- "qvalue"
sb <- addDefaultMetrics(sb)
```

Now, we'll plot the results.

```{r, GWAS1-plot, results="hide", width=15, height=15}
# plot nrejects by method overall and stratified by covariate
rejections_scatter(sb,
                   supplementary=FALSE)

rejection_scatter_bins(sb, covariate="ind_covar_N", bins=4,
                   supplementary=FALSE)

# upset plot 
plotFDRMethodsOverlap(sb, 
                   alpha=0.05, nsets=ncol(sb),
                   order.by="freq", decreasing=TRUE,
                   supplementary=FALSE)
```

#### Run buildBench for the allele frequency covariate and examine results

Now, we'll repeat the previous steps for the other covariate (AF):

```{r, GWAS1-sb2, results="hide", message=FALSE}
bd <- initializeBenchDesign()
resAF <- paste0(resdir, "bmi_AF_summarizedBenchmark_",
                  nrow(bmi), ".RData")

duration <- NA
if (!file.exists(resAF)){
  t1 <- proc.time()
  sb <- bd %>% buildBench(data=(bmi %>% mutate(ind_covariate=ind_covar_AF) %>%
                                        filter(!is.na(ind_covariate))), 
                          ftCols = c("ind_covar_AF"),
                          parallel=TRUE, BPPARAM=multicoreParam)
  metadata(sb)$data_download_link <- "http://portals.broadinstitute.org/collaboration/giant/images/3/3a/BMI.SNPadjSMK.zip"
  save(sb, file=resAF)
  duration <- round((proc.time()-t1)[3]/60,1)
}else{
  load(resAF)
}
```

```{r, echo=FALSE}
if(!is.na(duration)){
  message("This step took ", duration, " minutes for ", nrow(bmi), " SNPs using ",
        cores, " cores.")
}
```

Next, we'll add the default performance metric for q-value assays. First, we have
to rename the assay to 'qvalue'.

```{r, GWAS1-metrics2}
# rename assay to qvalue
assayNames(sb) <- "qvalue"
sb <- addDefaultMetrics(sb)
```

Now, we'll plot the results.

```{r, GWAS1-plot2, results="hide", width=15, height=15}
# plot nrejects by method overall and stratified by covariate
rejections_scatter(sb,
                   supplementary=FALSE)

rejection_scatter_bins(sb, covariate="ind_covar_AF", bins=4,
                   supplementary=FALSE)

# upset plot 
plotFDRMethodsOverlap(sb, 
                   alpha=0.05, nsets=ncol(sb),
                   order.by="freq", decreasing=TRUE,
                   supplementary=FALSE)
```

And we'll clean up the workspace before moving on to the next dataset.

```{r, GWAS1-cleanup}
rm(bmi)
rm(bd)
rm(sb)
rm(pf)
```