### 2017-12-08 datasets/GSEA/gsea.Rmd
GSEA

---
title: "GSEA datasets for FDR benchmarking"
author: "Alejandro Reyes (on behalf of the Rafalab Benchmarking Consortium)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Differential expression analysis with DESeq2

Downloading *DESeqDataSet* if not present locally,

```{r, message=FALSE}

file <- "dsd.RData"
if(!file.exists( file ) ){
    download.file("https://www.dropbox.com/s/rdmv3tyqt4g37b0/dsd.RData?dl=1",destfile="dsd.RData")
}
load(file)

library(DESeq2)
levels(colData(dsdObject)$individual) <- gsub("GTEX-", "",levels(colData(dsdObject)$individual))

```

Specify model matrix design and run DESeq2 to get differentially expressed genes.

```{r deseq2, message=FALSE}

library(EnsDb.Hsapiens.v86)
gnType <-  genes(EnsDb.Hsapiens.v86, column="gene_biotype")

dsdObject <- dsdObject[rownames(dsdObject) %in% gnType[gnType$gene_biotype == "protein_coding"]$gene_id,]
dsdObject <- DESeq(dsdObject)
res <- results( dsdObject, independentFiltering=FALSE )

```

Performing a gene set enrichment analysis with goseq.

```{r goseq, message=FALSE}

library(goseq)

file="goSets.RData"
if(!file.exists(file)){
  stopifnot( require( biomaRt ) )
  mart <- useMart( "ensembl", "hsapiens_gene_ensembl" )
  goSets <- getBM( c("ensembl_gene_id", "go_id"), mart=mart, filters="ensembl_gene_id", values=rownames(res))
  goSets <- goSets[!nchar( goSets$go_id ) == 0,]
  goSets <- with( goSets, split( go_id, ensembl_gene_id ) )
  save(goSets, file=file)
}else{
  load(file)
}

genes <- as.numeric( res$padj < 0.1 )
names( genes ) <- rownames( res )

### getting median transcript length
txByGene <- transcriptsBy(EnsDb.Hsapiens.v86, "gene")
geneLength <- sapply( width( txByGene ), median )
geneLength <- geneLength[names(genes)]
genes[is.na( genes )] <- 0

any( is.na(genes) )
stopifnot(!any(is.na(genes)))
pwf <- nullp(genes, bias.data=geneLength )
goRes <- goseq( pwf, gene2cat=goSets )

```

```{r informativeCovariate, message=FALSE, fig.height=3.5, message=FALSE, fig.width=4.5}

library(ggplot2)
library(magrittr)
library(cowplot)
library(R.utils)
library(scales)
sourceDirectory("../R/")

rank_scatter( dat=goRes, pval="over_represented_pvalue", 
              covariate="numInCat", bins=50, funx=log2, 
              funfill=log10_trans()) +
    ylim(0, 12) + ggtitle("Over-represented gene sets") +
    xlab( expression(log[10]~"(# of genes)")) +
    ylab( expression(-log[10]~"(p-value)") )

```

```{r assumptions, message=FALSE, fig.width=10, fig.height=3.2}

library(dplyr)

## filter numInCat below only used for plotting
## filter numDEInCat for cases where there are no DE genes in the category (these are actually not tested)
#dplyr:::filter( goRes, over_represented_pvalue == 1, numDEInCat > 0)

dplyr:::filter( goRes, numDEInCat > 0, numInCat > 15 ) %>%
  strat_hist( pval="over_represented_pvalue", covariate="numInCat", maxy=11)

```

 Generating the *SummarizedBenchmark* object:

```{r runningBenckmark, message=FALSE}
designBen <- initializeBenchDesign()
res <- dplyr:::select( goRes, c("over_represented_pvalue", "numInCat") ) %>%
  dplyr:::rename( pval = over_represented_pvalue, ind_covariate = numInCat )
res$pval <- pmin( res$pval, 1 )
sGSEA <- designBen %>% buildBench( res, ftCols="ind_covariate")
assayNames(sGSEA) <- "qvalue"
sGSEA <- addDefaultMetrics( sGSEA )
```

```{r plottingResults}
rejections_scatter( sGSEA, as_fraction=FALSE, supplementary=FALSE)

sGSEA <- estimatePerformanceMetrics(sGSEA, addColData=TRUE)

estimatePerformanceMetrics( sGSEA, alpha=0.1, tidy=TRUE ) %>%
  filter( blabel %in% c("lfdr","ihw-a10","bl-df03","qvalue","bh", "bonf") ) %>%
  select( blabel, value )

```

```{r}
plotFDRMethodsOverlap( sGSEA, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
```

# Gene set size of overlaps

```{r}

getOverlapIndexes <- function( labs=c("lfdr", "ihw-a10"), alpha=0.1, sGSEA ){
  rowSums( assays(sGSEA)[["qvalue"]][,labs,drop=FALSE] < alpha) == length(labs) &
  !(rowSums( assays(sGSEA)[["qvalue"]][,c("lfdr","ihw-a10","bl-df03","qvalue","bh", "bonf")] < alpha, na.rm=TRUE ) > (length(labs)))
}

set1 <- getOverlapIndexes( c("lfdr", "ihw-a10"), alpha=0.1, sGSEA=sGSEA )
set2 <- getOverlapIndexes( c("bl-df03"), alpha=0.1, sGSEA=sGSEA )
set3 <- getOverlapIndexes( c("lfdr","ihw-a10","bl-df03","qvalue","bh"), alpha=0.1, sGSEA=sGSEA )

sizesDf <- data.frame( 
  size=c( rowData( sGSEA )$ind_covariate[which(set1)], 
          rowData( sGSEA )$ind_covariate[which(set2)], 
          rowData( sGSEA )$ind_covariate[which(set3)] ),
  label=rep(c("set1", "set2", "set3"), c(sum(set1), sum(set2), sum(set3)) ) )

ggplot( sizesDf, aes( label, log2(size) )) +
  geom_boxplot()

```

# NULL comparisons

```{r, eval=FALSE }

nms <- names( genes )
xx <- seq_len(length(nms))
gns <- 1*( xx %in% sample( xx, 1000, replace=FALSE ) )
names( gns ) <- nms
pwf <- nullp( gns, bias.data=geneLength )
goRes <- goseq( pwf, gene2cat=goSets )
res <- goRes %>%
  dplyr:::filter( numDEInCat > 0 ) %>%
  dplyr:::select( c("over_represented_pvalue", "numInCat") ) %>%
  dplyr:::rename( pval = over_represented_pvalue, ind_covariate = numInCat )
pr <- designBen %>% buildBench( res, ftCols="ind_covariate")
assayNames(pr) <- "qvalue"
pr <- addDefaultMetrics(pr, metrics="rejections")

estimatePerformanceMetrics( pr, alpha=0.1, tidy=TRUE ) %>%
  filter( blabel %in% c("lfdr","ihw-a10","bl-df03","qvalue","bh", "bonf") ) %>%
  select( blabel, value )

```

# Same analysis but using the HSC-MPP dataset from Nina

```{r}

dsdMouse <- "dsdHSCMPP.RData"
if( !file.exists( dsdMouse ) ){
  download.file("https://www.dropbox.com/s/spo1pifum5k72rn/dseHSCMPP.RData?dl=1", destfile="dsdHSCMPP.RData")
}
load(dsdMouse)
dseHSCMPP <- DESeq( dseHSCMPP)

res <- results( dseHSCMPP, contrast=c("conditions", "HSC", "MPP1"), independentFiltering=FALSE )
genes <- as.numeric( res$padj < 0.1 )
names( genes ) <- rownames( res )
genes[is.na( genes )] <- 0
table(genes)

### preparing gene length info and goSets ###
library("EnsDb.Mmusculus.v75")
txByGene <- transcriptsBy(EnsDb.Mmusculus.v75, "gene")
geneLength <- sapply( width( txByGene ), median )
geneLength <- geneLength[names(genes)]

file="goSetsMm.RData"
if(!file.exists(file)){
  stopifnot( require( biomaRt ) )
  mart <- useMart( "ensembl", "mmusculus_gene_ensembl" )
  goSets <- getBM( c("ensembl_gene_id", "go_id"), mart=mart, filters="ensembl_gene_id", values=rownames(res))
  goSets <- goSets[!nchar( goSets$go_id ) == 0,]
  goSets <- with( goSets, split( go_id, ensembl_gene_id ) )
  save(goSets, file=file)
}else{
  load(file)
}

pwf <- nullp(genes, bias.data=geneLength )
goRes <- goseq( pwf, gene2cat=goSets )

res <- dplyr:::select( goRes, c("over_represented_pvalue", "numInCat") ) %>%
  dplyr:::rename( pval = over_represented_pvalue, ind_covariate = numInCat )
res$pval <- pmin( res$pval, 1 )
sGSEAMouse <- designBen %>% buildBench( res, ftCols="ind_covariate")
assayNames(sGSEAMouse) <- "qvalue"
sGSEAMouse <- addDefaultMetrics( sGSEAMouse )

rank_scatter( dat=goRes, pval="over_represented_pvalue", 
              covariate="numInCat", bins=50, funx=log2, 
              funfill=log10_trans()) +
    ylim(0, 12) + ggtitle("Over-represented gene sets") +
    xlab( expression(log[10]~"(# of genes)")) +
    ylab( expression(-log[10]~"(p-value)") )

library(dplyr)

dplyr:::filter( goRes, numDEInCat > 0, numInCat > 20)  %>%
  strat_hist( pval="over_represented_pvalue", covariate="numInCat", maxy=7)

rejections_scatter( sGSEAMouse, as_fraction=FALSE, supplementary=FALSE)

plotFDRMethodsOverlap( sGSEAMouse, supplementary=FALSE, order.by="freq")

set1 <- getOverlapIndexes( c("lfdr","ihw-a10","bl-df03","qvalue","bh", "bonf"), alpha=0.1, sGSEA=sGSEAMouse)
set2 <- getOverlapIndexes( c("lfdr", "ihw-a10"), alpha=0.1, sGSEA=sGSEAMouse )
set3 <- getOverlapIndexes( c("bl-df03"), alpha=0.1, sGSEA=sGSEAMouse )

sizesDf <- data.frame( 
  size=c( rowData( sGSEAMouse )$ind_covariate[which(set1)], 
          rowData( sGSEAMouse )$ind_covariate[which(set2)], 
          rowData( sGSEAMouse )$ind_covariate[which(set3)] ),
  label=rep(c("set1", "set2", "set3"), c(sum(set1), sum(set2), sum(set3)) ) )

ggplot( sizesDf, aes( label, log2(size) )) +
  geom_boxplot()

```