### 2018-01-05 datasets/RNA-Seq/polyester-benchmarkFDR.Rmd
add initial polyester RNA-Seq simulation R Markdown

---
title: "Simulation study using RNA-Seq counts with Polyester"
author: "Rafalab Journal Club Members"
date: "`r Sys.Date()`"
output: 
    html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
        code_folding: hide

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load packages needed to simulate RNA-seq counts with polyester
library(zebrafishRNASeq) 
library(genefilter)
library(limma)
library(polyester)

# Differential testing
library(DESeq2)

# Load supporting packages for data analyses
library(data.table)
library(magrittr)
library(dplyr)
library(ggplot2)
library(R.utils)
library(cowplot)

# benchmark methods
library(IHW)
library(ashr)
library(qvalue)
library(swfdr)
library(fdrtool)
library(FDRreg)

# comparison tools/functions
library(SummarizedBenchmark)

# load helper functions
working_dir <- list(project_home = normalizePath(getwd()))
sourceDirectory(file.path(working_dir$project_home,"datasets/R/"))

# set up parallel backend
library(BiocParallel)
cores <- 4
multicoreParam <- MulticoreParam(workers = cores)
```

### Simulate RNA-Seq counts using polyester

We use the [polyester](https://bioconductor.org/packages/release/bioc/html/polyester.html) Bioconductor package to simulate RNA-Seq reads. We start with the [zebrafish](https://bioconductor.org/packages/release/data/experiment/html/zebrafishRNASeq.html) Bioconductor data package.

```{r simulate-data}
data(zfGenes)
filter <- apply(zfGenes, 1, function(x) length(x[x>5])>=2)
counts <- zfGenes[filter,]

## Estimate the zero inflated negative binomial parameters
params = get_params(counts)

plot(rowMeans(log(counts+1)), rowVars(log(counts+1)), 
     pch=19, main="zebrafish data w/fit")
lines(params$fit, col = 2)

## Create some data from the model
dat0 = create_read_numbers(params$mu, params$fit, params$p0,
                           m=dim(counts)[1], n=dim(counts)[2])
```

### Build DESeq2 object and run DESeq2

```{r run-deseq2}
group = factor(rep(c(0,1), each=3))
design = model.matrix(~group)

## Buid DESeq object and run DESeq2
dds <- DESeqDataSetFromMatrix(countData=dat0, colData = DataFrame(group), design=~group)
dds <- DESeq(dds) 
res <- results(dds, independentFiltering = F)
```

### Create SummarizedBenchmark object

```{r create-summarizedbenchmark-object}
# Build input data frame for SummarizedBenchmark object
# Important to not change col names since SB is strict about that
geneExp <- tbl_df(data.frame(geneName=rownames(zfGenes)[filter], pval=res$pvalue, SE=res$lfcSE, 
                      ind_covariate = res$baseMean, effect_size=res$log2FoldChange, test_statistic=res$stat))

geneExp <- geneExp %>% na.omit() %>% dplyr::filter(ind_covariate>=1)
```

```{r plot-strat-hist}
strat_hist(geneExp, pvalue="pval", covariate="ind_covariate", maxy=2)
```

```{r plot-rank-scatter}
rank_scatter(geneExp, pvalue="pval", covariate="ind_covariate")
```

```{r}
bd <- initializeBenchDesign()
sb <- bd %>% buildBench(data=geneExp, parallel = T)

assayNames(sb) <- "qvalue"
sb <- addDefaultMetrics(sb)

```

```{r}
rejections_scatter(sb, supplementary=FALSE)
```

```{r}
plotFDRMethodsOverlap(sb, alpha=0.05, nsets=ncol(sb), order.by="freq", decreasing=TRUE, supplementary=FALSE)
```