### 2018-01-19 datasets/simulations/analysis-05-plot-psim-roc-upsetr.Rmd
Create function for generating aggregated UpSet plots

---
title: "Benchmarking FDR Methods - Simulation Summarized ROC / UpsetR Plots"
author: "Rafalab Journal Club Members"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    highlight: tango
    number_sections: true
    code_folding: hide
---

```{r, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, fig.width = 12, fig.height = 10)
```

<!-- ######################################################################## -->
## Workspace Setup
<!-- ######################################################################## -->

```{r, load-libraries, message=FALSE}
library("SummarizedBenchmark")
library("tidyverse")
```

Let's just start with a single set of simulations. These simulations include 20,000 test statistics simulated from a normal distribution, with 90% simulated from a standard normal, and 10% simulated from a mean-shifted normal distribution. The different settings correspond to different mean shifts for the alternative tests. For each setting, the simulation results are summarized over 100 replications.

```{r}
fl <- list.files("data-psim/M100/esize_fixed-bl-cubic-90-0.8",
                 pattern="M100\\.rds", full.names=TRUE)
```

<!-- ######################################################################## -->
## Average UpSet Plot Across Replicates
<!-- ######################################################################## -->

First, we consider generating UpSet plots for the average behavior over the 100 replications.

<!-- ######################################################################## -->
## Average ROC Plot Across Replicates
<!-- ######################################################################## -->

Next, we consider drawing average ROC curves for each setting...