### 2018-02-15 datasets/RNA-Seq/polyester-replicates-benchmarkFDR.Rmd

---
title: "Polyester RNA-seq simulation study"
author: "Stephanie Hicks"
output: 
    html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This is an analysis using [polyester](https://bioconductor.org/packages/release/bioc/html/polyester.html) Bioconductor package to simulate RNA-Seq reads. We will  
(1) implement null comparisons on samples simulated using polyester
with no differentially expressed (DE) genes, (2) compare FDR approaches 
on their ability to discover 'true positive' DE genes with log2 fold 
changes simulated from a normal distribution.

In this Rmd, we will perform a Monte Carlo simulation study 
and average over the replicates in plots.

## Set up workspace

```{r workspace-setup, results='hide', message=FALSE, warning=FALSE}
# Load packages needed to simulate RNA-seq counts with polyester
library(zebrafishRNASeq) 
library(genefilter)
library(limma)
library(polyester)

# Differential testing
library(DESeq2)

# Load supporting packages for data analyses
library(data.table)
library(magrittr)
library(dplyr)
library(ggplot2)
library(R.utils)
library(cowplot)
library(tibble)
library(ggthemes)

# benchmark methods
library(IHW)
library(ashr)
library(qvalue)
library(swfdr)
library(fdrtool)
# Installed BayesLogit from CRAN archive via `R CMD INSTALL BayesLogit_0.6.tar.gz`
# install_github('jgscott/FDRreg', subdir="R_pkg/")
library(FDRreg)

# comparison tools/functions
library(SummarizedBenchmark)

# load helper functions
working_dir <- list(project_home = "/users/shicks1/projects/benchmark-fdr")
sourceDirectory(file.path(working_dir$project_home,"datasets/R/"))

# # set up parallel backend:
library(doParallel)
nCores <- 1
multicoreParam <- MulticoreParam(workers = nCores)
```

## Set up Polyester

We use the [polyester](https://bioconductor.org/packages/release/bioc/html/polyester.html) Bioconductor package to simulate RNA-Seq reads. The package requires a count matrix as input which estimates the mean variance relationship. Here we start with the [zebrafish](https://bioconductor.org/packages/release/data/experiment/html/zebrafishRNASeq.html) Bioconductor data package.

Estimate mean and variance relationship

```{r estimate-parameters-NB}
data(zfGenes)
dim(zfGenes)

filter <- apply(zfGenes, 1, function(x) length(x[x>5])>=2)
counts <- as.matrix(zfGenes[filter,])

## Estimate the zero inflated negative binomial parameters
params = get_params(counts)
```

```{r}
bd <- initializeBenchDesign() # only needs to be done once
```

Try running `buildBench()` one time.
```{r}
sampleSize <- 5
nDE <- 0

groupPolyester = rep(c(-1,1),each=sampleSize)
modPolyester = model.matrix(~-1 + groupPolyester)

# things needed for differential testing using DESeq2
groupDESeq = factor(rep(c(0,1), each=sampleSize))

# NULL simulation
truth <- rep(FALSE, nrow(counts)) 
names(truth) <- rownames(counts)

# log2FC are all 0 in null simulation
log2FC = rep(0,nrow(counts))
dat_alt = create_read_numbers(params$mu, params$fit,
                              params$p0, m=dim(counts)[1], 
                              n=sampleSize*2, beta=cbind(log2FC), 
                              mod=modPolyester)

# run DESeq2
dds_alt <- DESeqDataSetFromMatrix(countData=dat_alt, 
                                  colData = DataFrame(groupDESeq), 
                                  design=~groupDESeq)
dds_alt <- DESeq(dds_alt, parallel = FALSE)
res_alt <- results(dds_alt, independentFiltering = F)

geneExp <- tbl_df(data.frame(geneName=rownames(counts),
                             pval=res_alt$pvalue, SE=res_alt$lfcSE, 
                             ind_covariate = res_alt$baseMean, 
                             effect_size=res_alt$log2FoldChange, 
                             test_statistic=res_alt$stat, qvalue=truth, 
                             pzero = rowSums(counts(dds_alt)==0)/
                                     ncol(counts(dds_alt))  ))

# filter NAs and those with less than 50% expressed 
geneExp  <- geneExp %>% na.omit() %>% dplyr::filter(pzero < 0.5)
```

```{r}
head(geneExp)
```

```{r}
# build Benchmark data
sb <- bd %>% buildBench(data=geneExp, parallel=FALSE, BPPARAM=multicoreParam, 
                        truthCols = "qvalue")
```

# Session information

```{r}
sessionInfo()
```