### 2018-02-19 datasets/microbiome/microbiome-schubert.Rmd
Start splitting microbiome analyses

---
title: "Comparison of FDR methods on 16S microbiome data with SummarizedBenchmark"
author: "Claire Duvallet"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
----

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

Here we download and analyze the CDI Schubert diarrhea (C. diff and non-C. diff diarrhea) dataset. 
We'll download the processed OTU tables from Zenodo and unzip them in the 
`benchmark-fdr/datasets/microbiome/DATA` folder.

In the previous document `cdi_schubert.Rmd`, we explored two microbiome datasets for FDR
benchmarking. Here we will make use of the new `SummarizedBenchmark` package to
perform the benchmarking in a standardized way.

# Workspace setup

```{r, wkspace-setup, results='hide', message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(magrittr)
library(SummarizedBenchmark)

## load helper functions
for (f in list.files("../R", "\\.(r|R)$", full.names = TRUE)) {
    source(f)
}

## project data/results folders
datdir <- "data"
resdir <- "results"
dir.create(datdir, showWarnings = FALSE)
dir.create(resdir, showWarnings = FALSE)
```

# Data Preparation

```{r, schubert-download}
cdi_data <- file.path(datdir, "cdi_schubert_results.tar.gz")
if (!file.exists(cdi_data)) {
    download.file("https://zenodo.org/record/840333/files/cdi_schubert_results.tar.gz",
                  destfile = cdi_data)
}

cdi_dir <- file.path(datdir, "cdi_schubert_results")
if (!file.exists(file.path(cdi_dir, "cdi_schubert.metadata.txt"))) {
    untar(cdi_data, exdir = datdir)
}
```

Next, we'll read in the unzipped OTU table and metadata files into R.

We'll need to manually define which disease labels are of interest and what metadata
column contains the sample IDs (i.e. the IDs in the OTU table).

```{r, schubert-readdata, eval=TRUE}
## load OTU table and metadata
otu <- read.table(file.path(cdi_dir, "RDP",
                            "cdi_schubert.otu_table.100.denovo.rdp_assigned"))
meta <- read.csv(file.path(cdi_dir, "cdi_schubert.metadata.txt"), sep = '\t')

# Keep only samples with the right DiseaseState metadata
meta <- filter(meta, DiseaseState %in% c("H", "nonCDI", "CDI"))

# Keep only samples with both metadata and 16S data
keep_samples <- intersect(colnames(otu), meta$sample_id)
otu <- otu[, keep_samples]
meta <- filter(meta, 'sample_id' %in% keep_samples)
```

Since we'll be using OTU-wise covariates, we shouldn't need to perform any
filtering/cleaning of the OTUs, apart from removing any that are all zeros. 
(This may happen after removing shallow samples, I think.)
However, let's make sure we get rid of any samples with too few reads. 
We define the minimum number of reads per sample in `sample_reads`.

After we've removed any shallow samples, we'll convert the OTU table to relative
abundances. I'll add a pseudo-count of 1e-6 to any zero entries, to avoid
problems with taking logs.

```{r, schubert-cleandata, eval=TRUE}
min_otu <- 10
min_sample <- 100

## Remove OTUs w/ <= min reads, samples w/ <= min reads
otu <- otu[, colSums(otu) > min_otu]
otu <- otu[rowSums(otu) > min_sample, ]

## Update metadata with new samples
meta <- filter(meta, sample_id %in% colnames(otu))

## Remove empty OTUs
otu <- otu[rowSums(otu) > 0, ]

## Convert to relative abundance
abun_otu <- t(t(otu) / rowSums(t(otu)))

## Add pseudo counts
zeroabun <- 0
abun_otu <- abun_otu + zeroabun
```

# Data Analysis

## Differential Testing

Next, we need to calculate the pvalues, effect size, and standard error for each OTU.
Here, we'll compare diarrhea vs. healthy. Diarrhea will include both CDI and nonCDI
patients. We'll put these results into a dataframe, and label the columns with the 
standardized names for downstream use (`pval`, `SE`, `effect_size`, `test_statistic`).
The test statistic is the one returned by `wilcox.test()`.

Note that the effect here is calculated as logfold change of mean abundance in controls
relative to cases (i.e. `log(mean_abun[controls]/mean_abun[cases])`)

While we're at it, we'll also calculate the mean abundance and ubiquity (detection rate)
of each OTU. Later, we can assign their values to a new column called `ind_covariate` 
for use in downstream steps.

```{r, schubert-pvals, eval=TRUE}
resfile <- file.path(resdir, paste0("schubert_results_", nrow(abun_otu), "_OTUs.rda"))
if (!file.exists(resfile)) {
    res <- test_microbiome(abundance = abun_otu, shift = zeroabun,
                           is_case = meta$DiseaseState %in% c("CDI", "nonCDI"))
    save(res, file=resfile)
} else {
  load(resfile)
}
head(res)
```

Finally, let's try to add phylogeny as covariates. Here we'll have columns for each separate taxonomic level.

```{r, schubert-addphylo, eval=TRUE}
res <- dplyr::separate(res, otu, 
                       c("kingdom", "phylum", "class", "order",
                         "family", "genus", "species", "denovo"), 
                       sep = ";", remove = FALSE)
```

## Covariate Diagnostics

Here we look to see if the covariates do indeed look informative.

### Ubiquity

```{r, schubert-ubi, fig.width=10, fig.height=3.5, eval=TRUE}
strat_hist(res, pvalue="pval", covariate="ubiquity", maxy=20, binwidth=0.05)
rank_scatter(res, pvalue="pval", covariate="ubiquity")
```

### Mean Abundance (across non-zero samples)

```{r, schubert-abun, fig.width=10, fig.height=3.5, eval=TRUE}
strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=8, binwidth=0.05)
rank_scatter(res, pvalue="pval", covariate="mean_abun_present")
```

### Phylogeny

Let's look at phylum-level stratification first. A priori, I *might* expect
Proteobacteria to be enriched for low p-values? But I don't know if that's
super legit, and Eric doesn't seem to think that phylogeny will be informative at all...

```{r, schubert-phylo, fig.width=10, fig.height=6, eval=FALSE}
ggplot(res, aes(x=pval)) +
    geom_histogram() +
    facet_wrap(~phylum, scales = "free")
```

## Multiple-Testing Correction

Let's use `ubiquity` as our `ind_covariate`.

```{r, schubert-labelcovariate, eval=TRUE}
res <- dplyr::mutate(res, ind_covariate = ubiquity)
```

First, we'll create an object of `BenchDesign` class to hold the data and 
add the benchmark methods to the `BenchDesign` object. We'll do this for a few
values of nmids, to explore its effects on the outputs.

Then, we'll construct the `SummarizedBenchmark` object, which will run
the functions specified in each method (these are actually sourced in from the
helper scripts).

```{r, schubert-sb, results="hide", message=FALSE, eval=TRUE}
for (nmids in c(5, 50, 150)) {
    print(nmids)
    bd <- initializeBenchDesign(nmids = nmids)
    ## remove ASH from comparison
    bd <- dropBMethod(bd, "ashs")

    resfile <- file.path(resdir, paste0("schubert_summarizedBenchmark_",
                                        nrow(res),  "_nmids_", nmids, ".rda"))
    duration <- NA
    if (!file.exists(resfile)) {
        sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
        metadata(sb)$data_download_link <-
                       "https://zenodo.org/record/840333/files/cdi_schubert_results.tar.gz"
        save(sb, file = resfile)
    } else {
        load(resfile)
    }
}
```

## Benchmark Metrics

Next, we'll add the default performance metric for q-value assays. First, we have
to rename the assay to 'qvalue'.

```{r, schubert-metrics, eval=TRUE}
assayNames(sb) <- "qvalue"
sb <- addDefaultMetrics(sb)
```

Now, we'll plot the results.

```{r, schubert-plot, width=15, height=15, eval=TRUE}
# plot nrejects by method overall and stratified by covariate
rejections_scatter(sb, as_fraction=FALSE, supplementary=FALSE)
rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
```

Hm, now the code runs. However, there are clearly still some issues:
- ashs rejects all hypotheses (all q-values are essentially 0).
- lfdr and scott-empirical are all NaN (I think this is likely related to the df error)

```{r, schubert-genus-covariates}
methods <- c("lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf")
plotCovariateBoxplots(sb, alpha = 0.1, nsets = 6, methods = methods)
```

```{r, eval=FALSE}
assays(sb)[["qvalue"]]["ashs"] %>% max()
sum(is.na(assays(sb)[["qvalue"]]["lfdr"]))
sum(is.na(assays(sb)[["qvalue"]]["scott-empirical"]))
sum(is.na(assays(sb)[["qvalue"]]["scott-theoretical"]))
```

Plotting methods are giving errors for some reason. Let's try to use Alejandro's code instead.

# Session Info

```{r}
sessionInfo()
```