### 2018-02-20 datasets/microbiome/microbiome-baxter.Rmd
Add more notes to Schubert; get through Baxter

---
title: "Case Study: 16S microbiome data (Baxter et al., 2016)"
author: "Claire Duvallet"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
----

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

Here we download and analyze the Baxter colorectal cancer (CRC) dataset
[(Baxter et al., 2016)](http://dx.doi.org/10.1186/s13073-016-0290-3).
The dataset includes stool samples from 172 patients with no colonic lesions,
and 120 patients with CRC. We'll download the processed OTU tables from
Zenodo and unzip them in the `benchmark-fdr/datasets/microbiome/data` folder.

For colorectal cancer, we do expect some amount of truly differentially abundant OTUs,
but not as many as in diarrhea (e.g. Schubert et al., 2014). This dataset will hopefully
provide an intermediate non-extreme case study.

# Workspace Setup

```{r, wkspace-setup, results='hide', message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(magrittr)
library(SummarizedBenchmark)

## load helper functions
for (f in list.files("../R", "\\.(r|R)$", full.names = TRUE)) {
    source(f)
}

## project data/results folders
datdir <- "data"
resdir <- "results"
dir.create(datdir, showWarnings = FALSE)
dir.create(resdir, showWarnings = FALSE)
```

# Data Preparation

```{r, baxter-download}
crc_data <- file.path(datdir, "crc_baxter_results.tar.gz")
if (!file.exists(crc_data)) {
    download.file("https://zenodo.org/record/840333/files/crc_baxter_results.tar.gz",
                  destfile = crc_data)
}

crc_dir <- file.path(datdir, "crc_baxter_results")
if (!file.exists(file.path(crc_dir, "crc_baxter.metadata.txt"))) {
    untar(crc_data, exdir = datdir)
}
```

Next, we'll read in the unzipped OTU table and metadata files into R.

```{r, baxter-readdata}
## load OTU table and metadata
otu <- read.table(file.path(crc_dir, "RDP",
                            "crc_baxter.otu_table.100.denovo.rdp_assigned"))
meta <- read.csv(file.path(crc_dir, "crc_baxter.metadata.txt"), sep='\t')

## Keep only samples with the right DiseaseState metadata
meta <- dplyr::filter(meta, DiseaseState %in% c("H", "CRC"))

## add "X" in front of sample IDs because of how R read in the OTU table
meta$Sample_Name_s <- paste0("X", meta$Sample_Name_s)

## keep only samples with both metadata and 16S data
keep_samples <- intersect(colnames(otu), meta$Sample_Name_s)
otu <- otu[, keep_samples]
meta <- dplyr::filter(meta, Sample_Name_s %in% keep_samples)
```

A brief aside: what's the count distribution of these OTUs?

```{r, eval=FALSE}
hist(colSums(otu))
```

Since we'll be using OTU-wise covariates, we shouldn't need to perform any
filtering/cleaning of the OTUs, apart from removing any that are all zeros.
(This may happen after removing shallow samples, I think.)
We still apply a minimum threshold of 10 reads per OTU across all samples.
After removing these shallow OTUs, we also get rid of any samples with too few reads. 
We define the minimum number of reads per OTU in `min_otu`, and
the minimum number of reads per sample in `min_sample`.

After we've removed any shallow OTUs and samples, we'll convert the OTU table to
relative abundances.

```{r, baxter-cleandata}
min_otu <- 10
minp_otu <- 0.01
min_sample <- 100

## Remove OTUs w/ <= min reads, w/ <= min prop, samples w/ <= min reads
otu <- otu[rowSums(otu) > min_otu, ]
out <- otu[rowSums(otu > 0) / nrow(otu) > minp_otu, ]
otu <- otu[, colSums(otu) > min_sample]

## Update metadata with new samples
meta <- dplyr::filter(meta, Sample_Name_s %in% colnames(otu))

## Remove empty OTUs
otu <- otu[rowSums(otu) > 0, ]

## Convert to relative abundance
abun_otu <- t(t(otu) / rowSums(t(otu)))

## Add pseudo counts
zeroabun <- 0
abun_otu <- abun_otu + zeroabun
```

# Data Analysis

## Differential Testing

Next, we need to calculate the pvalues, effect size, and standard error for each OTU.
Here, we'll compare CRC vs. healthy. We won't consider the nonCRC adenoma patients.
We'll put these results into a dataframe, and label the columns with the 
standardized names for downstream use (`pval`, `SE`, `effect_size`, `test_statistic`).
The test statistic is the one returned by `wilcox.test()`.

Note that the effect here is calculated as logfold change of mean abundance in controls
relative to cases (i.e. `log(mean_abun[controls]/mean_abun[cases])`)

While we're at it, we'll also calculate the mean abundance and ubiquity (detection rate)
of each OTU. Later, we can assign their values to a new column called `ind_covariate` 
for use in downstream steps.

```{r, baxter-pvals}
resfile <- file.path(resdir, paste0("baxter_results_", nrow(abun_otu), "_OTUs.RData"))
if (!file.exists(resfile)) {
    res <- test_microbiome(abundance = abun_otu, shift = zeroabun,
                           is_case = meta$DiseaseState == "CRC")
    save(res, file = resfile)
} else {
    load(resfile)
}
```

Finally, let's try to add phylogeny as covariates. Here we'll have columns for each separate taxonomic level.

```{r, baxter-addphylo}
res <- tidyr::separate(res, otu, 
                       c("kingdom", "phylum", "class", "order",
                         "family", "genus", "species", "denovo"), 
                       sep=";", remove = FALSE)
```

## Covariate Diagnostics

Here we look to see if the covariates do indeed look informative.

### Ubiquity

```{r, baxter-ubi, fig.width=10, fig.height=3.5}
strat_hist(res, pvalue="pval", covariate="ubiquity", maxy=10, binwidth=0.05)
rank_scatter(res, pvalue="pval", covariate="ubiquity")
```

Something weird is happening with p=0.20 and p=0.40 - maybe this has something to do with ties?
Either way, ubiquity looks to be a bit informative (from the scatter plot, but not really the histograms...)

### Mean Abundance (across non-zero samples)

```{r, baxter-abun, fig.width=10, fig.height=3.5}
strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=8, binwidth=0.05)
rank_scatter(res, pvalue="pval", covariate="mean_abun_present")
```

### Phylogeny

Let's look at phylum-level stratification first. A priori, I *might* expect
Proteobacteria to be enriched for low p-values? But I don't know if that's
super legit, and Eric doesn't seem to think that phylogeny will be informative at all...

```{r, baxter-phylo, fig.width=10, fig.height=6, eval=FALSE}
ggplot(res, aes(x=pval)) +
    geom_histogram() +
    facet_wrap(~phylum, scales = "free")
```

Not really informative either...

## Multiple-Testing Correction

Let's use `ubiquity` as our `ind_covariate`.

```{r, baxter-labelcovariate}
res <- dplyr::mutate(res, ind_covariate = ubiquity)
```

First, we'll create an object of `BenchDesign` class to hold the data and 
add the benchmark methods to the `BenchDesign` object. We'll do this for a few
values of nmids, to explore its effects on the outputs. We remove ASH from
the default comparison.

Then, we'll construct the `SummarizedBenchmark` object, which will run
the functions specified in each method (these are actually sourced in from the
helper scripts).

```{r, baxter-sb, results="hide", message=FALSE, eval=TRUE}
for (nmids in c(5, 50, 150)){
    print(nmids)
    bd <- initializeBenchDesign(nmids=nmids)
    ## remove ASH from comparison
    bd <- dropBMethod(bd, "ashs")

    resfile <- file.path(resdir, paste0("baxter_summarizedBenchmark_",
                                        nrow(res), "_nmids_", nmids, ".rda"))
    if (!file.exists(resfile)){
        sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
        metadata(sb)$data_download_link <-
                       "https://zenodo.org/record/840333/files/crc_baxter_results.tar.gz"
        save(sb, file = resfile)
    } else {
        load(resfile)
    }
}
```

## Benchmark Metrics

Next, we'll add the default performance metric for q-value assays. First, we have
to rename the assay to 'qvalue'.

```{r, baxter-metrics, eval=TRUE}
assayNames(sb) <- "qvalue"
sb <- addDefaultMetrics(sb)
```

Now, we'll plot the results.

```{r, baxter-plot, width=15, height=15, eval=TRUE}
## plot nrejects by method overall and stratified by covariate
rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
#plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
```

```{r, eval=FALSE}
assays(sb)[["qvalue"]][, "ashs"] %>% max()
sum(is.na(assays(sb)[["qvalue"]][, "scott-empirical"]))
sum(is.na(assays(sb)[["qvalue"]][, "lfdr"]))
```

# Data Analysis (genus-level)

## Collapse to Genus

```{r, baxter-genus}
## add column with otu names
genus_df <- as.data.frame(abun_otu)
genus_df <- dplyr::as_tibble(genus_df, rownames = "otu")

## just get genus information
genus_df <- dplyr::mutate(genus_df, genus = sapply(strsplit(otu, ";"), `[`, 6))

## gather into tidy format
genus_df <- tidyr::gather(genus_df, key = "sample", value = "abun", -otu, -genus)

## get rid of unannoated genera, and sum abundances for genera
genus_df <- genus_df %>% 
    dplyr::filter(genus != "g__") %>% 
    dplyr::group_by(genus, sample) %>% 
    dplyr::summarise(total_abun = sum(abun)) %>%
    ungroup()

## convert back to longform
genus_df <- tidyr::spread(genus_df, sample, total_abun)
genus_df <- as.data.frame(as.list(dplyr::select(genus_df, -genus)),
                          row.names = genus_df$genus)

## re-order columns to match metadata
genus_df <- genus_df[, match(meta$Sample_Name_s, colnames(genus_df))]

## use matrix
genus <- as.matrix(genus_df)
```

## Differential Testing

```{r, baxter-genus-pvals}
resfile <- file.path(resdir, paste0("baxter_results_", nrow(genus_df), "_genera.RData"))
if (!file.exists(resfile)) {
    res <- test_microbiome(abundance = genus, shift = zeroabun,
                           is_case = meta$DiseaseState == "CRC")
    save(res, file = resfile)
} else {
    load(resfile)
}
```

## Covariate Diagnostics

Here we look to see if the covariates do indeed look informative.

### Ubiquity

```{r, baxter-genus-ubi, fig.width=10, fig.height=3.5}
strat_hist(res, pvalue="pval", covariate="ubiquity", maxy=10, binwidth=0.05)
rank_scatter(res, pvalue="pval", covariate="ubiquity")
```

Something weird is happening with p=0.20 and p=0.40 - maybe this has something to do with ties?
Either way, ubiquity looks to be a bit informative (from the scatter plot, but not really the histograms...)

### Mean Abundance (across non-zero samples)

```{r, baxter-genus-abun, fig.width=10, fig.height=3.5}
strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=8, binwidth=0.05)
rank_scatter(res, pvalue="pval", covariate="mean_abun_present")
```

## Multiple-Testing Correction

Let's use `ubiquity` as our `ind_covariate`.

```{r, baxter-genus-labelcovariate}
res <- res %>% mutate(ind_covariate = ubiquity)
```

First, we'll create an object of `BenchDesign` class to hold the data and 
add the benchmark methods to the `BenchDesign` object. We'll do this for a few
values of nmids, to explore its effects on the outputs. We remove ASH from
the comparison.

Then, we'll construct the `SummarizedBenchmark` object, which will run
the functions specified in each method (these are actually sourced in from the
helper scripts).

```{r, baxter-genus-sb, results="hide", message=FALSE, eval=TRUE}
for (nmids in c(5, 50, 150)) {
    print(nmids)
    bd <- initializeBenchDesign(nmids=nmids)
    ## remove ASH from comparison
    bd <- dropBMethod(bd, "ashs")

    resfile <- file.path(resdir, paste0("baxter_summarizedBenchmark_", nrow(res),
                                        "_genera_nmids_", nmids, ".rda"))
    if (!file.exists(resfile)) {
        sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
        metadata(sb)$data_download_link <-
                       "https://zenodo.org/record/840333/files/crc_baxter_results.tar.gz"
        save(sb, file = resfile)
    } else {
        load(resfile)
    }
}
```

## Benchmark Metrics

Next, we'll add the default performance metric for q-value assays. First, we have
to rename the assay to 'qvalue'.

```{r, baxter-genus-metrics, eval=TRUE}
assayNames(sb) <- "qvalue"
sb <- addDefaultMetrics(sb)
```

Now, we'll plot the results.

```{r, baxter-genus-plot, width=15, height=15, eval=TRUE}
## plot nrejects by method overall and stratified by covariate
rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
##plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
```

Note: Benjamini-Hochberg (bh) overlaps exactly with the IHW results.

```{r, baxter-genus-overlap}
plotFDRMethodsOverlap(sb, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
```
```{r}
methods <- c( "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf" )
plotCovariateBoxplots( sb, alpha=0.1, nsets=6, methods=methods)
```

# Session Info

```{r}
sessionInfo()
```