### 2018-02-20 datasets/microbiome/microbiome-joint.Rmd
Rename/move microbiome joint analysis

---
title: "Exploration into microbiome qvalues"
author: "Claire Duvallet"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
---

In this notebook, I want to see what's going on with all the different methods for three microbiome datasets.
It seems that the two datasets I've used so far (ob_goodrich, with ~70,000 OTUs but very few true differentially abundant OTUs, and cdi_schubert, with ~18,000 OTUs and very many true differentially abundant OTUs) are getting weird results in different ways and with different methods.

My hypotheses so far are that:

- ash is rejecting all hypotheses
  - True
- lfdr is also being affected by the `nmids`/`df` [error](https://github.com/pkimes/benchmark-fdr/issues/16#issuecomment-345844180)
  - probably not a problem
- decreasing nmids doesn't fix the error/problem with Scott (FDRreg)
  - True

To look into these, I'll run SummarizedBenchmark for each value of nmids (5, 50, 150) for each dataset, and plot histograms of the resulting q values for each method.

Note that I ran SummarizedBenchmark for different values of nmids in `microbiome-SummarizedBenchmark.Rmd` and saved the results to files.

```{r, load-wkspc, results='hide', message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)

library(SummarizedBenchmark)

setwd("/Users/claire/github/benchmark-fdr/datasets/microbiome/")
datdir <- "/Users/claire/github/benchmark-fdr/datasets/microbiome/DATA/"
resdir <- "/Users/claire/github/benchmark-fdr/datasets/microbiome/RESULTS/"

```

# CDI Schubert

```{r, cdi-schubert}
dflst = list()
i <- 1
for (nmids in c(5, 50, 150)){
  print(nmids)
  resfile <- paste0(resdir, "schubert_summarizedBenchmark_",
                    "18703", "_nmids_", nmids, ".RData")
  load(resfile)

  # Extract qvalues and melt into tidy dataframe
  qvals <- assay(sb) %>% as.data.frame %>% gather("method", "qval")
  qvals$nmids <- nmids

  # TO DO: add/concatenate to other things?
  dflst[[i]] <- qvals
  i <- i + 1
}
allres <- bind_rows(dflst)
```
```{r, nmids-effect}
nmids_qvals <- c()
# First, check which methods are affected by nmids
for (met in unique(allres$method)){
  n5 <- allres %>% filter(method == met) %>% filter(nmids == 5)
  n50 <- allres %>% filter(method == met) %>% filter(nmids == 50)
  n150 <- allres %>% filter(method == met) %>% filter(nmids == 150)
  if (!(n5$qval == n50$qval && n50$qval == n150$qval) | is.na(n5$qval == n50$qval && n50$qval == n150$qval)){
    nmids_qvals <- c(nmids_qvals, met)
  }
}
nmids_qvals
```
## scott and lfdr (are messed up)

LFDR seems to work okay (though its values do depend on nmids), but Scott either returns all NaN's or all zeros...

```{r, plot-qvals}
allres %>% 
  filter(method %in% nmids_qvals) %>% 
  #mutate(logqval = log10(qval)) %>%
  ggplot(aes(x=qval)) + geom_histogram() + facet_grid(nmids ~ method)
```

```{r}
## Look at number of NaN q-values for each method
allres %>% filter(method %in% c(nmids_qvals, "ashs")) %>% group_by(method, nmids) %>% summarize(n_nan = sum(is.na(qval)))

```

### ashs (return all q-values = 0)

And one final note: all ashs qvalues are basically 0...

```{r}
allres %>% filter(method == "ashs") %>% group_by(nmids) %>% summarize(min = min(qval), max = max(qval))
```

### All other methods (seem OK)

Let's also plot the q-value distributions for all methods.

```{r}
allres %>% 
  filter(!method %in% nmids_qvals) %>% 
  filter(!method == "ashs") %>%
  #filter(!method == "unadjusted") %>%
  group_by(method, qval) %>% filter(row_number() == 1) %>%
  ggplot(aes(x=qval)) + geom_histogram() + facet_grid(method ~ .) + theme(strip.text.y = element_text(angle = 0))

```

# OB Goodrich

```{r, ob-goodrich}
dflst = list()
i <- 1
for (nmids in c(5, 50, 150)){
  print(nmids)
  resfile <- paste0(resdir, "goodrich_summarizedBenchmark_",
                    "71710", "_nmids_", nmids, ".RData")
  load(resfile)

  # Extract qvalues and melt into tidy dataframe
  qvals <- assay(sb) %>% as.data.frame %>% gather("method", "qval")
  qvals$nmids <- nmids

  # TO DO: add/concatenate to other things?
  dflst[[i]] <- qvals
  i <- i + 1
}
allres <- bind_rows(dflst)
```
```{r, nmids-effect-goodrich}
nmids_qvals <- c()
# First, check which methods are affected by nmids
for (met in unique(allres$method)){
  n5 <- allres %>% filter(method == met) %>% filter(nmids == 5)
  n50 <- allres %>% filter(method == met) %>% filter(nmids == 50)
  n150 <- allres %>% filter(method == met) %>% filter(nmids == 150)
  if (!(n5$qval == n50$qval && n50$qval == n150$qval) | is.na(n5$qval == n50$qval && n50$qval == n150$qval)){
    nmids_qvals <- c(nmids_qvals, met)
  }
}
nmids_qvals
```

## scott and lfdr (are messed up)

LFDR seems to work okay (though its values do depend on nmids), but Scott either returns all NaN's or all zeros...

```{r, plot-qvals-goodrich}
allres %>% 
  filter(method %in% nmids_qvals) %>% 
  #mutate(logqval = log10(qval)) %>%
  ggplot(aes(x=qval)) + geom_histogram() + facet_grid(nmids ~ method)
```
```{r}
## Look at number of NaN q-values for each method
allres %>% filter(method %in% c(nmids_qvals, "ashs")) %>% group_by(method, nmids) %>% summarize(n_nan = sum(is.na(qval)))

```

### ashs (return all q-values = 0)

And one final note: all ashs qvalues are basically 0...

```{r}
allres %>% filter(method == "ashs") %>% group_by(nmids) %>% summarize(min = min(qval), max = max(qval))
```

### All other methods (seem OK)

Let's also plot the q-value distributions for all methods.

```{r}
allres %>% 
  filter(!method %in% nmids_qvals) %>% 
  filter(!method == "ashs") %>%
  #filter(!method == "unadjusted") %>%
  group_by(method, qval) %>% filter(row_number() == 1) %>%
  ggplot(aes(x=qval)) + geom_histogram() + facet_grid(method ~ .) + theme(strip.text.y = element_text(angle = 0))

```

# CRC Baxter

```{r, crc-baxter}
dflst = list()
i <- 1
for (nmids in c(5, 50, 150)){
  print(nmids)
  resfile <- paste0(resdir, "baxter_summarizedBenchmark_",
                    "91843", "_nmids_", nmids, ".RData")
  load(resfile)

  # Extract qvalues and melt into tidy dataframe
  qvals <- assay(sb) %>% as.data.frame %>% gather("method", "qval")
  qvals$nmids <- nmids

  # TO DO: add/concatenate to other things?
  dflst[[i]] <- qvals
  i <- i + 1
}
allres <- bind_rows(dflst)
```

```{r, nmids-effect-baxter}
nmids_qvals <- c()
# First, check which methods are affected by nmids
for (met in unique(allres$method)){
  n5 <- allres %>% filter(method == met) %>% filter(nmids == 5)
  n50 <- allres %>% filter(method == met) %>% filter(nmids == 50)
  n150 <- allres %>% filter(method == met) %>% filter(nmids == 150)
  if (!(n5$qval == n50$qval && n50$qval == n150$qval) | is.na(n5$qval == n50$qval && n50$qval == n150$qval)){
    nmids_qvals <- c(nmids_qvals, met)
  }
}
nmids_qvals
```

## scott and lfdr

LFDR seems to work okay (though its values do depend on nmids), but Scott either returns all NaN's or all zeros...

```{r, plot-qvals-baxter}
allres %>% 
  filter(method %in% nmids_qvals) %>% 
  #mutate(logqval = log10(qval)) %>%
  ggplot(aes(x=qval)) + geom_histogram() + facet_grid(nmids ~ method)
```
```{r}
## Look at number of NaN q-values for each method
allres %>% filter(method %in% c(nmids_qvals, "ashs")) %>% group_by(method, nmids) %>% summarize(n_nan = sum(is.na(qval)))

```

### ashs

And one final note: all ashs qvalues are basically 0...

```{r}
allres %>% filter(method == "ashs") %>% group_by(nmids) %>% summarize(min = min(qval), max = max(qval))
```

### All other methods

Let's also plot the q-value distributions for all methods.

```{r}
allres %>% 
  filter(!method %in% nmids_qvals) %>% 
  filter(!method == "ashs") %>%
  #filter(!method == "unadjusted") %>%
  group_by(method, qval) %>% filter(row_number() == 1) %>%
  ggplot(aes(x=qval)) + geom_histogram() + facet_grid(method ~ .) + theme(strip.text.y = element_text(angle = 0))

```

# Conclusions:

- not sure what's going on with the scott-empirical and scott-theoretical. Seems like the same df error is being thrown, but it does run to completion *but also* all the q-values are zero (or NaN).
- ashs is also returning every q-value = 0. This is the one that needs stderr, right? Maybe I am calculating this wrong. Should check with Keegan/Patrick.