### 2018-02-21 datasets/scrnaseq/scRNAseq-latest.Rmd
AS:added 2 more datasets, helper fxn

---
title: "scRNAseq"
author: "Ayshwarya Subramanian"
date: "9/15/2017"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,cache=TRUE)

```
## R workspace setup

The downloaded plate data are stored in R objects called `MultiAssayExperiment`; this package requires BioConductor release 3.5 and R version 3.4 or greater for installation.

```{r, workspace-setup, results='hide', message=FALSE, warning=FALSE}
datdir <- "data/scrnaseq/"
resdir <- "results/scrnaseq/"
#installation
# source("https://bioconductor.org/biocLite.R")
# biocLite("swfdr")
# biocLite("IHW")
# biocLite("qvalue")
# install.packages("ashr")
# install.packages("fdrtool")
# biocLite("edgeR")
# source("https://bioconductor.org/biocLite.R")
# biocLite("scran")
#  Samseq installation
# source("https://bioconductor.org/biocLite.R")
# biocLite("impute") 
# install.packages("samr")
# source("https://bioconductor.org/biocLite.R")
#biocLite("MultiAssayExperiment")
#source("https://bioconductor.org/biocLite.R")
#biocLite("SummarizedExperiment")
#biocLite("BiocGenerics")
#biocLite("S4Vectors")
#biocLite("BiocParallel")
#install_github("areyesq89/SummarizedBenchmark")
#biocLite("EBSeq")
#biocLite("scDD")
#source("https://bioconductor.org/biocLite.R")
#biocLite("SingleCellExperiment")
#source("https://bioconductor.org/biocLite.R")
#biocLite("MAST")
#install.packages("FDRreg")

# packages for data analyses
suppressMessages(library(MultiAssayExperiment))
suppressMessages(library(dplyr))
suppressMessages(library(ggplot2))
suppressMessages(library(cowplot))
suppressMessages(library(UpSetR))
suppressMessages(library(hexbin))

# packages for differential expression testing
suppressMessages(library(impute))
#suppressMessages(library(samr)
suppressMessages(library(scDD))
suppressMessages(library(EBSeq))
suppressMessages(library(edgeR))
suppressMessages(library(MAST))

# benchmark methods
suppressMessages(library(IHW))
suppressMessages(library(ashr))
suppressMessages(library(qvalue))
suppressMessages(library(swfdr))
suppressMessages(library(fdrtool))
suppressMessages(library(FDRreg))

# summarizebenchmark
suppressMessages(library(SummarizedBenchmark))
suppressMessages(library(SingleCellExperiment))

# useful libraries
suppressMessages(library(R.utils))
suppressMessages(library(knitr))
library(BiocParallel)
cores = 6
multicoreParam <- MulticoreParam(workers = cores)
sourceDirectory('benchmark-fdr/datasets/R/')
```

## scRNAseq datasets

For the application of differential gene expression, we look at scRNAseq and bulk RNAseq data. Here we discuss the analyses of scRNAseq.

```{r echo=FALSE, results='asis'}
df=data.frame(datasets=c("GSE63818-GPL16791","GSE45719"),organism=c("human","human","mouse","mouse"),technology=c("smart-seq","smart-seq","smart-seq2","smart-seq2"))
kable(df,caption="scRNAseq Datasets")
df1=data.frame(DE_method=c("Wilcoxon-rank sum","ScDD","MAST"),input=c("tpm","count","tpm"),output=c("p-val","p-val","p-val"))
kable(df1,caption="DE methods")

```

We selected two datasets, one each from mouse and human spp, from the `Conquer` (Soneson et al., 2017) scRNAseq data collection. `Conquer` provides uniformly processed gene expression data summarized to both count (aggregated from transcript members), and length-scaled TPMs at the gene level. The authors use `salmon` to quantify raw reads and `tximport` to summarize to gene level. The data comes from Smartseq, a plate based sequencing method which results in deeper sequencing of reads at the cost of numbers of cells.

## Data Download

We first download the data in our working directory. First, let's download the Human data:GSE63818-GPL16791

```{bash, humandata-download, eval=FALSE}
# all data and results are on the wd: /net/irizarryfs01/srv/export/irizarryfs01_backed_up/share_root/fdrbenchmark-scrnaseq

#these look well separated on the tsne plots on the conquer manuscript supplement
curl -o data/scrnaseq/ http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE63818-GPL16791.rds
# qc report: 
curl -o data/scrnaseq/ http://imlspenticton.uzh.ch/robinson_lab/conquer/report-multiqc/GSE63818-GPL16791_multiqc_report.html
# scater report
curl -o data/scrnaseq/ http://imlspenticton.uzh.ch/robinson_lab/conquer/report-scater/GSE63818-GPL16791_scater.html

## human ds2
curl -o data/scrnaseq/ http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE84465.rds
```

Next, we download the Mouse data:GSE45719

```{bash, mousedata-download, eval=FALSE}
#another well-separated dataset on tsne: GSE60749−GPL13112
curl -o DATA/scrnaseq/ http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE45719.rds
```

## Data processing and filtering

For data processing, we follow practices and recommendations from Soneson et al., 2017. Briefly, we limit ourselves to 2-group comparisons, and  select the maximal size defined as the smallest number of cells in either group as the group-level sample size for the DE analyses. We do not filter genes as filtering may affect the subsequent false discovery control steps.

First, we get the human data for the two groups of interest.

```{r, dataload, eval=FALSE}
#load compressed mae
humandata=readRDS("GSE63818-GPL16791.rds")
#get gene-level data
genes=experiments(humandata)[1]
#TPMS
#get tpms
tpms=assays(genes$gene)[1]
#get tpm dataframe
human_tpms=tpms$TPM
# get column data/metadata for sample
cdata=colData(humandata)
cdata=as.data.frame(cdata)
# select groups of interest
g1 = cdata%>%subset(source_name_ch1=="Somatic Cells" & characteristics_ch1=="developmental stage: 7 week gestation")
g2 = cdata%>%subset(source_name_ch1=="Primordial Germ Cells" & characteristics_ch1=="developmental stage: 7 week gestation")
# select maximal sample size
sample_size = c(dim(g1)[1],dim(g2)[1])
ssize = which.min(sample_size)
g1_ids=sample.int(dim(g1)[1], size = sample_size[ssize], replace = FALSE)
g2_ids=sample.int(dim(g2)[1], size = sample_size[ssize], replace = FALSE)
# get group data of maximal size
g1_data=human_tpms[,colnames(human_tpms)%in%g1$geo_accession[g1_ids]]
g2_data=human_tpms[,colnames(human_tpms)%in%g2$geo_accession[g2_ids]]
#save data
saveRDS(g1_data,"h_g1_data_GSE63818-GPL16791.rds")
saveRDS(g2_data,"h_g2_data_GSE63818-GPL16791.rds")
#COUNTS
#get counts
count=assays(genes$gene)[2]
#get count dataframe
human_count=count$count
# get column data/metadata for sample
cdata=colData(humandata)
cdata=as.data.frame(cdata)
# get group data of maximal size
g1_cdata=human_count[,colnames(human_count)%in%g1$geo_accession[g1_ids]]
g2_cdata=human_count[,colnames(human_count)%in%g2$geo_accession[g2_ids]]
#save data
saveRDS(g1_cdata,"h_g1_cdata_GSE63818-GPL16791.rds")
saveRDS(g2_cdata,"h_g2_cdata_GSE63818-GPL16791.rds")

#load compressed mao
humandata2=readRDS("GSE84465.rds")
#get gene-level data
genes=experiments(humandata2)[1]
#TPMS
#get tpms
tpms=assays(genes$gene)[1]
#get tpm dataframe
human_tpms=tpms$TPM
# get column data/metadata for sample
cdata=colData(humandata2)
cdata=as.data.frame(cdata)
# select groups of interest
g1 = cdata%>%subset(characteristics_ch1.7=="neoplastic: Neoplastic" & characteristics_ch1.5=="tsne cluster: 11")
g2 = cdata%>%subset(characteristics_ch1.7=="neoplastic: Regular"& characteristics_ch1.5=="tsne cluster: 8")
# select maximal sample size
sample_size = c(dim(g1)[1],dim(g2)[1])
ssize = which.min(sample_size)
g1_ids=sample.int(dim(g1)[1], size = sample_size[ssize], replace = FALSE)
g2_ids=sample.int(dim(g2)[1], size = sample_size[ssize], replace = FALSE)
# get group data of maximal size
g1_data=human_tpms[,colnames(human_tpms)%in%g1$geo_accession[g1_ids]]
g2_data=human_tpms[,colnames(human_tpms)%in%g2$geo_accession[g2_ids]]
#save data
saveRDS(g1_data,"h_g1_data_GSE84465.rds")
saveRDS(g2_data,"h_g2_data_GSE84465rds")
#COUNTS
#get counts
count=assays(genes$gene)[2]
#get count dataframe
human_count=count$count
# get column data/metadata for sample
cdata=colData(humandata2)
cdata=as.data.frame(cdata)
# get group data of maximal size
g1_cdata=human_count[,colnames(human_count)%in%g1$geo_accession[g1_ids]]
g2_cdata=human_count[,colnames(human_count)%in%g2$geo_accession[g2_ids]]
#save data
saveRDS(g1_cdata,"h_g1_cdata_GSE84465.rds")
saveRDS(g2_cdata,"h_g2_cdata_GSE84465.rds")

```

For the human dataset1, each group has 26 cells and 65218 genes. We merge the data and remove all genes which have 0s in all samples

```{r, nofilter1, eval=FALSE}
g_data=cbind(g1_data,g2_data)
g_data_sub=g_data[rowSums(g_data==0)!=dim(g_data)[2],] 
saveRDS(g_data_sub,"h_g_nofilter_subdata_GSE63818-GPL16791.rds") # removing rows with all zeros reduces the number of genes to 33169

g_cdata=cbind(g1_cdata,g2_cdata)
g_cdata_sub=g_cdata[rowSums(g_cdata==0)!=dim(g_cdata)[2],]
saveRDS(g_cdata_sub,"h_g_nofilter_subcdata_GSE63818-GPL16791.rds") # removing rows with all zeros reduces the number of genes to 33169

saveRDS(g1_ids,"g1_ids_GSE63818-GPL16791.rds")
saveRDS(g2_ids,"g2_ids_GSE63818-GPL16791.rds")
```

For the human dataset2, each group has 695 cells and 65218 genes. We merge the data and remove all genes which have 0s in all samples

```{r, nofilter2, eval=FALSE}
g_data=cbind(g1_data,g2_data)
g_data_sub=g_data[rowSums(g_data==0)!=dim(g_data)[2],] 
saveRDS(g_data_sub,"h_g_nofilter_subdata_GSE84465.rds") # removing rows with all zeros reduces the number of genes to 52942

g_cdata=cbind(g1_cdata,g2_cdata)
g_cdata_sub=g_cdata[rowSums(g_cdata==0)!=dim(g_cdata)[2],]
saveRDS(g_cdata_sub,"h_g_nofilter_subcdata_GSE84465.rds") # removing rows with all zeros reduces the number of genes to 52942

saveRDS(g1_ids,"g1_ids_GSE84465.rds")
saveRDS(g2_ids,"g2_ids_GSE84465.rds")
```

Next, we repeat the processing steps for the mouse datasets.

```{r,mouse, eval=FALSE}
##mouse data
mousedata=readRDS("GSE45719.rds")
#get gene-level data
genes=experiments(mousedata)[1]
#TPMS
#get tpms
tpms=assays(genes$gene)[1]
#get tpm dataframe
mouse_tpms=tpms$TPM
# get column data/metadata for sample
cdata=colData(mousedata)
cdata=as.data.frame(cdata)
# select groups of interest
g1 = cdata%>%subset(source_name_ch1=="16-cell stage blastomere")
g2 = cdata%>%subset(source_name_ch1=="Mid blastocyst cell (92-94h post-fertilization)")
# select maximal sample size
sample_size = c(dim(g1)[1],dim(g2)[1])
ssize = which.min(sample_size)
g1_ids=sample.int(dim(g1)[1], size = sample_size[ssize], replace = FALSE)
g2_ids=sample.int(dim(g2)[1], size = sample_size[ssize], replace = FALSE)
# get group data of maximal size
g1_data=mouse_tpms[,colnames(mouse_tpms)%in%g1$geo_accession[g1_ids]]
g2_data=mouse_tpms[,colnames(mouse_tpms)%in%g2$geo_accession[g2_ids]]
#save data
saveRDS(g1_data,"m_g1_data_GSE45719.rds")
saveRDS(g2_data,"m_g2_data_GSE45719.rds")
#COUNTS
#get counts
count=assays(genes$gene)[2]
#get count dataframe
mouse_count=count$count
# get column data/metadata for sample
cdata=colData(mousedata)
cdata=as.data.frame(cdata)
# get group data of maximal size
g1_cdata=mouse_count[,colnames(mouse_count)%in%g1$geo_accession[g1_ids]]
g2_cdata=mouse_count[,colnames(mouse_count)%in%g2$geo_accession[g2_ids]]
#save data
saveRDS(g1_cdata,"m_g1_cdata_GSE45719.rds")
saveRDS(g2_cdata,"m_g2_cdata_GSE45719.rds")

mousedata2=readRDS("GSE94383.rds")
#get gene-level data
genes=experiments(mousedata2)[1]
#TPMS
#get tpms
tpms=assays(genes$gene)[1]
#get tpm dataframe
mouse_tpms=tpms$TPM
# get column data/metadata for sample
cdata=colData(mousedata2)
cdata=as.data.frame(cdata)
# select groups of interest
g1 = cdata%>%subset(characteristics_ch1.2=="condition: No stimulation")
g2 = cdata%>%subset(characteristics_ch1.2=="condition: LPS"&characteristics_ch1.1=="time point: 150 min")
# select maximal sample size
sample_size = c(dim(g1)[1],dim(g2)[1])
ssize = which.min(sample_size)
g1_ids=sample.int(dim(g1)[1], size = sample_size[ssize], replace = FALSE)
g2_ids=sample.int(dim(g2)[1], size = sample_size[ssize], replace = FALSE)
# get group data of maximal size
g1_data=mouse_tpms[,colnames(mouse_tpms)%in%g1$geo_accession[g1_ids]]
g2_data=mouse_tpms[,colnames(mouse_tpms)%in%g2$geo_accession[g2_ids]]
#save data
saveRDS(g1_data,"m_g1_GSE94383.rds")
saveRDS(g2_data,"m_g2_GSE94383.rds")
#COUNTS
#get counts
count=assays(genes$gene)[2]
#get count dataframe
mouse_count=count$count
# get column data/metadata for sample
cdata=colData(mousedata2)
cdata=as.data.frame(cdata)
# get group data of maximal size
g1_cdata=mouse_count[,colnames(mouse_count)%in%g1$geo_accession[g1_ids]]
g2_cdata=mouse_count[,colnames(mouse_count)%in%g2$geo_accession[g2_ids]]
#save data
saveRDS(g1_cdata,"m_g1_c_GSE94383.rds")
saveRDS(g2_cdata,"m_g2_c_GSE94383.rds")

```

For the mouse dataset1, each group has 50 cells and 45686 genes. We merge the data and remove all genes which have 0s in all samples

```{r, m_nofilter, eval=FALSE}
g_data=cbind(g1_data,g2_data)
g_data_sub=g_data[rowSums(g_data==0)!=dim(g_data)[2],] # 0.25*52=13
saveRDS(g_data_sub,"m_g_nofilter_subdata_GSE45719.rds") # removing rows with all zeros reduces the number of genes to 33169

g_cdata=cbind(g1_cdata,g2_cdata)
g_cdata_sub=g_cdata[rowSums(g_cdata==0)!=dim(g_data)[2],]
saveRDS(g_cdata_sub,"m_g_nofilter_subcdata_GSE45719.rds") # removing rows with all zeros reduces the number of genes to 33169

saveRDS(g1_ids,"m_g1_ids_GSE45719.rds")
saveRDS(g2_ids,"m_g2_ids_GSE45719.rds")
```

For the mouse dataset2, each group has 202 cells and 25179 genes after filtering for non-detected genes.

```{r, m_nofilter_ds2, eval=FALSE}
g_data=cbind(g1_data,g2_data)
g_data_sub=g_data[rowSums(g_data==0)!=dim(g_data)[2],] # 0.25*52=13
saveRDS(g_data_sub,"m_g_nofilter_subdata_GSE94383.rds") # removing rows with all zeros reduces the number of genes to 25179

g_cdata=cbind(g1_cdata,g2_cdata)
g_cdata_sub=g_cdata[rowSums(g_cdata==0)!=dim(g_data)[2],]
saveRDS(g_cdata_sub,"m_g_nofilter_subcdata_GSE94383.rds") # removing rows with all zeros reduces the number of genes to 25179

saveRDS(g1_ids,"m_g1_ids_GSE94383.rds")
saveRDS(g2_ids,"m_g2_ids_GSE94383.rds")
```

## Differential expression testing

We perform differential expression testing using 3 methods on the unfiltered data: non-parametric Wilcoxon, MAST which has a hurdle model to account for zeros and scDD which implicitly models any heterogeneity in expression. We will first assign names for all files

```{r, filenames}
h_wilcox_dat1="h_data_wilcox_p.rds"
h_wilcox_dat2="h_data_wilcox_p_GSE84465.rds"
m_wilcox_dat1="m_data_wilcox_p_GSE45719.rds"
m_wilcox_dat2="m_data_wilcox_p_GSE94383.rds"

h_scdd_dat1="humanplate_scDatEx_ks_GSE63818-GPL16791.rds"
h_scdd_dat2="humanplate_scDatEx_ks_GSE84465.rds"
m_scdd_dat1="mouseplate_scDatEx_ks_GSE45719.rds"
m_scdd_dat2="mouseplate_scDatEx_ks_GSE94383.rds"

m_mast_dat1="mouseplate_masttpm_GSE45719.rds"
h_mast_dat1="humanplate_masttpm_GSE63818-GPL16791.rds"
h_mast_dat2="humanplate_masttpm_GSE84465.rds"
m_mast_dat2="mouseplate_masttpm_GSE94383.rds"

```

### Wilcoxon rank-sum test

Per Soneson et al., we run TMM normalization to the TPM units before running Wilcoxon.

```{r, wilcoxon, eval=FALSE}

g_data_sub=readRDS("h_g_nofilter_subdata.rds")
groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
compute_wilcoxde_sc(g_data_sub,h_wilcox_dat1,groups)

g_data_sub=readRDS("h_g_nofilter_subdata_GSE84465.rds")
groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
compute_wilcoxde_sc(g_data_sub,h_wilcox_dat2,groups)

g_data_sub=readRDS("m_g_nofilter_subdata_GSE45719.rds")
groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
compute_wilcoxde_sc(g_data_sub,m_wilcox_dat1,groups)

g_data_sub=readRDS("m_g_nofilter_subdata_GSE94383.rds")
groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
compute_wilcoxde_sc(g_data_sub,m_wilcox_dat2,groups)
## warnings:
## In wilcox.test.default(x = structure(c(0, 0, 14.2107316559397,  ... :
## cannot compute exact p-value with ties

```

### Differences in distributions of gene expression in the 2 conditions

We apply scDD to log-transformed counts as below. Th helper functions set up the Single-Cell experiment object.

```{r, scdd, eval=FALSE}

## human data
g_cdata_sub=readRDS("h_g_nofilter_subcdata_GSE63818-GPL16791.rds")
compute_scdd_sc(g_cdata_sub,"h_sizefactors_GSE63818-GPL16791.rds","h_data_sce_GSE63818-GPL16791.rds",h_scdd_dat1)
##Notice: 16096 genes have less than 3 nonzero cells per condition.  Skipping these genes.
##Clustering observed expression data for each gene
##Setting up parallel back-end using 6 cores
# perm=1000
g_cdata_sub=readRDS("h_g_nofilter_subcdata_GSE84465.rds")
compute_scdd_sc(g_cdata_sub,"h_sizefactors_GSE84465.rds","h_data_sce_GSE84465.rds",h_scdd_dat2)
# Notice: 22390 genes have less than 3 nonzero cells per condition.  Skipping these genes.
# Clustering observed expression data for each gene
# Setting up parallel back-end using 6 cores

## mouse data
g_cdata_sub=readRDS("m_g_nofilter_subcdata_GSE45719.rds")
compute_scdd_sc(g_cdata_sub,"m_sizefactors_GSE45719.rds","m_data_sce_GSE45719.rds",m_scdd_dat1)
# Notice: 10246 genes have less than 3 nonzero cells per condition.  Skipping these genes.
# Clustering observed expression data for each gene
# Setting up parallel back-end using 6 cores
# Notice: Number of permutations is set to zero; using 
#             Kolmogorov-Smirnov to test for differences in distributions
#             instead of the Bayes Factor permutation test
# Classifying significant genes into patterns
g_cdata_sub=readRDS("m_g_nofilter_subcdata_GSE94383.rds")
compute_scdd_sc(g_cdata_sub,"m_sizefactors_GSE94383.rds","m_data_sce_GSE94383.rds",m_scdd_dat2)
# Notice: 8982 genes have less than 3 nonzero cells per condition.  Skipping these genes.
# Clustering observed expression data for each gene
# Setting up parallel back-end using 6 cores

```

### MAST

We run MAST on the TPMs following the approach in Soneson et al.

```{r, MAST,eval=FALSE}
g_data_sub=readRDS("m_g_nofilter_subdata_GSE45719.rds")
groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
names(groups)=colnames(g_cdata_sub)
compute_MASTtpmde(g_data_sub,groups,)

g_data_sub=readRDS("h_g_nofilter_subdata_GSE63818-GPL16791.rds")
groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
names(groups)=colnames(g_data_sub)
compute_MASTtpmde(g_data_sub,groups,h_mast_dat1)

g_data_sub=readRDS("h_g_nofilter_subdata_GSE84465.rds")
groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
names(groups)=colnames(g_data_sub)
compute_MASTtpmde(g_data_sub,groups,h_mast_dat2)

g_data_sub=readRDS("m_g_nofilter_subdata_GSE94383.rds")
groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
names(groups)=colnames(g_data_sub)
compute_MASTtpmde(g_data_sub,groups,m_mast_dat2)
```

## Covariate Diagnostics

In scrnaseq data, strength of the signal can be a function of level of gene expression. We quantify it as median or mean gene expression,and detection rate, defined as the percentage of cells expressing the gene at a level of TPM>0.

```{r,cov}
h_cov=compute_covariates_scrnaseq("h_g_nofilter_subdata.rds")
m_cov=compute_covariates_scrnaseq("m_g_nofilter_subdata.rds")
h2_cov=compute_covariates_scrnaseq("h_g_nofilter_subdata_GSE84465.rds")
m2_cov=compute_covariates_scrnaseq("m_g_nofilter_subdata_GSE94383.rds")
```

First we will look at the covariates for the Wilcoxon text

```{r, wilcoxcov, fig.width=10, fig.height=7}
print("human_plate_1")
hwilcox1 <- plot_covariate_diagnostics(paste0(h_wilcox_dat1),"wilcox",h_cov)
print(hwilcox1$plot)

print("human_plate_2")
hwilcox2 <- plot_covariate_diagnostics(paste0(h_wilcox_dat2),"wilcox",h2_cov)
print(hwilcox2$plot)

print("mouse_plate_1")
mwilcox1 <- plot_covariate_diagnostics(paste0(m_wilcox_dat1),"wilcox",m_cov)
print(mwilcox1$plot)

print("mouse_plate_2")
mwilcox2 <- plot_covariate_diagnostics(paste0(m_wilcox_dat2),"wilcox",m2_cov)
print(mwilcox2$plot)
```

Next, we look at the covariates for MAST

```{r, mastcov, fig.width=10, fig.height=7}
print("human_plate_1")
hmast1 <- plot_covariate_diagnostics(paste0(,h_mast_dat1),"mast",h_cov)
print(hmast1$plot)

print("human_plate_2")
hmast2 <- plot_covariate_diagnostics(paste0(h_mast_dat2),"mast",h2_cov)
print(hmast2$plot)

print("mouse_plate_1")
mmast1 <- plot_covariate_diagnostics(paste0(m_mast_dat1),"mast",m_cov)
print(mmast1$plot)

print("mouse_plate_2")
mmast2 <- plot_covariate_diagnostics(paste0(m_mast_dat2),"mast",m2_cov)
print(mmast2$plot)
```

Finally, we look at how informative the covariates are with results from ScDD.

```{r,scddcov, fig.width=10, fig.height=7}
print("human_plate_1")
hscdd1 <- plot_covariate_diagnostics(paste0(h_scdd_dat1),"scdd",h_cov)
print(hscdd1$plot)

print("human_plate_2")
hscdd2 <- plot_covariate_diagnostics(paste0(h_scdd_dat2),"scdd",h2_cov)
print(hscdd2$plot)

print("mouse_plate_1")
mscdd1 <- plot_covariate_diagnostics(paste0(m_scdd_dat1),"scdd",m_cov)
print(mscdd1$plot)

print("mouse_plate_2")
mscdd2 <- plot_covariate_diagnostics(paste0(m_scdd_dat2),"scdd",m2_cov)
print(mscdd2$plot)
```

### Benchmarking FDR methods

First we will create the benchdesign object for wilcox -- IN PROGRESS

```{r, benchmarking, eval=FALSE}

bd <- initializeBenchDesign()
resN <- paste0(resdir, "scrnaseq_wilcox_summarizedBenchmark_",nrow(m_scdd_cov), ".RData")

duration <- NA
if (!file.exists(resN)){
  t1 <- proc.time()
  scrna_sb <- bd %>% buildBench(data=(hscdd1$data%>%select(pval,mean_exp)%>%mutate(effect_size=0, test_statistic=0, SE=0) %>% mutate(ind_covariate=mean_exp)), 
                          ftCols = c("mean_exp"),
                          parallel=TRUE, BPPARAM=multicoreParam)

  save(scrna_sb, file=resN)
  duration <- round((proc.time()-t1)[3]/60,1)
}else{
  load(resN)
}

assayNames(scrna_sb) <- "qvalue"
scrna_sb <- addDefaultMetrics(scrna_sb)

# plot nrejects by method overall and stratified by covariate
rejections_scatter(scrna_sb,supplementary=FALSE)

rejection_scatter_bins(scrna_sb, covariate="ind_covar_N", bins=3,supplementary=FALSE)

# upset plot 
plotFDRMethodsOverlap(scrna_sb, 
                      alpha=0.05, nsets=ncol(scrna_sb),
                      order.by="freq", decreasing=TRUE,
                      supplementary=FALSE)

# properties of method-exclusive discoveries
methods <- c("lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf")
mcols(scrna_sb)$ind_covariate <- mcols(scrna_sb)$ind_covar_N
plotCovariateBoxplots(scrna_sb, alpha=0.1, nsets=6, methods=methods, trans=log2, maxNum=3)