### 2018-02-22 datasets/GSEA/gsea-human.Rmd
Rename GSEA analysis Rmd

---
title: "Case Study: Gene Set Enrichment Analysis (GSEA)"
author: "Alejandro Reyes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

**Some sort of summary of the analysis.**

# Workspace Setup

```{r}
library(dplyr)
library(ggplot2)
library(scales)
library(DESeq2)
library(EnsDb.Hsapiens.v86)
library(biomaRt)
library(goseq)

## load helper functions
for (f in list.files("../R", "\\.(r|R)$", full.names = TRUE)) {
    source(f)
}

## project data/results folders
datdir <- "data"
resdir <- "results"
dir.create(datdir, showWarnings = FALSE)
dir.create(resdir, showWarnings = FALSE)
```

# Data Preparation

Download *DESeqDataSet* if not present locally.

```{r, message=FALSE}
dsd_data <- file.path(datdir, "dsd.RData")
if (!file.exists(dsd_data)) {
    download.file("https://www.dropbox.com/s/rdmv3tyqt4g37b0/dsd.RData?dl=1",
                  destfile = dsd_data)
}
load(dsd_data)
```

**What are we doing here?**

```{r}
levels(colData(dsdObject)$individual) <- gsub("GTEX-", "", levels(colData(dsdObject)$individual))
```

We only perform differential tesing on protein coding genes, as specified in Ensembl release 86.

```{r}
gnType <-  genes(EnsDb.Hsapiens.v86, column = "gene_biotype")
dsdObject <- dsdObject[rownames(dsdObject) %in% gnType[gnType$gene_biotype == "protein_coding"]$gene_id, ]
```

# Data Analysis

## Differential Testing

Specify model matrix design and run DESeq2 to get differentially expressed genes.

```{r deseq2, message=FALSE}
dsdObject <- DESeq(dsdObject)
res <- results(dsdObject, independentFiltering = FALSE)
```

## Enrichment Analysis

**Add description of what's going on.**

```{r goseq, message=FALSE}
goset_data <- file.path(datdir, "goSets.RData")
if (!file.exists(goset_data)) {
    mart <- useMart("ensembl", "hsapiens_gene_ensembl")
    goSets <- getBM( c("ensembl_gene_id", "go_id"), mart = mart,
                    filters = "ensembl_gene_id", values = rownames(res))
    goSets <- goSets[!nchar( goSets$go_id ) == 0, ]
    goSets <- with(goSets, split(go_id, ensembl_gene_id))
    save(goSets, file = file)
} else {
    load(goset_data)
}

genes <- as.numeric(res$padj < 0.1)
names(genes) <- rownames(res)

### getting median transcript length
txByGene <- transcriptsBy(EnsDb.Hsapiens.v86, "gene")
geneLength <- sapply( width( txByGene ), median )
geneLength <- geneLength[names(genes)]
genes[is.na( genes )] <- 0
```

Performing a gene set enrichment analysis with `goseq`.

```{r}
pwf <- nullp(genes, bias.data = geneLength)
goRes <- goseq(pwf, gene2cat = goSets)
```

## Covariate Diagnostics

Here, we can see that the covariant variable is actually informative.

### Gene Set Size

**Add description of independent covariate.**

```{r informativeCovariate, message=FALSE, fig.height=3.5, message=FALSE, fig.width=4.5}
rank_scatter( dat=goRes, pval="over_represented_pvalue", 
              covariate="numInCat", bins=50, funx=log2, 
              funfill=log10_trans()) +
    ylim(0, 12) + ggtitle("Over-represented gene sets") +
    xlab( expression(log[10]~"(# of genes)")) +
    ylab( expression(-log[10]~"(p-value)") )

```

```{r assumptions, message=FALSE, fig.width=10, fig.height=3.2}
## filter numInCat below only used for plotting
## filter numDEInCat for cases where there are no DE genes in the category (these are actually not tested)
#dplyr:::filter( goRes, over_represented_pvalue == 1, numDEInCat > 0)
dplyr:::filter( goRes, numDEInCat > 0, numInCat > 10 ) %>%
  strat_hist( pval="over_represented_pvalue", covariate="numInCat", maxy=10)
```

## Multiple-Testing Correction

Generating the *SummarizedBenchmark* object:

```{r runningBenckmark, message=FALSE}
## rename columns and prepare for benchmarking
res <- dplyr:::select( goRes, c("over_represented_pvalue", "numInCat") ) %>%
    dplyr:::rename( pval = over_represented_pvalue, ind_covariate = numInCat )
res$pval <- pmin( res$pval, 1 )

## generate default BenchDesign
designBen <- initializeBenchDesign()

## run benchmark
sGSEA <- buildBench(designBen, res, ftCols="ind_covariate")
```

## Benchmark Metrics

```{r}
assayNames(sGSEA) <- "qvalue"
sGSEA <- addDefaultMetrics(sGSEA)
sGSEA <- estimatePerformanceMetrics(sGSEA, addColData=TRUE)
```

```{r plottingResults}
rejections_scatter(sGSEA, as_fraction=FALSE, supplementary=FALSE)
```

```{r}
plotFDRMethodsOverlap(sGSEA, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100)
```

Gene set size of overlaps.

```{r}
methods <- c( "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf" )
plotCovariateBoxplots(sGSEA, alpha=0.1, nsets=6, methods=methods, trans=log2)
plotCovariateBoxplots(sGSEA, alpha=0.1, nsets=7, methods=methods, trans=log2, maxNum=5)
```

# Data Analysis (Null Comparison)

**Add description.**

## Create Null Data

```{r, eval=FALSE }
nms <- names( genes )
xx <- seq_len(length(nms))
gns <- 1*( xx %in% sample( xx, 1000, replace=FALSE ) )
names( gns ) <- nms
pwf <- nullp( gns, bias.data=geneLength )
goRes <- goseq( pwf, gene2cat=goSets )
res <- goRes %>%
    dplyr:::filter( numDEInCat > 0 ) %>%
    dplyr:::select( c("over_represented_pvalue", "numInCat") ) %>%
    dplyr:::rename( pval = over_represented_pvalue, ind_covariate = numInCat )
```

## Multiple-Testing Correction

```{r}
pr <- designBen %>% buildBench( res, ftCols="ind_covariate")
assayNames(pr) <- "qvalue"
pr <- addDefaultMetrics(pr, metrics="rejections")
```

## Benchmark Metrics

```{r}
estimatePerformanceMetrics( pr, alpha=0.1, tidy=TRUE ) %>%
    filter( blabel %in% c("lfdr","ihw-a10","bl-df03","qvalue","bh", "bonf") ) %>%
    select( blabel, value )
```