### 2018-02-22 datasets/GSEA/gsea.Rmd
Standardize format of gsea

---
title: "GSEA datasets for FDR benchmarking"
author: "Alejandro Reyes (on behalf of the Rafalab Benchmarking Consortium)"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
    html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

Here we will make use of the `SummarizedBenchmark` package to benchmark two 
Gene Set Enrichment Analysis (GSEA) case studies.

# Workspace setup

```{r, wkspace-setup, results='hide', message=FALSE, warning=FALSE}
# data and results directories
system("mkdir -p DATA")
system("mkdir -p RESULTS")
datdir <- "DATA/"
resdir <- "RESULTS/"

# GSEA analysis tools
library(DESeq2)
library(goseq)
library(EnsDb.Hsapiens.v86)
library(EnsDb.Mmusculus.v75)
library(biomaRt)

# wrangling & plotting tools
library(ggplot2)
library(cowplot)
library(scales)
library(dplyr)

# comparison tools/functions
library(R.utils)
sourceDirectory("../R/")
```

# GTEX expression data

## Differential expression analysis with DESeq2

First we'll download the dataset from GTEX, which is saved as an `RData` file and
contains a *DESeqDataSet* object.

```{r, message=FALSE}

file <- paste0(datdir, "dsd.RData")
if(!file.exists( file ) ){
    download.file("https://www.dropbox.com/s/rdmv3tyqt4g37b0/dsd.RData?dl=1",
                  destfile=file)
}
load(file)

levels(colData(dsdObject)$individual) <- gsub("GTEX-", "",levels(colData(dsdObject)$individual))

```

Specify model matrix design and run **DESeq2** to get differentially expressed genes.

```{r deseq2, message=FALSE}
gnType <-  genes(EnsDb.Hsapiens.v86, column="gene_biotype")

dsdObject <- dsdObject[rownames(dsdObject) %in% gnType[gnType$gene_biotype == "protein_coding"]$gene_id,]
dsdObject <- DESeq(dsdObject)
res <- results( dsdObject, independentFiltering=FALSE )
```

## Gene set enrichment analysis with goseq

Next we'll use the 
**biomaRt** package to download GO gene set annotations (for humans since the
GTEX data contains human samples).

```{r biomart, message=FALSE}
file <- paste0(datdir, "goSets.RData")
if(!file.exists(file)){
  mart <- useMart( "ensembl", "hsapiens_gene_ensembl" )
  goSets <- getBM( c("ensembl_gene_id", "go_id"), mart=mart, 
                   filters="ensembl_gene_id", values=rownames(res))
  goSets <- goSets[!nchar( goSets$go_id ) == 0,]
  goSets <- with( goSets, split( go_id, ensembl_gene_id ) )
  save(goSets, file=file)
}else{
  load(file)
}
```

Finally, we'll perform a gene set enrichment analysis with **goseq**. We'll 
include all genes with a **DESeq2** adjusted p-value less than 0.10, and adjust
for gene length since not doing so may bias the results toward GO sets with 
longer genes.

```{r goseq, message=FALSE}
genes <- as.numeric( res$padj < 0.1 )
names( genes ) <- rownames( res )

### getting median transcript length
txByGene <- transcriptsBy(EnsDb.Hsapiens.v86, "gene")
geneLength <- sapply( width( txByGene ), median )
geneLength <- geneLength[names(genes)]
genes[is.na( genes )] <- 0

pwf <- nullp(genes, bias.data=geneLength )
goRes <- goseq( pwf, gene2cat=goSets )
```

## Check Covariate Diagnostics

```{r informativeCovariate, message=FALSE, fig.height=3.5, message=FALSE, fig.width=4.5}
rank_scatter( dat=goRes, pval="over_represented_pvalue", 
              covariate="numInCat", bins=50, funx=log2, 
              funfill=log10_trans()) +
    ylim(0, 12) + ggtitle("Over-represented gene sets") +
    xlab( expression(log[10]~"(# of genes)")) +
    ylab( expression(-log[10]~"(p-value)") )

```

```{r assumptions, message=FALSE, fig.width=10, fig.height=3.2}
## filter numInCat below only used for plotting
## filter numDEInCat for cases where there are no DE genes in the category (these are actually not tested)
#dplyr:::filter( goRes, over_represented_pvalue == 1, numDEInCat > 0)

dplyr:::filter( goRes, numDEInCat > 0, numInCat > 10 ) %>%
  strat_hist( pval="over_represented_pvalue", covariate="numInCat", maxy=10)

```

## Run benchmarks on set size covariate

Generating the *SummarizedBenchmark* object:

```{r runningBenckmark, message=FALSE}
designBen <- initializeBenchDesign()
res <- dplyr:::select( goRes, c("over_represented_pvalue", "numInCat") ) %>%
  dplyr:::rename( pval = over_represented_pvalue, ind_covariate = numInCat )
res$pval <- pmin( res$pval, 1 )
sGSEA <- designBen %>% buildBench( res, ftCols="ind_covariate")
assayNames(sGSEA) <- "qvalue"
sGSEA <- addDefaultMetrics( sGSEA )
```

```{r plottingResults}
rejections_scatter( sGSEA, as_fraction=FALSE, supplementary=FALSE)

sGSEA <- estimatePerformanceMetrics(sGSEA, addColData=TRUE)

```

```{r}
plotFDRMethodsOverlap( sGSEA, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
```

## Gene set size of overlaps

```{r}

methods <- c( "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf" )
plotCovariateBoxplots( sGSEA, alpha=0.1, nsets=6, methods=methods, trans=log2 )

```

## NULL comparisons

```{r, eval=FALSE }

nms <- names( genes )
xx <- seq_len(length(nms))
gns <- 1*( xx %in% sample( xx, 1000, replace=FALSE ) )
names( gns ) <- nms
pwf <- nullp( gns, bias.data=geneLength )
goRes <- goseq( pwf, gene2cat=goSets )
res <- goRes %>%
  dplyr:::filter( numDEInCat > 0 ) %>%
  dplyr:::select( c("over_represented_pvalue", "numInCat") ) %>%
  dplyr:::rename( pval = over_represented_pvalue, ind_covariate = numInCat )
pr <- designBen %>% buildBench( res, ftCols="ind_covariate")
assayNames(pr) <- "qvalue"
pr <- addDefaultMetrics(pr, metrics="rejections")

estimatePerformanceMetrics( pr, alpha=0.1, tidy=TRUE ) %>%
  filter( blabel %in% c("lfdr","ihw-a10","bl-df03","qvalue","bh", "bonf") ) %>%
  select( blabel, value )

```

# Same analysis but using the HSC-MPP dataset from Nina

```{r}

dsdMouse <- "dsdHSCMPP.RData"
if( !file.exists( dsdMouse ) ){
  download.file("https://www.dropbox.com/s/spo1pifum5k72rn/dseHSCMPP.RData?dl=1", destfile="dsdHSCMPP.RData")
}
load(dsdMouse)
dseHSCMPP <- DESeq( dseHSCMPP)

res <- results( dseHSCMPP, contrast=c("conditions", "HSC", "MPP1"), independentFiltering=FALSE )
genes <- as.numeric( res$padj < 0.1 )
names( genes ) <- rownames( res )
genes[is.na( genes )] <- 0
table(genes)

### preparing gene length info and goSets ###
txByGene <- transcriptsBy(EnsDb.Mmusculus.v75, "gene")
geneLength <- sapply( width( txByGene ), median )
geneLength <- geneLength[names(genes)]

file="goSetsMm.RData"
if(!file.exists(file)){
  stopifnot( require( biomaRt ) )
  mart <- useMart( "ensembl", "mmusculus_gene_ensembl" )
  goSets <- getBM( c("ensembl_gene_id", "go_id"), mart=mart, filters="ensembl_gene_id", values=rownames(res))
  goSets <- goSets[!nchar( goSets$go_id ) == 0,]
  goSets <- with( goSets, split( go_id, ensembl_gene_id ) )
  save(goSets, file=file)
}else{
  load(file)
}

pwf <- nullp(genes, bias.data=geneLength )
goRes <- goseq( pwf, gene2cat=goSets )

res <- dplyr:::select( goRes, c("over_represented_pvalue", "numInCat") ) %>%
  dplyr:::rename( pval = over_represented_pvalue, ind_covariate = numInCat )
res$pval <- pmin( res$pval, 1 )
sGSEAMouse <- designBen %>% buildBench( res, ftCols="ind_covariate")
assayNames(sGSEAMouse) <- "qvalue"
sGSEAMouse <- addDefaultMetrics( sGSEAMouse )

rank_scatter( dat=goRes, pval="over_represented_pvalue", 
              covariate="numInCat", bins=50, funx=log2, 
              funfill=log10_trans()) +
    ylim(0, 12) + ggtitle("Over-represented gene sets") +
    xlab( expression(log[10]~"(# of genes)")) +
    ylab( expression(-log[10]~"(p-value)") )

dplyr:::filter( goRes, numDEInCat > 0, numInCat > 20)  %>%
  strat_hist( pval="over_represented_pvalue", covariate="numInCat", maxy=7)

rejections_scatter( sGSEAMouse, as_fraction=FALSE, supplementary=FALSE)

plotFDRMethodsOverlap( sGSEAMouse, supplementary=FALSE, order.by="freq")

plotCovariateBoxplots( sGSEAMouse, alpha=0.1, nsets=4, methods=methods, trans=log2, maxNum=5)

plotCovariateBoxplots( sGSEA, alpha=0.1, nsets=7, methods=methods, trans=log2, maxNum=5)

```