### 2018-02-23 datasets/GSEA/gsea-mouse.Rmd

---
title: "Case Study: Gene Set Enrichment Analysis (GSEA), Mouse"
author: "Alejandro Reyes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

**Some sort of summary of the analysis.**
Same analysis but using the HSC-MPP dataset from Nina.  
**Who is Nina?**

# Workspace Setup

```{r}
library(dplyr)
library(ggplot2)
library(scales)
library(DESeq2)
library(EnsDb.Mmusculus.v75)
library(biomaRt)
library(goseq)

## load helper functions
for (f in list.files("../R", "\\.(r|R)$", full.names = TRUE)) {
    source(f)
}

## project data/results folders
datdir <- "data"
resdir <- "results"
dir.create(datdir, showWarnings = FALSE)
dir.create(resdir, showWarnings = FALSE)
```

# Data Preparation

Download *DESeqDataSet* if not present locally.

```{r, message=FALSE}
dsd_data <- "dsdHSCMPP.RData"
if(!file.exists(dsd_data)) {
    download.file("https://www.dropbox.com/s/spo1pifum5k72rn/dseHSCMPP.RData?dl=1",
                  destfile = dsd_data)
}
load(dsd_data)
```

**Why no filtering on gene_biotype == "protein_coding" for this data?**

# Data Analysis

## Differential Testing

Specify model matrix design and run DESeq2 to get differentially expressed genes.

```{r deseq2, message=FALSE}
dseHSCMPP <- DESeq(dseHSCMPP)
res <- results(dseHSCMPP, contrast = c("conditions", "HSC", "MPP1"),
               independentFiltering = FALSE)
```

## Enrichment Analysis

**Add description of what's going on.**

```{r goseq, message=FALSE}
goset_data <- file.path(datdir, "goSetsMm.RData")
if (!file.exists(goset_data)) {
    mart <- useMart("ensembl", "mmusculus_gene_ensembl")
    goSets <- getBM(c("ensembl_gene_id", "go_id"), mart = mart,
                    filters = "ensembl_gene_id", values = rownames(res))
    goSets <- goSets[!nchar(goSets$go_id) == 0, ]
    goSets <- with(goSets, split(go_id, ensembl_gene_id))
    save(goSets, file = goset_data)
} else {
   load(goset_data)
}

## getting median transcript length
genes <- as.numeric(res$padj < 0.1)
names(genes) <- rownames(res)

## getting median transcript length
txByGene <- transcriptsBy(EnsDb.Mmusculus.v75, "gene")
geneLength <- sapply(width(txByGene), median)
geneLength <- geneLength[names(genes)]
genes[is.na( genes )] <- 0
```

Performing a gene set enrichment analysis with `goseq`.

```{r}
pwf <- nullp(genes, bias.data = geneLength)
goRes <- goseq(pwf, gene2cat = goSets)
```

## Covariate Diagnostics

Here, we can see that the covariant variable is actually informative.

### Gene Set Size

**Add description of independent covariate.**

```{r informativeCovariate, message=FALSE, fig.height=3.5, message=FALSE, fig.width=4.5}
rank_scatter(dat = goRes, pval = "over_represented_pvalue", 
              covariate = "numInCat", bins = 50, funx = log2, 
              funfill = log10_trans()) +
    ylim(0, 12) +
    ggtitle("Over-represented gene sets") +
    xlab(expression(log[10]~"(# of genes)")) +
    ylab(expression(-log[10]~"(p-value)")) 
```

```{r assumptions, message=FALSE, fig.width=10, fig.height=3.2}
## filter numInCat below only used for plotting
## filter numDEInCat for cases where there are no DE genes in the category (these are actually not tested)
#dplyr::filter( goRes, over_represented_pvalue == 1, numDEInCat > 0)
dplyr::filter(goRes, numDEInCat > 0, numInCat > 10) %>%
    strat_hist(pval = "over_represented_pvalue", covariate = "numInCat", maxy=10)
```

## Multiple-Testing Correction

Generating the *SummarizedBenchmark* object:

```{r runningBenckmark, message=FALSE}
## rename columns and prepare for benchmarking
res <- dplyr::select(goRes, over_represented_pvalue, numInCat) %>%
    dplyr::rename(pval = over_represented_pvalue,
                  ind_covariate = numInCat) %>%
    dplyr::mutate(pval = pmin(pval, 1))

## generate default BenchDesign
designBen <- initializeBenchDesign()

## run benchmark
sGSEAMouse <- buildBench(designBen, res, ftCols = "ind_covariate")
```

## Benchmark Metrics

```{r}
assayNames(sGSEAMouse) <- "qvalue"
sGSEAMouse <- addDefaultMetrics(sGSEAMouse)
```

```{r plottingResults}
rejections_scatter(sGSEAMouse, as_fraction = FALSE, supplementary = FALSE)
```

```{r}
plotFDRMethodsOverlap(sGSEAMouse, alpha = 0.1, supplementary = FALSE,
                      order.by = "freq", nsets = 100)
```

Gene set size of overlaps.

```{r}
methods <- c( "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf" )
plotCovariateBoxplots(sGSEAMouse, alpha = 0.1, nsets = 4,
                      methods = methods, trans = log2, maxNum = 5)
```

# Session Info

```{r}
sessionInfo()
```