### 2018-02-23 datasets/GSEA/gsea.Rmd
Add details to gsea

---
title: "GSEA datasets for FDR benchmarking"
author: "Alejandro Reyes (on behalf of the Rafalab Benchmarking Consortium)"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
    html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

Here we will make use of the `SummarizedBenchmark` package to benchmark two 
Gene Set Enrichment Analysis (GSEA) case studies.

# Workspace setup

```{r, wkspace-setup, results='hide', message=FALSE, warning=FALSE}
# data and results directories
system("mkdir -p DATA")
system("mkdir -p RESULTS")
datdir <- "DATA/"
resdir <- "RESULTS/"

# GSEA analysis tools
library(DESeq2)
library(goseq)
library(EnsDb.Hsapiens.v86)
library(EnsDb.Mmusculus.v75)
library(biomaRt)

# wrangling & plotting tools
library(ggplot2)
library(cowplot)
library(scales)
library(dplyr)

# comparison tools/functions
library(R.utils)
sourceDirectory("../R/")
```

# GTEX expression data

First we'll download the dataset from GTEX, which is saved as an `RData` file and
contains a *DESeqDataSet* object.

```{r, message=FALSE}

file <- paste0(datdir, "dsd.RData")
if(!file.exists( file ) ){
    download.file("https://www.dropbox.com/s/rdmv3tyqt4g37b0/dsd.RData?dl=1",
                  destfile=file)
}
load(file)

levels(colData(dsdObject)$individual) <- gsub("GTEX-", "",levels(colData(dsdObject)$individual))

```

## Differential expression analysis with DESeq2

Check model matrix design and run **DESeq2** to get differentially expressed genes. 
The design includes a factor for tissue (here we have Cortex and Cerebellum), 
and adjusts for individual.

```{r deseq2, message=FALSE}
colData(dsdObject)
design(dsdObject)

gnType <-  genes(EnsDb.Hsapiens.v86, column="gene_biotype")

dsdObject <- dsdObject[rownames(dsdObject) %in% gnType[gnType$gene_biotype == "protein_coding"]$gene_id,]
dsdObject <- DESeq(dsdObject)
res <- results( dsdObject, contrast=c("tissue", "Cortex", "Cerebellum"),
                independentFiltering=FALSE)

```

## Gene set enrichment analysis with goseq

Next we'll use the 
**biomaRt** package to download GO gene set annotations (for humans since the
GTEX data contains human samples).

```{r biomart, message=FALSE}
file <- paste0(datdir, "goSets.RData")
if(!file.exists(file)){
  mart <- useMart( "ensembl", "hsapiens_gene_ensembl" )
  goSets <- getBM( c("ensembl_gene_id", "go_id"), mart=mart, 
                   filters="ensembl_gene_id", values=rownames(res))
  goSets <- goSets[!nchar( goSets$go_id ) == 0,]
  goSets <- with( goSets, split( go_id, ensembl_gene_id ) )
  save(goSets, file=file)
}else{
  load(file)
}
```

Finally, we'll perform a gene set enrichment analysis with **goseq**. We'll 
include all genes with a **DESeq2** adjusted p-value less than 0.10, and adjust
for gene length since not doing so may bias the results toward GO sets with 
longer genes. The bias is seen in the plot below, with gene length bin on the
x-axis and proportion of genes detected as DE on the y-axis.

```{r goseq, message=FALSE}
genes <- as.numeric( res$padj < 0.1 )
names( genes ) <- rownames( res )

### getting median transcript length
txByGene <- transcriptsBy(EnsDb.Hsapiens.v86, "gene")
geneLength <- sapply( width( txByGene ), median )
geneLength <- geneLength[names(genes)]
genes[is.na( genes )] <- 0

pwf <- nullp(genes, bias.data=geneLength )
goRes <- goseq( pwf, gene2cat=goSets )
```

## Check Covariate Diagnostics

First we'll look at a scatterplot of the relationship between the pvalue and the number 
of genes in the gene set.

```{r informativeCovariate, message=FALSE, fig.height=3.5, message=FALSE, fig.width=4.5}
rank_scatter( dat=goRes, pval="over_represented_pvalue", 
              covariate="numInCat", bins=50, funx=log2, 
              funfill=log10_trans()) +
    ylim(0, 12) + ggtitle("Over-represented gene sets") +
    xlab( expression(log[10]~"(# of genes)")) +
    ylab( expression(-log[10]~"(p-value)") )

```

Next we'll look at the histogram of p-values stratified by quantiles of the 
covariate (number of genes in the gene set).

```{r assumptions, message=FALSE, fig.width=10, fig.height=3.2}
## filter numInCat below only used for plotting
## filter numDEInCat for cases where there are no DE genes in the category (these are actually not tested)

dplyr:::filter( goRes, numDEInCat > 0, numInCat > 10 ) %>%
  strat_hist( pval="over_represented_pvalue", covariate="numInCat", maxy=10)

```

## Run benchmarks on set size covariate

Now we'll generate the *SummarizedBenchmark* object and run the benchmark methods.
Note that we cannot the ash method since we do not have an effect size for this
test. We also can't run the scott method since we don't have a test statistic.

```{r runningBenckmark, message=FALSE}
designBen <- initializeBenchDesign()
res <- dplyr:::select( goRes, c("over_represented_pvalue", "numInCat") ) %>%
  dplyr:::rename( pval = over_represented_pvalue, ind_covariate = numInCat )
res$pval <- pmin( res$pval, 1 )
sGSEA <- designBen %>% buildBench( res, ftCols="ind_covariate")
assayNames(sGSEA) <- "qvalue"
sGSEA <- addDefaultMetrics( sGSEA )
```

Finally, we'll visualize the results by looking at the number of rejections at
several alpha levels.

```{r plottingResults}
rejections_scatter( sGSEA, as_fraction=FALSE, supplementary=FALSE)

sGSEA <- estimatePerformanceMetrics(sGSEA, addColData=TRUE)

```

In addition, we'll look at the overlap of gene sets found by each method at the
alpha cutoff of 0.10.

```{r}
plotFDRMethodsOverlap( sGSEA, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100 )
```

Finally, we'll examine whether there are any preferences for methods to detect
gene sets with certain covariate values.

```{r}

methods <- c( "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf" )
plotCovariateBoxplots( sGSEA, alpha=0.1, nsets=6, methods=methods, trans=log2 )

```

## Null comparison

Here we'll evaluate specificity of the methods by investigating performance in 
the null setting. Ideally, no significant gene sets should be found when we 
perform the analysis on a set of random genes instead of a set of DE genes.
Here we sample 1000 random genes and perform a GSEA.

```{r, eval=FALSE }
set.seed(4885)
nms <- names( genes )
xx <- seq_len(length(nms))
gns <- 1*( xx %in% sample( xx, 1000, replace=FALSE ) )
names( gns ) <- nms
pwf <- nullp( gns, bias.data=geneLength )
goRes <- goseq( pwf, gene2cat=goSets )
```

We'll use the same covariate (gene set size) and run the benchmark methods
on the GSEA results.

```{r}
res <- goRes %>%
  dplyr:::filter( numDEInCat > 0 ) %>%
  dplyr:::select( c("over_represented_pvalue", "numInCat") ) %>%
  dplyr:::rename( pval = over_represented_pvalue, ind_covariate = numInCat )

pr <- designBen %>% buildBench( res, ftCols="ind_covariate")
assayNames(pr) <- "qvalue"
pr <- addDefaultMetrics(pr, metrics="rejections")

estimatePerformanceMetrics( pr, alpha=0.1, tidy=TRUE ) %>%
  filter( blabel %in% c("lfdr","ihw-a10","bl-df03","qvalue","bh", "bonf") ) %>%
  select( blabel, value )

```

# Same analysis but using the HSC-MPP dataset from Nina

Add some description of the HSC-MPP dataset.

```{r}
dsdMouse <- "dsdHSCMPP.RData"
if( !file.exists( dsdMouse ) ){
  download.file("https://www.dropbox.com/s/spo1pifum5k72rn/dseHSCMPP.RData?dl=1", destfile="dsdHSCMPP.RData")
}
load(dsdMouse)
```

## Differential expression analysis with DESeq2

```{r}
dseHSCMPP <- DESeq( dseHSCMPP)

res <- results( dseHSCMPP, contrast=c("conditions", "HSC", "MPP1"), independentFiltering=FALSE )
genes <- as.numeric( res$padj < 0.1 )
names( genes ) <- rownames( res )
genes[is.na( genes )] <- 0
table(genes)
```

## Gene set enrichment analysis with goseq

```{r}
### preparing gene length info and goSets ###
txByGene <- transcriptsBy(EnsDb.Mmusculus.v75, "gene")
geneLength <- sapply( width( txByGene ), median )
geneLength <- geneLength[names(genes)]

file="goSetsMm.RData"
if(!file.exists(file)){
  stopifnot( require( biomaRt ) )
  mart <- useMart( "ensembl", "mmusculus_gene_ensembl" )
  goSets <- getBM( c("ensembl_gene_id", "go_id"), mart=mart, filters="ensembl_gene_id", values=rownames(res))
  goSets <- goSets[!nchar( goSets$go_id ) == 0,]
  goSets <- with( goSets, split( go_id, ensembl_gene_id ) )
  save(goSets, file=file)
}else{
  load(file)
}

pwf <- nullp(genes, bias.data=geneLength )
goRes <- goseq( pwf, gene2cat=goSets )

res <- dplyr:::select( goRes, c("over_represented_pvalue", "numInCat") ) %>%
  dplyr:::rename( pval = over_represented_pvalue, ind_covariate = numInCat )
res$pval <- pmin( res$pval, 1 )
```
## Check Covariate Diagnostics

```{r, message=FALSE, fig.height=3.5, message=FALSE, fig.width=4.5}
rank_scatter( dat=goRes, pval="over_represented_pvalue", 
              covariate="numInCat", bins=50, funx=log2, 
              funfill=log10_trans()) +
    ylim(0, 12) + ggtitle("Over-represented gene sets") +
    xlab( expression(log[10]~"(# of genes)")) +
    ylab( expression(-log[10]~"(p-value)") )

```

```{r, message=FALSE, fig.width=10, fig.height=3.2}
## filter numInCat below only used for plotting
## filter numDEInCat for cases where there are no DE genes in the category (these are actually not tested)

dplyr:::filter( goRes, numDEInCat > 0, numInCat > 10 ) %>%
  strat_hist( pval="over_represented_pvalue", covariate="numInCat", maxy=10)

```

## Run benchmarks on set size covariate

```{r}
sGSEAMouse <- designBen %>% buildBench( res, ftCols="ind_covariate")
assayNames(sGSEAMouse) <- "qvalue"
sGSEAMouse <- addDefaultMetrics( sGSEAMouse )

rank_scatter( dat=goRes, pval="over_represented_pvalue", 
              covariate="numInCat", bins=50, funx=log2, 
              funfill=log10_trans()) +
    ylim(0, 12) + ggtitle("Over-represented gene sets") +
    xlab( expression(log[10]~"(# of genes)")) +
    ylab( expression(-log[10]~"(p-value)") )

dplyr:::filter( goRes, numDEInCat > 0, numInCat > 20)  %>%
  strat_hist( pval="over_represented_pvalue", covariate="numInCat", maxy=7)

rejections_scatter( sGSEAMouse, as_fraction=FALSE, supplementary=FALSE)

plotFDRMethodsOverlap( sGSEAMouse, supplementary=FALSE, order.by="freq")

plotCovariateBoxplots( sGSEAMouse, alpha=0.1, nsets=4, methods=methods, trans=log2, maxNum=5)

plotCovariateBoxplots( sGSEA, alpha=0.1, nsets=7, methods=methods, trans=log2, maxNum=5)

```