### 2018-02-26 datasets/scrnaseq/scRNAseq-keegan.Rmd
Add checks for precomputed objects; Add scDD evaluation

---
title: "scRNAseq"
author: "Ayshwarya Subramanian and Keegan Korthauer"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,cache=TRUE,cache.lazy = FALSE)
```
# Summary

For the application of differential gene expression, we look at scRNAseq and bulk RNAseq data. Here we discuss the analyses of scRNAseq.

```{r echo=FALSE, results='asis'}
library(knitr)
df=data.frame(datasets=c("GSE63818-GPL16791","GSE84465","GSE45719","GSE94383"),
              organism=c("human","human","mouse","mouse"),
              technology=c("smart-seq","smart-seq","smart-seq2","smart-seq2"),
              comparison=c("Somatic vs Primordial Germ Cells", 
                           "Neoplastic vs Regular",
                           "16-cell stage blastomere vs Mid blastocyst cell",
                           "LPS vs No stimulation"))
kable(df,caption="scRNAseq Datasets")
df1=data.frame(DE_method=c("Wilcoxon-rank sum","ScDD","MAST"),
               input=c("tpm","count","tpm"),
               output=c("p-val","p-val","p-val"))
kable(df1,caption="DE methods")
```

We selected four datasets, two each from mouse and human. 
These datasets come from the `Conquer` (Soneson et al., 2017) scRNAseq data collection. `Conquer` provides uniformly processed gene expression data summarized to both count (aggregated from transcript members), and length-scaled TPMs at the gene level. The authors used `salmon` to quantify raw reads and `tximport` to summarize to gene level. The data comes from Smartseq, a plate based sequencing method which results in deeper sequencing of reads at the cost of numbers of cells.

Here we are interested in performing differential expression analyses to determine
which

# Workspace setup

The downloaded plate data are stored in R objects called `MultiAssayExperiment`; this package requires BioConductor release 3.5 and R version 3.4 or greater for installation.

```{r, workspace-setup, results='hide', message=FALSE, warning=FALSE}
system("mkdir -p DATA")
system("mkdir -p RESULTS")
datdir <- "DATA/"
resdir <- "RESULTS/"

# packages for data analyses
library(MultiAssayExperiment)
library(dplyr)
library(ggplot2)
library(cowplot)
library(UpSetR)
library(hexbin)

# packages for differential expression testing
library(impute)
library(scDD)
library(EBSeq)
library(edgeR)
library(MAST)

# summarizebenchmark
library(SummarizedBenchmark)
library(SingleCellExperiment)

# useful libraries
library(R.utils)
library(BiocParallel)

cores <- 8
BiocParallel::register(MulticoreParam(workers = cores))

sourceDirectory("../R/")
```

# Data preparation

We first download the data in the data directory.`conquer` provides the processed datasets in convenient .rds files. First, let's download the Human datasets: GSE63818-GPL16791 and
GSE84465.

```{r, humandata-download}
#these look well separated on the tsne plots on the conquer manuscript supplement
download.file("http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE63818-GPL16791.rds",
  destfile=paste0(datdir, "GSE63818-GPL16791.rds"))

## human ds2
download.file("http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE84465.rds",
  destfile=paste0(datdir, "GSE84465.rds"))

```

Next, we download the Mouse data: GSE45719 and GSE94383

```{r, mousedata-download}
#another well-separated dataset on tsne: GSE60749−GPL13112
download.file("http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE45719.rds",
  destfile=paste0(datdir, "GSE45719.rds"))

download.file(url="http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE94383.rds",
  destfile=paste0(datdir,"GSE94383.rds"))
```

## Data processing and filtering

For data processing, we follow practices and recommendations from Soneson et al., 2017. Briefly, we limit ourselves to 2-group comparisons. We do not filter genes as filtering may affect the subsequent false discovery control steps.

First, we get the human data for the two groups of interest.
We'll extract two conditions: Primordial Germ Cells and Somatic cells at 
7 weeks gestation.

```{r, dataload}
datfile_h1 <- paste0(datdir, "h1_GSE63818-GPL16791.rds")

if (!file.exists(datfile_h1)){

   # load MultiAssayExperiment
  humandata <- readRDS(paste0(datdir,"GSE63818-GPL16791.rds"))

  # subset by groups of interest
  humandata <- humandata[, (colData(humandata)$source_name_ch1 %in% 
                         c("Somatic Cells", "Primordial Germ Cells")) &
                         (colData(humandata)$characteristics_ch1 == 
                         c("developmental stage: 7 week gestation"))]
  cdata=colData(humandata)
  cdata$group <- cdata$source_name_ch1

  # subset by gene expression TPM & count assays
  humandata <- experiments(humandata)[[1]]
  assays(humandata) <- assays(humandata)[c("TPM", "count")]
  colData(humandata) <- cdata

  # subset by genes not all zero
  humandata <- humandata[rowSums(assays(humandata)[["count"]]==0) < ncol(humandata),]

  # look at number of cells in each group 
  table(colData(humandata)$group)

  # save SummarizedExperiment 
  saveRDS(humandata, datfile_h1)
}else{
  humandata <- readRDS(datfile_h1)
}
```

For the human dataset1 there are `r ncol(humandata)` cells total and 
`r nrow(humandata)` genes.

Now humandata2. We'll extract two conditions: Neoplastic and Regular cells, as
defined by their tsne clusters.

```{r}
datfile_h2 <- paste0(datdir, "h2_GSE84465.rds")

if (!file.exists(datfile_h2)){

  # load MultiAssayExperiment
  humandata2 <- readRDS(paste0(datdir, "GSE84465.rds"))

  # subset by groups of interest
  humandata2 <- humandata2[, (colData(humandata2)$characteristics_ch1.7 %in% 
                         c("neoplastic: Neoplastic", "neoplastic: Regular")) &
                         (colData(humandata2)$characteristics_ch1.5 %in% 
                         c("tsne cluster: 11",  "tsne cluster: 8"))]
  cdata=colData(humandata2)
  cdata$group <- cdata$characteristics_ch1.7

  # subset by gene expression TPM & count assays
  humandata2 <- experiments(humandata2)[[1]]
  assays(humandata2) <- assays(humandata2)[c("TPM", "count")]
  colData(humandata2) <- cdata

  # subset by genes not all zero
  humandata2 <- humandata2[rowSums(assays(humandata2)[["count"]]==0) < ncol(humandata2),]

  # look at number of cells in each group
  table(colData(humandata2)$group)

  # save SummarizedExperiment 
  saveRDS(humandata2, datfile_h2)
}else{
  humandata2 <- readRDS(datfile_h2)
}
```

For the human dataset2, there are a total of `r nrow(humandata2)` cells and `r ncol(humandata2)` genes.

Next, we repeat the processing steps for the mouse datasets. For the first one,
we'll extract two conditions: 16-cell stage blastomere and mid blastocysts cells.

```{r,mouse}
datfile_m1 <- paste0(datdir, "m1_GSE45719.rds")

if (!file.exists(datfile_m1)){

  # load MultiAssayExperiment
  mousedata=readRDS(paste0(datdir, "GSE45719.rds"))

  # subset by groups of interest
  mousedata <- mousedata[, colData(mousedata)$source_name_ch1 %in% 
                         c("16-cell stage blastomere", 
                           "Mid blastocyst cell (92-94h post-fertilization)")]
  cdata=colData(mousedata)
  cdata$group <- cdata$source_name_ch1

  # subset by gene expression TPM & count assays
  mousedata <- experiments(mousedata)[[1]]
  assays(mousedata) <- assays(mousedata)[c("TPM", "count")]
  colData(mousedata) <- cdata

  # subset by genes not all zero
  mousedata <- mousedata[rowSums(assays(mousedata)[["count"]]==0) < ncol(mousedata),]

  # look at number of cells in each group
  table(as.character(colData(mousedata)$group))

  # save SummarizedExperiment 
  saveRDS(mousedata, datfile_m1)
}else{
  mousedata <- readRDS(datfile_m1)
}
```

For the mouse dataset1, we have a total of `r ncol(mousedata)` cells and 
`r nrow(mousedata)` genes.

Now mousedata2. We'll extract two conditions: No stimulation versus LPS at 150 min.

```{r}
datfile_m2 <- paste0(datdir, "m2_GSE94383.rds")

if (!file.exists(datfile_m2)){

  # load MultiAssayExperiment
  mousedata2=readRDS(paste0(datdir,"GSE94383.rds"))

  # subset by groups of interest
  mousedata2 <- mousedata2[, (colData(mousedata2)$characteristics_ch1.2 ==
                              "condition: No stimulation") |
                           (colData(mousedata2)$characteristics_ch1.2 ==
                              "condition: LPS" &
                            colData(mousedata2)$characteristics_ch1.1 %in% 
                              "time point: 150 min") ]
  cdata=colData(mousedata2)
  cdata$group <- cdata$characteristics_ch1.2

  # subset by gene expression TPM & count assays
  mousedata2 <- experiments(mousedata2)[[1]]
  assays(mousedata2) <- assays(mousedata2)[c("TPM", "count")]
  colData(mousedata2) <- cdata

  # subset by genes not all zero
  mousedata2 <- mousedata2[rowSums(assays(mousedata2)[["count"]]==0) < ncol(mousedata2),]

  # look at number of cells in each group
  table(as.character(colData(mousedata2)$group))

  # save SummarizedExperiment 
  saveRDS(mousedata2, datfile_m2)
}else{
  mousedata2 <- readRDS(datfile_m2) 
}
```

For the mouse dataset2, we have a total of `r ncol(mousedata2)` cells and 
`r nrow(mousedata2)` genes after filtering for non-detected genes.

## Differential expression testing

We perform differential expression testing using 3 methods on the unfiltered data: non-parametric Wilcoxon, MAST which has a hurdle model to account for zeros and scDD which implicitly models any heterogeneity in expression.

### Wilcoxon rank-sum test

Per Soneson et al., we run TMM normalization to the TPM units before running Wilcoxon.

```{r, wilcoxon}
# function to add wilcox p-values
compute_wilcoxde_sc <- function(dat){
  # check if already computed
  if (! "wilcox_p" %in% colnames(rowData(dat))){
    expr <- assays(dat)[["TPM"]] 
    tmm <- edgeR::calcNormFactors(expr)
    expr <- edgeR::cpm(expr, lib.size = tmm * colSums(expr))
    rowData(dat)$wilcox_p <- sapply(seq_len(nrow(expr)), 
      function(i) {
        wilcox.test(expr[i, ] ~ colData(dat)$group)$p.value
      })
  }else{
     message("Wilcox p-values already computed.")
  }
  return(dat)
}

humandata <- compute_wilcoxde_sc(humandata)

humandata2 <- compute_wilcoxde_sc(humandata2)

mousedata <- compute_wilcoxde_sc(mousedata)

mousedata2 <- compute_wilcoxde_sc(mousedata2)

# save results
saveRDS(humandata, datfile_h1)
saveRDS(humandata2, datfile_h2)
saveRDS(mousedata, datfile_m1)
saveRDS(mousedata2, datfile_m2)
```

### Differences in distributions of gene expression in the 2 conditions

We apply scDD to normalized log-transformed counts.

```{r, scdd, eval=FALSE}
# function to add scDD p-values
compute_scdd_sc <- function(dat){
  # check if already computed
  if (! "scdd_p" %in% colnames(rowData(dat))){
    sce <- SingleCellExperiment(assays=list(count=assays(dat)[["count"]]),
                                colData=data.frame(colData(dat)))
    # scran normalization
    libsize <- scran::computeSumFactors(assay(sce, "count"))
    normcounts(sce) <- GetNormalizedMat(assay(sce, "count"), libsize)

    rowData(dat)$scdd_p <- 
        results(scDD(sce, 
             condition = "group",
             prior_param = list(alpha=0.01, mu0=0, s0=0.01, a0=0.01, b0=0.01), 
             testZeroes = FALSE))$nonzero.pvalue
  }else{
     message("scDD p-values already computed.")
  }
  return(dat)
}

## human data

compute_scdd_sc(humandata)

compute_scdd_sc(humandata2)

compute_scdd_sc(mousedata)

compute_scdd_sc(mousedata2)

```

### MAST
** Still need to clean up from here on; nothing below is eval'd**

We run MAST on the TPMs following the approach in Soneson et al.

```{r, MAST,eval=FALSE, eval=FALSE}
g_data_sub=readRDS(paste0(datdir,"m_g_nofilter_subdata_GSE45719.rds"))
groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
names(groups)=colnames(g_cdata_sub)
compute_MASTtpmde(g_data_sub,groups,)

g_data_sub=readRDS(paste0(datdir,"h_g_nofilter_subdata_GSE63818-GPL16791.rds"))
groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
names(groups)=colnames(g_data_sub)
compute_MASTtpmde(g_data_sub,groups,h_mast_dat1)

g_data_sub=readRDS(paste0(datdir,"h_g_nofilter_subdata_GSE84465.rds"))
groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
names(groups)=colnames(g_data_sub)
compute_MASTtpmde(g_data_sub,groups,h_mast_dat2)

g_data_sub=readRDS(paste0(datdir,"m_g_nofilter_subdata_GSE94383.rds"))
groups=c(rep(1,(dim(g_data_sub)[2]/2)),rep(2,(dim(g_data_sub)[2]/2)))
names(groups)=colnames(g_data_sub)
compute_MASTtpmde(g_data_sub,groups,m_mast_dat2)
```

## Covariate Diagnostics

In scrnaseq data, strength of the signal can be a function of level of gene expression. We quantify it as median or mean gene expression,and detection rate, defined as the percentage of cells expressing the gene at a level of TPM>0.

```{r,cov, eval=FALSE}
h_cov=compute_covariates_scrnaseq(paste0(datdir,"h_g_nofilter_subdata_GSE63818-GPL16791.rds"))
m_cov=compute_covariates_scrnaseq(paste0(datdir,"m_g_nofilter_subdata_GSE45719.rds"))
h2_cov=compute_covariates_scrnaseq(paste0(datdir,"h_g_nofilter_subdata_GSE84465.rds"))
m2_cov=compute_covariates_scrnaseq(paste0(datdir,"m_g_nofilter_subdata_GSE94383.rds"))
```

First we will look at the covariates for the Wilcoxon text

```{r, wilcoxcov, fig.width=10, fig.height=7, eval=FALSE}
print("human_plate_1")
hwilcox1 <- plot_covariate_diagnostics(paste0(h_wilcox_dat1),"wilcox",h_cov)
print(hwilcox1$plot)

print("human_plate_2")
hwilcox2 <- plot_covariate_diagnostics(paste0(h_wilcox_dat2),"wilcox",h2_cov)
print(hwilcox2$plot)

print("mouse_plate_1")
mwilcox1 <- plot_covariate_diagnostics(paste0(m_wilcox_dat1),"wilcox",m_cov)
print(mwilcox1$plot)

print("mouse_plate_2")
mwilcox2 <- plot_covariate_diagnostics(paste0(m_wilcox_dat2),"wilcox",m2_cov)
print(mwilcox2$plot)
```

Next, we look at the covariates for MAST

```{r, mastcov, fig.width=10, fig.height=7, eval=FALSE}
print("human_plate_1")
hmast1 <- plot_covariate_diagnostics(paste0(h_mast_dat1),"mast",h_cov)
print(hmast1$plot)

print("human_plate_2")
hmast2 <- plot_covariate_diagnostics(paste0(h_mast_dat2),"mast",h2_cov)
print(hmast2$plot)

print("mouse_plate_1")
mmast1 <- plot_covariate_diagnostics(paste0(m_mast_dat1),"mast",m_cov)
print(mmast1$plot)

print("mouse_plate_2")
mmast2 <- plot_covariate_diagnostics(paste0(m_mast_dat2),"mast",m2_cov)
print(mmast2$plot)
```

Finally, we look at how informative the covariates are with results from ScDD.

```{r,scddcov, fig.width=10, fig.height=7, eval=FALSE}
print("human_plate_1")
hscdd1 <- plot_covariate_diagnostics(paste0(h_scdd_dat1),"scdd",h_cov)
print(hscdd1$plot)

print("human_plate_2")
hscdd2 <- plot_covariate_diagnostics(paste0(h_scdd_dat2),"scdd",h2_cov)
print(hscdd2$plot)

print("mouse_plate_1")
mscdd1 <- plot_covariate_diagnostics(paste0(m_scdd_dat1),"scdd",m_cov)
print(mscdd1$plot)

print("mouse_plate_2")
mscdd2 <- plot_covariate_diagnostics(paste0(m_scdd_dat2),"scdd",m2_cov)
print(mscdd2$plot)
```

### Benchmarking FDR methods

First we will create the benchdesign object for wilcox -- IN PROGRESS

```{r, benchmarking, eval=FALSE, eval=FALSE}

bd <- initializeBenchDesign()
resN <- paste0(resdir, "scrnaseq_wilcox_summarizedBenchmark_",nrow(m_scdd_cov), ".RData")

duration <- NA
if (!file.exists(resN)){
  t1 <- proc.time()
  scrna_sb <- bd %>% buildBench(data=(hscdd1$data%>%select(pval,mean_exp)%>%mutate(effect_size=0, test_statistic=0, SE=0) %>% mutate(ind_covariate=mean_exp)), 
                          ftCols = c("mean_exp"),
                          parallel=TRUE, BPPARAM=multicoreParam)

  save(scrna_sb, file=resN)
  duration <- round((proc.time()-t1)[3]/60,1)
}else{
  load(resN)
}

assayNames(scrna_sb) <- "qvalue"
scrna_sb <- addDefaultMetrics(scrna_sb)

# plot nrejects by method overall and stratified by covariate
rejections_scatter(scrna_sb,supplementary=FALSE)

rejection_scatter_bins(scrna_sb, covariate="ind_covar_N", bins=3,supplementary=FALSE)

# upset plot 
plotFDRMethodsOverlap(scrna_sb, 
                      alpha=0.05, nsets=ncol(scrna_sb),
                      order.by="freq", decreasing=TRUE,
                      supplementary=FALSE)

# properties of method-exclusive discoveries
methods <- c("lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf")
mcols(scrna_sb)$ind_covariate <- mcols(scrna_sb)$ind_covar_N
plotCovariateBoxplots(scrna_sb, alpha=0.1, nsets=6, methods=methods, trans=log2, maxNum=3)
```