### 2018-02-27 datasets/scrnaseq/scrnaseq-mouse-mast.Rmd

---
title: "Case Study: scRNA-seq (Mouse data, MAST method)"
author: "Ayshwarya Subramanian and Keegan Korthauer"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,cache=TRUE,cache.lazy = FALSE)
```
# Summary

For the application of differential gene expression, we look at scRNAseq and bulk RNAseq data. Here we discuss the analyses of scRNAseq.

```{r echo=FALSE, results='asis'}
library(knitr)
df=data.frame(datasets=c("GSE84465","GSE94383"),
              organism=c("human","mouse"),
              technology=c("smart-seq","smart-seq2"),
              comparison=c("Neoplastic vs Regular",
                           "LPS vs No stimulation"))
kable(df,caption="scRNAseq Datasets")
df1=data.frame(DE_method=c("ScDD","MAST"),
               input=c("count","tpm"),
               output=c("p-val","p-val"))
kable(df1,caption="DE methods")
```

We selected two datasets, one each from mouse and human. 
These datasets come from the `Conquer` (Soneson et al., 2017) scRNAseq data collection. `Conquer` provides uniformly processed gene expression data summarized to both count (aggregated from transcript members), and length-scaled TPMs at the gene level. The authors used `salmon` to quantify raw reads and `tximport` to summarize to gene level. The data comes from Smartseq, a plate based sequencing method which results in deeper sequencing of reads at the cost of numbers of cells.

Here we are interested in performing differential expression analyses to determine
which genes have different expression among the two biological groups in each comparison.
We also apply two different analysis methods: MAST and scDD. In this vignette, we will
explore the mouse dataset analysed with MAST.

# Workspace setup

The downloaded plate data are stored in R objects called `MultiAssayExperiment`; this package requires BioConductor release 3.5 and R version 3.4 or greater for installation.

```{r, workspace-setup, results='hide', message=FALSE, warning=FALSE}
system("mkdir -p DATA")
system("mkdir -p RESULTS")
datdir <- "DATA/"
resdir <- "RESULTS/"

# packages for data analyses
library(MultiAssayExperiment)
library(dplyr)
library(ggplot2)
library(cowplot)
library(UpSetR)
library(hexbin)

# packages for differential expression testing
library(impute)
library(scDD)
library(EBSeq)
library(edgeR)
library(MAST)

# summarizebenchmark
library(SummarizedBenchmark)
library(SingleCellExperiment)

# useful libraries
library(R.utils)
library(BiocParallel)

cores <- 6
BiocParallel::register(MulticoreParam(workers = cores))

sourceDirectory("../R/")
```

# Data preparation

We first download the data in the data directory.`conquer` provides the processed datasets in convenient .rds files. We download the Mouse data from GSE94383

```{r, mousedata-download}
download.file(url="http://imlspenticton.uzh.ch/robinson_lab/conquer/data-mae/GSE94383.rds",
  destfile=paste0(datdir,"GSE94383.rds"))
```

## Data processing and filtering

For data processing, we follow practices and recommendations from Soneson et al., 2017. Briefly, we limit ourselves to 2-group comparisons. We do not filter genes as filtering may affect the subsequent false discovery control steps.

We'll extract two conditions: No stimulation versus LPS at 150 min.

```{r}
datfile <- paste0(datdir, "mouse_GSE94383.rds")

if (!file.exists(datfile)){

  # load MultiAssayExperiment
  mousedata=readRDS(paste0(datdir,"GSE94383.rds"))

  # subset by groups of interest
  mousedata <- mousedata[, (colData(mousedata)$characteristics_ch1.2 ==
                              "condition: No stimulation") |
                           (colData(mousedata)$characteristics_ch1.2 ==
                              "condition: LPS" &
                            colData(mousedata)$characteristics_ch1.1 %in% 
                              "time point: 150 min") ]
  cdata=colData(mousedata)
  cdata$group <- cdata$characteristics_ch1.2

  # subset by gene expression TPM & count assays
  mousedata <- experiments(mousedata)[[1]]
  assays(mousedata) <- assays(mousedata)[c("TPM", "count")]
  colData(mousedata) <- cdata

  # subset by genes not all zero
  mousedata <- mousedata2[rowSums(assays(mousedata)[["count"]]==0) < ncol(mousedata),]

  # look at number of cells in each group
  table(as.character(colData(mousedata2)$group))

  # save SummarizedExperiment 
  saveRDS(mousedata, datfile)
}else{
  mousedata <- readRDS(datfile) 
}
```

For the mouse dataset2, we have a total of `r ncol(mousedata2)` cells and 
`r nrow(mousedata2)` genes after filtering for non-detected genes.

## Differential expression testing

We perform differential expression testing using 
MAST which has a hurdle model to account for zeroes.

We apply MAST to the TPMs following the approach in Soneson et al.

```{r, MAST}
# function to add MAST p-values
compute_MAST_sc <- function(dat){
  # check if already computed
  if (! "mast_p" %in% colnames(rowData(dat))){
    sca <- FromMatrix(exprsArray = log2(assays(dat)[["TPM"]] + 1), 
                    cData = data.frame(wellKey=colnames(dat), colData(dat)))
    zlmdata <- zlm.SingleCellAssay(~group, sca)

    rowData(dat)$mast_p <- 
        lrTest(zlmdata, "group")[, "hurdle", "Pr(>Chisq)"]
  }else{
     message("MAST p-values already computed.")
  }
  return(dat)
}

compute_MAST_sc(mousedata)

# save results
saveRDS(mousedata, datfile)
```

## Covariate Diagnostics

In scrnaseq data, strength of the signal can be a affected by the of level of gene expression. 
We will explore two potential covariates related to expression level: mean expression 
and detection rate (defined as the proportion of cells expressing the gene at a nonzero level). 
These will also be added to the `rowData` slot of the `SummarizedExperiments`.

```{r,cov}
add_covariates_scrnaseq <- function(dat){
  rowData(dat)$meanExp <- rowMeans(assays(dat)[["count"]])
  rowData(dat)$detection <- rowMeans(assays(dat)[["count"]]==0)
  return(dat)
}

mousedata <- add_covariates_scrnaseq(mousedata)

# save results
saveRDS(mousedata, datfile)
```

### Covariate one: Mean Expression

For each covariate, we'll examine the plots for each of the two analysis methods.

```{r, fig.width=4.5, fig.height=3.5, message=FALSE}
rank_scatter(data.frame(rowData(mousedata)), 
             pvalue="mast_p", covariate="meanExp") + 
  ggtitle("MAST, Covariate 1: Mean Expression")
```

```{r, fig.width=10, fig.height=3.2, message=FALSE}
strat_hist(data.frame(rowData(mousedata)), 
           pvalue="mast_p", covariate="meanExp", maxy=25) 
```

The mean expression appears to be informative and approximately satisfies the 
assumptions

### Covariate two: Detection Rate

Next we look at the detection rate covariate.

```{r, fig.width=4.5, fig.height=3.5, message=FALSE}
rank_scatter(data.frame(rowData(mousedata)), 
             pvalue="mast_p", covariate="detection") + 
  ggtitle("MAST, Covariate 2: Detection Rate")
```

```{r, fig.width=10, fig.height=3.2, message=FALSE}
strat_hist(data.frame(rowData(mousedata)), 
           pvalue="mast_p", covariate="detection", maxy=25) 
```

### Benchmarking FDR methods

** Still need to clean up from here on; nothing below is eval'd**

```{r, benchmarking, eval=FALSE}

bd <- initializeBenchDesign()
resN <- paste0(resdir, "scrnaseq_wilcox_summarizedBenchmark_",nrow(m_scdd_cov), ".RData")

duration <- NA
if (!file.exists(resN)){
  t1 <- proc.time()
  scrna_sb <- bd %>% buildBench(data=(hscdd1$data%>%select(pval,mean_exp)%>%mutate(effect_size=0, test_statistic=0, SE=0) %>% mutate(ind_covariate=mean_exp)), 
                          ftCols = c("mean_exp"),
                          parallel=TRUE, BPPARAM=multicoreParam)

  save(scrna_sb, file=resN)
  duration <- round((proc.time()-t1)[3]/60,1)
}else{
  load(resN)
}

assayNames(scrna_sb) <- "qvalue"
scrna_sb <- addDefaultMetrics(scrna_sb)

# plot nrejects by method overall and stratified by covariate
rejections_scatter(scrna_sb,supplementary=FALSE)

rejection_scatter_bins(scrna_sb, covariate="ind_covar_N", bins=3,supplementary=FALSE)

# upset plot 
plotFDRMethodsOverlap(scrna_sb, 
                      alpha=0.05, nsets=ncol(scrna_sb),
                      order.by="freq", decreasing=TRUE,
                      supplementary=FALSE)

# properties of method-exclusive discoveries
methods <- c("lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf")
mcols(scrna_sb)$ind_covariate <- mcols(scrna_sb)$ind_covar_N
plotCovariateBoxplots(scrna_sb, alpha=0.1, nsets=6, methods=methods, trans=log2, maxNum=3)
```