### 2018-02-27 datasets/ChIP-seq/ChIPseq-promotors.Rmd

---
title: "Case Study: ChIP-seq Differential Peak Calling (predefined regions)"
author: "Mingxiang Teng"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

In differential peak calling, classical pipelines often begin by first calling peaks
in samples, followed by performing differential testing using one of a few methods,
e.g. _csaw_ or _DiffBind_. However, with these approaches, testing is only performed on a
filtered set of genomic regions, namely those with significance in at least one condition,
resulting in a potentially limited number of null tests cases. An analysis using this
classical approach is performed in the accompanying `ChIPseq-peaks.Rmd` case study.

To avoid this limitation, in this case study we examine ChIP-seq data of the histone
H3K4Me3, an active marker of gene expression, and focus our analysis on promoter regions
of all genes for this histone marker. By doing so, we avoid limiting the set of tests
to only regions significant in at least one sample, while also keeping most regions of
interest for this histone marker.

As in `ChIPseq-peaks.Rmd`, we test for differential peaks using publicly available ChIP-seq
data for histone H3K4Me3 from two cell lines (K562 and Gm12878). Data are obtained from the
ENCODE Project. Differential peaks are tested using DESeq2.

# Workspace Setup

```{r, wkspace-setup, results='hide', message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(SummarizedBenchmark)
library(BiocParallel)
library(GenomicRanges)
library(TxDb.Hsapiens.UCSC.hg19.knownGene)
library(DESeq2)
library(Rsubread)

## load helper functions
for (f in list.files("../R", "\\.(r|R)$", full.names = TRUE)) {
    source(f)
}

## project data/results folders
datdir <- "data"
resdir <- "results"
dir.create(datdir, showWarnings = FALSE)
dir.create(resdir, showWarnings = FALSE)

## set up parallel backend
cores <- as.numeric(Sys.getenv("SLURM_NTASKS"))
multicoreParam <- MulticoreParam(workers = cores)
```

# Data Preparation

We downloaded the bam files directly from UCSC ENCODE portal.

```{r, encode-download}
broad_url <- "http://hgdownload.soe.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeBroadHistone/"
broad_bams <- c("wgEncodeBroadHistoneGm12878H3k4me3StdAlnRep1.bam",
                "wgEncodeBroadHistoneGm12878H3k04me3StdAlnRep2V2.bam",
                "wgEncodeBroadHistoneK562H3k4me3StdAlnRep1.bam",
                "wgEncodeBroadHistoneK562H3k4me3StdAlnRep2.bam")

for (i_bam in broad_bams) {
    if (!file.exists(file.path(datdir, i_bam))) {
        download.file(paste0(broad_url, i_bam),
                      destfile = file.path(datdir, i_bam))
    }
    i_bai <- paste0(i_bam, ".bai")
    if (!file.exists(file.path(datdir, i_bai))) {
        download.file(paste0(broad_url, i_bai),
                      destfile = file.path(datdir, i_bai))
    }
}

uw_url <- "http://hgdownload.soe.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeUwHistone/"
uw_bams <- c("wgEncodeUwHistoneGm12878H3k4me3StdAlnRep1.bam",
             "wgEncodeUwHistoneGm12878H3k4me3StdAlnRep2.bam",
             "wgEncodeUwHistoneK562H3k4me3StdAlnRep1.bam",
             "wgEncodeUwHistoneK562H3k4me3StdAlnRep2.bam")

for (i_bam in uw_bams) {
    if (!file.exists(file.path(datdir, i_bam))) {
        download.file(paste0(uw_url, i_bam),
                      destfile = file.path(datdir, i_bam))
    }
    i_bai <- paste0(i_bam, ".bai")
    if (!file.exists(file.path(datdir, i_bai))) {
        download.file(paste0(uw_url, i_bai),
                      destfile = file.path(datdir, i_bai))
    }
}
```

We determine sample metadata from the file names.

```{r}
bamfiles <- list.files(datdir, pattern = 'bam$', full = TRUE)
labs <- gsub("wgEncode(.*)Histone.*", "\\1", basename(bamfiles))
cells <- gsub("wgEncode.*Histone(.*)H3k.*", "\\1", basename(bamfiles))

meta <- data.frame(cellline = cells, lab = labs, file = bamfiles)
```

Next, we identify promotor regions using the UCSC "Known Gene" annotations for
human genome assembly hg19 (GRCh37).

```{r}
prom <- promoters(genes(TxDb.Hsapiens.UCSC.hg19.knownGene))
seqlevels(prom, force = TRUE) <- seqlevels(prom)[!grepl("_", seqlevels(prom))]
```

# Data Analysis

## Differential Testing

We count sequencing reads for all promoter regions here, and test for differential peaks using these counts.

```{r count, results='hide', message=FALSE, warning=FALSE}
deseq_res  <- file.path(resdir, "h3k4me3-promotors-deseq.rda")
if (file.exists(deseq_res)) {
    load(deseq_res)
} else {
    ## count reads in each promotor region
    anno <- data.frame(GeneID = seq_len(length(prom)),
                       Chr = seqnames(prom),
                       Start = start(prom),
                       End = end(prom),
                       Strand = strand(prom))
    rc <- featureCounts(files = bamfiles,
                        annot.ext = anno,
                        allowMultiOverlap = TRUE,
                        minOverlap = 50,
                        readExtension3 = 150,
                        ignoreDup = TRUE)$counts

    ## perform differential testing
    dds <- DESeqDataSetFromMatrix(countData = rc,
                                  colData = as.data.frame(meta),
                                  design= ~ lab + cellline)
    dds <- DESeq(dds, fitType = "mean")
    dat <- as.data.frame(results(dds))
    colnames(dat) <- c('ind_covariate', 'effect_size', 'SE',
                       'stat', 'pval', 'padj')

    ## compute test-statistic from p-value for ASH
    dat <- dplyr::mutate(dat,
                         test_statistic = qnorm(exp(log(pval)-log(2)),
                                                lower.tail=FALSE),
                         test_statistic = test_statistic * sign(effect_size))

    save(deseq, dat, file = deseq_res)
}

## remove regions with effect sizes = NA
dat <- dat[!is.na(dat[, 2]), ]
```

## Covariate Diagnostics

Here, we can see that the covariant variable is actually informative.

### Mean Coverage

We consider the average coverage across conditions and replicates as a
potential covariate of interest. This is motivated by our prior assumption that
ChIP-seq data always involves two sets of regions: signal and background.
Sometimes, there are a few regions with signal coverage between signal and background.
In this example, we see that p-values are differently distributed under low, 
median and high coverage.

```{r, chipseq-diag-scatter, fig.width=4.5, fig.height=3.5}
rank_scatter(dat, pvalue = "pval", covariate = "ind_covariate")
```

```{r, chipseq-diag-hist, fig.width=10, fig.height=3.2}
strat_hist(dat, pvalue = "pval", covariate = "ind_covariate", maxy = 2)
```

## Multiple-Testing Correction

We use the common `BenchDesign` with the set of multiple testing correction
methods already included.

```{r, chipseq-benchdesign}
bd <- initializeBenchDesign()
```

Now, we're ready to construct the `SummarizedBenchmark` object, which will run
the functions specified in each method.

```{r, chipseq-sb, results="hide", message=FALSE}
bench_res  <- file.path(resdir, "h3k4me3-promotors-sb.rda")
if (file.exists(bench_res)) {
    load(bench_res)
} else {
    sb <- buildBench(bd, data = dat, ftCols = c("ind_covariate"),
                     parallel = TRUE, BPPARAM = multicoreParam)
    save(sb, file = bench_res)
}
```

## Benchmark Metrics

Next, we'll add the default performance metric for q-value assays. First, we have
to rename the assay to 'qvalue'.

```{r, chipseq-metrics}
assayNames(sb) <- "qvalue"
sb <- addDefaultMetrics(sb)
```

Now, we'll plot the results.

```{r, chipseq-plot, results="hide"}
rejections_scatter(sb, supplementary = FALSE)
rejection_scatter_bins(sb, covariate = "ind_covariate",
                       bins = 4, supplementary = FALSE)

plotFDRMethodsOverlap(sb,  alpha = 0.05, nsets = ncol(sb),
                      order.by = "freq", decreasing = TRUE,
                      supplementary = FALSE)

methods <- c("ashs", "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf")
plotCovariateBoxplots(sb, alpha = 0.1, nsets = 6,
                      methods = methods, trans = log2, maxNum = 5)
```

# Session Info

```{r}
sessionInfo()
```