### 2018-03-08 datasets/ChIP-seq/ChIPseq-promoters.Rmd

---
title: "Case Study: ChIP-seq Differential Peak Calling (pre-defined regions)"
author: "Mingxiang Teng, Patrick Kimes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

In this case study, we perform differential peak calling on ChIP-seq data of the histone
H3K4Me3 for samples from two cell lines (K562 and Gm12878) using publicly available data
generated by the ENCODE Project. The same data set is used for all ChIP-seq differential
testing case studies.

While several approaches have been proposed for differential peak calling, in this case
study we examine a simple approach of testing for differential peaks in the promoter 
regions of all genes. While using pre-defined regions can reduce the power of the test to
detect differences in non-promoter regions, since H3K4Me3 is an active marker of gene
expression, restricting tests to promoter regions is not unreasonable. Differential peaks
are tested using DESeq2.

# Workspace Setup

```{r, wkspace-setup, results='hide', message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(SummarizedBenchmark)
library(BiocParallel)
library(GenomicRanges)
library(TxDb.Hsapiens.UCSC.hg19.knownGene)
library(DESeq2)
library(Rsubread)

## load helper functions
for (f in list.files("../R", "\\.(r|R)$", full.names = TRUE)) {
    source(f)
}

## project data/results folders
datdir <- "data"
resdir <- "results"
dir.create(datdir, showWarnings = FALSE)
dir.create(resdir, showWarnings = FALSE)

## intermediary files we create below
count_file <- file.path(resdir, "h3k4me3-promotors-counts.rda")
result_file <- file.path(resdir, "h3k4me3-promotors-results.rda")
bench_file <- file.path(resdir, "h3k4me3-promotors-benchmark.rda")

## set up parallel backend
cores <- as.numeric(Sys.getenv("SLURM_NTASKS"))
multicoreParam <- MulticoreParam(workers = cores)
```

# Data Preparation

We downloaded the bam files directly from UCSC ENCODE portal.

```{r, encode-download}
broad_url <- "http://hgdownload.soe.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeBroadHistone/"
broad_bams <- c("wgEncodeBroadHistoneGm12878H3k4me3StdAlnRep1.bam",
                "wgEncodeBroadHistoneGm12878H3k04me3StdAlnRep2V2.bam",
                "wgEncodeBroadHistoneK562H3k4me3StdAlnRep1.bam",
                "wgEncodeBroadHistoneK562H3k4me3StdAlnRep2.bam")

for (i_bam in broad_bams) {
    if (!file.exists(file.path(datdir, i_bam))) {
        download.file(paste0(broad_url, i_bam),
                      destfile = file.path(datdir, i_bam))
    }
    i_bai <- paste0(i_bam, ".bai")
    if (!file.exists(file.path(datdir, i_bai))) {
        download.file(paste0(broad_url, i_bai),
                      destfile = file.path(datdir, i_bai))
    }
}

uw_url <- "http://hgdownload.soe.ucsc.edu/goldenPath/hg19/encodeDCC/wgEncodeUwHistone/"
uw_bams <- c("wgEncodeUwHistoneGm12878H3k4me3StdAlnRep1.bam",
             "wgEncodeUwHistoneGm12878H3k4me3StdAlnRep2.bam",
             "wgEncodeUwHistoneK562H3k4me3StdAlnRep1.bam",
             "wgEncodeUwHistoneK562H3k4me3StdAlnRep2.bam")

for (i_bam in uw_bams) {
    if (!file.exists(file.path(datdir, i_bam))) {
        download.file(paste0(uw_url, i_bam),
                      destfile = file.path(datdir, i_bam))
    }
    i_bai <- paste0(i_bam, ".bai")
    if (!file.exists(file.path(datdir, i_bai))) {
        download.file(paste0(uw_url, i_bai),
                      destfile = file.path(datdir, i_bai))
    }
}
```

We determine sample metadata from the file names.

```{r}
bamfiles <- list.files(datdir, pattern = 'bam$', full = TRUE)
labs <- gsub("wgEncode(.*)Histone.*", "\\1", basename(bamfiles))
cells <- gsub("wgEncode.*Histone(.*)H3k.*", "\\1", basename(bamfiles))

meta <- data.frame(cellline = cells, lab = labs, file = bamfiles)
```

Next, we identify promotor regions using the UCSC "Known Gene" annotations for
human genome assembly hg19 (GRCh37).

```{r}
prom <- promoters(genes(TxDb.Hsapiens.UCSC.hg19.knownGene))
seqlevels(prom, force = TRUE) <- seqlevels(prom)[!grepl("_", seqlevels(prom))]
```

## Read Counting

We count reads for all promoter regions.

```{r count, results='hide', message=FALSE, warning=FALSE}
if (file.exists(count_file)) {
    load(count_file)
} else {
    anno <- data.frame(GeneID = seq_len(length(prom)),
                       Chr = seqnames(prom),
                       Start = start(prom),
                       End = end(prom),
                       Strand = strand(prom))
    rc <- featureCounts(files = bamfiles,
                        annot.ext = anno,
                        allowMultiOverlap = TRUE,
                        minOverlap = 50,
                        readExtension3 = 150,
                        ignoreDup = TRUE)$counts
    save(rc, file = count_file)
}
```

# Data Analysis

## Differential Testing

We test for differential peaks at each of the .

```{r count, results='hide', message=FALSE, warning=FALSE}
if (file.exists(result_file)) {
    load(result_file)
} else {
    dds <- DESeqDataSetFromMatrix(countData = rc,
                                  colData = as.data.frame(meta),
                                  design = ~ lab + cellline)
    dds <- DESeq(dds, fitType = "mean")
    dat <- as.data.frame(results(dds))
    colnames(dat) <- c('ind_covariate', 'effect_size', 'SE',
                       'stat', 'pval', 'padj')
    ## compute test-statistic from p-value for ASH
    dat <- dplyr::mutate(dat,
                         test_statistic = qnorm(exp(log(pval)-log(2)),
                                                lower.tail=FALSE),
                         test_statistic = test_statistic * sign(effect_size))
    ## remove regions with effect sizes = NA
    dat <- dat[!is.na(dat$effect_size), ]
    save(dat, file = result_file)
}
```

## Covariate Diagnostics

Here, we can check diagnostic plots to assess whether candidate covariates are actually informative.

### Mean Coverage

We consider the average coverage across conditions and replicates as a
potential covariate of interest. This is motivated by our prior assumption that
ChIP-seq data always involves two sets of regions: signal and background.
Sometimes, there are a few regions with signal coverage between signal and background.
In this example, we see that p-values are differently distributed under low, 
median and high coverage.

```{r, chipseq-diag-scatter, fig.width=4.5, fig.height=3.5}
rank_scatter(dat, pvalue = "pval", covariate = "ind_covariate")
```

```{r, chipseq-diag-hist, fig.width=10, fig.height=3.2}
strat_hist(dat, pvalue = "pval", covariate = "ind_covariate", maxy = 2)
```

## Multiple-Testing Correction

We use the common `BenchDesign` with the set of multiple testing correction
methods already included.

```{r, chipseq-benchdesign}
bd <- initializeBenchDesign()
```

Now, we're ready to construct the `SummarizedBenchmark` object, which will run
the functions specified in each method.

```{r, chipseq-sb, results="hide", message=FALSE}
if (file.exists(bench_file)) {
    load(bench_file)
} else {
    sb <- buildBench(bd, data = dat, ftCols = c("ind_covariate"),
                     parallel = TRUE, BPPARAM = multicoreParam)
    save(sb, file = bench_file)
}
```

## Benchmark Metrics

Next, we'll add the default performance metric for q-value assays. First, we have
to rename the assay to 'qvalue'.

```{r, chipseq-metrics}
assayNames(sb) <- "qvalue"
sb <- addDefaultMetrics(sb)
```

Now, we'll plot the results.

```{r, chipseq-plot, results="hide"}
rejections_scatter(sb, supplementary = FALSE)
rejection_scatter_bins(sb, covariate = "ind_covariate",
                       bins = 4, supplementary = FALSE)

plotFDRMethodsOverlap(sb,  alpha = 0.05, nsets = ncol(sb),
                      order.by = "freq", decreasing = TRUE,
                      supplementary = FALSE)

methods <- c("ashs", "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf")
plotCovariateBoxplots(sb, alpha = 0.1, nsets = 6,
                      methods = methods, trans = log2, maxNum = 5)
```

# Session Info

```{r}
sessionInfo()
```