### 2018-03-22 datasets/GWAS/GWAS.Rmd
Fix a couple typos in RNAseq Rmds

---
title: "Case study: Genome-Wide Association Analyses (GWAS)"
author: "Keegan Korthauer"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
   html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

Here we will make use of the `SummarizedBenchmark` package to benchmark a 
Genome-Wide Association Study (GWAS) dataset. We will examine a GWAS 
meta-analysis for Body Mass Index (BMI), which was also included in the 
Boca-Leek manuscript.

# Workspace setup

This analysis requires that the software PLINK (version 1.9) 
is installed on your system.
PLINK is freely available for download 
[here](http://zzz.bwh.harvard.edu/plink/plink2.shtml).

```{r, wkspace-setup, results='hide', message=FALSE, warning=FALSE}
library(data.table)
library(readxl)
library(readr)
library(dplyr)
library(ggplot2)
library(cowplot)
library(BiocParallel)

## load helper functions
for (f in list.files("../R", "\\.(r|R)$", full.names = TRUE)) {
    source(f)
}

# data and results directories
datdir <- "data"
resdir <- "results"
dir.create(datdir, showWarnings = FALSE)
dir.create(resdir, showWarnings = FALSE)

# set up parallel backend
cores <- 8
multicoreParam <- MulticoreParam(workers = cores)
```

# Data preparation

We'll first download the GWAS summary statistics dataset from the 
GIANT consortium data portal.

## Data download

After downloading, we unzip the file, and delete the files we won't use (they provide results
subsetted by ancestry and sex), keeping only the file with European ancestry and
both sexes `BMI.SNPadjSMK.CombinedSexes.EuropeanOnly.txt`. Note that
the Boca-Leek paper also used the 'EuropeanOnly' subset, likely because using 
all ancestries would require that we assess the impact of population stratification
and adjust for it if present. For simplicity, we follow suit and use the homogeneous 
subset to avoid the impact of population stratification on our results.

```{r, GWAS1-download}
if (!file.exists(file.path(datdir, "BMI.SNPadjSMK.CombinedSexes.EuropeanOnly.txt"))) {
  download.file(url = "http://portals.broadinstitute.org/collaboration/giant/images/3/3a/BMI.SNPadjSMK.zip", 
             destfile = file.path(datdir, "BMI.SNPadjSMK.zip")) 
  unzip(file.path(datdir, "BMI.SNPadjSMK.zip"), exdir = datdir)
  file.remove(file.path(datdir,"BMI.SNPadjSMK.zip"))

  dfiles <- list.files(path = datdir, pattern = "BMI.SNPadjSMK.*.txt", 
                       full.names = TRUE)
  dfiles <- dfiles[!grepl("BMI.SNPadjSMK.CombinedSexes.EuropeanOnly.txt", dfiles)]
  file.remove(dfiles)
}
```

We also download the 1000 genomes CEU reference population data, which will be used to 
determine which SNPs are LD buddies for the purposes of obtaining an independent set of SNPs.

```{r, GWAS1-refdata}
reffile <- file.path(datdir, "1000G_20101123_v3_GIANT_chr1_23_minimacnamesifnotRS_CEU_MAF0.01")
if (!file.exists(paste0(reffile, ".fam"))) {
    download.file("http://neurogenetics.qimrberghofer.edu.au/iSECA/1000G_20101123_v3_GIANT_chr1_23_minimacnamesifnotRS_CEU_MAF0.01.zip", 
                  destfile = paste0(reffile, ".zip"))
    unzip(paste0(reffile, ".zip"), exdir = datdir)
    file.remove(paste0(reffile, ".zip"))
}
```

Next, we'll read in the unzipped `.txt` file into R and verify that it contains
the necessary inputs to run the FDR benchmark comparisons.

```{r, GWAS1-verify-contents}
bmi <- fread(file.path(datdir, "BMI.SNPadjSMK.CombinedSexes.EuropeanOnly.txt"),
             header = TRUE)
dim(bmi)
head(bmi)
```

It looks like we have:

- `p_value`:p-value
- `effect`:effect size
- `stderr`:standard error
- additional covariates: 
  - `N`: Number of samples with this SNP - BMI association measured
- `Freq_Allele1_HapMapCEU`: Allele Frequency of Allele1, as measured in HapMapCEU

for `r nrow(bmi)` SNPs.

## Prune SNPs that are in Linkage Disequilibrium

Since nearby SNPs in linkage disequilibrium may not constitute separate discoveries, we carry out 
a pruning step. This commonly done by examining correlation of genotypes among nearby markers, but
since we don't have the genotype data, we'll use the linkage disequilibrium results of common SNPs and 
previously observed LD blocks. Specifically, we'll use the phase 3 data from 1000 genomes, restricted
to the CEU population (since this most closely matches our population of European ancestry). This 
reference data was downloaded in a previous code chunk. Here we read in the list of more than
9 million SNPs with data in 1000 genomes and create a list of SNPs that overlap our set. We'll write 
this list to a file that matches PLINK format.

```{r, GWAS1-snplist, eval=TRUE}
# load reference data 
onekg <- fread(file.path(datdir,  "1000G_20101123_v3_GIANT_chr1_23_minimacnamesifnotRS_CEU_MAF0.01.bim"))$V2

# construct input list of SNPs that overlap - 97.5% overlap
bmi$BP <- as.numeric(unlist(lapply(strsplit(bmi$markername, ":"), function(x) x[[2]])))
write_delim(bmi %>%
            dplyr::rename(CHR=chromosome, SNP=rs_id, A1=allele_1, A2=allele_2) %>%
            mutate(P=0.5, F_A=Freq_Allele1_HapMapCEU, F_U=F_A, CHISQ=1, OR=1) %>% 
            select(CHR, SNP, BP, A1, F_A, F_U, A2, CHISQ, P, OR) %>%
            filter(SNP %in% onekg), 
            path=file.path(datdir,  "rslist.assoc"), delim="\t")

ol <- sum(bmi$rs_id %in% onekg)/nrow(bmi)
rm(onekg)
```

It turns out that `r signif(ol*100, 3)` percent of SNPs in our set are included in the 1000 genomes 
reference data. Next we'll use the PLINK software to read in our list of SNPs and the 1000 genomes
linkage disequilibrium data to 'clump' the SNPs into independent sets of SNPs that have LD correlation (r squared)
less than 0.2 with all neighbors within 250 kilobases of distance. We write the results to a file called `plink_clump.clumped`. We have to jump out of R here since PLINK is a command line tool.

```{r, clump}
## use PLINK clumping tool to obtain a set of independent SNPs by LD
if (!dir.exists(plink_clump)) {
    system(paste("plink",
                 "--bfile DATA/1000G_20101123_v3_GIANT_chr1_23_minimacnamesifnotRS_CEU_MAF0.01",
                 "--clump DATA/rslist.assoc",
                 "--clump-field P",
                 "--clump-p1 0.9999",
                 "--clump-p2 0.9999",
                 "--clump-r2 0.2",
                 "--clump-kb 250",
                 "--out DATA/plink_clump"))
}
```

Next, we'll read in the clumped results and subset our data by our list of independent SNPs.

```{r, GWAS1-prune}
clump <- fread(file.path(datdir, "plink_clump.clumped"))

# subset to include only the SNPs in the ld.list
bmi <- bmi[bmi$rs_id %in% clump$SNP, ]

rm(clump)
```

After pruning, we are left with for `r nrow(bmi)` approximately independent SNPs.

Before moving on, we'll rename the relevant columns so that they will use common terms
across the different datasets. We'll also add a 'test_statistic' column.

```{r, GWAS1-reformat}
bmi <- bmi %>% 
    dplyr::rename(pval = p_value,
                  SE = stderr,
                  effect_size = effect,
                  ind_covar_N = N,
                  ind_covar_AF = Freq_Allele1_HapMapCEU) %>%
    mutate(test_statistic = qnorm(exp(log(pval)-log(2)), lower.tail=FALSE) * sign(effect_size)) %>%
    select(pval, SE, effect_size, ind_covar_N, ind_covar_AF, test_statistic) 
```

# Data Analysis

## Differential Testing
Since we do not have access to the individual genotype information and since
our data already contains the results of performing hypothesis tests of
association of BMI with each SNP (including p-values and effect sizes), we 
don't need to carry out any additional testing. We move on to checking the
covariate diagnostics and adjusting for multiple comparisons.

## Covariate Diagnostics

Here we look to see if the covariates do indeed look informative. First we look at 
the sample size covariate.

### Covariate one: Sample Size

```{r, fig.width=4.5, fig.height=3.5, message=FALSE}
rank_scatter(bmi, pvalue="pval", covariate="ind_covar_N") + 
  ggtitle("Covariate 1: Sample Size")
```

```{r, fig.width=10, fig.height=3.2, message=FALSE}
strat_hist(bmi, pvalue="pval", covariate="ind_covar_N", maxy=2) 
```

### Covariate two: Minor Allele Frequency

Next we look at the allele frequency covariate.

```{r, fig.width=4.5, fig.height=3.5, message=FALSE}
rank_scatter(bmi, pvalue="pval", covariate="ind_covar_AF") +
  ggtitle("Covariate 1: Minor allele frequency")
```

```{r, fig.width=10, fig.height=3.2, message=FALSE}
strat_hist(bmi, pvalue="pval", covariate="ind_covar_AF", maxy=2)
```

## Multiple-Testing Correction

First, we'll create an object of `BenchDesign` class to hold the data and 
add the benchmark methods to the `BenchDesign` object.

```{r, GWAS1-benchdesign, message=FALSE}
bd <- initializeBenchDesign()
```

Now, we're ready to construct the `SummarizedBenchmark` object, which will run
the functions specified in each method (these are actually sourced in from the
                                        helper scripts).

### Covariate one: Sample size

First we'll include the sample size covariate.

```{r, GWAS1-sb, results="hide", message=FALSE}
resN <- paste0(resdir, "bmi_N_summarizedBenchmark_",
               nrow(bmi), ".RData")

duration <- NA
if (!file.exists(resN)) {
  t1 <- proc.time()
  sbN <- bd %>% buildBench(data=(bmi %>% mutate(ind_covariate=ind_covar_N)), 
                          ftCols = c("ind_covar_N"),
                          parallel=TRUE, BPPARAM=multicoreParam)
  metadata(sbN)$data_download_link <- 
    "http://portals.broadinstitute.org/collaboration/giant/images/3/3a/BMI.SNPadjSMK.zip"
  saveRDS(sbN, file=resN)
  duration <- round((proc.time()-t1)[3]/60,1)
}else{
  sbN <- readRDS(resN)
}
```

```{r, echo=FALSE}
if (!is.na(duration)) {
  message("This step took ", duration, " minutes for ", nrow(bmi), " SNPs using ",
          cores, " cores.")
}
```

### Covariate two: Minor Allele Frequency

Now, we'll repeat the multiple testing correction using the other covariate (AF):

```{r, GWAS1-sb2, results="hide", message=FALSE}
resAF <- paste0(resdir, "bmi_AF_summarizedBenchmark_",
nrow(bmi), ".RData")

duration <- NA
if (!file.exists(resAF)){
  t1 <- proc.time()
  sbAF <- bd %>% buildBench(data=(bmi %>% mutate(ind_covariate=ind_covar_AF) %>%
                                  filter(!is.na(ind_covariate))), 
                            ftCols = c("ind_covar_AF"),
                            parallel=TRUE, BPPARAM=multicoreParam)
  metadata(sbAF)$data_download_link <-    
    "http://portals.broadinstitute.org/collaboration/giant/images/3/3a/BMI.SNPadjSMK.zip"
  saveRDS(sbAF, file=resAF)
  duration <- round((proc.time()-t1)[3]/60,1)
}else{
  sbAF <- readRDS(resAF)
}
```

```{r, echo=FALSE}
if(!is.na(duration)){
  message("This step took ", duration, " minutes for ", nrow(bmi), " SNPs using ",
  cores, " cores.")
}
```

## Benchmark Metrics

Next, we'll add the default performance metric for q-value assays and 
plot the results. We'll start with covariate one.

### Covariate one: Sample Size

First, we have
to rename the assay to 'qvalue'.

```{r, GWAS1-metrics}
# rename assay to qvalue
assayNames(sbN) <- "qvalue"
sbN <- addDefaultMetrics(sbN)
```

Now, we'll plot the results.

```{r, GWAS1-plot, results="hide"}
# plot nrejects by method overall and stratified by covariate
rejections_scatter(sbN, supplementary=FALSE) +
    ggtitle("Covariate 1: Sample size")

rejection_scatter_bins(sbN, covariate="ind_covar_N", bins=4,
                       supplementary=FALSE) +
    ggtitle("Covariate 1: Sample size")

# upset plot 
plotFDRMethodsOverlap(sbN, 
                      alpha=0.05, nsets=ncol(sb),
                      order.by="freq", decreasing=TRUE,
                      supplementary=FALSE)

# properties of method-exclusive discoveries
methods <- c("ashs", "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf",
             "scott-empirical", "scott-theoretical")
mcols(sbN)$ind_covariate <- mcols(sbN)$ind_covar_N
plotCovariateBoxplots(sbN, alpha=0.1, nsets=6, methods=methods, trans=log2, maxNum=3) +
    ggtitle("Covariate 1: Sample size")
```

```{r, fig.width=8, fig.height=3.5}
mcols(sbN)$tstat <- bmi %>% filter(!is.na(ind_covar_N)) %>% pull(effect_size)
covariateHeatmap(sbN, alpha=0.05, covname="tstat", trans = "log1p")
covariateHeatmap(sbN, alpha=0.05, covname="ind_covar_N", trans = "log1p")

covariateHeatmap(sbN, alpha=0.05, covname="tstat", nbins=25, 
                 linePlot = TRUE, trans="log1p")
covariateHeatmap(sbN, alpha=0.05, covname="ind_covar_N", nbins=25, 
                 linePlot = TRUE, trans="log1p")
```

### Covariate two: Minor Allele Frequency

Next, we'll look at the performance metrics for the other covariate (minor
allele frequency).

```{r, GWAS1-metrics2}
# rename assay to qvalue
assayNames(sbAF) <- "qvalue"
sbAF <- addDefaultMetrics(sbAF)
```

Now, we'll plot the results.

```{r, GWAS1-plot2, results="hide"}
# plot nrejects by method overall and stratified by covariate
rejections_scatter(sbAF, supplementary=FALSE) +
  ggtitle("Covariate 2: Minor allele frequency")

rejection_scatter_bins(sbAF, covariate="ind_covar_AF", bins=4,
                       supplementary=FALSE) +
  ggtitle("Covariate 2: Minor allele frequency")

# upset plot 
plotFDRMethodsOverlap(sbAF, 
                      alpha=0.05, nsets=ncol(sbAF),
                      order.by="freq", decreasing=TRUE,
                      supplementary=FALSE)

# properties of method-exclusive discoveries
methods <- c("ashs", "lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf",
             "scott-empirical", "scott-theoretical")
mcols(sbAF)$ind_covariate <- abs(mcols(sbAF)$ind_covar_AF-0.5)
plotCovariateBoxplots(sbAF, alpha=0.1, nsets=6, methods=methods, maxNum=3)+
  ggtitle("Covariate 2: Minor allele frequency")
```

```{r, fig.width=8, fig.height=3.5}
mcols(sbAF)$tstat <- bmi %>% filter(!is.na(ind_covar_AF)) %>% pull(effect_size)
covariateHeatmap(sbAF, alpha=0.05, covname="tstat", trans = "log1p")
covariateHeatmap(sbAF, alpha=0.05, covname="ind_covar_AF")

covariateHeatmap(sbAF, alpha=0.05, covname="tstat", nbins=25, 
                 linePlot = TRUE, trans = "log1p")
covariateHeatmap(sbAF, alpha=0.05, covname="ind_covar_AF", nbins=25, 
                 linePlot = TRUE)
```

# Session Information

```{r}
sessionInfo()
```