### 2018-03-23 datasets/simulations/parameter-simulation.Rmd

---
title: "Simulation Study"
author: "Patrick Kimes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
    html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
        code_folding: hide
---

```{r, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, fig.width = 14, fig.height = 6)
```

# Summary

For each simulation, a collection (20,000) of test statistics were directly
sampled from univariate distributions (Normal, non-central-t, or
non-central-chi-squared). Test statistics corresponding to null tests are
sampled with non-centrality parameter zero, and non-null tests are sampled
with some non-zero non-centrality parameter, determined based on the
simulation setting. In these simulations, the magnitude of the non-centrality
parameter for each test can be understood as the effect size or signal in
that test. The univariate distributions (Normal, t, chi-squared) are the variyng
sampling distributions that may be encountered under different testing scenarios
(e.g. Student's t-test or Pearson's goodness-of-fit).

Since ASH requires standard errors, the true standard errors is provided to the method.

All simulation settings were carried out with an independent covariate. Both
settings where the covariate was informative and uninformative were considered.
Additionally, several different informative covariates were considered. Following
the approach of Boca and Leek (2015) and Chen, Robinson and Storey (2017), we
sampled the independent covariate for each test uniformly from the interval [0, 1].
Whether the covariate was informative or uninformative was determined by a functional
relationship between the value of the covariate and the probability of the
corresponding test being null. The following relationships were considered:

- pi0(x) = pi0 (`uniform-80`)
- pi0(x) = pi0 (`uniform-90`)
- pi0(x) = pi0 (`uniform-95`)
- pi0(x) = 0.25*I(x < 0.25) + ... (`step-more`, 80% pi0)
- pi0(x) = 0.25*I(x < 0.25) + ... (`step-90`, 90% pi0)
- pi0(x) = 0.25*I(x < 0.25) + ... (`step-95`, 95% pi0)
- pi0(x) = 1 - x^3 ... (`cubic`, 80% pi0)
- pi0(x) = 1 - x^3 ... (`cubic-90`, 90% pi0)
- pi0(x) = 1 - x^3 ... (`cubic-95`, 95% pi0)

Here, we only consider a subset of the entire set of simulation settings.
Namely, we look at: [1] null simulations, [2] non-null simulations with an
informative covariate (cubic), and [3] non-null simulations with an uninformative
covariate. For each of the 3 simulation sets, we further consider three different
sampling distributions of the test statistics (Normal, t, chi-squared).

Performance is evaluated with: the false positive rate (FPR), true positive rate (TPR),
and method overlaps (upset plots). For null simulations, since FPR and TPR are useless,
we instead plot the proportion of rejected tests.

# Workspace Setup

```{r, wkspace-setup, results='hide', message=FALSE, warning=FALSE}
library("SummarizedBenchmark")
library("tidyverse")

source(file.path("..", "R", "plotsim_aggupset.R"))
```

# Helper Functions

```{r helper-preclean}
preclean <- function(res, merge_ihw = TRUE, supplementary = FALSE) {
    ## set of method names
    method_names <- unique(res$blabel)
    ## remove unadjusted p-values - too extreme
    if ("unadjusted" %in% method_names) {
        res <- res %>%
            filter(blabel != "unadjusted")
    }
    ## drop ihw not matching 'alpha'
    if (merge_ihw) {
        res <- res %>%
            mutate(ihw_alpha = gsub("ihw-a(.*)", "\\1", blabel),
                   ihw_alpha = ifelse(grepl("^ihw-a", blabel), ihw_alpha, alpha*100),
                   ihw_alpha = as.numeric(ihw_alpha) / 100) %>%
            filter(abs(alpha - ihw_alpha) < 1e-6) %>%
            mutate(blabel = ifelse(grepl("^ihw-a", blabel), "ihw", blabel))
    }
    ## drop excess Boca-Leek results
    if(!supplementary) {
        methods_bl <- grep("^bl-", method_names, value = TRUE)
        res <- res %>%
            filter(!(blabel %in% setdiff(methods_bl, "bl-df03")))
    }
    return(res)
}
```

```{r helper-myplot}
myplot <- function(res, metrics = c("TPR", "FPR"), alpha_max = 1) {
    gp <- res %>%
        filter(performanceMetric %in% metrics) %>%
        filter (alpha <= alpha_max) %>%
        group_by(alpha, blabel, performanceMetric) %>%
        summarize(value50 = median(value, na.rm = TRUE),
                  value10 = quantile(value, 0.10, na.rm = TRUE),
                  value90 = quantile(value, 0.90, na.rm = TRUE)) %>%
        ungroup %>%
        ggplot(aes(x = alpha, y = value50, ymin = value10, ymax = value90,
                   group = blabel, color = blabel))
    if ("FPR" %in% metrics) {
        gp <- gp +
            geom_abline(aes(intercept = y, slope = s), alpha = 1/2,
                        data = data.frame(y = 0, s = 1, performanceMetric = "FPR"))
    }
    gp <- gp +
        geom_line(alpha = 1/4,
                  position = position_dodge(width = 0.003)) +
        geom_errorbar(width = .01,
                      position = position_dodge(width = 0.003)) + 
        geom_point(alpha = 1/2, position = position_dodge(width = 0.003)) +
        facet_wrap(~ performanceMetric) +
        scale_color_discrete("Method") + 
        ylab("performance across replicates") +
        xlab("alpha cutoff") +
        theme_bw()
    gp 
}
```

# Null Simulations

First, we look at the null simulation settings to assess which methods control
false discoveries when no truly significant tests are present. The TPR is not
reported here because there are no "true positivies" in the null setting.

- *null* (pi0 = 100) simulations across *noise distributions*
  1. *null* (pi0 = 100) simulations across *Gaussian*
  2. *null* (pi0 = 100) simulations across *t*
  3. *null* (pi0 = 100) simulations across *chi-squared*

## Normal

```{r, fig.width = 7, fig.height = 6}
## common suffix 
rds_name <- "M100/allnull-uniform/results-altnoise_normal_null-M100.rds"

## read in performance metrics, plot metric
mets <- readRDS(file.path("data-psim-parsed", rds_name))
mets <- preclean(mets)
myplot(mets, metrics = "rejectprop") + 
    ggtitle("Distribution: normal, pi0: 100% (100 replications)")
```

## t (10 dof)

```{r, fig.width = 7, fig.height = 6}
## common suffix 
rds_name <- "M100/allnull-uniform/results-altnoise_t_df10_null-M100.rds"

## read in performance metrics, plot metric
mets <- readRDS(file.path("data-psim-parsed", rds_name))
mets <- preclean(mets)
myplot(mets, metrics = "rejectprop") +
    ggtitle("Distribution: t (10 dof), pi0: 100% (100 replications)")
```

## chi-sq (4 dof)

```{r, fig.width = 7, fig.height = 6}
## common suffix 
rds_name <- "M100/allnull-uniform/results-altnoise_chisq_df4_null-M100.rds"

## read in performance metrics, plot metric
mets <- readRDS(file.path("data-psim-parsed", rds_name))
mets <- preclean(mets)
myplot(mets, metrics = "rejectprop") +
    ggtitle("Distribution: chi-squared (4 dof), pi0: 100% (100 replications)")
```

# Informative Covariate Simulations

Next, we consider simulations where a truly informative covariate is present.

- *cubic-informative*, *non-null* (pi0 = 80,90,95) simulations across *noise distributions*
  1. *cubic-informative*, *non-null* (pi0 = 80,90,95) simulations across *Gaussian*
  2. *cubic-informative*, *non-null* (pi0 = 80,90,95) simulations across *t*
  3. *cubic-informative*, *non-null* (pi0 = 80,90,95) simulations across *chi-squared*

## Normal

```{r}
## common suffix 
rds_name <- "M100/esize_fixed-bl-cubic-90-0.8/results-alteff_30-M100.rds"

## read in performance metrics, plot metric
mets <- readRDS(file.path("data-psim-parsed", rds_name))
mets <- preclean(mets)
myplot(mets) + 
    ggtitle("Distribution: normal, pi0: 90% (100 replications)")

## read in SB objects, plot aggregate upset plot
res <- readRDS(file.path("data-psim", rds_name))
aggupset(res, alpha = 0.05)
```

Also looking at varying pi0s (only showing FPR/TPR).

```{r}
"M100/esize_fixed-bl-cubic/results-alteff_30-M100.rds" %>%
    file.path("data-psim-parsed", .) %>%
    readRDS() %>%
    preclean() %>%
    myplot() + 
    ggtitle("Distribution: normal, pi0: 80% (100 replications)")

"M100/esize_fixed-bl-cubic-95-0.8/results-alteff_30-M100.rds" %>%
    file.path("data-psim-parsed", .) %>%
    readRDS() %>%
    preclean() %>%
    myplot() + 
    ggtitle("Distribution: normal, pi0: 95% (100 replications)")
```

## t (10 dof)

```{r}
rds_name <- "M100/altnoise-bl-cubic-90-0.8/results-altnoise_t_df10_bimodal-M100.rds"

## read in performance metrics, plot metric
mets <- readRDS(file.path("data-psim-parsed", rds_name))
mets <- preclean(mets)
myplot(mets) +
    ggtitle("Distribution: t (10 dof), pi0: 90% (100 replications)")

## read in SB objects, plot aggregate upset plot
res <- readRDS(file.path("data-psim", rds_name))
aggupset(res, alpha = 0.05)
```

Also looking at varying pi0s (only showing FPR/TPR).

```{r}
"M100/altnoise-bl-cubic/results-altnoise_t_df10_bimodal-M100.rds" %>%
    file.path("data-psim-parsed", .) %>%
    readRDS() %>%
    preclean() %>%
    myplot() + 
    ggtitle("Distribution: t (10 dof), pi0: 80% (100 replications)")

"M100/altnoise-bl-cubic-95-0.8/results-altnoise_t_df10_bimodal-M100.rds" %>%
    file.path("data-psim-parsed", .) %>%
    readRDS() %>%
    preclean() %>%
    myplot() + 
    ggtitle("Distribution: t (10 dof), pi0: 95% (100 replications)")
```

## chi-sq (4 dof)

```{r}
## common suffix 
rds_name <- "M100/altnoise-bl-cubic-90-0.8/results-altnoise_chisq_df4_shift3sq-M100.rds"

## read in performance metrics, plot metric
mets <- readRDS(file.path("data-psim-parsed", rds_name))
mets <- preclean(mets)
myplot(mets) +
    ggtitle("Distribution: chi-squared (4 dof), pi0: 90% (100 replications)")

## read in SB objects, plot aggregate upset plot
res <- readRDS(file.path("data-psim", rds_name))
aggupset(res, alpha = 0.05)
```

Also looking at varying pi0s (only showing FPR/TPR).

```{r}
"M100/altnoise-bl-cubic/results-altnoise_chisq_df4_shift3sq-M100.rds" %>%
    file.path("data-psim-parsed", .) %>%
    readRDS() %>%
    preclean() %>%
    myplot() + 
    ggtitle("Distribution: chi-squared (4 dof), pi0: 80% (100 replications)")

"M100/altnoise-bl-cubic-95-0.8/results-altnoise_chisq_df4_shift3sq-M100.rds" %>%
    file.path("data-psim-parsed", .) %>%
    readRDS() %>%
    preclean() %>%
    myplot() + 
    ggtitle("Distribution: chi-squared (4 dof), pi0: 95% (100 replications)")
```

# Uninformative Covariate Simulations

Finally, we consider simulations with non-null (alternative) tests are present.

- *non-informative*, *non-null* (pi0 = 80,90,95) simulations across *noise distributions*
  1. *uninformative*, *non-null* (pi0 = 80,90,95) simulations across *Gaussian*
  2. *uninformative*, *non-null* (pi0 = 80,90,95) simulations across *t*
  3. *uninformative*, *non-null* (pi0 = 80,90,95) simulations across *chi-squared*

## Normal

```{r}
## common suffix 
rds_name <- "M100/esize_fixed-uniform-0.9/results-alteff_30-M100.rds"

## read in performance metrics, plot metric
mets <- readRDS(file.path("data-psim-parsed", rds_name))
mets <- preclean(mets)
myplot(mets) + 
    ggtitle("Distribution: normal, pi0: 90% (100 replications)")

## read in SB objects, plot aggregate upset plot
res <- readRDS(file.path("data-psim", rds_name))
aggupset(res, alpha = 0.05)
```

Also looking at varying pi0s (only showing FPR/TPR).

```{r}
"M100/esize_fixed-uniform/results-alteff_30-M100.rds" %>%
    file.path("data-psim-parsed", .) %>%
    readRDS() %>%
    preclean() %>%
    myplot() + 
    ggtitle("Distribution: normal, pi0: 80% (100 replications)")

"M100/esize_fixed-uniform-0.95/results-alteff_30-M100.rds" %>%
    file.path("data-psim-parsed", .) %>%
    readRDS() %>%
    preclean() %>%
    myplot() + 
    ggtitle("Distribution: normal, pi0: 95% (100 replications)")

"M100/esize_fixed-uniform-0.99/results-alteff_30-M100.rds" %>%
    file.path("data-psim-parsed", .) %>%
    readRDS() %>%
    preclean() %>%
    myplot() + 
    ggtitle("Distribution: normal, pi0: 99% (100 replications)")
```

## t (10 dof)

```{r}
rds_name <- "M100/altnoise-uniform/results-altnoise_t_df10_bimodal-M100.rds"

## read in performance metrics, plot metric
mets <- readRDS(file.path("data-psim-parsed", rds_name))
mets <- preclean(mets)
myplot(mets) +
    ggtitle("Distribution: t (10 dof), pi0: 80% (100 replications)")

## read in SB objects, plot aggregate upset plot
res <- readRDS(file.path("data-psim", rds_name))
aggupset(res, alpha = 0.05)
```

*Need to run with different pi0 values.*

## chi-sq (4 dof)

```{r}
rds_name <- "M100/altnoise-uniform/results-altnoise_chisq_df4_shift3sq-M100.rds"

## read in performance metrics, plot metric
mets <- readRDS(file.path("data-psim-parsed", rds_name))
mets <- preclean(mets)
myplot(mets) +
    ggtitle("Distribution: chi-squared (4 dof), pi0: 80% (100 replications)")

## read in SB objects, plot aggregate upset plot
res <- readRDS(file.path("data-psim", rds_name))
aggupset(res, alpha = 0.05)
```

*Need to run with different pi0 values.*