### 2018-03-26 datasets/GSEA/gsea-human.Rmd

---
title: "Case Study: Gene Set Enrichment Analysis (Human Data Set)"
author: "Alejandro Reyes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

The objective of this document is to test different multiple testing methods in the context of Gene Set Enrichment Analysis (GSEA). In order to get a list of genes, we will use RNA-seq data consisting on cortex and cerebellum samples of a subset of the GTEx data.

# Workspace Setup

```{r}
library(dplyr)
library(ggplot2)
library(scales)
library(DESeq2)
library(EnsDb.Hsapiens.v86)
library(biomaRt)
library(goseq)

## load helper functions
for (f in list.files("../R", "\\.(r|R)$", full.names = TRUE)) {
    source(f)
}

## project data/results folders
datdir <- "data"
resdir <- "results"
dir.create(datdir, showWarnings = FALSE)
dir.create(resdir, showWarnings = FALSE)

## intermediary files we create below
count_file <- file.path(datdir, "human-counts.rds")
deseq_file <- file.path(datdir, "human-deseq.rds")
goset_file <- file.path(datdir, "human-gosets.rds")
result_file <- file.path(resdir, "human-results.rds")
bench_file <- file.path(resdir, "human-benchmark.rds")
```

# Data Preparation

The dataset consists of 10 samples (6 cerebellum and 4 cortex samples), all from male individuals. We download a *DESeqDataSet* if not present locally and modify the design to test for differences in gene expression between cerebellums and cortex.

```{r, message=FALSE}
if (!file.exists(count_file)) {
    download.file("https://www.dropbox.com/s/itrs2a6cnmurnck/gsea-human-counts.rds?dl=1",
                  destfile = count_file)
}
dsdObject <- readRDS(count_file)
design( dsdObject ) <- ~ tissue
```

To keep running times short, we only perform differential tesing on protein coding genes, as specified in Ensembl release 86.

```{r}
gnType <- genes(EnsDb.Hsapiens.v86, column = "gene_biotype")
protGenes <- gnType$gene_id[gnType$gene_biotype == "protein_coding"]
dsdObject <- dsdObject[rownames(dsdObject) %in% protGenes, ]
```

# Data Analysis

## Enrichment Analysis

Next, we run DESeq2 to retrieve a list of differentially expressed genes at a FDR of 10%.

```{r deseq2, message=FALSE}
if (!file.exists(deseq_file)) {
    dsdObject <- DESeq(dsdObject)
    res <- results(dsdObject, independentFiltering = FALSE)
    saveRDS(res, file = deseq_file)
} else {
    res <- readRDS(deseq_file)
}

genes <- as.numeric(res$padj < 0.1)
names(genes) <- rownames(res)
```

Next we’ll use the biomaRt package to download and associate GO annotations for each gene.

```{r goseq, message=FALSE}
if (!file.exists(goset_file)) {
    mart <- useMart("ensembl", "hsapiens_gene_ensembl")
    goSets <- getBM(c("ensembl_gene_id", "go_id"), mart = mart,
                    filters = "ensembl_gene_id", values = rownames(res))
    goSets <- goSets[!nchar( goSets$go_id ) == 0, ]
    goSets <- with(goSets, split(go_id, ensembl_gene_id))
    saveRDS(goSets, file = goset_file)
} else {
    goSets <- readRDS(goset_file)
}
```

Finally, we’ll perform a gene set enrichment analysis with goseq. goseq performs GSEA and adjusts for gene length biases. The bias is represented in the plot below, with gene length bin on the x-axis and proportion of genes detected to be differentially expressed on the y-axis.

```{r}
if (!file.exists(result_file)) {
    ## get median transcript length for bias adjustment
    txByGene <- transcriptsBy(EnsDb.Hsapiens.v86, "gene")
    geneLength <- sapply(width(txByGene), median)
    geneLength <- geneLength[names(genes)]
    genes[is.na(genes)] <- 0

    ## perform gsea
    pwf <- nullp(genes, bias.data = geneLength)
    goRes <- goseq(pwf, gene2cat = goSets)

    saveRDS(goRes, file = result_file)
} else {
    goRes <- readRDS(result_file)
}
```

## Covariate Diagnostics

Here, we can see that the covariant variable is actually informative.

### Gene Set Size

We will explore whether the size of the gene set can be used as a covariate for modern multiple-testing correction methods in the context of GSEA. In the plot below, the log10 of the p-values is plotted as a function of the size of the gene set. There is a pattern in which gene sets with a higher number of genes tend to have smaller p-values, which is indicative that gene set size is an informative covariate.

```{r informativeCovariate, message=FALSE, fig.height=3.5, message=FALSE, fig.width=4.5}
rank_scatter(dat = goRes, pval = "over_represented_pvalue", 
              covariate = "numInCat", bins = 50, funx = log2, 
              funfill = log10_trans()) +
    ylim(0, 12) +
    ggtitle("Over-represented gene sets") +
    xlab(expression(log[10]~"(# of genes)")) +
    ylab(expression(-log[10]~"(p-value)") ) 
```

We can also explore if the covariate seems to be independent under the null.

```{r assumptions, message=FALSE, fig.width=10, fig.height=3.2}
## filter numInCat below only used for plotting
## filter numDEInCat for cases where there are no DE genes in the category (these are actually not tested)
#dplyr::filter( goRes, over_represented_pvalue == 1, numDEInCat > 0)
dplyr::filter( goRes, numDEInCat > 0, numInCat > 10 ) %>%
    strat_hist( pval="over_represented_pvalue", covariate="numInCat", maxy=10)
```

## Multiple-Testing Correction

We then execute the benchDesign and generate a *SummarizedBenchmark* object containing multiple-testing corrections using several methods.

```{r runningBenckmark, message=FALSE}
## rename columns and prepare for benchmarking
res <- dplyr:::select(goRes, c("over_represented_pvalue", "numInCat")) %>%
    dplyr:::rename(pval = over_represented_pvalue, ind_covariate = numInCat)
res$pval <- pmin(res$pval, 1)

## generate default BenchDesign
bd <- initializeBenchDesign()

## run benchmark
if (!file.exists(bench_file)) {
    sGSEA <- buildBench(bd, res, ftCols = "ind_covariate")
    saveRDS(sGSEA, file = bench_file)
} else {
    sGSEA <- readRDS(bench_file)
}
```

## Benchmark Metrics

```{r}
assayNames(sGSEA) <- "qvalue"
sGSEA <- addDefaultMetrics(sGSEA)
sGSEA <- estimatePerformanceMetrics(sGSEA, addColData=TRUE)
```

```{r plottingResults}
rejections_scatter(sGSEA, as_fraction=FALSE, supplementary=FALSE)
```

```{r}
plotFDRMethodsOverlap(sGSEA, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100)
```

Gene set size of overlaps.

```{r}
methods <- c("lfdr", "ihw-a10", "bl-df03", "qvalue", "bh", "bonf")
plotCovariateBoxplots(sGSEA, alpha=0.1, nsets=6, methods=methods, trans=log2)
plotCovariateBoxplots(sGSEA, alpha=0.1, nsets=7, methods=methods, trans=log2, maxNum=5)
```

## Benchmark Metrics

```{r}
estimatePerformanceMetrics( pr, alpha=0.1, tidy=TRUE ) %>%
    dplyr::filter( blabel %in% c("lfdr","ihw-a10","bl-df03","qvalue","bh", "bonf") ) %>%
    dplyr::select( blabel, value )
```

# Session Info

```{r}
sessionInfo()
```