### 2018-04-06 datasets/ChIP-seq/ChIPseq-CBP-csaw.Rmd
Fix typos in ChIPseq cpb Rmd code causing errors

---
title: "Case Study: ChIP-seq Differential Peak Calling (csaw)"
author: "Patrick Kimes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
    html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

In this case study, we perform differential peak calling on ChIP-seq data for a
trnscription factor, CREB-binding protein (CBP), from Kasper et al. (2014), used in
Lun et al. (2015) to demonstrate the usage of _csaw_. As described in a separate case
study analyzing ChIP-seq data for H3K4Me3, _csaw_ is a package for differential peak
calling based on counting reads in sliding windows across the genome. Code and steps
for initial processing and analysis of this data with _csaw_ are taken directly
from Lun et al. (2015).

# Workspace Setup

```{r, wkspace-setup, results='hide', message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(SummarizedBenchmark)
library(BiocParallel)
library(rtracklayer)
library(Rsamtools)
library(Rsubread)
library(csaw)
library(edgeR)

## load helper functions
for (f in list.files("../R", "\\.(r|R)$", full.names = TRUE)) {
    source(f)
}

## project data/results folders
datdir <- "data"
resdir <- "results"
dir.create(datdir, showWarnings = FALSE)
dir.create(resdir, showWarnings = FALSE)

## intermediary files we create below
count_file <- file.path(resdir, "cbp-csaw-counts.rds")
filtered_file <- file.path(resdir, "cbp-csaw-counts-filtered.rds")
normfacs_file <- file.path(resdir, "cbp-csaw-counts-normfacs.rds")
result_file <- file.path(resdir, "cbp-csaw-results.rds")
bench_file <- file.path(resdir, "cbp-csaw-benchmark.rds")

## set up parallel backend
cores <- as.numeric(Sys.getenv("SLURM_NTASKS"))
multicoreParam <- MulticoreParam(workers = cores)
```

# Data Preparation

We download the fastq files directly from the European Nucleotide Archive (ENA).

```{r download-fastqs}
fqurls <- c("007/SRR1145787/SRR1145787.fastq.gz",
            "008/SRR1145788/SRR1145788.fastq.gz",
            "009/SRR1145789/SRR1145789.fastq.gz",
            "000/SRR1145790/SRR1145790.fastq.gz")
fqurls <- paste0("ftp://ftp.sra.ebi.ac.uk/vol1/fastq/SRR114/", fqurls)

for (i_fq in fqurls) {
    out_fq <- file.path(datdir, basename(i_fq))
    if (!file.exists(out_fq)) {
        download.file(i_fq, destfile = out_fq)
    }
}
```

We download the mm10 reference genome used for ENCODE3 and build the
index for alignment if not already available.

```{r, download-ref}
ref_url <- paste0("https://www.encodeproject.org/files/",
                  "mm10_no_alt_analysis_set_ENCODE/@@download/",
                  "mm10_no_alt_analysis_set_ENCODE.fasta.gz")
ref_fastagz <- file.path(datdir, basename(ref_url))
ref_fasta <- gsub("\\.gz$", "", ref_fastagz)

if (!file.exists(file.path(datdir, "mm_ref_index.00.b.tab"))) {
    if (!file.exists(ref_fasta)) {
        download.file(ref_url, destfile = ref_fastagz)
        system(paste("gunzip", ref_fastagz))
    }
    buildindex(basename = file.path(datdir, "mm_ref_index"),
               reference = ref_fasta)
}
```

Sample metadata is stored in a data.frame.

```{r}
fqfiles <- file.path(datdir, basename(fqurls))
meta <- data.frame(genotype = factor(c("wt", "wt", "ko", "ko")),
                   fqurl = fqurls, fqfile = fqfiles,
                   bamfile = gsub("\\.fastq\\.gz", ".bam", fqfiles),
                   sortedfile = gsub("\\.fastq\\.gz", ".sorted.bam", fqfiles),
                   stringsAsFactors = FALSE)
```

We also download blacklisted regions for mm10 from ENCODE (https://www.encodeproject.org/annotations/ENCSR636HFF/).

```{r}
blacklist_url <- "https://www.encodeproject.org/files/ENCFF547MET/@@download/ENCFF547MET.bed.gz"
if (!file.exists(file.path(datdir, basename(blacklist_url)))) {
    download.file(blacklist_url, destfile = file.path(datdir, basename(blacklist_url)))
}

bl <- import(file.path(datdir, basename(blacklist_url)))
```

Standard set of parameters are defined for the analysis.
Only the canonical set of chromosomes are used in the analysis.

```{r}
std_chr <- paste0("chr", c(1:19, "X", "Y"))
param <- readParam(minq = 20, discard = bl, restrict = std_chr)
```

## Read Counting

We count reads within sliding windows across the genome.

```{r count, results = 'hide', message = FALSE, warning = FALSE}
if (file.exists(count_file)) {
    win_cnts <- readRDS(count_file)
} else {
    unaligned <- !file.exists(meta$bamfile)
    if (any(unaligned)) {
        align(index = file.path(datdir, "mm_ref_index"),
              readfile1 = meta$fqfile[unaligned],
              type = 1, phredOffset = 64,
              input_format = "FASTQ",
              output_file = meta$bamfile[unaligned])
    }

    ## sort bam files
    for (i in 1:nrow(meta)) {
        if (!file.exists(meta$sortedfile[i])) {
            sortBam(meta$bamfile[i], gsub("\\.bam$", "", meta$sortedfile[i]))
        }
    }

    ## mark duplicates w/ picard and index bam files
    if (any(!file.exists(paste0(meta$sortedfile, ".bai")))) {
        temp_bam <- file.path(datdir, "temp_dups_mm.bam")
        temp_file <- file.path(datdir, "temp_mets_mm.txt")
        temp_dir <- file.path(datdir, "temp_dups_mm")
        dir.create(temp_dir)
        for (i_bam in meta$sortedfile) {
            code <- paste0("java -jar ${PICARD_TOOLS_HOME}/picard.jar ",
                           "MarkDuplicates I=%s O=%s M=%s ", 
                           "TMP_DIR=%s AS=true REMOVE_DUPLICATES=false ",
                           "VALIDATION_STRINGENCY=SILENT")
            code <- sprintf(code, i_bam, temp_bam, temp_file, temp_dir)
            code <- system(code)
            stopifnot(code == 0L)
            file.rename(temp_bam, i_bam)
        }
        unlink(temp_file)
        unlink(temp_dir, recursive = TRUE)

        indexBam(meta$sortedfile)
    }

    ## use correlateReads to determine fragment length (remove dups)
    x <- correlateReads(meta$sortedfile, param = reform(param, dedup = TRUE))
    frag_len <- which.max(x) - 1

    ## count reads in sliding windows (keep dups)
    win_cnts <- windowCounts(meta$sortedfile, param = param, width = 10, ext = frag_len)

    ## save unfiltered counts
    saveRDS(win_cnts, file = count_file)
}
```

We can apply prefiltering on windows with low abundance as described in Lun and Smyth (2016).

```{r}
if (file.exists(filtered_file)) {
    filtered_cnts <- readRDS(filtered_file)
    normfacs <- readRDS(normfacs_file)
} else {
    ## filter windows by abundance
    bins <- windowCounts(meta$sortedfile, bin = TRUE, width = 10000, param = param)
    filter_stat <- filterWindows(win_cnts, bins, type = "global")
    keep <- filter_stat$filter > log2(3)
    filtered_cnts <- win_data[keep, ]

    ## calculate normalizing offsets based on larger windows
    normfacs <- normOffsets(bins)

    ## save filtered counts, normalizing factors
    saveRDS(filtered_cnts, file = filtered_file)
    saveRDS(normfacs, file = normfacs_file)
}
```

# Data Analysis

## Differential Testing

We use _edgeR_ to test for differential binding with the filtered data.

```{r test, results='hide', message=FALSE, warning=FALSE}
if (file.exists(result_file)) {
    res_ranges <- readRDS(result_file)
} else {
    ## create model diesign
    design <- model.matrix(~ 0 + genotype, data = meta)
    colnames(design) <- levels(meta$genotype)

    ## run quasi-likelihood F-test
    y <- asDGEList(filtered_cnts, norm.factors = normfacs)
    y <- estimateDisp(y, design)
    fit <- glmQLFit(y, design, robust = TRUE)
    res <- glmQLFTest(fit, contrast = makeContrasts(wt-ko, levels = design))

    ## merge p-values across regions
    merged <- mergeWindows(rowRanges(filtered_cnts), tol = 100, max.width = 5000)
    tab_comb <- combineTests(merged$id, res$table)
    tab_best <- getBestTest(merged$id, res$table)

    ## save results
    res_ranges <- merged$region
    elementMetadata(res_ranges) <-
        data.frame(tabcom,
                   best_pos = mid(ranges(rowRanges(filtered_cnts[tab_best$best]))),
                   best_logFC = tab_best$logFC)
    saveRDS(res_ranges, file = result_file)
}

## covert to df for downstream analysis
res_df <- as.data.frame(res_ranges)
```

## Covariate Diagnostics

### Number of Windows

```{r, nwin-diag-scatter, fig.width=4.5, fig.height=3.5}
rank_scatter(res_df, pvalue = "PValue", covariate = "nWindows") +
    ggtitle("Number of windows as independent covariate") +
    xlab("Number of Windows")
```

```{r, nwin-diag-hist, fig.width=10, fig.height=3.2}
strat_hist(res_df, pvalue = "PValue", covariate = "nWindows", maxy = 35, numQ = 3)
```

### Region Width

```{r, width-diag-scatter, fig.width=4.5, fig.height=3.5}
rank_scatter(res_df, pvalue = "PValue", covariate = "width") +
    ggtitle("Region width as independent covariate") +
    xlab("Region Width")
```

```{r, meancov-diag-hist, fig.width=10, fig.height=3.2}
strat_hist(res_df, pvalue = "PValue", covariate = "width", maxy = 35, numQ = 3)
```

## Multiple-Testing Correction

We use the common `BenchDesign` with the set of multiple testing correction
methods already included. We use the width of the regions as the independent
covariate.

```{r}
if (!file.exists(bench_file)) {
    res_df$ind_covariate <- res_df$width
    res_df$pval <- res_df$PValue

    bd <- initializeBenchDesign()
    sb <- buildBench(bd, data = res_df, ftCols = "ind_covariate")
    saveRDS(sb, file = bench_file)
} else {
    sb <- readRDS(bench_file)
}
```

## Benchmark Metrics

Next, we'll add the default performance metric for q-value assays. First, we have
to rename the assay to 'qvalue'.

```{r, chipseq-metrics}
assayNames(sb) <- "qvalue"
sb <- addDefaultMetrics(sb)
```

Now, we'll plot the results.

```{r, chipseq-scatter, results="hide"}
rejections_scatter(sb, supplementary = FALSE)
rejection_scatter_bins(sb, covariate = "ind_covariate",
                       bins = 4, supplementary = FALSE)
```

```{r, chipseq-overlap}
plotFDRMethodsOverlap(sb, alpha = 0.05, nsets = ncol(sb),
                      order.by = "freq", decreasing = TRUE,
                      supplementary = FALSE)
```

```{r, rnaseq-lineplots}
covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
```

# Session Info

```{r}
sessionInfo()
```