### 2018-04-23 datasets/microbiome/microbiome-enigma.Rmd

---
title: "Case Study: 16S microbiome data (ENIGMA)"
author: "Claire Duvallet"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

Here we analyze a subset of the ENIGMA data ([Smith, M. and Rocha, A. et al 2015](http://dx.doi.org/10.1128/mBio.00326-15)).
The dataset includes water samples from 67 wells with well-characterized biochemical metadata (i.e. pH, heavy metal concentrations, etc).
We acquired the data from a collaborator and have put it in the `benchmark-fdr/datasets/microbiome/DATA/enigma` folder.

Note that I've uploaded this to our repo as a gzipped file, make sure to gunzip the OTU table before knitting this.

We did some preliminary correlation analysis and found that the metadata variables show a range of correlation with different OTUs: some metadata variables don't seem correlated with any OTUs, whereas others seem correlated with basically the entire community.
We'll pick three of these metadata variables to illustrate our FDR correction methods in cases where we expect different numbers of true significant associations.

# Workspace Setup

```{r, wkspace-setup, results='hide', message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(magrittr)
library(SummarizedBenchmark)

## load helper functions
for (f in list.files("../R", "\\.(r|R)$", full.names = TRUE)) {
    source(f)
}

## project data/results folders
datdir <- "data"
resdir <- "results"
enigma_dir <- file.path(datdir, "enigma_results")

# From their paper, we expect SO4 to have little correlation, pH
# to have high correlations, and Al to be in the middle
so4_result_file <- file.path(resdir, "enigma-so4-otus-results.rds")
so4_bench_file <- file.path(resdir, "enigma-so4-otus-benchmark.rds")

ph_result_file <- file.path(resdir, "enigma-ph-otus-results.rds")
ph_bench_file <- file.path(resdir, "enigma-ph-otus-benchmark.rds")

al_result_file <- file.path(resdir, "enigma-al-otus-results.rds")
al_bench_file <- file.path(resdir, "enigma-al-otus-benchmark.rds")
```

# Data Preparation

We download the data from Dropbox if not already available.

```{r, enigma-download-data}
if (!file.exists(file.path(enigma_dir, "enigma.metadata.txt"))) {
    download.file("https://www.dropbox.com/s/0r1zjp26u10l0p4/enigma.metadata.txt?dl=1",
                  destfile = file.path(enigma_dir, "enigma.metadata.txt"))
}

if (!file.exists(file.path(enigma_dir, "enigma.otu_table_resampled_updated_r.txt.gz"))) {
    download.file("https://www.dropbox.com/s/u9qh86cd7c0vbn9/enigma.otu_table_resampled_updated_r.txt.gz?dl=1",
                  destfile = file.path(enigma_dir, "enigma.otu_table_resampled_updated_r.txt.gz"))
}
```

```{r, enigma-readdata}
## load OTU table and metadata
print(file.path(enigma_dir, "enigma.otu_table_resampled_updated_r.txt.gz"))
otu <- read.table(file.path(enigma_dir, "enigma.otu_table_resampled_updated_r.txt.gz"), 
                  sep='\t', header = TRUE, row.names = 1)
meta <- read.csv(file.path(enigma_dir, "enigma.metadata.txt"), sep='\t')

# Select only the 0.02 um samples
meta <- meta[endsWith(as.character(meta$Sample), "02"), ]
# Make metadata sample ID match OTU table
meta$Sample <- gsub("_", "\\.", as.character(meta$Sample))
meta$Sample <- substr(meta$Sample, 1, nchar(meta$Sample)-3)

## keep only samples with both metadata and 16S data
keep_samples <- intersect(colnames(otu), meta$Sample)
otu <- otu[, keep_samples]
meta <- dplyr::filter(meta, Sample %in% keep_samples)

## Make sure samples in metadata and OTU table match
meta <- meta[match(colnames(otu), meta$Sample), ]
```

We'll do a little filtering here before calculating correlations.
We'll apply a minimum threshold of 10 reads per OTU across all samples.
After removing these shallow OTUs, we also get rid of any samples with too few reads. 
We define the minimum number of reads per OTU in `min_otu`, and
the minimum number of reads per sample in `min_sample`.

After we've removed any shallow OTUs and samples, we'll convert the OTU table to
relative abundances.

```{r, enigma-cleandata}
min_otu <- 10
minp_otu <- 0.01
min_sample <- 100

## Remove OTUs w/ <= min reads, w/ <= min prop, samples w/ <= min reads
otu <- otu[rowSums(otu) > min_otu, ]
otu <- otu[rowSums(otu > 0) / ncol(otu) > minp_otu, ]
otu <- otu[, colSums(otu) > min_sample]

## Update metadata with new samples
meta <- dplyr::filter(meta, Sample %in% colnames(otu))

## Remove empty OTUs
otu <- otu[rowSums(otu) > 0, ]

## Convert to relative abundance
abun_otu <- t(t(otu) / rowSums(t(otu)))

## Add pseudo counts
zeroabun <- 0
abun_otu <- abun_otu + zeroabun
```

# Data Analysis

## pH

Next, we need to calculate the pvalues for correlations between abundances of each OTU and the metadata variables of interest. (i.e. we're correlating each OTU's abundances across wells with the metadata variable across wells).
Let's start with "pH".
We'll store the pvalue results in a dataframe.

While we're at it, we'll also calculate the mean abundance and ubiquity (detection rate)
of each OTU. Later, we can assign their values to a new column called `ind_covariate` 
for use in downstream steps.

```{r, ph-pvals}
if (!file.exists(ph_result_file)) {
    res <- test_microbiome_corr(abundance = abun_otu, meta_col = meta$pH,
                                shift = zeroabun)
    saveRDS(res, file = ph_result_file)
} else {
    res <- readRDS(ph_result_file)
}
```

## Covariate Diagnostics

Here we look to see if the covariates do indeed look informative.

### Ubiquity

```{r, ph-ubi-hist, fig.width=10, fig.height=3.2}
strat_hist(res, pvalue="pval", covariate="ubiquity", maxy=10, binwidth=0.05)
```

```{r, ph-ubi-scatter, fig.width=4.5, fig.height=3.5}
rank_scatter(res, pvalue="pval", covariate="ubiquity")
```

### Mean Abundance (across non-zero samples)

```{r, ph-abun-hist, fig.width=10, fig.height=3.2}
strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=8, binwidth=0.05)
```

```{r, ph-abun-scatter, fig.width=4.5, fig.height=3.5}
rank_scatter(res, pvalue="pval", covariate="mean_abun_present")
```

## Multiple-Testing Correction

Let's use `ubiquity` as our `ind_covariate`.

```{r, enigma-labelcovariate}
res <- dplyr::mutate(res, ind_covariate = ubiquity)
```

First, we'll create an object of `BenchDesign` class to hold the data and 
add the benchmark methods to the `BenchDesign` object. We remove ASH from
the default comparison.

Then, we'll construct the `SummarizedBenchmark` object, which will run
the functions specified in each method (these are actually sourced in from the
helper scripts).

```{r, ph-sb, results="hide", message=FALSE, eval=TRUE}
if (!file.exists(ph_bench_file)) {
    bd <- initializeBenchDesign()
    bd <- dropBMethod(bd, "ashq")
    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
    saveRDS(sb, file = ph_bench_file)
} else {
    sb <- readRDS(ph_bench_file)
}
```

## Benchmark Metrics

Next, we'll add the default performance metric for q-value assays. First, we have
to rename the assay to 'qvalue'.

```{r, ph-metrics, eval=TRUE}
assayNames(sb) <- "qvalue"
sb <- addDefaultMetrics(sb)
```

Now, we'll plot the results.

```{r, ph-plot, width=15, height=15, eval=TRUE}
rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
```

## SO4

Next, we'll look at SO4 levels

```{r, so4-pvals}
if (!file.exists(so4_result_file)) {
    res <- test_microbiome_corr(abundance = abun_otu, meta_col = meta$SO4_mgL,
                                shift = zeroabun)
    saveRDS(res, file = so4_result_file)
} else {
    res <- readRDS(so4_result_file)
}
```

## Covariate Diagnostics

Here we look to see if the covariates do indeed look informative.

### Ubiquity

```{r, so4-ubi-hist, fig.width=10, fig.height=3.2}
strat_hist(res, pvalue="pval", covariate="ubiquity", maxy=10, binwidth=0.05)
```

```{r, so4-ubi-scatter, fig.width=4.5, fig.height=3.5}
rank_scatter(res, pvalue="pval", covariate="ubiquity")
```

### Mean Abundance (across non-zero samples)

```{r, so4-abun-hist, fig.width=10, fig.height=3.2}
strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=8, binwidth=0.05)
```

```{r, so4-abun-scatter, fig.width=4.5, fig.height=3.5}
rank_scatter(res, pvalue="pval", covariate="mean_abun_present")
```

## Multiple-Testing Correction

Let's use `ubiquity` as our `ind_covariate`.

```{r, so4-labelcovariate}
res <- dplyr::mutate(res, ind_covariate = ubiquity)
```

First, we'll create an object of `BenchDesign` class to hold the data and 
add the benchmark methods to the `BenchDesign` object. We remove ASH from
the default comparison.

Then, we'll construct the `SummarizedBenchmark` object, which will run
the functions specified in each method (these are actually sourced in from the
helper scripts).

```{r, so4-sb, results="hide", message=FALSE, eval=TRUE}
if (!file.exists(so4_bench_file)) {
    bd <- initializeBenchDesign()
    bd <- dropBMethod(bd, "ashq")
    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
    saveRDS(sb, file = so4_bench_file)
} else {
    sb <- readRDS(so4_bench_file)
}
```

## Benchmark Metrics

Next, we'll add the default performance metric for q-value assays. First, we have
to rename the assay to 'qvalue'.

```{r, so4-metrics, eval=TRUE}
assayNames(sb) <- "qvalue"
sb <- addDefaultMetrics(sb)
```

Now, we'll plot the results.

```{r, so4-plot, width=15, height=15, eval=TRUE}
rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
```

## Al

Finally, we'll look at Al levels

```{r, al-pvals}
if (!file.exists(al_result_file)) {
    res <- test_microbiome_corr(abundance = abun_otu, meta_col = meta$Al_mgL,
                                shift = zeroabun)
    saveRDS(res, file = al_result_file)
} else {
    res <- readRDS(al_result_file)
}
```

## Covariate Diagnostics

Here we look to see if the covariates do indeed look informative.

### Ubiquity

```{r, al-ubi-hist, fig.width=10, fig.height=3.2}
strat_hist(res, pvalue="pval", covariate="ubiquity", maxy=10, binwidth=0.05)
```

```{r, al-ubi-scatter, fig.width=4.5, fig.height=3.5}
rank_scatter(res, pvalue="pval", covariate="ubiquity")
```

### Mean Abundance (across non-zero samples)

```{r, al-abun-hist, fig.width=10, fig.height=3.2}
strat_hist(res, pvalue="pval", covariate="mean_abun_present", maxy=8, binwidth=0.05)
```

```{r, al-abun-scatter, fig.width=4.5, fig.height=3.5}
rank_scatter(res, pvalue="pval", covariate="mean_abun_present")
```

## Multiple-Testing Correction

Let's use `ubiquity` as our `ind_covariate`.

```{r, al-labelcovariate}
res <- dplyr::mutate(res, ind_covariate = ubiquity)
```

First, we'll create an object of `BenchDesign` class to hold the data and 
add the benchmark methods to the `BenchDesign` object. We remove ASH from
the default comparison.

Then, we'll construct the `SummarizedBenchmark` object, which will run
the functions specified in each method (these are actually sourced in from the
helper scripts).

```{r, al-sb, results="hide", message=FALSE, eval=TRUE}
if (!file.exists(al_bench_file)) {
    bd <- initializeBenchDesign()
    bd <- dropBMethod(bd, "ashq")
    sb <- buildBench(bd, data = res, ftCols = "ind_covariate")
    saveRDS(sb, file = al_bench_file)
} else {
    sb <- readRDS(al_bench_file)
}
```

## Benchmark Metrics

Next, we'll add the default performance metric for q-value assays. First, we have
to rename the assay to 'qvalue'.

```{r, al-metrics, eval=TRUE}
assayNames(sb) <- "qvalue"
sb <- addDefaultMetrics(sb)
```

Now, we'll plot the results.

```{r, al-plot, width=15, height=15, eval=TRUE}
rejections_scatter( sb, as_fraction=FALSE, supplementary=FALSE)
rejection_scatter_bins(sb, covariate="ind_covariate", supplementary=FALSE)
```

# Session Info

```{r}
sessionInfo()
```