### 2018-04-27 manuscript_figures/mansuscript_figures.Rmd
started Rmd to create manuscript figures

---
title: "Manuscript Figures for Genome Biology"
author: "Rafalab"
output: 
    html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Figure 1

This is a schema of Inputs, Assumptions and Outputs for the various FDR methods. Claire is leading this charge.

# Figure 2

This is a faceted figure with results using the yeast data with 48 biological replicates analyzed in this [publication](https://www.ncbi.nlm.nih.gov/pubmed/26206307/). A complete analysis 
can be found in the `/datasets/RNA-Seq/yeast-simulation.Rmd`. This is the 5v5 comparison.

```{r Figure2-load-data}
# Load packages and source benchmark FDR
library(data.table)
library(tidyr)
library(dplyr)
library(readr)
library(ggplot2)
library(magrittr)
library(cowplot)
library(purrr)
library(DESeq2)
library(tibble)
library(ggthemes)
library(R.utils)

## load helper functions
for (f in list.files("../R", "\\.(r|R)$", full.names = TRUE)) {
    source(f)
}

# set up data / results directories
resdir <- "yeast-results"
```

```{r Figure2}
rseed <- 198
sampleSize <- 5
nDE <- 500
resfile <- file.path(resdir, "de5.rds")
de5 <- readRDS(file=resfile)
res5d <- plotsim_standardize(de5, alpha = seq(0.01, 0.10, 0.01))

# Figure 2A: Not all methods control FDR at nominal alpha level
p1 <- plotsim_average(res5d, met="FDR",filter_set = excludeSet,
                merge_ihw = TRUE, errorBars=TRUE)

# Figure 2B: TPR of methods across nominal alpha level
p2 <- plotsim_average(res5d, met="TPR",filter_set = excludeSet,
                merge_ihw = TRUE, errorBars=TRUE)

# Figure 2C: Total number of discoveries varied across method
p3 <- plotsim_average(res5d, met="rejections",filter_set = excludeSet,
                merge_ihw = TRUE, errorBars=TRUE)

# Figure 2D: Some methods are tailored to detecting smaller effect sizes
p4 <- covariateHeatmap(de5, alpha=0.05, covname="log2FC", trans="log1p")

# maybe 2E? 
# p5 <- aggupset(de5, alpha=ualpha, supplementary = FALSE, return_list = FALSE)

# Figure 2 (ALL)
plot_grid(p1, p2, p3, p4, ncol = 2, 
          labels = LETTERS[1:4], label_size = 25)
```

# Session information

```{r}
sessionInfo()
```