### 2018-07-23 datasets/RNA-Seq/polyester-simulation.Rmd
Add flexibility to simulate arbitrary number of genes

---
title: "Polyester RNA-seq simulation study"
author: "Stephanie Hicks"
output: 
    html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This is an analysis using [polyester](https://bioconductor.org/packages/release/bioc/html/polyester.html) Bioconductor package to simulate RNA-Seq reads. We will  
(1) implement null comparisons on samples simulated using polyester
with no differentially expressed (DE) genes, (2) compare FDR approaches 
on their ability to discover 'true positive' DE genes with log2 fold 
changes simulated from a normal distribution.

In this Rmd, we will perform a Monte Carlo simulation study 
and average over the replicates in plots.

## Set up workspace

```{r workspace-setup, results='hide', message=FALSE, warning=FALSE}
# Load packages needed to simulate RNA-seq counts with polyester
library(SummarizedBenchmark)
library(genefilter)
library(limma)
library(polyester)
library(DESeq2)
library(data.table)
library(magrittr)
library(dplyr)
library(ggplot2)
library(cowplot)
library(tibble)
library(ggthemes)
library(readr)

# load helper functions
for (f in list.files("../R", "\\.(r|R)$", full.names = TRUE)) {
    source(f)
}

# set up results directories 
datdir <- "yeast-data"
resdir <- "../../results/RNAseq"
dir.create(datdir, showWarnings = FALSE)
dir.create(resdir, showWarnings = FALSE)

# results files that will be generated below
resfile_n5 <- file.path(resdir, "polyester-results-null5.rds")
resfile_n10 <- file.path(resdir, "polyester-results-null10.rds")
resfile_d5 <- file.path(resdir, "polyester-results-de5.rds")
resfile_d10 <- file.path(resdir, "polyester-results-de10.rds")

resfile_n5_uninfCov <- file.path(resdir, "polyester-results-null5-uninfCov.rds")
resfile_n10_uninfCov <- file.path(resdir, "polyester-results-null10-uninfCov.rds")
resfile_d5_uninfCov <- file.path(resdir, "polyester-results-de5-uninfCov.rds")
resfile_d10_uninfCov <- file.path(resdir, "polyester-results-de10-uninfCov.rds")

# # set up parallel backend:
library(BiocParallel)
nCores <- 20
register(MulticoreParam(workers = nCores))
```

## Set up Polyester

We use the [polyester](https://bioconductor.org/packages/release/bioc/html/polyester.html) Bioconductor package to simulate RNA-Seq reads. The package requires a count matrix as input which estimates the mean variance relationship. Here we start with the yeast data with 48 biological replicates in each of 
two conditions (analyzed in this [publication](https://www.ncbi.nlm.nih.gov/pubmed/26206307/)). 
We'll use the samples passing QC in the original study that belong to the WT group to 
estimate mean and variance relationship. We assume the data has already been downloaded
(this is carried out in the `yeast-simulation.Rmd` vignette).

```{r, readin, results='hide', message=FALSE}
files <- dir(path = datdir, pattern = "*.bam.gbgout", full.names = TRUE)
sample_names <- sapply(strsplit(dir(path = datdir, pattern = "*.bam.gbgout"), "_MID"),
                       function(x) x[[1]])
badreps <- read_tsv(file.path(datdir, "badreps.txt"), col_names = FALSE)$X1
badreps <- unlist(lapply(strsplit(badreps, "_MID"), function(x) x[1]))

counts <- files %>%
  purrr::map(read_tsv, col_names = FALSE) %>% # read in all the files individually
  purrr::map2(sample_names, ~ dplyr:::rename(.x, !! .y := X2, feature = X1) ) %>% # add sample names
  purrr::reduce(left_join, by = "feature") %>% # reduce with rbind into one dataframe
  dplyr::select(-badreps ) %>% # remove badreps
  data.frame() %>%
  dplyr::select(contains("WT"))  %>%
  as.matrix

# filter low count genes
counts <- counts[rowMeans(counts) > 1,]

## Estimate the zero inflated negative binomial parameters
paramPolyester = get_params(counts)
```

## Helper functions for simulation

Next, we'll simulate RNA-Seq samples, both with and without the 
addition of simulated DE genes. Here we'll create a function that we can use to 
run one replicate given sample size and number of DE gene settings. This will be
looped over many replications and results averaged over them.

```{r}
#' @param sampleSize is the number of samples in each condition
#' @param nDE is the number of DE genes
#' @param bd is the bench design object
#' @param uninformativeCovariate logical indicating whether to replace overall mean 
#'  independent covariate with a randomly generated (uninformative) covariate. 
#'  Default is FALSE.
#' @param rseed random seed
simulateOneSplit <- function(X, rseed, 
                             nDE, sampleSize, bd, uninformativeCovariate = FALSE){
  ngenes <- 10000

  uninf_covar <- rnorm(ngenes)

  # set random seed
  set.seed(as.numeric(X)*as.numeric(rseed))

  # things needed to simulate with Polyester
  groupPolyester = rep(c(-1,1),each=sampleSize)
  modPolyester = model.matrix(~-1 + groupPolyester)

  # things needed for differential testing using DESeq2
  groupDESeq = factor(rep(c(0,1), each=sampleSize))

  gs <- sample(1:nrow(counts), ngenes, replace = TRUE)
  mu <- paramPolyester$mu[gs]
  p0 <- paramPolyester$p0[gs]

  # Null genes
  log2FC = rep(0, ngenes)
  dat_alt = create_read_numbers(mu,
                                paramPolyester$fit,
                                p0,
                                n=sampleSize*2, beta=cbind(log2FC), 
                                mod=modPolyester)

  # filter genes with mostly zeroes
  pzero = rowSums(dat_alt==0)/ncol(dat_alt)
  dat_alt <- dat_alt[pzero < 0.5,]
  truth <- rep(FALSE, nrow(dat_alt))
  log2FC <- log2FC[pzero < 0.5]
  mu <- mu[pzero < 0.5]
  p0 <- p0[pzero < 0.5]
  uninf_covar <- uninf_covar[pzero < 0.5]

  # DE simulation
  if(nDE > 0){
    # pick random set of nDE genes to add signal to
    DE <- sample(1:nrow(dat_alt), nDE)
    truth[DE] <- TRUE

    # randomly sample a log2FC from N(0,1) 
    log2FC[DE] <- rnorm(nDE, sd = 1)
    dat_alt[DE,] = create_read_numbers(mu[DE], paramPolyester$fit, p0[DE], 
                                  n=sampleSize*2, beta=cbind(log2FC[DE]), 
                                  mod=modPolyester)
  }

  # run DESeq2
  dds_alt <- DESeqDataSetFromMatrix(countData=dat_alt, 
                                    colData = DataFrame(groupDESeq), 
                                    design=~groupDESeq)
  dds_alt <- DESeq(dds_alt, parallel = FALSE)
  res_alt <- results(dds_alt, independentFiltering = F)

  geneExp <- tbl_df(data.frame(pval=res_alt$pvalue, SE=res_alt$lfcSE, 
                               ind_covariate = res_alt$baseMean, 
                               effect_size=res_alt$log2FoldChange, 
                               test_statistic=res_alt$stat, qvalue=truth))

  if (uninformativeCovariate)
    geneExp <- mutate(geneExp, ind_covariate = uninf_covar)

  # filter NAs 
  geneExp <- geneExp %>% na.omit()

  # built Benchmark data
  sb <- bd %>% buildBench(data=geneExp, parallel = FALSE, 
                          truthCols = "qvalue", 
                          ftCols = "ind_covariate")
  sb <- addDefaultMetrics(sb)
  rowData(sb)$log2FC <- geneExp$effect_size

  return(sb)
}

```

We'll also set some parameters that will be common to all simulations. These
include the number of replications, the bench design object, the set of 
methods to exclude in the results plots, and the alpha cutoff level to 
be used when plotting the aggregated Upset results.

```{r}
B <- 100
bd <- initializeBenchDesign() # only needs to be done once
excludeSet <- c("unadjusted", "bl-df02", "bl-df04", "bl-df05") 
ualpha <- 0.05
```

We also add in Scott's FDR Regression (both
`nulltype = "empirical"` and `nulltype = "theoretical"`)
since our test statistics are approximately t-distributed.

```{r}
bd <- addBMethod(bd, "scott-theoretical",
                     FDRreg::FDRreg,
                     function(x) { x$FDR },
                     z = test_statistic,
                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
                     nulltype = 'theoretical',
                     control = list(lambda = 0.01))
bd <- addBMethod(bd, "scott-empirical",
                     FDRreg::FDRreg,
                     function(x) { x$FDR },
                     z = test_statistic,
                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
                     nulltype = 'empirical',
                     control = list(lambda = 0.01))
```

Here's a helper function to return the number of methods with rejections at
a particular alpha level (this helps us determine whether or not to plot the
aggregated upset plot - if there aren't at least 2 methods it will throw an
error, which is a problem for the null simulations).

```{r}
# To be included in the upset agg plot, method needs to have found on average
# at least one rejection per replicate. To create an upset plot, require that
# at least two methods rejected at this threshold.
#' @param res standardized metric data.table generated using
#'        standardize_results.
#' @param alpha alpha cutoff
#' @param filterSet which methods to exclude from consideration 
numberMethodsReject <- function(res, alphacutoff, filterSet){
  res <- res %>% 
    filter(is.na(param.alpha) | (param.alpha == alphacutoff)) %>%
    filter(!(blabel %in% filterSet)) %>%
    filter(alpha == alphacutoff) %>%
    filter(performanceMetric == "rejections") %>%
    select(blabel, performanceMetric, value) %>%
    group_by(blabel) %>%
    summarize(mean_value = mean(value)) %>%
    filter(mean_value > 1)
  return(nrow(res))
}
```

# Sim N5: Null Comparison 5v5

Here we run a null comparison 5 versus 5 RNA-Seq samples. 
This will be done for 100 simulations.

## Generate a list of SB results objects

```{r, results='hide', message=FALSE}
sampleSize <- 5
nDE <- 0
rseed <- 225

if (!file.exists(resfile_n5)){
  null5 <- bplapply(X=1:B, rseed=rseed, FUN=simulateOneSplit,  
                    nDE=nDE, sampleSize=sampleSize, bd=bd)
  saveRDS(null5, file=resfile_n5)
} else {
  null5 <- readRDS(file=resfile_n5)
}
```

## Plot average results over replications

Plot results.

```{r}
# Check for missing results (if any methods threw an error for relevant metrics).
rowSums(sapply(null5, function(x) colSums(is.na(assays(x)$qvalue)) > 0))

res5 <- plotsim_standardize(null5, alpha = seq(0.01, 0.10, 0.01))

# metrics = ""TPR"   "FPR"   "TNR"   "FNR"   "rejections"  "FWER"   "rejectprop"
plotsim_average(res5, met="rejections",filter_set = excludeSet,
                merge_ihw = TRUE, errorBars=TRUE) 
plotsim_average(res5, met="TNR",filter_set = excludeSet,
                merge_ihw = TRUE, errorBars=TRUE)

covariateLinePlot(null5, alpha=0.05, covname="log2FC", nbins=25, 
                 trans="log1p")
covariateLinePlot(null5, alpha=0.05, covname="ind_covariate", nbins=25, 
                 trans="log1p")

if (numberMethodsReject(res5, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
  aggupset(null5, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
} else {
  message("Not enough methods found rejections at alpha ", ualpha, 
          "; skipping upset plot")
}

```

# Sim N10: Null Comparison 10v10

Here we'll repeat the above, but for a null comparison 10 versus 10 samples. 
This will be done for 100 simulations.

## Generate a list of SB results objects

```{r, results='hide', message=FALSE}
sampleSize <- 10
nDE <- 0
rseed <- 837

if (!file.exists(resfile_n10)){
  null10 <- bplapply(X=1:B, rseed=837, FUN=simulateOneSplit,
                     nDE=nDE, sampleSize=sampleSize, bd=bd)
  saveRDS(null10, file=resfile_n10)
} else {
  null10 <- readRDS(file=resfile_n10)
}
```

## Plot average results over replications

Plot results.

```{r}
# Check for missing results (if any methods threw an error for relevant metrics).
rowSums(sapply(null10, function(x) colSums(is.na(assays(x)$qvalue)) > 0))

res10 <- plotsim_standardize(null10, alpha = seq(0.01, 0.10, 0.01))

plotsim_average(res10, met="rejections",filter_set = excludeSet,
                merge_ihw = TRUE, errorBars=TRUE) 
plotsim_average(res10, met="TNR",filter_set = excludeSet,
                merge_ihw = TRUE, errorBars=TRUE)

covariateLinePlot(null10, alpha=0.05, covname="log2FC", nbins=25, 
                 trans="log1p")
covariateLinePlot(null10, alpha=0.05, covname="ind_covariate", nbins=25, 
                 trans="log1p")

if (numberMethodsReject(res10, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
  aggupset(null10, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
} else {
  message("Not enough methods found rejections at alpha ", ualpha, "; skipping upset plot")
}

```

# Sim D5: DE Comparison 5v5

Here we'll repeat the above, but for a DE comparison 5 versus 5 samples
and 2000 DE genes are added. This will be done for 100 simulations.

## Generate a list of SB results objects

```{r, results='hide', message=FALSE}
sampleSize <- 5
nDE <- 2000
rseed <- 198

if (!file.exists(resfile_d5)){
  de5 <- bplapply(X=1:B, rseed = rseed, FUN=simulateOneSplit, 
                  nDE=nDE, sampleSize=sampleSize, bd=bd)
  saveRDS(de5, file=resfile_d5)
} else {
  de5 <- readRDS(file=resfile_d5)
}
```

## Plot average results over replications

Plot results.

```{r}
# Check for missing results (if any methods threw an error for relevant metrics).
rowSums(sapply(de5, function(x) colSums(is.na(assays(x)$qvalue)) > 0))

res5d <- plotsim_standardize(de5, alpha = seq(0.01, 0.20, 0.01))

plotsim_average(res5d, met="rejections",filter_set = excludeSet,
                merge_ihw = TRUE, errorBars=TRUE) 
plotsim_average(res5d, met="FDR",filter_set = excludeSet,
                merge_ihw = TRUE, errorBars=TRUE) 
plotsim_average(res5d, met="TPR",filter_set = excludeSet,
                merge_ihw = TRUE, errorBars=TRUE)

covariateLinePlot(de5, alpha=0.05, covname="log2FC", nbins=25, 
                 trans="log1p")
covariateLinePlot(de5, alpha=0.05, covname="ind_covariate", nbins=25, 
                 trans="log1p")

if (numberMethodsReject(res5d, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
  aggupset(de5, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
} else {
  message("Not enough methods found rejections at alpha ", ualpha, "; skipping upset plot")
}

```

# Sim D10: DE Comparison 10v10

Here we'll repeat the above, but for a DE comparison 10 versus 10 samples
and 2000 DE genes are added. This will be done for 100 simulations.

## Generate a list of SB results objects

```{r, results='hide', message=FALSE}
sampleSize <- 10
nDE <- 2000
rseed <- 961

if (!file.exists(resfile_d10)){
  de10 <- bplapply(X=1:B, rseed = rseed, FUN=simulateOneSplit, 
                   nDE=nDE, sampleSize=sampleSize, bd=bd)
  saveRDS(de10, file=resfile_d10)
} else {
  de10 <- readRDS(file=resfile_d10)
}

```

## Plot average results over replications

Plot results.

```{r}
# Check for missing results (if any methods threw an error for relevant metrics).
rowSums(sapply(de10, function(x) colSums(is.na(assays(x)$qvalue)) > 0))

res10d <- plotsim_standardize(de10, alpha = seq(0.01, 0.20, 0.01))

plotsim_average(res10d, met="rejections",filter_set = excludeSet, 
                merge_ihw = TRUE, errorBars=TRUE) 
plotsim_average(res10d, met="FDR",filter_set = excludeSet,
                merge_ihw = TRUE, errorBars=TRUE) 
plotsim_average(res10d, met="TPR",filter_set = excludeSet,
                merge_ihw = TRUE, errorBars=TRUE)

covariateLinePlot(de10, alpha=0.05, covname="log2FC", nbins=25, 
                 trans="log1p")
covariateLinePlot(de10, alpha=0.05, covname="ind_covariate", nbins=25, 
                 trans="log1p")

if (numberMethodsReject(res10d, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
  aggupset(de10, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
} else {
  message("Not enough methods found rejections at alpha ", ualpha, "; skipping upset plot")
}
```

# Sample size comparison

Here we compare the method ranks for the different sample sizes at alpha = 0.10.

```{r}
plotMethodRanks(c(resfile_d5, resfile_d10), 
                colLabels = c("DE 5", "DE 10"), 
                alpha = 0.10, xlab = "Comparison",
                excludeMethods = NULL)
```

# Uninformative Covariate

## Sim N5: Null Comparison 5v5

Here we run a null comparison 5 versus 5 RNA-Seq samples. 
This will be done for 100 simulations.

### Generate a list of SB results objects

```{r, results='hide', message=FALSE}
sampleSize <- 5
nDE <- 0
rseed <- 225

if (!file.exists(resfile_n5_uninfCov)){
  null5 <- bplapply(X=1:B, rseed=rseed, FUN=simulateOneSplit, 
                    nDE=nDE, sampleSize=sampleSize, 
                   bd=bd, uninformativeCovariate = TRUE)
  saveRDS(null5, file=resfile_n5_uninfCov)
} else {
  null5 <- readRDS(file=resfile_n5_uninfCov)
}
```

### Plot average results over replications

Plot results.

```{r}
# Check for missing results (if any methods threw an error for relevant metrics).
rowSums(sapply(null5, function(x) colSums(is.na(assays(x)$qvalue)) > 0))

res5 <- plotsim_standardize(null5, alpha = seq(0.01, 0.10, 0.01))

# metrics = ""TPR"   "FPR"   "TNR"   "FNR"   "rejections"  "FWER"   "rejectprop"
plotsim_average(res5, met="rejections",filter_set = excludeSet,
                merge_ihw = TRUE, errorBars=TRUE) 
plotsim_average(res5, met="TNR",filter_set = excludeSet,
                merge_ihw = TRUE, errorBars=TRUE)

covariateLinePlot(null5, alpha=0.05, covname="log2FC", nbins=25, 
                 trans="log1p")
covariateLinePlot(null5, alpha=0.05, covname="ind_covariate", nbins=25, 
                 trans="log1p")

if (numberMethodsReject(res5, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
  aggupset(null5, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
} else {
  message("Not enough methods found rejections at alpha ", ualpha, 
          "; skipping upset plot")
}

```

## Sim N10: Null Comparison 10v10

Here we'll repeat the above, but for a null comparison 10 versus 10 samples. 
This will be done for 100 simulations.

### Generate a list of SB results objects

```{r, results='hide', message=FALSE}
sampleSize <- 10
nDE <- 0
rseed <- 837

if (!file.exists(resfile_n10_uninfCov)){
  null10 <- bplapply(X=1:B, rseed=rseed, FUN=simulateOneSplit, 
                     nDE=nDE, sampleSize=sampleSize, 
                   bd=bd, uninformativeCovariate = TRUE)
  saveRDS(null10, file=resfile_n10_uninfCov)
} else {
  null10 <- readRDS(file=resfile_n10_uninfCov)
}
```

### Plot average results over replications

Plot results.

```{r}
# Check for missing results (if any methods threw an error for relevant metrics).
rowSums(sapply(null10, function(x) colSums(is.na(assays(x)$qvalue)) > 0))

res10 <- plotsim_standardize(null10, alpha = seq(0.01, 0.10, 0.01))

plotsim_average(res10, met="rejections",filter_set = excludeSet,
                merge_ihw = TRUE, errorBars=TRUE) 
plotsim_average(res10, met="TNR",filter_set = excludeSet,
                merge_ihw = TRUE, errorBars=TRUE)

covariateLinePlot(null10, alpha=0.05, covname="log2FC", nbins=25, 
                 trans="log1p")
covariateLinePlot(null10, alpha=0.05, covname="ind_covariate", nbins=25, 
                 trans="log1p")

if (numberMethodsReject(res10, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
  aggupset(null10, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
} else {
  message("Not enough methods found rejections at alpha ", ualpha, "; skipping upset plot")
}

```

## Sim D5: DE Comparison 5v5

Here we'll repeat the above, but for a DE comparison 5 versus 5 samples
and 2000 DE genes are added. This will be done for 100 simulations.

### Generate a list of SB results objects

```{r, results='hide', message=FALSE}
sampleSize <- 5
nDE <- 2000
rseed <- 198

if (!file.exists(resfile_d5_uninfCov)){
  de5 <- bplapply(X=1:B, rseed=rseed, FUN=simulateOneSplit, 
                  nDE=nDE, sampleSize=sampleSize, 
                  bd=bd, uninformativeCovariate = TRUE)
  saveRDS(de5, file=resfile_d5_uninfCov)
} else {
  de5 <- readRDS(file=resfile_d5_uninfCov)
}
```

### Plot average results over replications

Plot results.

```{r}
# Check for missing results (if any methods threw an error for relevant metrics).
rowSums(sapply(de5, function(x) colSums(is.na(assays(x)$qvalue)) > 0))

res5d <- plotsim_standardize(de5, alpha = seq(0.01, 0.20, 0.01))

plotsim_average(res5d, met="rejections",filter_set = excludeSet,
                merge_ihw = TRUE, errorBars=TRUE) 
plotsim_average(res5d, met="FDR",filter_set = excludeSet,
                merge_ihw = TRUE, errorBars=TRUE) 
plotsim_average(res5d, met="TPR",filter_set = excludeSet,
                merge_ihw = TRUE, errorBars=TRUE)

covariateLinePlot(de5, alpha=0.05, covname="log2FC", nbins=25, 
                 trans="log1p")
covariateLinePlot(de5, alpha=0.05, covname="ind_covariate", nbins=25, 
                 trans="log1p")

if (numberMethodsReject(res5d, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
  aggupset(de5, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
} else {
  message("Not enough methods found rejections at alpha ", ualpha, "; skipping upset plot")
}

```

## Sim D10: DE Comparison 10v10

Here we'll repeat the above, but for a DE comparison 10 versus 10 samples
and 2000 DE genes are added. This will be done for 100 simulations.

### Generate a list of SB results objects

```{r, results='hide', message=FALSE}
sampleSize <- 10
nDE <- 2000
rseed <- 961

if (!file.exists(resfile_d10_uninfCov)){
  de10 <- bplapply(X=1:B, rseed=rseed, FUN=simulateOneSplit, 
                   nDE=nDE, sampleSize=sampleSize, 
                   bd=bd, uninformativeCovariate = TRUE)
  saveRDS(de10, file=resfile_d10_uninfCov)
} else {
  de10 <- readRDS(file=resfile_d10_uninfCov)
}

```

### Plot average results over replications

Plot results.

```{r}
# Check for missing results (if any methods threw an error for relevant metrics).
rowSums(sapply(de10, function(x) colSums(is.na(assays(x)$qvalue)) > 0))

res10d <- plotsim_standardize(de10, alpha = seq(0.01, 0.20, 0.01))

plotsim_average(res10d, met="rejections",filter_set = excludeSet, 
                merge_ihw = TRUE, errorBars=TRUE) 
plotsim_average(res10d, met="FDR",filter_set = excludeSet,
                merge_ihw = TRUE, errorBars=TRUE) 
plotsim_average(res10d, met="TPR",filter_set = excludeSet,
                merge_ihw = TRUE, errorBars=TRUE)

covariateLinePlot(de10, alpha=0.05, covname="log2FC", nbins=25, 
                 trans="log1p")
covariateLinePlot(de10, alpha=0.05, covname="ind_covariate", nbins=25, 
                 trans="log1p")

if (numberMethodsReject(res10d, alphacutoff=ualpha, filterSet=excludeSet) >= 2){
  aggupset(de10, alpha=ualpha, supplementary = FALSE, return_list = FALSE) 
} else {
  message("Not enough methods found rejections at alpha ", ualpha, "; skipping upset plot")
}
```

## Sample size comparison

Here we compare the method ranks for the different sample sizes at alpha = 0.10.

```{r}
# uninformative covariate
plotMethodRanks(c(resfile_d5_uninfCov, resfile_d10_uninfCov), 
                colLabels = c("DE 5", "DE 10"), 
                alpha = 0.10, xlab = "Comparison",
                excludeMethods = NULL)
```

# Session information

```{r}
sessionInfo()
```