### 2018-07-26 datasets/simulations/simulations-null.Rmd
Add info vs uninfo comparisons to all sims

---
title: "Simulation Study: Null Settings"
author: "Patrick Kimes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

In this set of simulations, we only consider settings where all tests are null.

# Workspace Setup

```{r, wkspace-setup, results='hide', message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(SummarizedBenchmark)
library(parallel)

## load helper functions
for (f in list.files("../R", "\\.(r|R)$", full.names = TRUE)) {
    source(f)
}

## project data/results folders
resdir <- "results"
dir.create(resdir, showWarnings = FALSE, recursive = TRUE)

## intermediary files we create below
gauss_file <- file.path(resdir, "null-benchmark-gaussian.rds")
tdist_file <- file.path(resdir, "null-benchmark-t5.rds")
tdist11_file <- file.path(resdir, "null-benchmark-t11.rds")
chisq_file <- file.path(resdir, "null-benchmark-chisq4.rds")

## number of cores for parallelization
cores <- 20
B <- 100

## define bechmarking design
bd <- initializeBenchDesign()
```

We include Scott's FDR Regression in the analysis. We include both `nulltype = "empirical"` and
`nulltype = "theoretical"`. In both cases, test statistics are modeled as being
sampled from mean-shifted normal distributions with equal variance under the null and alternative.
While `nulltype = "theoretical"` makes the additional assumption that the null is standard normal,
`nulltype = "empirical"` attempts to estimate the mean and variance under the null empirically.
While the assumptions of `nulltype = "theoretical"` are met for a subset of the simulations, we
include `nulltype = "empirical"` as this is not true for all simulations, and not necessarily
reflective of all real data sets.

```{r}
bdplus <- bd
bdplus <- addBMethod(bdplus, "fdrreg-t",
                     FDRreg::FDRreg,
                     function(x) { x$FDR },
                     z = test_statistic,
                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
                     nulltype = 'theoretical',
                     control = list(lambda = 0.01))
bdplus <- addBMethod(bdplus, "fdrreg-e",
                     FDRreg::FDRreg,
                     function(x) { x$FDR },
                     z = test_statistic,
                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
                     nulltype = 'empirical',
                     control = list(lambda = 0.01))
```

All simulation settings will share the following parameters.

```{r parameters-shared}
m <- 20000                        # integer: number of hypothesis tests
pi0 <- 1                          # numeric: proportion of null hypotheses
tstat <- function(x) { 0 }        # functional: dist of alternative test stats
icovariate <- runif               # functional: independent covariate
```

Simulation results will be presented excluding a subset of methods, and
for certain plots (upset plots), a single alpha cutoff will be used.

```{r}
excludeSet <- c("unadjusted", "bl-df02", "bl-df04", "bl-df05")
ualpha <- 0.05
```

# Gaussian Setting

First, we consider the null setting with gaussian noise.

## Data Simulation

```{r gauss-parameters}
tstat_dist <- rnorm_perturber(1)  # functional: sampling dist/noise for test stats
null_dist <- rnorm_2pvaluer(1)    # functional: dist to calc p-values
seed <- 608
```

We next run the simulations (including Scott's FDR Regression).

```{r gauss-run-simulation}
if (file.exists(gauss_file)) {
    res <- readRDS(gauss_file)
} else {
    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdplus, m = m,
                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
                    tstat_dist = tstat_dist, null_dist = null_dist,
                    seed = seed, mc.cores = cores)
    saveRDS(res, file = gauss_file)
}
res_i <- lapply(res, `[[`, "informative")
res_u <- lapply(res, `[[`, "uninformative")
```

## Covariate Diagnostics

Here, we show the relationship between the independent covariate and p-values for a
single replication of the experiment.

```{r gauss-one-simulation}
onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
```

```{r, gauss-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
```

```{r, gauss-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
```

## Benchmark Metrics

We plot the average results across `r B` simulations.

```{r gauss-metrics-averages, results = "hide"}
resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))

plotsim_average(resdf, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE) 
```

We also take a look at the distribution of rejects for each method as a function of
the effect size and independent covariate.

```{r gauss-metrics-covlineplot, results = "hide"}
covariateLinePlot(res_i, alpha = 0.05, covname = "effect_size", trans = "log1p")

covariateLinePlot(res_i, alpha = 0.05, covname = "ind_covariate", trans = "log1p")
```

Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
the overlap of rejections between methods.

```{r gauss-metrics-upset, results = "hide"}
if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
    aggupset(res_i, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
} else {
    message("Not enough methods found rejections at alpha ", ualpha, 
            "; skipping upset plot")
}
```

We also compare the simulation results with and without an informative covariate.

```{r gauss-metrics-differences, results = "hide"}
resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))

resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
                                   performanceMetric, alpha, value),
                            select(resdfu, rep, blabel, param.alpha, key,
                                   performanceMetric, alpha, value),
                            by = c("rep", "blabel", "param.alpha", "key",
                                   "performanceMetric", "alpha"),
                            suffix = c(".info", ".uninfo"))
resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)

plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)

plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)

plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)

plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
```

# Student's t Setting (df = 5)

Next, we consider the null setting with t-distributed noise.

## Data Simulation

```{r t5-parameters}
tstat_dist <- rt_perturber(5)  # functional: sampling dist/noise for test stats
null_dist <- rt_2pvaluer(5)    # functional: dist to calc p-values
seed <- 815
```

For the t-distributed noise setting, we must specify the number of degrees of freedom
for ASH. We add an additional parameter to the `ashq` method with the corresponding
degrees of freedom of the noise distribution.

```{r}
bdplust <- modifyBMethod(bdplus, "ashq", df = 5)
```

We next run the simulations (including Scott's FDR Regression and ASH with degrees of
freedom specified).

```{r t5-run-simulation}
if (file.exists(tdist_file)) {
    res <- readRDS(tdist_file)
} else {
    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdplust, m = m,
                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
                    tstat_dist = tstat_dist, null_dist = null_dist,
                    seed = seed, mc.cores = cores)
    saveRDS(res, file = tdist_file)
}
res_i <- lapply(res, `[[`, "informative")
res_u <- lapply(res, `[[`, "uninformative")
```

## Covariate Diagnostics

Here, we show the relationship between the independent covariate and p-values for a
single replication of the experiment.

```{r t5-one-simulation}
onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
```

```{r, t5-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
```

```{r, t5-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
```

## Benchmark Metrics

We plot the average results across `r B` simulations.

```{r t5-metrics-averages, results = "hide"}
resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))

plotsim_average(resdf, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE) 
```

We also take a look at the distribution of rejects for each method as a function of
the effect size and independent covariate.

```{r t5-metrics-covlineplot, results = "hide"}
covariateLinePlot(res_i, alpha = 0.05, covname = "effect_size", trans = "log1p")

covariateLinePlot(res_i, alpha = 0.05, covname = "ind_covariate", trans = "log1p")
```

Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
the overlap of rejections between methods.

```{r t5-metrics-upset, results = "hide"}
if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
    aggupset(res_i, alpha = ualpha, supplementary = FALSE, return_list = FALSE) 
} else {
    message("Not enough methods found rejections at alpha ", ualpha, 
            "; skipping upset plot")
}
```

We also compare the simulation results with and without an informative covariate.

```{r t5-metrics-differences, results = "hide"}
resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))

resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
                                   performanceMetric, alpha, value),
                            select(resdfu, rep, blabel, param.alpha, key,
                                   performanceMetric, alpha, value),
                            by = c("rep", "blabel", "param.alpha", "key",
                                   "performanceMetric", "alpha"),
                            suffix = c(".info", ".uninfo"))
resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)

plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)

plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)

plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)

plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
```

# Student's t Setting (df = 11)

Next, we consider the null setting with t-distributed noise.

## Data Simulation

```{r t11-parameters}
tstat_dist <- rt_perturber(11)  # functional: sampling dist/noise for test stats
null_dist <- rt_2pvaluer(11)    # functional: dist to calc p-values
seed <- 9158
```

For the t-distributed noise setting, we must specify the number of degrees of freedom
for ASH. We add an additional parameter to the `ashq` method with the corresponding
degrees of freedom of the noise distribution.

```{r}
bdplust <- modifyBMethod(bdplus, "ashq", df = 11)
```

We next run the simulations (including Scott's FDR Regression and ASH with degrees of
freedom specified).

```{r t11-run-simulation}
if (file.exists(tdist11_file)) {
    res <- readRDS(tdist11_file)
} else {
    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdplust, m = m,
                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
                    tstat_dist = tstat_dist, null_dist = null_dist,
                    seed = seed, mc.cores = cores)
    saveRDS(res, file = tdist11_file)
}
res_i <- lapply(res, `[[`, "informative")
res_u <- lapply(res, `[[`, "uninformative")
```

## Covariate Diagnostics

Here, we show the relationship between the independent covariate and p-values for a
single replication of the experiment.

```{r t11-one-simulation}
onerun <- simIteration(bdplus, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
```

```{r, t11-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
```

```{r, t11-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
```

## Benchmark Metrics

We plot the average results across `r B` simulations.

```{r t11-metrics-averages, results = "hide"}
resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))

plotsim_average(resdf, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE) 
```

We also take a look at the distribution of rejects for each method as a function of
the effect size and independent covariate.

```{r t11-metrics-covlineplot, results = "hide"}
covariateLinePlot(res_i, alpha = 0.05, covname = "effect_size", trans = "log1p")

covariateLinePlot(res_i, alpha = 0.05, covname = "ind_covariate", trans = "log1p")
```

Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
the overlap of rejections between methods.

```{r t11-metrics-upset, results = "hide"}
if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
    aggupset(res_i, alpha = ualpha, supplementary = FALSE, return_list = FALSE) 
} else {
    message("Not enough methods found rejections at alpha ", ualpha, 
            "; skipping upset plot")
}
```

We also compare the simulation results with and without an informative covariate.

```{r t11-metrics-differences, results = "hide"}
resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))

resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
                                   performanceMetric, alpha, value),
                            select(resdfu, rep, blabel, param.alpha, key,
                                   performanceMetric, alpha, value),
                            by = c("rep", "blabel", "param.alpha", "key",
                                   "performanceMetric", "alpha"),
                            suffix = c(".info", ".uninfo"))
resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)

plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)

plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)

plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)

plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
```

# Chi-Squared Setting

Finally, we consider the null setting with chi-squared distributed noise.

## Data Simulation

```{r chisq4-parameters}
tstat_dist <- rchisq_perturber(4)  # functional: sampling dist/noise for test stats
null_dist <- rchisq_pvaluer(4)     # functional: dist to calc p-values
seed <- 1023
```

For the chi-squared distributed noise setting, we must change the "mode" setting for ASH
from the default of 0 to `"estimate"` since the mode of null and alternative test statistics
are no longer centered at 0. While both approximate Normality and unimodality of effects are
violated in this simulation setting, by allowing the mode to be estimated, rather than forced
to 0, should return more comparable results.

```{r}
bdchi <- modifyBMethod(bd, "ashq", mode = "empirical")
```

We next run the simulations. We do not include FDR Regression because the test statistics
are not approximately normally distributed.

```{r chisq4-run-simulation}
if (file.exists(chisq_file)) {
    res <- readRDS(chisq_file)
} else {
    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdchi, m = m,
                    pi0 = pi0, tstat = tstat, icovariate = icovariate,
                    tstat_dist = tstat_dist, null_dist = null_dist,
                    seed = seed, mc.cores = cores)
    saveRDS(res, file = chisq_file)
}
res_i <- lapply(res, `[[`, "informative")
res_u <- lapply(res, `[[`, "uninformative")
```

## Covariate Diagnostics

Here, we show the relationship between the independent covariate and p-values for a
single replication of the experiment.

```{r chisq4-one-simulation}
onerun <- simIteration(bdchi, m = m, pi0 = pi0, tstat = tstat, tstat_dist = tstat_dist,
                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
```

```{r, chisq4-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
```

```{r, chisq4-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
```

## Benchmark Metrics

We plot the average results across `r B` simulations.

```{r chisq4-metrics-averages, results = "hide"}
resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))

plotsim_average(resdf, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE) 
```

Since `ashq` is not appropriate when the test statistics are not approximately normal or t,
we also plot the results without this method.

```{r chisq4-metrics-averages-noash, results = "hide"}
plotsim_average(resdf, met="rejections", filter_set = c(excludeSet, "ashq"),
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="FDR", filter_set = c(excludeSet, "ashq"),
                merge_ihw = TRUE, errorBars = TRUE) 
```

We also take a look at the distribution of rejects for each method as a function of
the effect size and independent covariate. Again, we plot these without `ashq`.

```{r chisq4-metrics-covlineplot, results = "hide"}
covariateLinePlot(lapply(res_i, function(x) { x[, -which(colnames(x) == "ashq")] }),
                  alpha = ualpha, covname = "effect_size")

covariateLinePlot(lapply(res_i, function(x) { x[, -which(colnames(x) == "ashq")] }),
                  alpha = ualpha, covname = "ind_covariate")
```

Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
the overlap of rejections between methods. Again, without `ashq`.

```{r chisq4-metrics-upset, results = "hide"}
if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = c(excludeSet, "ashq")) >= 3) {
    aggupset(lapply(res_i, function(x) { x[, -which(colnames(x) == "ashq")] }),
             alpha = ualpha, supplementary = FALSE, return_list = FALSE)
} else {
    message("Not enough methods found rejections at alpha ", ualpha, 
            "; skipping upset plot")
}
```

We also compare the simulation results with and without an informative covariate.

```{r chisq4-metrics-differences, results = "hide"}
resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))

resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
                                   performanceMetric, alpha, value),
                            select(resdfu, rep, blabel, param.alpha, key,
                                   performanceMetric, alpha, value),
                            by = c("rep", "blabel", "param.alpha", "key",
                                   "performanceMetric", "alpha"),
                            suffix = c(".info", ".uninfo"))
resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)

plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)

plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)

plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)

plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
```

# Session Info

```{r}
sessionInfo()
```