### 2018-08-09 datasets/simulations/simulations-summary.Rmd
Move simulation summary plots to dataset dir; not yet manuscript figures

---
title: "Simulation Summary Figures (simulations)"
author: "Rafalab"
output: 
    html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Figure2-load-data}
# Load packages and source benchmark FDR
library(tidyr)
library(dplyr)
library(ggplot2)
library(magrittr)
library(cowplot)
library(tibble)
library(ggthemes)
library(grid)
library(SummarizedBenchmark)

## load helper functions
for (f in list.files("../datasets/R", "\\.(r|R)$", full.names = TRUE)) {
    source(f)
}

# assumes sb objects are in the following location
path <- file.path("..", "datasets", "simulations", "results")

# set up results directory
outdir <- "./figures"
dir.create(outdir, showWarnings = FALSE)

methodset <- c("bonf", "bh", "ihw", "qvalue", "bl-df03", "lfdr",
               "fdrreg-e", "fdrreg-t", "ashq", "adapt-glm")
```

# Data Preparation

First, we load summaries for all simulations results at significance thresholds
between 0.01 and 0.10. This process can take a long time (approximately 40 minutes
per alpha cutoff) because of the large number of simulation settings that must be
read in and summarized. Ideally, this initial step should be run externally separate
from this Rmd (by default, the following code chunk will not be run).

```{r, eval = FALSE}
for (ia in c(1:10, 20, 30, 40, 50)) {
    summary_file <- file.path("results", paste0("summary_alpha", ia, ".rds"))
    if (!file.exists(summary_file)) {
        res_files <- list.files(path, "\\.rds$", full.names = TRUE)
        res <- lapply(res_files, function(x) {
            zz <- readRDS(x)
            zz_i <- lapply(zz, `[[`, "informative")
            zz_u <- lapply(zz, `[[`, "uninformative")
            zz_i <- plotsim_standardize(zz_i, alpha = ia / 100L)
            zz_i <- dplyr::select(zz_i, rep, blabel, param.alpha, key, performanceMetric, alpha, value)
            zz_u <- plotsim_standardize(zz_u, alpha = ia / 100L)
            zz_u <- dplyr::select(zz_u, rep, blabel, param.alpha, key, performanceMetric, alpha, value)
            dplyr::full_join(zz_i, zz_u,
                             by = c("rep", "blabel", "param.alpha", "key", "performanceMetric", "alpha"),
                             suffix = c(".info", ".uninfo"))
        })
        names(res) <- gsub("\\.rds$", "", basename(res_files))
        saveRDS(res, summary_file)
    }
}
```

Assuming that the summaries have been saved for all simulation settings at
the various FDR thresholds using the code above, we load the summarized results
and merge them into a single results table.

```{r}
alphas <- 1:10
res <- vector("list", length(alphas))
for (ia in seq_len(length(alphas))) {
    imethodset <- methodset
    imethodset[which(imethodset == "ihw")] <- paste0("ihw-a", sprintf("%02d", alphas[ia]))
    ires <- readRDS(file.path("results", paste0("summary_alpha", alphas[ia], ".rds")))
    ires <- bind_rows(ires, .id = "setting")
    ires <- dplyr::mutate(ires, dist = gsub(".*?-benchmark-(.*)", "\\1", setting),
                         setting = gsub("(.*?)-benchmark-.*", "\\1", setting))
    ires <- dplyr::filter(ires, blabel %in% imethodset)
    ires <- dplyr::mutate(ires, blabel = gsub("(-a)(.*)", "", blabel))
    res[[ia]] <- ires
}
res <- bind_rows(res)
res <- dplyr::select(res, -param.alpha)
res <- dplyr::rename(res, Method = blabel)
res <- dplyr::mutate(res, Method = gsub("-df03", "", Method))
res <- dplyr::mutate(res, Method = factor(Method, levels = gsub("-df03", "", methodset)))
```

We compute summary metrics for each method, setting, and cutoff.

```{r}
resmet <- dplyr::mutate(res, value.info = ifelse(is.na(value.info) & grepl("R$", key), 0, value.info),
                        value.uninfo = ifelse(is.na(value.uninfo) & grepl("R$", key), 0, value.uninfo))
resmet <- dplyr::group_by(resmet, setting, dist, Method, key, performanceMetric, alpha)
resmet <- dplyr::summarize(resmet,
                           mean.info = mean(value.info, na.rm = TRUE),
                           se.info = sd(value.info, na.rm = TRUE) / sqrt(sum(!is.na(value.info))),
                           nNA.info = sum(is.na(value.info)),
                           mean.diff = mean(value.info - value.uninfo, na.rm = TRUE),
                           se.diff = sd(value.info - value.uninfo, na.rm = TRUE) /
                               sqrt(sum(!is.na(value.info - value.uninfo))))
resmet <- dplyr::ungroup(resmet)
```

We clean up the annotations in the table for some of the more complex simulation settings.

```{r}
resmet <- dplyr::mutate(resmet, inform = ifelse(grepl("varyinginfo", setting), dist, NA))
resmet <- dplyr::mutate(resmet, inform = as.numeric(gsub("level", "", inform)))
resmet <- dplyr::mutate(resmet, dist = ifelse(grepl("varyinginfo", setting), "gaussian", dist))

resmet <- dplyr::mutate(resmet, pi0 = ifelse(grepl("varyingpi0", setting), dist, 90))
resmet <- dplyr::mutate(resmet, pi0 = as.numeric(gsub("nullprop", "", pi0)))
resmet <- dplyr::mutate(resmet, dist = ifelse(grepl("varyingpi0", setting), "gaussian", dist))

resmet <- dplyr::mutate(resmet, signal = ifelse(grepl("varyinges", setting), dist, NA))
resmet <- dplyr::mutate(resmet, dist = ifelse(grepl("^varyinges$", setting), "gaussian", dist))
resmet <- dplyr::mutate(resmet, dist = ifelse(grepl("^varyinges-t$", setting), "t11", dist))
resmet <- dplyr::mutate(resmet, dist = ifelse(grepl("^varyinges-chisq$", setting), "chisq4", dist))
resmet <- dplyr::mutate(resmet, setting = ifelse(grepl("varyinges", setting), "varyinges", setting))
```

We also replace the sampling noise labels with more plotting-friendly labels.

```{r}
resmet <- dplyr::mutate(resmet, dist = factor(dist, levels = c("gaussian", "t11", "t5", "chisq4"),
                                              labels = c(expression("Noise: " * N(0, 1)),
                                                         expression("Noise: " * t[11]),
                                                         expression("Noise: " * t[5]),
                                                         expression("Noise: " * {chi^2}[4]))))
```

We use the standardize candy color scheme and line types for the plots.

```{r}
col <- as.character(candycols$col)
names(col) <- as.character(candycols$Method)
lty <- as.character(candycols$lty)
names(lty) <- as.character(candycols$Method)
```

There are several different simulation settings that can be investigated.

```{r}
dplyr::distinct(resmet, setting)
```

Namely, simulation settings were run with to investigate the impact of different
informative covariates, noninformative covariates, varying null proportions,
varying covariate informativeness, and varying effect size distributions (including
unimodal effects as described in the ASH manuscript). Many of these simulation
settings were also run with different sampling noise distributions, including
Normal/Gaussian, t with 11 degrees of freedom, t with 5 degrees of freedom, and
Chi-squared with 4 degrees of freedom.

Results are reproted for the Chi-Squared simulation settings without `ashq`,
`fdrreg-e` and `fdrreg-t`, since the assumptions of these methods are clearly
violated under these settings, and it is neither appropriate nor fair to include
them in the comparisons.

```{r}
resmet <- dplyr::filter(resmet, !( grepl("chi", dist) & Method == "ashq" ))
resmet <- dplyr::filter(resmet, !( grepl("chi", dist) & grepl("fdrreg", Method) ))
```

# Null Simulations

We subset on simulation results that look at informative and non-informative covariates across
different noise distributions.

For null simulations, rather than the false discovery rate, we take a look at the true negative
rate (equivalently, the proportion of false positives) since the FDR will be 1 if any rejections
are made. Additionally, we take a look at the FWER (the proportion of simulation replicates out
of 100 with at least one rejection).

```{r, fig.width = 18, fig.height = 8}
res_nullset <- dplyr::filter(resmet, setting == "null")

null_tnr <- dplyr::filter(res_nullset, key == "TNR") %>%
    ggplot(aes(x = alpha, y = mean.info, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 0.0025, alpha=0.5) +
    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("TNR", labels = scales::percent) +
    ggtitle("Mean TNR Over 100 Replications") +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

null_fwer <- dplyr::filter(res_nullset, key == "FWER") %>%
    ggplot(aes(x = alpha, y = mean.info, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("FWER", labels = scales::percent) +
    ggtitle("FWER Over 100 Replications") +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

gp <- plot_grid(null_tnr + guides(color = FALSE, linetype = FALSE),
                null_fwer + guides(color = FALSE, linetype = FALSE),
                labels = LETTERS[1:2], ncol = 1)
gp <- plot_grid(gp, get_legend(null_tnr), rel_widths = c(1, 1/16))
gp
```

We also look at the difference between informative and uninformative covariates.

```{r, fig.width = 18, fig.height = 8}
null_tnr_diff <- dplyr::filter(res_nullset, key == "TNR") %>%
    ggplot(aes(x = alpha, y = mean.diff, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 0.0025, alpha=0.5) +
    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("TNR\n(informative - uninformative)", labels = scales::percent) +
    ggtitle("Mean TNR Difference Over 100 Replications") +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

null_fwer_diff <- dplyr::filter(res_nullset, key == "FWER") %>%
    ggplot(aes(x = alpha, y = mean.diff, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 0.0025, alpha=0.5) +
    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("FWER\n(informative - uninformative)", labels = scales::percent) +
    ggtitle("FWER Difference Over 100 Replications") +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

gp <- plot_grid(null_tnr_diff + guides(color = FALSE, linetype = FALSE),
                null_fwer_diff + guides(color = FALSE, linetype = FALSE),
                labels = LETTERS[1:2], ncol = 1)
gp <- plot_grid(gp, get_legend(null_tnr_diff), rel_widths = c(1, 1/16))
gp
```

# Sampling Distributions

We just take a look at different noise distributions for a single informative covariate.
We take a look at the false discovery proportion controlled along a grid of nominal alpha cutoffs
between 0.01 and 0.10 for each method, as well as the true positive rate for each method along
the set of cutoffs.

```{r, fig.width = 18, fig.height = 8}
res_noise <- dplyr::filter(resmet, setting == "informative-cosine")

noise_fdr <- dplyr::filter(res_noise, key == "FDR") %>%
    ggplot(aes(x = alpha, y = mean.info, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 0.0025, alpha=0.5) +
    geom_abline(intercept = 0, slope = 1, lty = 2, color = "blue", alpha = 1/2) + 
    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("FDR", labels = scales::percent) +
    ggtitle("Mean FDR Over 100 Replications") +
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

noise_tpr <- dplyr::filter(res_noise, key == "TPR") %>%
    ggplot(aes(x = alpha, y = mean.info, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 0.0025, alpha=0.5) +
    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("TPR", labels = scales::percent) +
    ggtitle("Mean TPR Over 100 Replications") +
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

gp <- plot_grid(noise_fdr + guides(color = FALSE, linetype = FALSE),
                noise_tpr + guides(color = FALSE, linetype = FALSE),
                labels = LETTERS[1:2], ncol = 1)
gp <- plot_grid(gp, get_legend(noise_tpr), rel_widths = c(1, 1/16))
gp

gp <- plot_grid(noise_fdr + guides(color = FALSE, linetype = FALSE) + coord_cartesian(ylim = c(0, .2)),
                noise_tpr + guides(color = FALSE, linetype = FALSE),
                labels = LETTERS[1:2], ncol = 1)
gp <- plot_grid(gp, get_legend(noise_tpr), rel_widths = c(1, 1/16))
gp
```

We also look at the difference between informative and uninformative covariates.

```{r, fig.width = 18, fig.height = 8}
noise_fdr_diff <- dplyr::filter(res_noise, key == "FDR") %>%
    ggplot(aes(x = alpha, y = mean.diff, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 0.0025, alpha=0.5) +
    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("FDR\n(informative - uninformative)", labels = scales::percent) +
    ggtitle("Mean FDR Difference Over 100 Replications") +
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

noise_tpr_diff <- dplyr::filter(res_noise, key == "TPR") %>%
    ggplot(aes(x = alpha, y = mean.diff, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 0.0025, alpha=0.5) +
    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
    facet_grid(. ~ dist, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("TPR\n(informative - uninformative)", labels = scales::percent) +
    ggtitle("Mean TPR Difference Over 100 Replications") +
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

gp <- plot_grid(noise_fdr_diff + guides(color = FALSE, linetype = FALSE),
                noise_tpr_diff + guides(color = FALSE, linetype = FALSE),
                labels = LETTERS[1:2], ncol = 1)
gp <- plot_grid(gp, get_legend(noise_tpr_diff), rel_widths = c(1, 1/16))
gp
```

# Informative Covariates

We also take a look across different informative covariates under just the normal
sampling noise setting. As before, we take a look at the false discovery proportion
controlled along a grid of nominal alpha cutoffs between 0.01 and 0.10 for each method,
as well as the true positive rate for each method along the set of cutoffs.

```{r, fig.width = 18, fig.height = 8}
res_icov <- dplyr::filter(resmet, grepl("^informative", setting), grepl("N\\(0, 1\\)", dist))

icov_fdr <- dplyr::filter(res_icov, key == "FDR") %>%
    ggplot(aes(x = alpha, y = mean.info, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 0.0025, alpha=0.5) +
    geom_abline(intercept = 0, slope = 1, lty = 2, color = "blue", alpha = 1/2) + 
    facet_grid(. ~ setting, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("FDR", labels = scales::percent) +
    ggtitle("Mean FDR Over 100 Replications") +
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

icov_tpr <- dplyr::filter(res_icov, key == "TPR") %>%
    ggplot(aes(x = alpha, y = mean.info, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 0.0025, alpha=0.5) +
    facet_grid(. ~ setting, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("TPR", labels = scales::percent) +
    ggtitle("Mean TPR Over 100 Replications") +
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

gp <- plot_grid(icov_fdr + guides(color = FALSE, linetype = FALSE),
                icov_tpr + guides(color = FALSE, linetype = FALSE),
                labels = LETTERS[1:2], ncol = 1)
gp <- plot_grid(gp, get_legend(icov_tpr), rel_widths = c(1, 1/12))
gp
```

We also look at the difference between informative and uninformative covariates.

```{r, fig.width = 18, fig.height = 8}
icov_fdr_diff <- dplyr::filter(res_icov, key == "FDR") %>%
    ggplot(aes(x = alpha, y = mean.diff, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 0.0025, alpha=0.5) +
    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
    facet_grid(. ~ setting, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("FDR\n(informative - uninformative)", labels = scales::percent) +
    ggtitle("Mean FDR Difference Over 100 Replications") +
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

icov_tpr_diff <- dplyr::filter(res_icov, key == "TPR") %>%
    ggplot(aes(x = alpha, y = mean.diff, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 0.0025, alpha=0.5) +
    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
    facet_grid(. ~ setting, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("TPR\n(informative - uninformative)", labels = scales::percent) +
    ggtitle("Mean TPR Difference Over 100 Replications") +
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

gp <- plot_grid(icov_fdr_diff + guides(color = FALSE, linetype = FALSE),
                icov_tpr_diff + guides(color = FALSE, linetype = FALSE),
                labels = LETTERS[1:2], ncol = 1)
gp <- plot_grid(gp, get_legend(icov_tpr_diff), rel_widths = c(1, 1/12))
gp
```

# Null Proportions

We take a look at the impact of the proportion of null hypotheses on the performance
of the different methods under just the normal sampling noise setting. We plot the
false discovery proportion and true positive rate at a nominal alpha cutoff of 0.05
for each method. The sine informative covariate was used in these simulation settings.

```{r, fig.width = 10, fig.height = 4}
res_pi0 <- dplyr::filter(resmet, setting == "varyingpi0-lowsignal", alpha == 0.05)

pi0_fdr <- dplyr::filter(res_pi0, key == "FDR") %>%
    ggplot(aes(x = pi0, y = mean.info, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 2.5, alpha=0.5) +
    geom_hline(aes(yintercept = alpha), lty = 2, color = "blue", alpha = 1/2) + 
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("pi0", breaks = seq(0, 100, 10)) +
    scale_y_continuous("FDR", labels = scales::percent) +
    ggtitle("Mean FDR at alpha = 0.05 Over 100 Replications") +
    expand_limits(x =0, y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

pi0_tpr <- dplyr::filter(res_pi0, key == "TPR") %>%
    ggplot(aes(x = pi0, y = mean.info, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 2.5, alpha=0.5) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("pi0", breaks = seq(0, 100, 10)) +
    scale_y_continuous("TPR", labels = scales::percent) +
    ggtitle("Mean TPR at alpha = 0.05 Over 100 Replications") +
    expand_limits(x =0, y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

gp <- plot_grid(pi0_fdr + guides(color = FALSE, linetype = FALSE),
                pi0_tpr + guides(color = FALSE, linetype = FALSE),
                labels = LETTERS[1:2], ncol = 2)
gp <- plot_grid(gp, get_legend(pi0_tpr), rel_widths = c(1, 1/8))
gp
```

# Informativeness

We take a look at the impact of the informativeness of a covariate on the performance
of the different methods under just the normal sampling noise setting. Here, the informative
covariate is sampled uniformly between 0 and 1 and the probability of the corresponding test
being null is one value if the covariate is less than 0.8, and another value if the covariate
is above 0.8. The "informativeness" is measured by the difference between these two probabilities,
with informativeness 0 and 1 corresponding to no difference and maximal difference (i.e. 0 and 1),
respectively. We plot the false discovery proportion and true positive rate at a nominal alpha
cutoff of 0.05 for each method.

```{r, fig.width = 10, fig.height = 4}
res_info <- dplyr::filter(resmet, setting == "varyinginfo-lowsignal", alpha == 0.05)

info_fdr <- dplyr::filter(res_info, key == "FDR") %>%
    ggplot(aes(x = inform, y = mean.info, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 2.5, alpha=0.5) +
    geom_hline(aes(yintercept = alpha), lty = 2, color = "blue", alpha = 1/2) + 
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("informativeness", breaks = seq(0, 100, 10)) +
    scale_y_continuous("FDR", labels = scales::percent) +
    ggtitle("Mean FDR at alpha = 0.05 Over 100 Replications") +
    expand_limits(x =0, y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

info_tpr <- dplyr::filter(res_info, key == "TPR") %>%
    ggplot(aes(x = inform, y = mean.info, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 2.5, alpha=0.5) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("informativeness", breaks = seq(0, 100, 10)) +
    scale_y_continuous("TPR", labels = scales::percent) +
    ggtitle("Mean TPR at alpha = 0.05 Over 100 Replications") +
    expand_limits(x =0, y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

gp <- plot_grid(info_fdr + guides(color = FALSE, linetype = FALSE),
                info_tpr + guides(color = FALSE, linetype = FALSE),
                labels = LETTERS[1:2], ncol = 2)
gp <- plot_grid(gp, get_legend(info_tpr), rel_widths = c(1, 1/8))
gp
```

We also look at the difference between informative and uninformative covariates.

```{r, fig.width = 10, fig.height = 4}
info_fdr_diff <- dplyr::filter(res_info, key == "FDR") %>%
    ggplot(aes(x = inform, y = mean.diff, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 2.5, alpha=0.5) +
    geom_hline(aes(yintercept = 0), lty = 2, color = "blue", alpha = 1/2) + 
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("informativeness", breaks = seq(0, 100, 10)) +
    scale_y_continuous("FDR", labels = scales::percent) +
    ggtitle("Mean FDR Difference at alpha = 0.05 Over 100 Replications") +
    expand_limits(x =0, y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

info_tpr_diff <- dplyr::filter(res_info, key == "TPR") %>%
    ggplot(aes(x = inform, y = mean.diff, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 2.5, alpha=0.5) +
    geom_hline(aes(yintercept = 0), lty = 2, color = "blue", alpha = 1/2) + 
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("informativeness", breaks = seq(0, 100, 10)) +
    scale_y_continuous("TPR", labels = scales::percent) +
    ggtitle("Mean TPR Difference at alpha = 0.05 Over 100 Replications") +
    expand_limits(x =0, y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

gp <- plot_grid(info_fdr_diff + guides(color = FALSE, linetype = FALSE),
                info_tpr_diff + guides(color = FALSE, linetype = FALSE),
                labels = LETTERS[1:2], ncol = 2)
gp <- plot_grid(gp, get_legend(info_tpr_diff), rel_widths = c(1, 1/8))
gp
```

# Unimodal Assumption

We also take a look at the results for simulation settings where the effect sizes follow a unimodal
distribution (null and alternative, combined) as proposed in the ASH paper. We implement some of the
unimodal effect size distributions described in the paper, "flattop", "skew", and "spiky", as well
as a single bimodal setting also described in the paper, "bimodal". As before, we plot the the
false discovery proportion and true positive rate for each method along a grid of nominal FDR cutoffs
between 0.01 and 0.10. The cubic informative covariate was used in these simulation settings.

## Normal Noise

First, we look at the unimodal setting under Normal sampling noise.

```{r, fig.width = 18, fig.height = 8}
res_ua <- dplyr::filter(resmet, setting == "varyinges", grepl("N\\(0, 1\\)", dist))

noise_fdr <- dplyr::filter(res_ua, key == "FDR") %>%
    ggplot(aes(x = alpha, y = mean.info, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 0.0025, alpha=0.5) +
    geom_abline(intercept = 0, slope = 1, lty = 2, color = "blue", alpha = 1/2) + 
    facet_grid(. ~ signal, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("FDR", labels = scales::percent) +
    ggtitle("Mean FDR Over 100 Replications") +
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

noise_tpr <- dplyr::filter(res_ua, key == "TPR") %>%
    ggplot(aes(x = alpha, y = mean.info, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 0.0025, alpha=0.5) +
    facet_grid(. ~ signal, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("TPR", labels = scales::percent) +
    ggtitle("Mean TPR Over 100 Replications") +
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

gp <- plot_grid(noise_fdr + guides(color = FALSE, linetype = FALSE),
                noise_tpr + guides(color = FALSE, linetype = FALSE),
                labels = LETTERS[1:2], ncol = 1)
gp <- plot_grid(gp, get_legend(noise_tpr), rel_widths = c(1, 1/16))
gp
```

We also look at the difference between informative and uninformative covariates.

```{r, fig.width = 18, fig.height = 8}
noise_fdr_diff <- dplyr::filter(res_ua, key == "FDR") %>%
    ggplot(aes(x = alpha, y = mean.diff, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 0.0025, alpha=0.5) +
    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
    facet_grid(. ~ signal, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("FDR\n(informative - uninformative)", labels = scales::percent) +
    ggtitle("Mean FDR Difference Over 100 Replications") +
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

noise_tpr_diff <- dplyr::filter(res_ua, key == "TPR") %>%
    ggplot(aes(x = alpha, y = mean.diff, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 0.0025, alpha=0.5) +
    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
    facet_grid(. ~ signal, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("TPR\n(informative - uninformative)", labels = scales::percent) +
    ggtitle("Mean TPR Difference Over 100 Replications") +
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

gp <- plot_grid(noise_fdr_diff + guides(color = FALSE, linetype = FALSE),
                noise_tpr_diff + guides(color = FALSE, linetype = FALSE),
                labels = LETTERS[1:2], ncol = 1)
gp <- plot_grid(gp, get_legend(noise_tpr_diff), rel_widths = c(1, 1/16))
gp
```

## t Noise

Next, we look at the unimodal settings under t (11 degrees of freedom) distributed
sampling noise.

```{r, fig.width = 18, fig.height = 8}
res_ua <- dplyr::filter(resmet, setting == "varyinges", grepl("t\\[11\\]", dist))

noise_fdr <- dplyr::filter(res_ua, key == "FDR") %>%
    ggplot(aes(x = alpha, y = mean.info, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 0.0025, alpha=0.5) +
    geom_abline(intercept = 0, slope = 1, lty = 2, color = "blue", alpha = 1/2) + 
    facet_grid(. ~ signal, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("FDR", labels = scales::percent) +
    ggtitle("Mean FDR Over 100 Replications") +
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

noise_tpr <- dplyr::filter(res_ua, key == "TPR") %>%
    ggplot(aes(x = alpha, y = mean.info, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.info - se.info, ymax = mean.info + se.info), width = 0.0025, alpha=0.5) +
    facet_grid(. ~ signal, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("TPR", labels = scales::percent) +
    ggtitle("Mean TPR Over 100 Replications") +
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

gp <- plot_grid(noise_fdr + guides(color = FALSE, linetype = FALSE),
                noise_tpr + guides(color = FALSE, linetype = FALSE),
                labels = LETTERS[1:2], ncol = 1)
gp <- plot_grid(gp, get_legend(noise_tpr), rel_widths = c(1, 1/16))
gp
```

We also look at the difference between informative and uninformative covariates.

```{r, fig.width = 18, fig.height = 8}
noise_fdr_diff <- dplyr::filter(res_ua, key == "FDR") %>%
    ggplot(aes(x = alpha, y = mean.diff, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 0.0025, alpha=0.5) +
    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
    facet_grid(. ~ signal, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("FDR\n(informative - uninformative)", labels = scales::percent) +
    ggtitle("Mean FDR Difference Over 100 Replications") +
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

noise_tpr_diff <- dplyr::filter(res_ua, key == "TPR") %>%
    ggplot(aes(x = alpha, y = mean.diff, color = Method)) +
    geom_line(aes(linetype = Method), alpha = 0.85) +
    geom_errorbar(aes(ymin = mean.diff - se.diff, ymax = mean.diff + se.diff), width = 0.0025, alpha=0.5) +
    geom_hline(yintercept = 0, lty = 2, color = "blue", alpha = 1/2) + 
    facet_grid(. ~ signal, scales = "free_x", labeller = label_parsed) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_continuous("alpha", breaks = seq(0, 1, .01)) +
    scale_y_continuous("TPR\n(informative - uninformative)", labels = scales::percent) +
    ggtitle("Mean TPR Difference Over 100 Replications") +
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14))

gp <- plot_grid(noise_fdr_diff + guides(color = FALSE, linetype = FALSE),
                noise_tpr_diff + guides(color = FALSE, linetype = FALSE),
                labels = LETTERS[1:2], ncol = 1)
gp <- plot_grid(gp, get_legend(noise_tpr_diff), rel_widths = c(1, 1/16))
gp
```