### 2018-08-28 datasets/GSEA/fgsea-mouse.Rmd

---
title: "Case Study: Gene Set Enrichment Analysis (Mouse Data Set)"
author: "Alejandro Reyes and Keegan Korthauer"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

The objective of this vignette is to test different multiple testing methods in the context of Gene Set Enrichment Analysis (GSEA). To do this, we will use mouse data from the paper by [Cabezas-Wallscheid et al. (Cell stem Cell, 2014)](https://www.sciencedirect.com/science/article/pii/S1934590914003014?via%3Dihub). The dataset consists of RNA-seq data from mouse hematopoietic stem cells and 4 multipotent progenitor lineages. We will test for difference in gene expression between two different cell fractions and perform GSEA in the resulting list of genes.

Here we use the `fgsea` Bioconductor package to implement the GSEA method. This is 
a Functional Class Scoring approach, which does not require setting an arbitrary threshold
for Differential Expression, but instead relies on the gene's rank (here we rank by
DESeq2 test statistic).

# Workspace Setup

```{r}
library(dplyr)
library(ggplot2)
library(scales)
library(DESeq2)
library(EnsDb.Mmusculus.v75)
library(fgsea)

## load helper functions
for (f in list.files("../R", "\\.(r|R)$", full.names = TRUE)) {
    source(f)
}

## project data/results folders
datdir <- "data"
resdir <- "results"
sbdir <- "../../results/GSEA"
dir.create(datdir, showWarnings = FALSE)
dir.create(resdir, showWarnings = FALSE)
dir.create(sbdir, showWarnings = FALSE)

## intermediary files we create below
count_file <- file.path(datdir, "mouse-counts.rds")
deseq_file <- file.path(datdir, "mouse-deseq.rds")
goset_file <- file.path(datdir, "mouse-gosets.rds")
result_file <- file.path(resdir, "mouse-results.rds")
bench_file <- file.path(sbdir, "mouse-benchmark.rds")
bench_file_uninf <- file.path(sbdir, "mouse-uninf-benchmark.rds")
```

# Data Preparation

The data has been preprocessed and saved as a DESeqDataset object. The following lines of code download this *DESeqDataSet* if it is not present locally.

```{r, message=FALSE}
if (!file.exists(count_file)) {
    download.file("https://www.dropbox.com/s/jq6giqeclxcdypq/gsea-mouse-counts.rds?dl=1",
                  destfile = count_file)
}
dseHSCMPP <- readRDS(count_file)
```

# Data Analysis

## Enrichment Analysis

In order to rank the list of genes for GSEA, we will test each gene for 
differential gene expression between hematopoietic stem cells and multipotent 
progenitors (fraction 1). To do this, we will run DESeq2 to obtain a statistic 
for differential expression.

```{r deseq2, message=FALSE}
if (!file.exists(deseq_file)) {
    dseHSCMPP <- DESeq(dseHSCMPP)
    res <- results(dseHSCMPP, contrast = c("conditions", "HSC", "MPP1"),
                   independentFiltering = FALSE)
    saveRDS(res, file = deseq_file)
} else {
    res <- readRDS(deseq_file)
}

genes <- as.numeric(res$padj < 0.1)
names(genes) <- rownames(res)
sum(genes, na.rm=TRUE)
```

We next use biomaRt to get the relations between GO categories and genes.

```{r gocat, message=FALSE}
if (!file.exists(goset_file)) {
    mart <- useMart("ensembl", "mmusculus_gene_ensembl")
    goSets <- getBM(c("ensembl_gene_id", "go_id"), mart = mart,
                    filters = "ensembl_gene_id", values = rownames(res))
    goSets <- goSets[!nchar(goSets$go_id) == 0, ]
    goSets <- with(goSets, split(go_id, ensembl_gene_id))
    saveRDS(goSets, file = goset_file)
} else {
   goSets <- readRDS(goset_file)
}
```

Now we use the `fgsea` package to perform the gene set enrichment analysis and 
obtain a enrichment p-value for each pathway.

```{r}
# invert the list so each item is a pathway instead of a gene
goSets <- split(rep(names(goSets), lengths(goSets)), unlist(goSets))
stats <- res$stat
names(stats) <- rownames(res)
stats <- stats[!is.na(stats)]

if (!file.exists(result_file)) {
  goRes <- fgsea(goSets, 
                  stats, 
                  nperm=10000, 
                  maxSize=500,
                  minSize=5)  
  saveRDS(goRes, file = result_file)
} else {
  goRes <- readRDS(result_file)
}
```

Add a random (uninformative covariate) to the dataset.

```{r}
## Add random (uninformative) covariate
set.seed(7476)
goRes$rand_covar <- rnorm(nrow(goRes))

```

## Covariate Diagnostics

Here, we want to check whether the size of the gene set is actually informative and 
independent under the null.

### Gene Set Size

In the following plot, we explore the relationship between the p-value and the gene set size. We can see that this covariate is actually informative.

```{r informativeCovariate, message=FALSE, fig.height=3.5, message=FALSE, fig.width=4.5}
rank_scatter(dat = goRes, pval = "pval", 
             covariate = "size", bins = 50, funx = log2, 
             funfill = log10_trans()) +
    ggtitle("Enriched gene sets") +
    xlab(expression(log[10]~"(# of genes)")) +
    ylab(expression(-log[10]~"(p-value)")) 
```

And at the same time, it seems to be fair to assume that it is independent under the null hypothesis.

```{r assumptions, message=FALSE, fig.width=10, fig.height=3.2}
strat_hist(goRes, pval = "pval", covariate = "size", maxy=12)
```

### Random

We will explore whether the random covariate can be used as a covariate for modern multiple-testing correction methods in the context of GSEA. In the plot below, the log10 of the p-values is plotted as a function of the 
random covariate. This covariate looks independent of the p-values.

```{r informativeCovariate-rand, message=FALSE, fig.height=3.5, message=FALSE, fig.width=4.5}
rank_scatter(dat = goRes, pval = "pval", 
              covariate = "rand_covar", bins = 50, 
              funfill = log10_trans()) +
    ggtitle("Enriched gene sets") +
    ylab(expression(-log[10]~"(p-value)") ) 
```

We can also explore if the covariate seems to be independent under the null.

```{r assumptions-rand, message=FALSE, fig.width=10, fig.height=3.2}
strat_hist(goRes, pval="pval", covariate="rand_covar", maxy=10)
```

## Multiple-Testing Correction

Generating the *SummarizedBenchmark* object:

```{r setup-benckmark, message=FALSE}
## rename columns and prepare for benchmarking
res <- dplyr::select(goRes, pval, size, rand_covar) %>%
    dplyr::rename(pval = pval,
                  ind_covariate = size)

## generate default BenchDesign
bd <- initializeBenchDesign()
```

We don't include `ashq`, `fdrreg-e` and `fdrreg-t` from the analysis because
the necessary assumptions are not met in the current case study. Namely, effect sizes
and standard errors are not available for ASH, and the test statistics are not normally
distributed under the null and alternative, as required by Scott's FDR regression methods.

```{r run-benchmark}
if (!file.exists(bench_file)) {
    sGSEAMouse <- buildBench(bd, res, ftCols = "ind_covariate")
    saveRDS(sGSEAMouse, file = bench_file)
} else {
    sGSEAMouse <- readRDS(bench_file)
}
```

We'll also compare the results to an uninformative (random) covariate.

```{r run-benchmark-rand}
if (!file.exists(bench_file_uninf)) {
  res$ind_covariate <- res$rand_covar
  sGSEAMouse_rand <- buildBench(bd, res, ftCols = "ind_covariate")
  saveRDS(sGSEAMouse_rand, file = bench_file_uninf)
} else {
  sGSEAMouse_rand <- readRDS(bench_file_uninf)
}
```

## Benchmark Metrics

### Gene Set Size

```{r}
assayNames(sGSEAMouse) <- "qvalue"
sGSEAMouse <- addDefaultMetrics(sGSEAMouse)
```

```{r plottingResults}
rejections_scatter(sGSEAMouse, as_fraction = FALSE, supplementary = FALSE)
```

```{r}
plotFDRMethodsOverlap(sGSEAMouse, alpha = 0.1, supplementary = FALSE,
                      order.by = "freq", nsets = 100)
```

```{r}
covariateLinePlot(sGSEAMouse, alpha = 0.1, covname = "ind_covariate", trans="log1p")
```

### Random

```{r}
assayNames(sGSEAMouse_rand) <- "qvalue"
sGSEAMouse_rand <- addDefaultMetrics(sGSEAMouse_rand)
sGSEAMouse_rand <- estimatePerformanceMetrics(sGSEAMouse_rand, addColData=TRUE)
```

```{r plottingResults-rand}
rejections_scatter(sGSEAMouse_rand, as_fraction=FALSE, supplementary=FALSE)
```

```{r}
plotFDRMethodsOverlap(sGSEAMouse_rand, alpha=0.1, supplementary=FALSE, order.by="freq", nsets=100)
```

```{r}
covariateLinePlot(sGSEAMouse_rand, alpha = 0.1, covname = "ind_covariate", trans = "log1p")
```

## Covariate comparison

Here we compare the method ranks for the two covariates at alpha = 0.10.

```{r}
plotMethodRanks(c(bench_file, bench_file_uninf), 
                colLabels = c("Set Size", "Random"), 
                alpha = 0.10, xlab = "Covariate", 
                excludeMethods = NULL)
```

# Session Info

```{r}
sessionInfo()
```