### 2018-08-29 datasets/ChIP-seq/ChIPseq-CBP-csaw.Rmd
Update ChiPseq csaw analyses
Add mean coverage (across all samples) as an informative covariate.
In addition, update call to csaw::normOffsets to reflect changes in
current release version that change the default return object to a
SE.

---
title: "Case Study: ChIP-seq Differential Peak Calling (csaw)"
author: "Patrick Kimes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
    html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

In this case study, we perform differential peak calling on ChIP-seq data for a
trnscription factor, CREB-binding protein (CBP), from Kasper et al. (2014), used in
Lun et al. (2015) to demonstrate the usage of _csaw_. As described in a separate case
study analyzing ChIP-seq data for H3K4Me3, _csaw_ is a package for differential peak
calling based on counting reads in sliding windows across the genome. Code and steps
for initial processing and analysis of this data with _csaw_ are taken directly
from Lun et al. (2015).

# Workspace Setup

```{r, wkspace-setup, results='hide', message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(SummarizedBenchmark)
library(BiocParallel)
library(rtracklayer)
library(Rsamtools)
library(Rsubread)
library(csaw)
library(edgeR)
library(GenomicAlignments)

## load helper functions
for (f in list.files("../R", "\\.(r|R)$", full.names = TRUE)) {
    source(f)
}

## project data/results folders
datdir <- "data"
resdir <- "results"
sbdir <- "../../results/ChIPseq"
dir.create(datdir, showWarnings = FALSE)
dir.create(resdir, showWarnings = FALSE)
dir.create(sbdir, showWarnings = FALSE)

## intermediary files we create below
count_file <- file.path(resdir, "cbp-csaw-counts.rds")
filtered_file <- file.path(resdir, "cbp-csaw-counts-filtered.rds")
normfacs_file <- file.path(resdir, "cbp-csaw-counts-normfacs.rds")
result_file <- file.path(resdir, "cbp-csaw-results.rds")
bench_file <- file.path(sbdir, "cbp-csaw-benchmark.rds")
bench_file_cov <- file.path(sbdir, "cbp-csaw-cov-benchmark.rds")
bench_file_uninf <- file.path(sbdir, "cbp-csaw-uninf-benchmark.rds")

## set up parallel backend
cores <- as.numeric(Sys.getenv("SLURM_NTASKS"))
multicoreParam <- MulticoreParam(workers = cores)
```

# Data Preparation

We download the fastq files directly from the European Nucleotide Archive (ENA).

```{r download-fastqs}
fqurls <- c("007/SRR1145787/SRR1145787.fastq.gz",
            "008/SRR1145788/SRR1145788.fastq.gz",
            "009/SRR1145789/SRR1145789.fastq.gz",
            "000/SRR1145790/SRR1145790.fastq.gz")
fqurls <- paste0("ftp://ftp.sra.ebi.ac.uk/vol1/fastq/SRR114/", fqurls)

for (i_fq in fqurls) {
    out_fq <- file.path(datdir, basename(i_fq))
    if (!file.exists(out_fq)) {
        download.file(i_fq, destfile = out_fq)
    }
}
```

We download the mm10 reference genome used for ENCODE3 and build the
index for alignment if not already available.

```{r, download-ref}
ref_url <- paste0("https://www.encodeproject.org/files/",
                  "mm10_no_alt_analysis_set_ENCODE/@@download/",
                  "mm10_no_alt_analysis_set_ENCODE.fasta.gz")
ref_fastagz <- file.path(datdir, basename(ref_url))
ref_fasta <- gsub("\\.gz$", "", ref_fastagz)

if (!file.exists(file.path(datdir, "mm_ref_index.00.b.tab"))) {
    if (!file.exists(ref_fasta)) {
        download.file(ref_url, destfile = ref_fastagz)
        system(paste("gunzip", ref_fastagz))
    }
    buildindex(basename = file.path(datdir, "mm_ref_index"),
               reference = ref_fasta)
}
```

Sample metadata is stored in a data.frame.

```{r}
fqfiles <- file.path(datdir, basename(fqurls))
meta <- data.frame(genotype = factor(c("wt", "wt", "ko", "ko")),
                   fqurl = fqurls, fqfile = fqfiles,
                   bamfile = gsub("\\.fastq\\.gz", ".bam", fqfiles),
                   sortedfile = gsub("\\.fastq\\.gz", ".sorted.bam", fqfiles),
                   stringsAsFactors = FALSE)
```

We also download blacklisted regions for mm10 from ENCODE (https://www.encodeproject.org/annotations/ENCSR636HFF/).

```{r}
blacklist_url <- "https://www.encodeproject.org/files/ENCFF547MET/@@download/ENCFF547MET.bed.gz"
if (!file.exists(file.path(datdir, basename(blacklist_url)))) {
    download.file(blacklist_url, destfile = file.path(datdir, basename(blacklist_url)))
}

bl <- import(file.path(datdir, basename(blacklist_url)))
```

Standard set of parameters are defined for the analysis.
Only the canonical set of chromosomes are used in the analysis.

```{r}
std_chr <- paste0("chr", c(1:19, "X", "Y"))
param <- readParam(minq = 20, discard = bl, restrict = std_chr)
```

## Read Counting

We count reads within sliding windows across the genome.

```{r count, results = 'hide', message = FALSE, warning = FALSE}
if (file.exists(count_file)) {
    win_cnts <- readRDS(count_file)
} else {
    unaligned <- !file.exists(meta$bamfile)
    if (any(unaligned)) {
        align(index = file.path(datdir, "mm_ref_index"),
              readfile1 = meta$fqfile[unaligned],
              type = 1, phredOffset = 64,
              input_format = "FASTQ",
              output_file = meta$bamfile[unaligned])
    }

    ## sort bam files
    for (i in 1:nrow(meta)) {
        if (!file.exists(meta$sortedfile[i])) {
            sortBam(meta$bamfile[i], gsub("\\.bam$", "", meta$sortedfile[i]))
        }
    }

    ## mark duplicates w/ picard and index bam files
    if (any(!file.exists(paste0(meta$sortedfile, ".bai")))) {
        temp_bam <- file.path(datdir, "temp_dups_mm.bam")
        temp_file <- file.path(datdir, "temp_mets_mm.txt")
        temp_dir <- file.path(datdir, "temp_dups_mm")
        dir.create(temp_dir)
        for (i_bam in meta$sortedfile) {
            code <- paste0("java -jar ${PICARD_TOOLS_HOME}/picard.jar ",
                           "MarkDuplicates I=%s O=%s M=%s ", 
                           "TMP_DIR=%s AS=true REMOVE_DUPLICATES=false ",
                           "VALIDATION_STRINGENCY=SILENT")
            code <- sprintf(code, i_bam, temp_bam, temp_file, temp_dir)
            code <- system(code)
            stopifnot(code == 0L)
            file.rename(temp_bam, i_bam)
        }
        unlink(temp_file)
        unlink(temp_dir, recursive = TRUE)

        indexBam(meta$sortedfile)
    }

    ## use correlateReads to determine fragment length (remove dups)
    x <- correlateReads(meta$sortedfile, param = reform(param, dedup = TRUE))
    frag_len <- which.max(x) - 1

    ## count reads in sliding windows (keep dups)
    win_cnts <- windowCounts(meta$sortedfile, param = param, width = 10, ext = frag_len)

    ## save unfiltered counts
    saveRDS(win_cnts, file = count_file)
}
```

We can apply prefiltering on windows with low abundance as described in Lun and Smyth (2016).

```{r}
if (file.exists(filtered_file)) {
    filtered_cnts <- readRDS(filtered_file)
    normfacs <- readRDS(normfacs_file)
} else {
    ## filter windows by abundance
    bins <- windowCounts(meta$sortedfile, bin = TRUE, width = 10000, param = param)
    filter_stat <- filterWindows(win_cnts, bins, type = "global")
    keep <- filter_stat$filter > log2(3)
    filtered_cnts <- win_cnts[keep, ]

    ## calculate normalizing offsets based on larger windows
    normfacs <- normOffsets(bins, se.out = FALSE)

    ## save filtered counts, normalizing factors
    saveRDS(filtered_cnts, file = filtered_file)
    saveRDS(normfacs, file = normfacs_file)
}
```

# Data Analysis

## Differential Testing

We use _edgeR_ to test for differential binding with the filtered data.

```{r test, results='hide', message=FALSE, warning=FALSE}
if (file.exists(result_file)) {
    res_ranges <- readRDS(result_file)
} else {
    ## create model diesign
    design <- model.matrix(~ 0 + genotype, data = meta)
    colnames(design) <- levels(meta$genotype)

    ## run quasi-likelihood F-test
    y <- asDGEList(filtered_cnts, norm.factors = normfacs)
    y <- estimateDisp(y, design)
    fit <- glmQLFit(y, design, robust = TRUE)
    res <- glmQLFTest(fit, contrast = makeContrasts(wt-ko, levels = design))

    ## merge p-values across regions
    merged <- mergeWindows(rowRanges(filtered_cnts), tol = 100, max.width = 5000)
    tab_comb <- combineTests(merged$id, res$table)
    tab_best <- getBestTest(merged$id, res$table)

    ## save results
    res_ranges <- merged$region
    elementMetadata(res_ranges) <-
        data.frame(tab_comb,
                   best_pos = mid(ranges(rowRanges(filtered_cnts[tab_best$best]))),
                   best_logFC = tab_best$logFC)

    ## get overall mean counts for each window
    merged_cnts <- summarizeOverlaps(res_ranges, BamFileList(meta$sortedfile))
    res_ranges$meancnt <- rowMeans(assays(merged_cnts)$counts)

    saveRDS(res_ranges, file = result_file)
}

## covert to df for downstream analysis
res_df <- as.data.frame(res_ranges)

## add random covariate 
set.seed(11245)
res_df$uninf_covar = rnorm(nrow(res_df))
```

## Covariate Diagnostics

### Number of Windows

```{r, nwin-diag-scatter, fig.width=4.5, fig.height=3.5}
rank_scatter(res_df, pvalue = "PValue", covariate = "nWindows") +
    ggtitle("Number of windows as independent covariate") +
    xlab("Number of Windows")
```

```{r, nwin-diag-hist, fig.width=10, fig.height=3.2}
strat_hist(res_df, pvalue = "PValue", covariate = "nWindows", maxy = 15)
```

### Region Width

```{r, width-diag-scatter, fig.width=4.5, fig.height=3.5}
rank_scatter(res_df, pvalue = "PValue", covariate = "width") +
    ggtitle("Region width as independent covariate") +
    xlab("Region Width")
```

```{r, width-diag-hist, fig.width=10, fig.height=3.2}
strat_hist(res_df, pvalue = "PValue", covariate = "width", maxy = 15)
```

### Mean Coverage

```{r, cov-diag-scatter, fig.width=4.5, fig.height=3.5}
rank_scatter(res_df, pvalue = "PValue", covariate = "meancnt") +
    ggtitle("Mean coverage as independent covariate") +
    xlab("Mean coverage")
```

```{r, cov-diag-hist, fig.width=10, fig.height=3.2}
strat_hist(res_df, pvalue = "PValue", covariate = "meancnt", maxy = 15)
```

### Random

```{r, rand-diag-scatter, fig.width=4.5, fig.height=3.5}
rank_scatter(res_df, pvalue = "PValue", covariate = "uninf_covar") +
    ggtitle("Random independent covariate") +
    xlab("Region Width")
```

```{r, rand-diag-hist, fig.width=10, fig.height=3.2}
strat_hist(res_df, pvalue = "PValue", covariate = "uninf_covar", maxy = 15)
```

## Multiple-Testing Correction

We use the common `BenchDesign` with the set of multiple testing correction
methods already included. We investigate both the width of the regions and their
mean coverage as the independent covariate. We won't assess the number of windows,
since this is very tightly correlated with the width of the region.

```{r}
cor(res_df[,c("width", "nWindows", "meancnt")])
```

First, we'll use the region width covariate.

```{r}
if (!file.exists(bench_file)) {
    res_df$ind_covariate <- res_df$width
    res_df$pval <- res_df$PValue

    bd <- initializeBenchDesign()
    sb <- buildBench(bd, data = res_df, ftCols = "ind_covariate")
    saveRDS(sb, file = bench_file)
} else {
    sb <- readRDS(bench_file)
}
```

Next, we'll use the mean coverage covariate.

```{r}
if (!file.exists(bench_file_cov)) {
    res_df$ind_covariate <- res_df$meancnt
    res_df$pval <- res_df$PValue

    bd <- initializeBenchDesign()
    sbC <- buildBench(bd, data = res_df, ftCols = "ind_covariate")
    saveRDS(sbC, file = bench_file_cov)
} else {
    sbC <- readRDS(bench_file_cov)
}
```

We'll also compare to the random covariate.

```{r}
if (!file.exists(bench_file_uninf)) {
    res_df$ind_covariate <- res_df$uninf_covar
    res_df$pval <- res_df$PValue

    bd <- initializeBenchDesign()
    sbU <- buildBench(bd, data = res_df, ftCols = "ind_covariate")
    saveRDS(sbU, file = bench_file_uninf)
} else {
    sbU <- readRDS(bench_file_uninf)
}
```

## Benchmark Metrics

### Region width

Next, we'll add the default performance metric for q-value assays. First, we have
to rename the assay to 'qvalue'.

```{r, chipseq-metrics}
assayNames(sb) <- "qvalue"
sb <- addDefaultMetrics(sb)
```

Now, we'll plot the results.

```{r, chipseq-scatter, results="hide"}
rejections_scatter(sb, supplementary = FALSE)
rejection_scatter_bins(sb, covariate = "ind_covariate",
                       bins = 4, supplementary = FALSE)
```

```{r, chipseq-overlap}
plotFDRMethodsOverlap(sb, alpha = 0.05, nsets = ncol(sb),
                      order.by = "freq", decreasing = TRUE,
                      supplementary = FALSE)
```

```{r, rnaseq-lineplots}
covariateLinePlot(sb, alpha = 0.05, covname = "ind_covariate")
```

### Mean Coverage

We'll do the same for the mean coverage covariate.

```{r, chipseq-metrics-cov}
assayNames(sbC) <- "qvalue"
sbC <- addDefaultMetrics(sbC)
```

Now, we'll plot the results.

```{r, chipseq-scatter-cov, results="hide"}
rejections_scatter(sbC, supplementary = FALSE)
rejection_scatter_bins(sbC, covariate = "ind_covariate",
                       bins = 4, supplementary = FALSE)
```

```{r, chipseq-overlap-cov}
plotFDRMethodsOverlap(sbC, alpha = 0.05, nsets = ncol(sbC),
                      order.by = "freq", decreasing = TRUE,
                      supplementary = FALSE)
```

```{r, rnaseq-lineplot-cov}
covariateLinePlot(sbC, alpha = 0.05, covname = "ind_covariate")
```

### Random

We'll do the same for the random (uninformative covariate).

```{r, chipseq-metrics-rand}
assayNames(sbU) <- "qvalue"
sbU <- addDefaultMetrics(sbU)
```

Now, we'll plot the results.

```{r, chipseq-scatter-rand, results="hide"}
rejections_scatter(sbU, supplementary = FALSE)
rejection_scatter_bins(sbU, covariate = "ind_covariate",
                       bins = 4, supplementary = FALSE)
```

```{r, chipseq-overlap-rand}
plotFDRMethodsOverlap(sbU, alpha = 0.05, nsets = ncol(sbU),
                      order.by = "freq", decreasing = TRUE,
                      supplementary = FALSE)
```

```{r, rnaseq-lineplot-rand}
covariateLinePlot(sbU, alpha = 0.05, covname = "ind_covariate")
```

## Covariate comparison

Here we compare the method ranks for the two covariates at alpha = 0.10.

```{r}
plotMethodRanks(c(bench_file, bench_file_cov, bench_file_uninf), 
                colLabels = c("Region width", "Mean Coverage", "Random"), 
                alpha = 0.10, xlab = "Covariate", 
                excludeMethods = NULL)
```

# Session Info

```{r}
sessionInfo()
```