### 2018-09-04 datasets/simulations/simulations-informative-cosine-nullAdaPT.Rmd
Add supplementary analyses of AdaPT with null option

---
title: "Simulation Study: AdaPT with null model option (cosine covariate)" 
author: "Patrick Kimes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

# Summary

In this set of simulations, we consider settings with both null and non-null
tests with an informative covariate. The covariate is sampled uniformly from
the interval [0, 1], and the conditional probability of a test being non-null
is a smooth (cosine) function of the covariate. We include simulation results
again with Gaussian, t-distributed, and Chi-Squared distributed noise.

This set of simulations differs from `simulations-informative-cosine.Rmd` only
in the implementation of the AdaPT method for multiple testing correction. This
is a follow up analysis similar to that described and performed in
`simulations-informative-step-nullAdaPT.Rmd`. For more details on the parameter settings
of the two additional AdaPT methods included in the comparison, please check the
other `-nullAdaPT.Rmd` analysis.

The analysis in `simulations-informative-step-nullAdaPT.Rmd` was carried out
to determine whether the additional `adapt-withnull` and `adapt-onlynull` methods
could provide an improvement over the `adapt-glm` parameter setting in the case
of low informativeness in the independent covariate. This analysis aims to examine
how the `adapt-withnull` and `adalpt-onlynull` parameter settings perform in
comparison to the default `adapt-glm` settings when the covariate is more
informative.

# Workspace Setup

```{r, wkspace-setup, results='hide', message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(SummarizedBenchmark)
library(parallel)

## load helper functions
for (f in list.files("../R", "\\.(r|R)$", full.names = TRUE)) {
    source(f)
}

## project data/results folders
resdir <- "results"
dir.create(resdir, showWarnings = FALSE, recursive = TRUE)

## intermediary files we create below
gauss_file <- file.path(resdir, "nullAdaPT-informative-cosine-benchmark-gaussian.rds")
tdist_file <- file.path(resdir, "nullAdaPT-informative-cosine-benchmark-t5.rds")
tdist11_file <- file.path(resdir, "nullAdaPT-informative-cosine-benchmark-t11.rds")
chisq_file <- file.path(resdir, "nullAdaPT-informative-cosine-benchmark-chisq4.rds")

## number of cores for parallelization
cores <- 20
B <- 100

## define bechmarking design
bd <- initializeBenchDesign()
```

As described in `simulations-null.Rmd`, we include Scott's FDR Regression in the analysis
for simulations with Gaussian or t-distributed noise. Again, we include both
`nulltype = "empirical"` and `nulltype = "theoretical"`.

```{r}
bdplus <- bd
bdplus <- addBMethod(bdplus, "fdrreg-t",
                     FDRreg::FDRreg,
                     function(x) { x$FDR },
                     z = test_statistic,
                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
                     nulltype = 'theoretical',
                     control = list(lambda = 0.01))
bdplus <- addBMethod(bdplus, "fdrreg-e",
                     FDRreg::FDRreg,
                     function(x) { x$FDR },
                     z = test_statistic,
                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
                     nulltype = 'empirical',
                     control = list(lambda = 0.01))
```

We add a couple alternative calls to AdaPT to test whether the method performance improves in cases
of low "informativeness" if given the option to not use the informative covariate. This option is
coded by the null model, `"pi0 ~ 1"`, where only an intercept term is included in the model. Specifically,
we add two alternative calls to the benchmark - one call with the null model added as one of the candidate
models to `pi_formulas=` and `mu_formulas=`, and a second call with the null model given as the only
candidate model to `pi_formulas=` and `mu_formulas=`.

```{r}
bdplus <- addBMethod(bdplus, "adapt-withnull",
                     adaptMT::adapt_glm,
                     function(x) { x$qvals },
                     pvals = pval,
                     x = data.frame(icov = ind_covariate),
                     pi_formulas = c("~ 1", paste0("splines::ns(icov, df = ", seq(2, 10, 2), ")")),
                     mu_formulas = c("~ 1", paste0("splines::ns(icov, df = ", seq(2, 10, 2), ")")),
                     alphas = 0)

bdplus <- addBMethod(bdplus, "adapt-onlynull",
                     adaptMT::adapt_glm,
                     function(x) { x$qvals },
                     pvals = pval,
                     x = data.frame(icov = ind_covariate),
                     pi_formulas = "~ 1",
                     mu_formulas = "~ 1",
                     alphas = 0)

candycols2 <- rbind(candycols,
                    data.frame(Method = c("adapt-withnull", "adapt-onlynull"),
                               col = c("darkgoldenrod3", "darkgoldenrod3"),
                               lty = c("dashed", "dotted"),
                               stringsAsFactors = FALSE))
```

All simulation settings will share the following parameters.

```{r parameters-shared}
m <- 20000                        # integer: number of hypothesis tests
pi0 <- pi0_cosine(0.90)           # numeric: proportion of null hypotheses
icovariate <- runif               # functional: independent covariate
```

Simulation results will be presented excluding a subset of methods, and
for certain plots (upset plots), a single alpha cutoff will be used.

```{r}
excludeSet <- c("unadjusted", "bl-df02", "bl-df04", "bl-df05")
ualpha <- 0.05
```

# Gaussian Setting

First, we consider the setting with Gaussian noise.

## Data Simulation

```{r gauss-parameters}
es_dist <- rnorm_generator(3)       # functional: dist of alternative test stats
ts_dist <- rnorm_perturber(1)  # functional: sampling dist/noise for test stats
null_dist <- rnorm_2pvaluer(1)    # functional: dist to calc p-values
seed <- 608
```

We next run the simulations (including Scott's FDR Regression).

```{r gauss-run-simulation}
if (file.exists(gauss_file)) {
    res <- readRDS(gauss_file)
} else {
    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdplus, m = m,
                    pi0 = pi0, es_dist = es_dist, icovariate = icovariate,
                    ts_dist = ts_dist, null_dist = null_dist,
                    seed = seed, mc.cores = cores)
    saveRDS(res, file = gauss_file)
}
res_i <- lapply(res, `[[`, "informative")
res_u <- lapply(res, `[[`, "uninformative")
```

## Covariate Diagnostics

Here, we show the relationship between the independent covariate and p-values for a
single replication of the experiment.

```{r gauss-one-simulation}
onerun <- simIteration(1, bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
```

```{r, gauss-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
```

```{r, gauss-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 5, numQ = 3)
```

## Benchmark Metrics

We plot the averaged results across `r B` replications.

```{r gauss-metrics-averages, results = "hide"}
resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))

plotsim_average(resdf, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, 
                palette = candycols2)

plotsim_average(resdf, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, 
                palette = candycols2)

plotsim_average(resdf, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, 
                palette = candycols2)

plotsim_average(resdf, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)
```

Since we are primarily interested in comparing just the AdaPT methods, we plot these
metrics excluding all other methods.

```{r gauss-metrics-averages-adapt, results = "hide"}
plotsim_average(resdf, met="rejections",
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="FDR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="TPR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="TNR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)
```

We also take a look at the distribution of rejects for each method as a function of
the effect size and independent covariate.

```{r gauss-metrics-covlineplot, results = "hide"}
covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size",
                  palette = candycols2)

covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate",
                  palette = candycols2)
```

We also look at the FDR as a function of the independent covariate.

```{r gauss-metrics-covlineplotFDR, results = "hide"}
covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate", metric = "FDR",
                  palette = candycols2)
```

We also compare the simulation results with and without an informative covariate.

```{r gauss-metrics-differences, results = "hide"}
resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))

resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
                                   performanceMetric, alpha, value),
                            select(resdfu, rep, blabel, param.alpha, key,
                                   performanceMetric, alpha, value),
                            by = c("rep", "blabel", "param.alpha", "key",
                                   "performanceMetric", "alpha"),
                            suffix = c(".info", ".uninfo"))
resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)

plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)
```

Again, we recreate the plots after subsetting on AdaPT methods.

```{r gauss-metrics-differences-adapt, results = "hide"}
plotsim_average(resdfiu, met="rejections", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="FDR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="TPR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="TNR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)
```

# Student's t Setting (df = 5)

Next, we consider the setting with t-distributed noise.

## Data Simulation

```{r t5-parameters}
es_dist <- rnorm_generator(6)    # functional: dist of alternative test stats
ts_dist <- rt_perturber(5)  # functional: sampling dist/noise for test stats
null_dist <- rt_2pvaluer(5)    # functional: dist to calc p-values
seed <- 815
```

For the t-distributed noise setting, we must specify the number of degrees of freedom
for ASH. We add an additional parameter to the `ashq` method with the corresponding
degrees of freedom of the noise distribution.

```{r}
bdplust <- modifyBMethod(bdplus, "ashq", df = 5)
```

We next run the simulations (including Scott's FDR Regression and ASH with degrees of
freedom specified).

```{r t5-run-simulation}
if (file.exists(tdist_file)) {
    res <- readRDS(tdist_file)
} else {
    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdplust, m = m,
                    pi0 = pi0, es_dist = es_dist, icovariate = icovariate,
                    ts_dist = ts_dist, null_dist = null_dist,
                    seed = seed, mc.cores = cores)
    saveRDS(res, file = tdist_file)
}
res_i <- lapply(res, `[[`, "informative")
res_u <- lapply(res, `[[`, "uninformative")
```

## Covariate Diagnostics

Here, we show the relationship between the independent covariate and p-values for a
single replication of the experiment.

```{r t5-one-simulation}
onerun <- simIteration(1, bdplust, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
```

```{r, t5-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
```

```{r, t5-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 10, numQ = 3)
```

## Benchmark Metrics

We plot the averaged results across `r B` replications.

```{r t5-metrics-averages, results = "hide"}
resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))

plotsim_average(resdf, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)
```

Since we are primarily interested in comparing just the AdaPT methods, we plot these
metrics excluding all other methods.

```{r t5-metrics-averages-adapt, results = "hide"}
plotsim_average(resdf, met="rejections",
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="FDR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="TPR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="TNR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)
```

We also take a look at the distribution of rejects for each method as a function of
the effect size and independent covariate.

```{r t5-metrics-covlineplot, results = "hide"}
covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size",
                  palette = candycols2)

covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate",
                  palette = candycols2)
```

We also look at the FDR as a function of the independent covariate.

```{r t5-metrics-covlineplotFDR, results = "hide"}
covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate", metric = "FDR",
                  palette = candycols2)
```

We also compare the simulation results with and without an informative covariate.

```{r t5-metrics-differences, results = "hide"}
resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))

resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
                                   performanceMetric, alpha, value),
                            select(resdfu, rep, blabel, param.alpha, key,
                                   performanceMetric, alpha, value),
                            by = c("rep", "blabel", "param.alpha", "key",
                                   "performanceMetric", "alpha"),
                            suffix = c(".info", ".uninfo"))
resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)

plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)
```

Again, we recreate the plots after subsetting on AdaPT methods.

```{r t5-metrics-differences-adapt, results = "hide"}
plotsim_average(resdfiu, met="rejections", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="FDR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="TPR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="TNR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)
```

# Student's t Setting (df = 11)

Next, we consider the setting with t-distributed noise.

## Data Simulation

```{r t11-parameters}
es_dist <- rnorm_generator(3)    # functional: dist of alternative test stats
ts_dist <- rt_perturber(11)  # functional: sampling dist/noise for test stats
null_dist <- rt_2pvaluer(11)    # functional: dist to calc p-values
seed <- 9158
```

For the t-distributed noise setting, we must specify the number of degrees of freedom
for ASH. We add an additional parameter to the `ashq` method with the corresponding
degrees of freedom of the noise distribution.

```{r}
bdplust <- modifyBMethod(bdplus, "ashq", df = 11)
```

We next run the simulations (including Scott's FDR Regression and ASH with degrees of
freedom specified).

```{r t11-run-simulation}
if (file.exists(tdist11_file)) {
    res <- readRDS(tdist11_file)
} else {
    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdplust, m = m,
                    pi0 = pi0, es_dist = es_dist, icovariate = icovariate,
                    ts_dist = ts_dist, null_dist = null_dist,
                    seed = seed, mc.cores = cores)
    saveRDS(res, file = tdist11_file)
}
res_i <- lapply(res, `[[`, "informative")
res_u <- lapply(res, `[[`, "uninformative")
```

## Covariate Diagnostics

Here, we show the relationship between the independent covariate and p-values for a
single replication of the experiment.

```{r t11-one-simulation}
onerun <- simIteration(1, bdplust, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
```

```{r, t11-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
```

```{r, t11-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 10, numQ = 3)
```

## Benchmark Metrics

We plot the averaged results across `r B` replications.

```{r t11-metrics-averages, results = "hide"}
resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))

plotsim_average(resdf, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)
```

Since we are primarily interested in comparing just the AdaPT methods, we plot these
metrics excluding all other methods.

```{r t11-metrics-averages-adapt, results = "hide"}
plotsim_average(resdf, met="rejections",
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="FDR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="TPR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="TNR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)
```

We also take a look at the distribution of rejects for each method as a function of
the effect size and independent covariate.

```{r t11-metrics-covlineplot, results = "hide"}
covariateLinePlot(res_i, alpha = ualpha, covname = "effect_size",
                  palette = candycols2)

covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate",
                  palette = candycols2)
```

We also look at the FDR as a function of the independent covariate.

```{r t11-metrics-covlineplotFDR, results = "hide"}
covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate", metric = "FDR",
                  palette = candycols2)
```

We also compare the simulation results with and without an informative covariate.

```{r t11-metrics-differences, results = "hide"}
resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))

resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
                                   performanceMetric, alpha, value),
                            select(resdfu, rep, blabel, param.alpha, key,
                                   performanceMetric, alpha, value),
                            by = c("rep", "blabel", "param.alpha", "key",
                                   "performanceMetric", "alpha"),
                            suffix = c(".info", ".uninfo"))
resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)

plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)
```

Again, we recreate the plots after subsetting on AdaPT methods.

```{r t11-metrics-differences-adapt, results = "hide"}
plotsim_average(resdfiu, met="rejections", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="FDR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="TPR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="TNR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)
```

# Chi-Squared Setting

Finally, we consider the setting with chi-squared distributed noise.

## Data Simulation

```{r chisq4-parameters}
es_dist <- rnorm_generator(15)       # functional: dist of alternative test stats
ts_dist <- rchisq_perturber(4)  # functional: sampling dist/noise for test stats
null_dist <- rchisq_pvaluer(4)     # functional: dist to calc p-values
seed <- 1023
```

For the chi-squared distributed noise setting, we must change the "mode" setting for ASH
from the default of 0 to `"estimate"` since the mode of null and alternative test statistics
are no longer centered at 0. While both approximate Normality and unimodality of effects are
violated in this simulation setting, by allowing the mode to be estimated, rather than forced
to 0, should return more comparable results.

```{r}
bdchi <- modifyBMethod(bdplus, "ashq", mode = "empirical")
```

We also drop the FDRreg methods from the chi-squared setting.

```{r}
bdchi <- dropBMethod(bdchi, "fdrreg-t")
bdchi <- dropBMethod(bdchi, "fdrreg-e")
```

We next run the simulations. We do not include FDR Regression because the test statistics
are not approximately normally distributed.

```{r chisq4-run-simulation}
if (file.exists(chisq_file)) {
    res <- readRDS(chisq_file)
} else {
    res <- mclapply(X = 1:B, FUN = simIteration, bench = bdchi, m = m,
                    pi0 = pi0, es_dist = es_dist, icovariate = icovariate,
                    ts_dist = ts_dist, null_dist = null_dist,
                    seed = seed, mc.cores = cores)
    saveRDS(res, file = chisq_file)
}
res_i <- lapply(res, `[[`, "informative")
res_u <- lapply(res, `[[`, "uninformative")
```

## Covariate Diagnostics

Here, we show the relationship between the independent covariate and p-values for a
single replication of the experiment.

```{r chisq4-one-simulation}
onerun <- simIteration(1, bdchi, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                       icovariate = icovariate, null_dist = null_dist, execute = FALSE)
```

```{r, chisq4-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
```

```{r, chisq4-diag-hist, results = "hide", fig.width=10, fig.height=3.2}
strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 10, numQ = 3)
```

## Benchmark Metrics

We plot the averaged results across `r B` replications.

```{r chisq4-metrics-averages, results = "hide"}
resdf <- plotsim_standardize(res_i, alpha = seq(0.01, 0.10, 0.01))

plotsim_average(resdf, met = "rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met = "FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met = "TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met = "TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE) 
```

Since we are primarily interested in comparing just the AdaPT methods, we plot these
metrics excluding all other methods.

```{r chisq4-metrics-averages-adapt, results = "hide"}
plotsim_average(resdf, met="rejections",
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="FDR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="TPR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)

plotsim_average(resdf, met="TNR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE,
                palette = candycols2)
```

We also take a look at the distribution of rejects for each method as a function of
the effect size and independent covariate. Again, we plot these without `ashq`.

```{r chisq4-metrics-covlineplot, results = "hide"}
covariateLinePlot(lapply(res_i, function(x) { x[, -which(colnames(x) == "ashq")] }),
                  alpha = ualpha, covname = "effect_size")

covariateLinePlot(lapply(res_i, function(x) { x[, -which(colnames(x) == "ashq")] }),
                  alpha = ualpha, covname = "ind_covariate")
```

We also look at the FDR as a function of the independent covariate.

```{r chisq4-metrics-covlineplotFDR, results = "hide"}
covariateLinePlot(res_i, alpha = ualpha, covname = "ind_covariate", metric = "FDR")
```

We also compare the simulation results with and without an informative covariate.

```{r chisq4-metrics-differences, results = "hide"}
resdfu <- plotsim_standardize(res_u, alpha = seq(0.01, 0.10, 0.01))

resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
                                   performanceMetric, alpha, value),
                            select(resdfu, rep, blabel, param.alpha, key,
                                   performanceMetric, alpha, value),
                            by = c("rep", "blabel", "param.alpha", "key",
                                   "performanceMetric", "alpha"),
                            suffix = c(".info", ".uninfo"))
resdfiu <- dplyr::mutate(resdfiu, value = value.info - value.uninfo)

plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)

plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)

plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)

plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE)
```

Again, we recreate the plots after subsetting on AdaPT methods.

```{r chisq4-metrics-differences-adapt, results = "hide"}
plotsim_average(resdfiu, met="rejections", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="FDR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="TPR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)

plotsim_average(resdfiu, met="TNR", 
                filter_set = grep("adapt", unique(resdf$blabel),
                                  invert = TRUE, value = TRUE), 
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE,
                palette = candycols2)
```

# Session Info

```{r}
sessionInfo()
```