### 2018-10-22 manuscript_figures/manuscript_ratings.Rmd
Add figures with tables for manuscript-ratings

---
title: "Final Ratings for Genome Biology"
author: "Rafalab"
output: 
    html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This Rmd is used to determine the ratings for each method shown in Figure 4.

# Set up workspace

```{r load-workspace, message = FALSE}
## Load packages and source benchmark FDR
library(tidyverse)
library(cowplot)
library(ggthemes)
library(grid)
##library(SummarizedBenchmark)
devtools::load_all("../../SummarizedBenchmark")

## load helper functions
for (f in list.files("../datasets/R", "\\.(r|R)$", full.names = TRUE)) {
    source(f)
}

# Assumes sb objects for the case studies and in silico experiments are in 
# the following location, which contains subfolders
# for each casestudy (if this isn't true, then parsing the case study and 
# dataset names later on will be incorrect)
path <- "../results"

# Also assumes that simulation results summary file is in the following location
# The `result-metrics.rds` was generated in the 
# `datasets/simulations/simulations-summary.Rmd` file.
simres_file <- file.path("..", "datasets", "simulations", "results-summary", "result-metrics.rds")

# set up results directory
outdir <- "./figures"
dir.create(outdir, showWarnings = FALSE)

# set alpha cutoff for plots that are fixed at a certain value of alpha
alpha.thresh <- 0.05

# methods to include in all figures (exclude bonferroni and fdrreg-e)
methodset <- c("bh", "ihw", "qvalue", "bl", "lfdr",
               "fdrreg-t", "ashq", "adapt-glm")
```

We use the standardize candy color scheme and line types for the plots. We'll
add the "lfdr*" method, which indicates lfdr was applied with fewer than 200 
tests per bin (out of 20 bins).

```{r}
col <- as.character(candycols$col)
names(col) <- as.character(candycols$Method)
lty <- as.character(candycols$lty)
names(lty) <- as.character(candycols$Method)
```

To generate the figures in this document, the simulation results must first be aggregated by running the code at `datasets/simulations/simulations-summary.Rmd`.

```{r load-sim-metrics}
simres <- readRDS(simres_file)
```

We similarly load all other results from the following files.

```{r}
objects <- list.files(path, recursive = TRUE, pattern = "rds", full.names = TRUE)
```

# Summarize Data Sets

## Yeast

```{r}
objects_yeast <- grep("yeast", objects, value = TRUE)

## read just yeast samples
if (file.exists("yeastres.rds")) {
   yeastres <- readRDS("yeastres.rds")
} else {
    yeastres <- lapply(objects_yeast, readRDS)
    yeastres <- lapply(yeastres, plotsim_standardize, alpha = alpha.thresh)

    names(yeastres) <- basename(objects_yeast)
    yeastres <- bind_rows(yeastres, .id = "series")
    saveRDS(yeastres, "yeastres.rds")
}

## clean up the results
yeast <- dplyr::select(yeastres, series, rep, blabel, param.alpha, key, alpha, value)
yeast <- dplyr::filter(yeast, as.numeric(param.alpha) == alpha | is.na(param.alpha))
yeast <- dplyr::mutate(yeast, blabel = ifelse(grepl("ihw", blabel), "ihw", blabel))
yeast <- dplyr::mutate(yeast, blabel = ifelse(blabel == "bl-df03", "bl", blabel))
yeast <- dplyr::filter(yeast, blabel %in% methodset)
yeast <- dplyr::select(yeast, -param.alpha, -alpha)
yeast <- dplyr::rename(yeast, Method = blabel)

## use 0s for "rates" that are reported as NAs
yeast <- dplyr::mutate(yeast, value = ifelse(is.na(value) & grepl("R$", key), 0, value))

## compute mean and standard errors for each method in each sim for each metric
yeast <- dplyr::group_by(yeast, series, Method, key)
yeast <- dplyr::summarize(yeast, mean = mean(value, na.rm = TRUE),
                          se = sd(value, na.rm = TRUE) / sqrt(sum(!is.na(value))))
yeast <- dplyr::ungroup(yeast)

## clean up the simulation names
yeast <- dplyr::mutate(yeast, series = gsub("\\.rds", "", series))
yeast <- tidyr::separate(yeast, series, c("setting_base", "setting_res", "setting_de", "setting_cov"),
                         fill = "right")
yeast <- dplyr::select(yeast, -setting_res)
yeast <- dplyr::mutate(yeast, setting_mode = ifelse(grepl("II", setting_base), "bimodal", "unimodal"))
yeast <- dplyr::mutate(yeast, setting_pi0 = ifelse(grepl("H", setting_base), "7.5%", "30%"))

yeast <- dplyr::mutate(yeast, setting_cov = ifelse(is.na(setting_cov), "strongInfoCov", setting_cov))
yeast <- dplyr::mutate(yeast, setting_cov = ifelse(grepl("W", setting_base), "weakInfoCov", setting_cov))
yeast <- dplyr::select(yeast, -setting_base)
```

We also rank each method in each setting according to FDR and TPR.

```{r}
yeast <- dplyr::group_by(yeast, setting_de, setting_cov, setting_mode, setting_pi0, key)
yeast <- dplyr::mutate(yeast, rank = rank(-mean))
yeast <- dplyr::ungroup(yeast)
```

## Case Studies

```{r}
objects_dat <- objects[!grepl("polyester", objects) & !grepl("yeast", objects)]

## remove microbiome objects from 'null' comparisons
## removing baxter, goodrich, papa OTU-level, and So4-enigma (since they mostly found
## nothing for all methods and the genus-level analysis is also present)
## Also remove'log' ubiquity since redundant
null <- grepl(c("goodrich|papa"), objects_dat) &
    grepl(c("otu"), objects_dat) |
    grepl("baxter", objects_dat) |
    grepl("log", objects_dat) |
    grepl("so4", objects_dat)
objects_dat <- objects_dat[!null]

## remove chip-seq promoter analysis (since it is the only one that uses
## ash and we also include the csaw analysis on the same dataset).
promot <- grepl("promot", objects_dat)
objects_dat <- objects_dat[!promot]

# remove random covariate analyses
uninf <- grepl("uninf", objects_dat)
objects_dat <- objects_dat[!uninf]

## read just real data samples
if (file.exists("datres.rds")) {
    datres <- readRDS("datres.rds")
} else {
    datres <- lapply(objects_dat, readRDS)

    ## gather all results
    datres <- lapply(datres, assay, "bench")
    datres <- lapply(datres, as_tibble)
    names(datres) <- gsub("^\\.\\./results/(.*?)-benchmark(.*)\\.rds", "\\1\\2", objects_dat)
    datres <- bind_rows(datres, .id = "cs")
    datres <- tidyr::separate(datres, cs, c("casestudy", "dataset"), sep = "/")
    datres <- tidyr::gather(datres, Method, adjp, -casestudy, -dataset)

    ## summarize results
    datres <- dplyr::group_by(datres, casestudy, dataset, Method)
    datres <- dplyr::summarize(datres, ntotal = n(), nNA = sum(is.na(adjp)),
                               nrej = sum(adjp < alpha.thresh, na.rm = TRUE))
    datres <- dplyr::ungroup(datres)

    saveRDS(datres, "datres.rds")
}

## NAs only appear in cases where the method wasn't applied so we can go ahead
## and replace the nrej values with NA for these methods
datres <- dplyr::mutate(datres, nrej = ifelse(nNA/ntotal > .99, NA, nrej))

## clean up method names
iihw <- paste0("ihw-a", sprintf("%02d", 100*alpha.thresh))
datres <- dplyr::mutate(datres, Method = ifelse(grepl(iihw, Method), "ihw", Method))
datres <- dplyr::mutate(datres, Method = ifelse(Method == "bl-df03", "bl", Method))
datres <- dplyr::filter(datres, Method %in% methodset)

## add method rank in each category
datres <- dplyr::group_by(datres, casestudy, dataset)
datres <- dplyr::mutate(datres, rank = rank(-nrej)) 
datres <- dplyr::mutate(datres, rank = ifelse(is.na(nrej), NA, rank))
datres <- dplyr::ungroup(datres)
```

## Simulations

We can use the results stored in `simres` for simulation settings.
For easier plotting, we replace create a shorter distribution string column.

```{r}
simres <- dplyr::mutate(simres, sdist = factor(dist, sort(levels(dist)), c("chisq", "normal", "t11", "t5")))
```

We only look at results at the desired alpha cutoff, and remove `fdrreg-e` from this comparison.

```{r}
simres <- dplyr::filter(simres, Method != "fdrreg-e", Method != "bonf", alpha == alpha.thresh)
```

We also rank each method in each setting according to FDR and TPR.

```{r}
simres <- dplyr::group_by(simres, setting, dist, inform, pi0, signal, ntests, sdist, key)
simres <- dplyr::mutate(simres, rank = rank(-mean.info))
simres <- dplyr::ungroup(simres)
```

For consistency across plotting of settings, we relabel some settings.

```{r}
simres <- dplyr::mutate(simres, setting = ifelse(setting == "varyingpi0-t", "varyingpi0", setting))
```

# Exploratory Figures

After summarizing all yeast experiments, simulations, and case studies, we take a look at the
aggregated results.

## Yeast

```{r, fig.width = 8, fig.height = 12}
dplyr::filter(yeast, key == "FDR") %>%
    ggplot(aes(x = interaction(setting_mode, setting_cov, setting_pi0, setting_de),
               y = mean, ymin = mean - se, ymax = mean+se, color = Method)) +
    geom_point(size = 1/2, position = position_dodge(width = 1/2)) +
    geom_errorbar(position = position_dodge(width = 1/2)) +
    facet_grid(Method ~ setting_de, scales = "free_x", space = "free_x") +
    expand_limits(y = 0) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    ylab("FDR") +
    xlab(NULL) +
    theme_bw() +
    geom_hline(yintercept = 0.05, linetype = 2, color = "blue", alpha = 1/2) + 
    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1, size = 6)) +
    ggtitle(paste0("Yeast Experiments: FDR control across settings, alpha = ", alpha.thresh))

dplyr::filter(yeast, key == "TPR", !grepl("null", setting_de)) %>%
    ggplot(aes(x = interaction(setting_mode, setting_pi0),
               y = mean, ymin = mean - se, ymax = mean+se, color = Method)) +
    geom_point(size = 1/2, position = position_dodge(width = 1/2)) +
    geom_errorbar(position = position_dodge(width = 1/2)) +
    facet_grid(setting_de ~ setting_cov) +
    expand_limits(y = 0) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    ylab("TPR") +
    xlab(NULL) +
    theme_bw() +
    geom_hline(yintercept = 0.05, linetype = 2, color = "blue", alpha = 1/2) + 
    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1, size = 6)) +
    ggtitle(paste0("Yeast Experiments: TPR across settings, alpha = ", alpha.thresh))
```

Now we also look at the ranks of methods.

```{r, fig.width = 8, fig.height = 12}
dplyr::filter(yeast, key == "FDR") %>%
    ggplot(aes(x = interaction(setting_mode, setting_cov, setting_pi0, setting_de),
               y = rank, color = Method)) +
    geom_point(size = 4, position = position_dodge(width = 1/2)) +
    facet_grid(Method ~ setting_de, scales = "free_x", space = "free_x") +
    expand_limits(y = 0) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    ylab("FDR rank") +
    xlab(NULL) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1, size = 6)) +
    ggtitle(paste0("Yeast Experiments: method rank for TPR, alpha = ", alpha.thresh))

dplyr::filter(yeast, key == "TPR", !grepl("null", setting_de)) %>%
    ggplot(aes(x = interaction(setting_mode, setting_cov, setting_pi0, setting_de),
               y = rank, color = Method)) +
    geom_point(size = 4, position = position_dodge(width = 1/2)) +
    facet_grid(Method ~ setting_de, scales = "free_x", space = "free_x") +
    expand_limits(y = 0) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    ylab("TPR rank") +
    xlab(NULL) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1, size = 6)) +
    ggtitle(paste0("Yeast Experiments: method rank for TPR, alpha = ", alpha.thresh))
```

## Case Studies

```{r, fig.width = 10, fig.height = 5}
ggplot(datres, aes(x = dataset, y = nrej / ntotal, color = Method)) +
    geom_point(size = 3/4, position = position_dodge(width = 1/2)) +
    facet_grid(. ~ casestudy, scales = "free_x", space = "free_x") +
    expand_limits(y = c(0, 1)) +
    scale_color_manual(values = col) +
    scale_y_continuous("Percent Rejections", labels = scales::percent) +
    xlab(NULL) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1)) +
    ggtitle(paste0("Case Study: method rejection proportion, alpha = ", alpha.thresh))
```

```{r, fig.width = 10, fig.height = 14}
ggplot(datres, aes(x = dataset, y = rank, color = Method, group = Method)) +
    geom_line(size = 1/2, alpha = 1/4, position = position_dodge(width = 1/4)) +
    geom_point(size = 1, position = position_dodge(width = 1/4)) +
    facet_grid(Method ~ casestudy, scales = "free_x", space = "free_x") +
    scale_color_manual(values = col) +
    scale_y_continuous("Rank", breaks = 1:8) +
    xlab(NULL) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1)) +
    ggtitle(paste0("Case Study: method rank for total rejections, alpha = ", alpha.thresh))
```

```{r, fig.width = 10, fig.height = 6}
ggplot(datres, aes(x = dataset, y = rank, color = Method, group = Method)) +
    geom_point(size = 4, position = position_dodge(width = 1/2)) +
    facet_grid(. ~ casestudy, scales = "free_x", space = "free_x") +
    scale_color_manual(values = col) +
    scale_y_continuous("Rank", breaks = 1:8) +
    xlab(NULL) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1)) +
    ggtitle(paste0("Case Study: method rank for total rejections, alpha = ", alpha.thresh))
```

## Simulations

```{r}
dplyr::filter(simres, key == "FDR") %>%
    tidyr::unite(subsetting, sdist, inform, pi0, signal, ntests) %>%
    ggplot(aes(x = subsetting,
               y = mean.info, ymin = mean.info - se.info, ymax = mean.info + se.info, color = Method)) +
    geom_point(size = 1/2, position = position_dodge(width = 1/2)) +
    geom_errorbar(position = position_dodge(width = 1/2)) +
    facet_grid(Method ~ setting, space = "free_x", scales = "free_x") +
    expand_limits(y = 0) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    ylab("FDR") +
    xlab(NULL) +
    theme_bw() +
    geom_hline(yintercept = 0.05, linetype = 2, color = "blue", alpha = 1/2) + 
    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1, size = 6)) +
    ggtitle(paste0("Simulations: FDR control across settings, alpha = ", alpha.thresh))

dplyr::filter(yeast, key == "TPR", !grepl("null", setting_de)) %>%
    ggplot(aes(x = interaction(setting_mode, setting_pi0),
               y = mean, ymin = mean - se, ymax = mean+se, color = Method)) +
    geom_point(size = 1/2, position = position_dodge(width = 1/2)) +
    geom_errorbar(position = position_dodge(width = 1/2)) +
    facet_grid(setting_de ~ setting_cov) +
    expand_limits(y = 0) +
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    ylab("TPR") +
    xlab(NULL) +
    theme_bw() +
    geom_hline(yintercept = 0.05, linetype = 2, color = "blue", alpha = 1/2) + 
    theme(axis.text.x = element_text(angle = 90, vjust = 1/2, hjust = 1, size = 6)) +
    ggtitle(paste0("Yeast Experiments: TPR across settings, alpha = ", alpha.thresh))
```

# Ratings

We determine cutoffs using the summarized results.

## FDR Control

FDR control ratings are determined using only the yeast experiments and simulation
results where ground truth is known.

```{r}
## FDR control across simulation settings
fdrcontrol_sims <- simres %>%
    dplyr::filter(key == "FDR", !grepl("null", setting), sdist != "chisq") %>%
    dplyr::group_by(Method) %>%
    dplyr::summarize(nExceeds = sum((mean.info - se.info) > alpha.thresh),
                     pExceeds = round(nExceeds / n(), 4)) %>%
    dplyr::ungroup()

## FDR control across yeast settings
fdrcontrol_yeast <- yeast %>%
    dplyr::filter(key == "FDR", !grepl("null", setting_de)) %>%
    dplyr::group_by(Method) %>%
    dplyr::summarize(nExceeds = sum((mean - se) > alpha.thresh),
                     pExceeds = round(nExceeds / n(), 4)) %>%
    dplyr::ungroup()

## combine
fdrcontrol <- dplyr::left_join(fdrcontrol_sims, fdrcontrol_yeast, by = "Method",
                               suffix = c(".sims", ".yeast"))
dplyr::arrange(fdrcontrol, pExceeds.sims + pExceeds.yeast)
```

We plot the results as well.

```{r, fig.width = 6, fig.height = 4}
bind_rows(yeast = fdrcontrol_yeast, sims = fdrcontrol_sims, .id = "analysis") %>%
    ggplot(aes(x = analysis, y = pExceeds, color = Method, group = Method)) +
    geom_line(aes(linetype = Method), alpha = 1/2, position = position_dodge(width = 1/4)) +
    geom_point(position = position_dodge(width = 1/4)) + 
    scale_linetype_manual(values = lty) + 
    scale_color_manual(values = col) +
    scale_x_discrete(NULL) + 
    scale_y_continuous("Percent Settings Exceeding nominal FDR", labels = scales::percent) + 
    expand_limits(y = 0) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14)) +
    ggtitle("FDR control across simulations and yeast experiments") 
```

## Power

Similarly, power ratings are only determined using only the yeast experiments and
simulation results where ground truth is known.

```{r}
## power across simulation settings
power_sims <- simres %>%
    dplyr::filter(key == "TPR", alpha == alpha.thresh,
                  !grepl("null", setting), sdist != "chisq") %>%
    dplyr::group_by(Method) %>%
    dplyr::summarize(meanRank = mean(rank),
                     medianRank = median(rank)) %>%
    dplyr::ungroup()

## power across yeast settings
power_yeast <- yeast %>%
    dplyr::filter(key == "TPR", !grepl("null", setting_de)) %>%
    dplyr::group_by(Method) %>%
    dplyr::summarize(meanRank = mean(rank),
                     medianRank = median(rank)) %>%
    dplyr::ungroup()

## combine
power <- dplyr::left_join(power_sims, power_yeast, by = "Method",
                          suffix = c(".sims", ".yeast"))
dplyr::arrange(power, meanRank.sims + meanRank.yeast)
```

We plot the results as well.

```{r, fig.width = 6, fig.height = 4}
bind_rows(yeast = power_yeast, sims = power_sims, .id = "analysis") %>%
    dplyr::mutate(Method = reorder(Method, meanRank)) %>%
    ggplot(aes(x = Method, y = meanRank, color = Method, group = analysis)) +
    geom_line(aes(linetype = analysis), color = 'black', alpha = 1/4) +
    geom_point() + 
    scale_color_manual(values = col, guide = FALSE) +
    scale_x_discrete(NULL) + 
    scale_y_continuous("Mean TPR rank", breaks = 0:10) + 
    expand_limits(y = 1) +
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14)) +
    ggtitle("TPR rank across simulations and yeast experiments") 
```

## Robustness

Next, we define metrics for the robustness metric, which quantifies the amount of consistency,
or robustness, of each method across simulations, yeast experiments, and case studies. For
this metric, we only look at modern methods, i.e. exclude Bonferroni correction, BH, and
Storey's q-value. We do this, as we quantify "robustness" as the relative gain of each method
over the average of BH and q-value in FDR and TPR (and proportion of rejections for case studies).

```{r}
## robustness across simulation settings
robustness_sims <- simres %>%
    dplyr::filter(key == "TPR" | key == "FDR", alpha == alpha.thresh,
                  !grepl("null", setting), sdist != "chisq",
                  Method != "bonf") %>%
    dplyr::select(Method, key, mean.info, setting, sdist, inform, pi0, signal, ntests) %>%
    tidyr::spread(Method, mean.info) %>%
    tidyr::gather(Method, mean.info, -key, -setting, -sdist, -inform,
                  -pi0, -signal, -ntests, -qvalue, -bh) %>%
    dplyr::group_by(Method, key) %>%
    dplyr::summarize(meanGain = mean(mean.info / (bh + qvalue) * 2),
                     sdGainBH = sd(mean.info / bh),
                     sdGainQ = sd(mean.info / qvalue),
                     sdGain = sd(mean.info / (bh + qvalue) * 2),
                     cvGain = sdGain / meanGain) %>%
    dplyr::ungroup()

## robustness across yeast experiments
robustness_yeast <- yeast %>%
    dplyr::filter(key == "TPR" | key == "FDR", !grepl("null", setting_de)) %>%
    dplyr::select(Method, mean, key, setting_de, setting_cov, setting_mode, setting_pi0) %>%
    tidyr::spread(Method, mean) %>%
    tidyr::gather(Method, mean, -key, -setting_de, -setting_cov, -setting_pi0,
                  -setting_mode, -qvalue, -bh) %>%
    dplyr::group_by(Method, key) %>%
    dplyr::summarize(meanGain = mean(mean / (bh + qvalue) * 2),
                     sdGainBH = sd(mean / bh),
                     sdGainQ = sd(mean / qvalue),
                     sdGain = sd(mean / (bh + qvalue) * 2),
                     cvGain = sdGain / meanGain) %>%
    dplyr::ungroup()

## robustness across case studies
robustness_data <- datres %>%
    dplyr::mutate(prej = nrej / ntotal) %>%
    dplyr::select(-rank, -nrej, -ntotal, -nNA) %>%
    tidyr::spread(Method, prej) %>%
    tidyr::gather(Method, prej, -casestudy, -dataset, -qvalue, -bh) %>%
    dplyr::group_by(Method) %>%
    dplyr::summarize(meanGain = mean(prej / (bh + qvalue) * 2, na.rm = TRUE),
                     sdGainBH = sd(prej / bh, na.rm = TRUE),
                     sdGainQ = sd(prej / qvalue, na.rm = TRUE),
                     sdGain = sd(prej / (bh + qvalue) * 2, na.rm = TRUE),
                     cvGain = sdGain / meanGain) %>%
    dplyr::ungroup() %>%
    dplyr::mutate(key = "PropRej")

## combine
robustness <- dplyr::full_join(dplyr::select(robustness_sims, Method, key, cvGain),
                               dplyr::select(robustness_yeast, Method, key, cvGain),
                               by = c("key", "Method"),
                               suffix = c(".sims", ""))
robustness <- dplyr::full_join(robustness,
                               dplyr::select(robustness_data, Method, key, cvGain),
                               by = c("key", "Method"),
                               suffix = c(".yeast", ".data"))
dplyr::arrange(robustness, key, cvGain.sims + cvGain.yeast, cvGain.data)
```

We plot the results as well.

```{r, fig.width = 6, fig.height = 8}
bind_rows(yeast = robustness_yeast, sims = robustness_sims,
          casestudies = robustness_data, .id = "analysis") %>%
    dplyr::mutate(Method = reorder(Method, sdGain)) %>%
    ggplot(aes(x = Method, y = sdGain, color = Method, group = analysis)) +
    geom_line(aes(linetype = analysis), color = 'black', alpha = 1/4) +
    geom_point() + 
    scale_color_manual(values = col, guide = FALSE) +
    scale_x_discrete(NULL) + 
    scale_y_continuous("SD of percent relative to BH/qvalue") + 
    expand_limits(y = 1) +
    facet_grid(key ~ ., scales = "free") + 
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14)) +
    ggtitle("Variation across simulations, yeast experiments,\ncase studies")

bind_rows(yeast = robustness_yeast, sims = robustness_sims,
          casestudies = robustness_data, .id = "analysis") %>%
    dplyr::mutate(Method = reorder(Method, cvGain)) %>%
    ggplot(aes(x = Method, y = cvGain, color = Method, group = analysis)) +
    geom_line(aes(linetype = analysis), color = 'black', alpha = 1/4) +
    geom_point() + 
    scale_color_manual(values = col, guide = FALSE) +
    scale_x_discrete(NULL) + 
    scale_y_continuous("CV of percent relative to BH/qvalue") + 
    expand_limits(y = 0) +
    facet_grid(key ~ ., scales = "free") + 
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14)) +
    ggtitle("Variation across simulations, yeast experiments,\ncase studies")

bind_rows(yeast = robustness_yeast, sims = robustness_sims,
          casestudies = robustness_data, .id = "analysis") %>%
    dplyr::mutate(Method = reorder(Method, cvGain)) %>%
    ggplot(aes(x = Method, y = cvGain, color = Method, group = key)) +
    geom_line(aes(linetype = key), color = 'black', alpha = 1/4) +
    geom_point() + 
    scale_color_manual(values = col, guide = FALSE) +
    scale_x_discrete(NULL) + 
    scale_y_continuous("CV of percent relative to BH/qvalue") +
    scale_linetype("Metric") + 
    expand_limits(y = 0) +
    facet_grid(analysis ~ ., scales = "free_y") + 
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14)) +
    ggtitle("Variation across simulations, yeast experiments,\ncase studies")

bind_rows(yeast = robustness_yeast, sims = robustness_sims,
          casestudies = robustness_data, .id = "analysis") %>%
    dplyr::mutate(Method = reorder(Method, sdGain)) %>%
    ggplot(aes(x = Method, y = sdGain, color = Method, group = key)) +
    geom_line(aes(linetype = key), color = 'black', alpha = 1/4) +
    geom_point() + 
    scale_color_manual(values = col, guide = FALSE) +
    scale_x_discrete(NULL) + 
    scale_y_continuous("SD of percent relative to BH/qvalue") +
    scale_linetype("Metric") + 
    expand_limits(y = 0) +
    facet_grid(analysis ~ ., scales = "free_y") + 
    theme_classic() +
    theme(axis.title = element_text(face = "bold", size = 10),
          plot.title = element_text(face = "bold", size = 14)) +
    ggtitle("Variation across simulations, yeast experiments,\ncase studies") 
```

## Applicability

Finally, for "applicability" we take a look at the number of times each method was not applied
in the case studies.

```{r}
applicability_data <- datres %>%
    dplyr::group_by(casestudy, Method) %>%
    dplyr::summarize(meanNA = mean(is.na(nrej))) %>%
    dplyr::group_by(Method) %>%
    dplyr::summarize(meanNA = mean(meanNA)) %>%
    dplyr::ungroup()
dplyr::arrange(applicability_data, meanNA)
```