### 2019-04-05 datasets/simulations/nulldependent/simulations-nulldependent-cubic.Rmd
Add simulation looking at dependence under null

---
title: "Simulation Study: Null Dependent Covariate (cubic)"
author: "Patrick Kimes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
   html_document:
        toc: true
        toc_float: true
        highlight: tango
        number_sections: true
---

# Summary

In this set of simulations, we consider settings with both null and non-null
tests with an informative and uninformative covariate. The informative covariate is sampled uniformly from the
interval [0, 1], and the conditional probability of a test being non-null
is a smooth (cubic) function of the covariate. The uninformative covariate is simply
uninformly sampled independently from the interval [0, 1]. The uninformative covariate is included
as a baseline to compare the informative covariate against. We include simulation results
again with Gaussian distributed test statistics.

Additionally, with the informative covariate, a dependence is introduced between the coviarate 
the p-value under the null distribution.

# Workspace Setup

```{r, wkspace-setup, results='hide', message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(parallel)
library(cowplot)

## load helper functions
for (f in list.files("../../R", "\\.(r|R)$", full.names = TRUE)) {
    source(f)
}

## new helper function
source("helpers-simulations-nulldependent.R")

## project data/results folders
resdir <- "results"
dir.create(resdir, showWarnings = FALSE, recursive = TRUE)

## intermediary files we create below
weak_file <- file.path(resdir, "nulldependent-cubic-benchmark-weaker.rds")
strong_file <- file.path(resdir, "nulldependent-cubic-benchmark-stronger.rds")

## number of cores for parallelization
cores <- 20
B <- 100

## define bechmarking design
bd <- initializeBenchDesign()
```

As described in `simulations-null.Rmd`, we include Scott's FDR Regression in the analysis
since all settings here use Gaussian test statistics. Again, we include both
`nulltype = "empirical"` and `nulltype = "theoretical"`.

```{r}
bdplus <- bd
bdplus <- addBMethod(bdplus, "fdrreg-t",
                     FDRreg::FDRreg,
                     function(x) { x$FDR },
                     z = test_statistic,
                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
                     nulltype = 'theoretical',
                     control = list(lambda = 0.01))
bdplus <- addBMethod(bdplus, "fdrreg-e",
                     FDRreg::FDRreg,
                     function(x) { x$FDR },
                     z = test_statistic,
                     features = model.matrix( ~  splines::bs(ind_covariate, df = 3) - 1),
                     nulltype = 'empirical',
                     control = list(lambda = 0.01))
```

All simulation settings will share the following parameters.

```{r parameters-shared}
m <- 20000                        # integer: number of hypothesis tests
pi0 <- pi0_cubic(0.90)            # numeric: proportion of null hypotheses
icovariate <- runif               # functional: independent covariate

es_dist <- rnorm_generator(3)     # functional: dist of alternative test stats
ts_dist <- rnorm_perturber(1)     # functional: sampling dist/noise for test stats
null_dist <- rnorm_2pvaluer(1)    # functional: dist to calc p-values
null_inv <- function(p) { (2*rbinom(length(p), 1, .5) - 1) * qnorm(p / 2, 0, 1) }
seed <- 608
```

Simulation results will be presented excluding a subset of methods, and
for certain plots (upset plots), a single alpha cutoff will be used.

```{r}
excludeSet <- c("unadjusted", "bl-df02", "bl-df04", "bl-df05")
ualpha <- 0.05
```

# Null Dependence

Here, we show null dependence on the variate by just showing the distribution of p-values when
all tests are null (pi0 = 1). We show the distribution of p-values for two settings:
weaker and stronger dependence between the covariate and p-values under the null.

## Weaker Dependence

```{r weak-null-one-simulation}
onerun_w <- simNullDependent(bdplus, m = 20000, pi0 = 1, es_dist = es_dist, ts_dist = ts_dist,
                             icovariate = icovariate, null_dist = null_dist, null_inv = null_inv,
                             null_dependence = 0.3, execute = FALSE)
```

```{r, weak-null-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
rank_scatter(onerun_w, pvalue = "pval", covariate = "ind_covariate") +
    ggtitle("Dist. of null p-values w/ weaker dependence")
```

```{r, weak-null-diag-hist, results = "hide", fig.width=10, fig.height=4}
strat_hist(onerun_w, pvalue = "pval", covariate = "ind_covariate", maxy = 7, numQ = 5) +
    ggtitle("Dist. of null p-values w/ weaker dependence")
```

## Stronger Dependence

```{r strong-null-one-simulation}
onerun_s <- simNullDependent(bdplus, m = 20000, pi0 = 1, es_dist = es_dist, ts_dist = ts_dist,
                             icovariate = icovariate, null_dist = null_dist, null_inv = null_inv,
                             null_dependence = 0.7, execute = FALSE)
```

```{r, strong-null-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
rank_scatter(onerun_s, pvalue = "pval", covariate = "ind_covariate") +
    ggtitle("Dist. of null p-values w/ stronger dependence")
```

```{r, strong-null-diag-hist, results = "hide", fig.width=10, fig.height=4}
strat_hist(onerun_s, pvalue = "pval", covariate = "ind_covariate", maxy = 7, numQ = 5) +
    ggtitle("Dist. of null p-values w/ stronger dependence")
```

## Multipanel Plot

We also create the above plots as a multipanel plot.

```{r weak-4-panels, fig.width = 9, fig.height = 3.5}
gp_oW <- rank_scatter(onerun_w, pvalue = "pval", covariate = "ind_covariate") +
    ggtitle("Dist. of null p-values w/ weaker dependence")
gp_oS <- rank_scatter(onerun_s, pvalue = "pval", covariate = "ind_covariate") +
    ggtitle("Dist. of null p-values w/ stronger dependence")

gp_o <- plot_grid(gp_oW + expand_limits(y = c(0, 5)),
                  gp_oS + expand_limits(y = c(0, 5)),
                  labels = LETTERS[1:2], nrow = 1)
gp_o
```

We next run the simulations.

# Weaker Dependence

First, we take a look at the results with weaker dependence.

```{r weak-run-simulation}
if (file.exists(weak_file)) {
    res <- readRDS(weak_file)
} else {
    res <- mclapply(X = 1:B, FUN = simNullDependent, bench = bdplus, m = m,
                    pi0 = pi0, es_dist = es_dist, icovariate = icovariate,
                    ts_dist = ts_dist, null_dist = null_dist, null_inv = null_inv,
                    null_dependence = 0.3,
                    seed = seed, mc.cores = cores)
    saveRDS(res, file = weak_file)
}

res_dep <- lapply(res, `[[`, "null_dependent")
res_indep <- lapply(res, `[[`, "null_independent")
```

## Covariate Diagnostics

Here, we show the relationship between the independent covariate and p-values for a
single replication of the experiment. Unlike the plots above, these include both
null and non-null tests.

```{r weak-one-simulation}
onerun <- simNullDependent(bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                           icovariate = icovariate, null_dist = null_dist, null_inv = null_inv,
                           null_dependence = 0.3, execute = FALSE)
```

```{r, weak-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
```

```{r, weak-diag-hist, results = "hide", fig.width=10, fig.height=4}
strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 7, numQ = 5)
```

## Benchmark Metrics

We plot the averaged results across `r B` replications.

```{r weak-metrics-averages, results = "hide"}
resdf <- plotsim_standardize(res_dep, alpha = seq(0.01, 0.10, 0.01))

plotsim_average(resdf, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE) 
```

We also take a look at the distribution of rejects for each method as a function of
the effect size and independent covariate.

```{r weak-metrics-covlineplot, results = "hide"}
covariateLinePlot(res_dep, alpha = ualpha, covname = "effect_size")

covariateLinePlot(res_dep, alpha = ualpha, covname = "ind_covariate")
```

We also look at the FDR as a function of the independent covariate.

```{r weak-metrics-covlineplotFDR, results = "hide"}
covariateLinePlot(res_dep, alpha = ualpha, covname = "ind_covariate", metric = "FDR")
```

Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
the overlap of rejections between methods.

```{r weak-metrics-upset, results = "hide"}
if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
    aggupset(res_dep, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
} else {
    message("Not enough methods found rejections at alpha ", ualpha, 
            "; skipping upset plot")
}
```

We also compare the simulation results with and without dependence between the covariate
under the nul.

```{r weak-metrics-differences, results = "hide"}
resdfu <- plotsim_standardize(res_indep, alpha = seq(0.01, 0.10, 0.01))

resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
                                   performanceMetric, alpha, value),
                            select(resdfu, rep, blabel, param.alpha, key,
                                   performanceMetric, alpha, value),
                            by = c("rep", "blabel", "param.alpha", "key",
                                   "performanceMetric", "alpha"),
                            suffix = c(".dep", ".indep"))
resdfiu <- dplyr::mutate(resdfiu, value = value.dep - value.indep)

plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE) +
    ylab("rejections (dep - indep)")

plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE) +
    scale_y_continuous("FDR (dep - indep)", labels = scales::percent)

plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE) +
    scale_y_continuous("TPR (dep - indep)", labels = scales::percent)

plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE) +
    scale_y_continuous("TNR (dep - indep)", labels = scales::percent)
```

## Multipanel Plot

Here, we create a multipanel plot to consolidate the results to a single figure.

```{r weak-4-panels, fig.width = 11, fig.height = 7.5}
gp_wA <- plotsim_average(resdf, met="FDR", filter_set = excludeSet,
                         merge_ihw = TRUE, errorBars = TRUE) 
gp_wB <- plotsim_average(resdf, met="TPR", filter_set = excludeSet,
                         merge_ihw = TRUE, errorBars = TRUE) 
gp_wC <- plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
                         merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE) +
    scale_y_continuous("FDR (dep - indep)", labels = scales::percent)
gp_wD <- plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
                         merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE) +
    scale_y_continuous("TPR (dep - indep)", labels = scales::percent)

gp_w <- plot_grid(gp_wA + theme(legend.position = "none"),
                  gp_wB + theme(legend.position = "none"),
                  gp_wC + theme(legend.position = "none"),
                  gp_wD + theme(legend.position = "none"),
                  labels = LETTERS[1:4], nrow = 2, ncol = 2)
gp_w <- plot_grid(gp_w, get_legend(gp_wA),
                  rel_widths = c(1, .2))
gp_w
```

# Stronger Dependence

Next, we take a look at the results with stronger dependence.

```{r strong-run-simulation}
if (file.exists(strong_file)) {
    res <- readRDS(strong_file)
} else {
    res <- mclapply(X = 1:B, FUN = simNullDependent, bench = bdplus, m = m,
                    pi0 = pi0, es_dist = es_dist, icovariate = icovariate,
                    ts_dist = ts_dist, null_dist = null_dist, null_inv = null_inv,
                    null_dependence = 0.7,
                    seed = seed, mc.cores = cores)
    saveRDS(res, file = strong_file)
}

res_dep <- lapply(res, `[[`, "null_dependent")
res_indep <- lapply(res, `[[`, "null_independent")
```

## Covariate Diagnostics

Here, we show the relationship between the independent covariate and p-values for a
single replication of the experiment. Unlike the plots above, these include both
null and non-null tests.

```{r strong-one-simulation}
onerun <- simNullDependent(bdplus, m = m, pi0 = pi0, es_dist = es_dist, ts_dist = ts_dist,
                           icovariate = icovariate, null_dist = null_dist, null_inv = null_inv,
                           null_dependence = 0.3, execute = FALSE)
```

```{r, strong-diag-scatter, results = "hide", fig.width=4.5, fig.height=3.5}
rank_scatter(onerun, pvalue = "pval", covariate = "ind_covariate")
```

```{r, strong-diag-hist, results = "hide", fig.width=10, fig.height=4}
strat_hist(onerun, pvalue = "pval", covariate = "ind_covariate", maxy = 7, numQ = 5)
```

## Benchmark Metrics

We plot the averaged results across `r B` replications.

```{r strong-metrics-averages, results = "hide"}
resdf <- plotsim_standardize(res_dep, alpha = seq(0.01, 0.10, 0.01))

plotsim_average(resdf, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE)

plotsim_average(resdf, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE) 
```

We also take a look at the distribution of rejects for each method as a function of
the effect size and independent covariate.

```{r strong-metrics-covlineplot, results = "hide"}
covariateLinePlot(res_dep, alpha = ualpha, covname = "effect_size")

covariateLinePlot(res_dep, alpha = ualpha, covname = "ind_covariate")
```

We also look at the FDR as a function of the independent covariate.

```{r strong-metrics-covlineplotFDR, results = "hide"}
covariateLinePlot(res_dep, alpha = ualpha, covname = "ind_covariate", metric = "FDR")
```

Finally, (if enough methods produce rejections at `r ualpha`) we take a look at
the overlap of rejections between methods.

```{r strong-metrics-upset, results = "hide"}
if (numberMethodsReject(resdf, alphacutoff = ualpha, filterSet = excludeSet) >= 3) {
    aggupset(res_dep, alpha = ualpha, supplementary = FALSE, return_list = FALSE)
} else {
    message("Not enough methods found rejections at alpha ", ualpha, 
            "; skipping upset plot")
}
```

We also compare the simulation results with and without dependence between the covariate
under the nul.

```{r strong-metrics-differences, results = "hide"}
resdfu <- plotsim_standardize(res_indep, alpha = seq(0.01, 0.10, 0.01))

resdfiu <- dplyr::full_join(select(resdf, rep, blabel, param.alpha, key,
                                   performanceMetric, alpha, value),
                            select(resdfu, rep, blabel, param.alpha, key,
                                   performanceMetric, alpha, value),
                            by = c("rep", "blabel", "param.alpha", "key",
                                   "performanceMetric", "alpha"),
                            suffix = c(".dep", ".indep"))
resdfiu <- dplyr::mutate(resdfiu, value = value.dep - value.indep)

plotsim_average(resdfiu, met="rejections", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE) +
    ylab("rejections (dep - indep)")

plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE) +
    scale_y_continuous("FDR (dep - indep)", labels = scales::percent)

plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE) +
    scale_y_continuous("TPR (dep - indep)", labels = scales::percent)

plotsim_average(resdfiu, met="TNR", filter_set = excludeSet,
                merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE) +
    scale_y_continuous("TNR (dep - indep)", labels = scales::percent)
```

## Multipanel Plot

Here, we create a multipanel plot to consolidate the results to a single figure.

```{r strong-4-panels, fig.width = 11, fig.height = 7.5}
gp_sA <- plotsim_average(resdf, met="FDR", filter_set = excludeSet,
                         merge_ihw = TRUE, errorBars = TRUE) 
gp_sB <- plotsim_average(resdf, met="TPR", filter_set = excludeSet,
                         merge_ihw = TRUE, errorBars = TRUE) 
gp_sC <- plotsim_average(resdfiu, met="FDR", filter_set = excludeSet,
                         merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE) +
    scale_y_continuous("FDR (dep - indep)", labels = scales::percent)
gp_sD <- plotsim_average(resdfiu, met="TPR", filter_set = excludeSet,
                         merge_ihw = TRUE, errorBars = TRUE, diffplot = TRUE) +
    scale_y_continuous("TPR (dep - indep)", labels = scales::percent)

gp_s <- plot_grid(gp_sA + theme(legend.position = "none"),
                  gp_sB + theme(legend.position = "none"),
                  gp_sC + theme(legend.position = "none"),
                  gp_sD + theme(legend.position = "none"),
                  labels = LETTERS[1:4], nrow = 2, ncol = 2)
gp_s <- plot_grid(gp_s, get_legend(gp_sA),
                  rel_widths = c(1, .2))
gp_s
```

# Session Info

```{r}
sessionInfo()
```